---
title: "Governance"
author: "Rhett D Harrison"
format:
        html:
                css: custom.css
        pdf: default
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| #| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ggwordcloud)

load("cfmg.RData")
```

## Summary

The implementation of community forest management processes is patchy. The legal processes required for establishment of a CF are mostly complete across our sample, but later processes such as the registration certificate, forest-use permits and benefit-sharing agreement have not been completed for a substantial number of CFs. While most CFMGs have held elections, have term-limits for executive officers and hold meetings with reasonable frequency, a large proportion have not opened a bank account, do not have a receipt book and have not completed annual financial and technical reports. Most CFMGs are overdue elections and hence are not compliant with their Bylaws & Constitutions. Business processes are even less well implemented. Many do not have a business plan, annual budgets or workplans. Many also do not have defined user groups. Most CFMGs are only rarely visited by the District Forest Officers, but most are supported by conservation NGOs or carbon trading enterprises, who do visit more frequently. Interest in expanding the CF, an indicator of satisfaction with CFM, was highest among CFMG executives and lowest among community youth, a difference that was significant. Among individual interview respondents, NGO representatives indicated they would like to expand the CF significantly more often than traditional leaders (usually headmen or headwomen) and community members. For those CFMGs that recognised forest user groups, these user-groups harvested a very wide variety of resources, but honey and mushrooms were particularly frequently mentioned.

## Introduction

In this chapter, we consider questions concerning the governance and management of CFMGs. As our questions pertain to the running of the CFMGs, our analysis focuses on the responses of the CFMG Executives. However, the responses from other FGDs and from the individual interviews are included for comparison where appropriate.

## CFMG legal processes

We asked CFMG executives what required legal processes had been completed, including the recognition letter, forest management agreement, forest management plan, community forest registration certificate, forest user-rights permit, CFMG bylaws & constitution, and benefit sharing agreement (@tbl-legal_processes).

```{r}
#| label: tbl-legal_processes
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "CFMG completed legal processes by province."

cfmg |> filter(interview_type == "cfmg_executive") |>
        group_by(province) |>
        summarise(
                `CFMG recongnition` = mean(
                        cfmg_governance_legal_documents_recognition_letter,
                        na.rm = TRUE),
                `CF agreement` = mean(
                        cfmg_governance_legal_documents_community_forest_management_agreement,
                        na.rm = TRUE),
                `Management plan` = mean(
                        cfmg_governance_legal_documents_forest_management_plan,
                        na.rm = TRUE),
                `Registration certificate` = mean(
                        cfmg_governance_legal_documents_community_forest_registration_certificate,
                        na.rm = TRUE),
                `User-rights permit` = mean(
                        cfmg_governance_legal_documents_forest_user_rights_license_permit,
                        na.rm = TRUE),
                `Bylaws & constitution` = mean(
                        cfmg_governance_legal_documents_bylaws_and_constitution,
                        na.rm = TRUE),
                `Benefit sharing agreement` = mean(
                        cfmg_governance_legal_documents_benefit_sharing_agreement,
                        na.rm = TRUE)
        ) |>
        gt() |>
        fmt_percent(columns = -province, decimals = 0) |>
        tab_header(title = "CFMG legal processes",
                   subtitle = "Values represent the % of CFMGs with each legal document by province. n = 75 CFMGs.") |>
        opt_align_table_header(align = "left") |>
        cols_align(align = "center", columns = -province) |>
        cols_align(align = "left", columns = province) |>
        tab_style(
          style     = cell_text(align = "center"),
          locations = cells_column_labels(columns = province)
        ) |>
        sub_missing()


```

Lusaka Province was the only province in which all processes had been completed across all CFMGs, while Eastern Province had the highest proportion of incomplete legal processes. Overall, completion of legal processes was quite patchy. Some processes, such as the Recognition letter, CF Agreement, Management Plan and Bylaws & Constitution seem to have been completed for most CFMGs, but the Registration Certificate, User-rights Permit and Benefit-sharing Agreement were less often completed. This assessment is based on the responses of the CFMG Executives, who should be informed. Nevertheless, it may be that more processes have been completed, especially where the CFMG is being supported by external agencies such as NGOs or carbon trading agencies.

## CFMG governance processes

Through the survey, we asked CFMG executives what governance processes were being implemented, such as holding CFMG elections, issuing permits and annual reporting (@tbl-governance_processes).

```{r}
#| label: tbl-governance_processes
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "CFMG governance processes conducted by province"

cfmg |> filter(interview_type == "cfmg_executive") |>
        mutate(
               cfmg_elections = case_match(
                       cfmg_governance_cfmg_held_election,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_term_limits = case_match(
                       cfmg_governance_term_limits_exist,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_HFO = case_match(
                       cfmg_governance_yn_employ_honorary_forest_officers,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_receipt_book = case_match(
                       cfmg_governance_receipt_book_maintained,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_bank_acc = case_match(
                       cfmg_governance_bank_account,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_financial_rep = case_match(
                       cfmg_governance_financial_reports_to_community,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_annual_rep = case_match(
                       cfmg_governance_annual_report_to_community,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_permits = case_match(
                       cfmg_governance_issue_licenses_permits,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_)

        ) |>
        group_by(province) |>
        summarise(
                `CFMG elections` = mean(
                        cfmg_elections,
                        na.rm = TRUE),
                `CFMG term-limits` = mean(
                        cfmg_term_limits,
                        na.rm = TRUE),
                `Receipt book` = mean(
                        cfmg_receipt_book,
                        na.rm = TRUE),
                `Bank acc` = mean(
                        cfmg_bank_acc,
                        na.rm = TRUE),
                `Permits issued` = mean(
                        cfmg_permits,
                        na.rm = TRUE),
                `Financial reporting` = mean(
                        cfmg_financial_rep,
                        na.rm = TRUE),
                `Annual reporting` = mean(
                        cfmg_annual_rep,
                        na.rm = TRUE),
                `HFOs appointed` = mean(
                        cfmg_HFO,
                        na.rm = TRUE)
        ) |>
        gt() |>
        fmt_percent(columns = -province, decimals = 0) |>
        tab_header(title = "CFMG governance processes",
                   subtitle = "Values represent the % of CFMGs completing/conducting each governance process by province. n = 75 CFMGs.") |>
        opt_align_table_header(align = "left") |>
        cols_align(align = "center", columns = -province) |>
        cols_align(align = "left", columns = province) |>
        tab_style(
          style     = cell_text(align = "center"),
          locations = cells_column_labels(columns = province)
        ) |>
        sub_missing()


```

While most CFMGs have held elections, have term limits for executive officers and employ honorary forest officers (HFOs), other governance processes were not as consistently implemented. Most CFMGs have not opened bank accounts, do not maintain a receipt book, or issue permits. Financial and annual reporting to community members was also apparently irregularly implemented.

It appears that most CFMGs have not held elections since they were formed, up to five years ago (@fig-last_election).

```{r}
#| label: fig-last_election
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Number of years since the last CFMG executive election. The length of each bar segment represents the proportion of CFMGs in that category. The numbers superimposed on the bars represent the actual number of CFMGs in each category."
#| fig-height: 2.5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    election = case_match(
      cfmg_governance_last_cfmg_election_when,
      "Less than a month"     ~ "<1 year ago",
      "less_than_1_year_ago"  ~ "<1 year ago",
      "1_2_years_ago"         ~ "1 to 2 years ago",
      "2 years"               ~ "1 to 2 years ago",
      "2 years ago"           ~ "1 to 2 years ago",
      "3_5_years_ago"         ~ "3 to 5 years ago",
      "3 years"               ~ "3 to 5 years ago",
      "3 years ago"           ~ "3 to 5 years ago",
      "4 years ago"           ~ "3 to 5 years ago",
      "more_than_5_years_ago" ~ ">5 years ago",
      .default = NA
    ),
    election = factor(election,
      levels = c(">5 years ago", "3 to 5 years ago", "1 to 2 years ago", "<1 year ago"))
  ) |>
  drop_na(election) |>
  count(election) |>
  mutate(prop = n / sum(n), group = "All CFMGs") |>
  ggplot(aes(y = group, x = prop, fill = election)) +
  geom_col(width = 0.4, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.06, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(add = c(0, 0.02)),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      "<1 year ago"      = "#2A9D8F",
      "1 to 2 years ago" = "#76B7A0",
      "3 to 5 years ago" = "#E07B54",
      ">5 years ago"     = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  scale_y_discrete(expand = expansion(add = c(0.3, 0.3))) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(2, 25, 10, 10)
  )
```

Forty-three out of 75 CFMGs have not held elections for over 3 yrs, although term limits for executive officers are typically 2-3 years (@fig-tenure). This would suggest that most CFMGs are not complying with their Bylaws & Constitutions.

```{r}
#| label: fig-tenure
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Tenure of CFMG executive positions. The length of each bar segment represents the proportion of CFMGs with that tenure length. The numbers superimposed on the bars represent the actual number of CFMGs in each category."
#| fig-height: 2.5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    tenure = case_match(
      cfmg_governance_cfmg_postions_tenure,
      "2_years"  ~ "2 years",
      "3_years"  ~ "3 years",
      "5 time"   ~ "5 years",
      "5years"   ~ "5 years",
      "6 years"  ~ ">5 years",
      "7 years"  ~ ">5 years",
      "dont_know" ~ NA,
      .default   = NA
    ),
    tenure = factor(tenure, levels = c(">5 years", "5 years", "3 years", "2 years"))
  ) |>
  drop_na(tenure) |>
  count(tenure) |>
  mutate(prop = n / sum(n), group = "All CFMGs") |>
  ggplot(aes(y = group, x = prop, fill = tenure)) +
  geom_col(width = 0.4, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.06, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(add = c(0, 0.02)),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      "2 years"  = "#2A9D8F",
      "3 years"  = "#76B7A0",
      "5 years"  = "#E07B54",
      ">5 years" = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  scale_y_discrete(expand = expansion(add = c(0.3, 0.3))) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(2, 25, 10, 10)
  )
```

In contrast, most CFMGs do appear to be holding meetings regularly (@fig-meetings). However, there is a small proportion of CFMGs that holds meetings rarely or never.

```{r}
#| label: fig-meetings
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Number of CFMG meetings per year. The length of each bar segment represents the proportion of CFMGs with the specified meeting frequency. The numbers superimposed on the bars represent the actual number of CFMGs in each category."
#| fig-height: 2.5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    meetings = case_match(
      cfmg_governance_cfmg_meetings,
      "Bi monthly"                                                            ~ ">4 yr-1",
      "monthly"                                                               ~ ">4 yr-1",
      "Quarterly"                                                             ~ "4 yr-1",
      "quarterly"                                                             ~ "4 yr-1",
      "yearly"                                                                ~ "1 yr-1",
      "only_when_needed_adhoc_meetings "                                      ~ "rarely",
      "Due to lack of support the Executive does not meet often"              ~ "rarely",
      "Quarterly but the executive has not had meetings in the last 2 years"  ~ "never",
      "Quarterly but they don't meet"                                         ~ "never",
      .default = NA
    ),
    meetings = factor(meetings,
      levels = c("never", "rarely", "1 yr-1", "4 yr-1", ">4 yr-1"))
  ) |>
  drop_na(meetings) |>
  count(meetings) |>
  mutate(prop = n / sum(n), group = "All CFMGs") |>
  ggplot(aes(y = group, x = prop, fill = meetings)) +
  geom_col(width = 0.4, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.06, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(add = c(0, 0.02)),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      ">4 yr-1" = "#2A9D8F",
      "4 yr-1"  = "#76B7A0",
      "1 yr-1"  = "#F0C040",
      "rarely"  = "#E07B54",
      "never"   = "#C0392B"
    ),
    labels = c(
      ">4 yr-1" = expression(">4 yr"^{-1}),
      "4 yr-1"  = expression("4 yr"^{-1}),
      "1 yr-1"  = expression("1 yr"^{-1}),
      "rarely"  = "rarely",
      "never"   = "never"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  scale_y_discrete(expand = expansion(add = c(0.3, 0.3))) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(2, 25, 10, 10)
  )
```

## CFMG business processes

As with the legal processes, we asked CFMG Executives about the state of their business processes. Specifically, we asked if they had developed annual workplans and budgets, and whether these plans were being implemented. We also asked if they had developed a business plan and if any forest user groups had been established (@tbl-business_processes).

```{r}
#| label: tbl-business_processes
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "CFMG completed business process by province."

cfmg |> filter(interview_type == "cfmg_executive") |>
        mutate(
               cfmg_workplan = case_match(
                       cfmg_governance_cfmg_workplan,
                       "yes" ~ 1,
                       "no" ~ 0,
                       "don't_know" ~ NA),
               cfmg_budget = case_match(
                       cfmg_governance_cfmg_budget,
                       "yes" ~ 1,
                       "no" ~ 0,
                       "don't_know" ~ NA),
               cfmg_business_plan = case_match(
                       cfmg_governance_cfma_business_plan,
                       "yes" ~ 1,
                       "no" ~ 0,
                       "don't_know" ~ NA),
               cfmg_user_groups = case_match(
                       cfmg_governance_any_user_groups,
                       "yes" ~ 1,
                       "no" ~ 0,
                       "don't_know" ~ NA)
        ) |>
        group_by(province) |>
        summarise(
                `CFMG workplan` = mean(
                        cfmg_workplan,
                        na.rm = TRUE),
                `CFMG budget` = mean(
                        cfmg_budget,
                        na.rm = TRUE),
                `Implementing plan` = mean(
                        cfmg_governance_implementing_plans_cfmg_executive,
                        na.rm = TRUE),
                `Business plan` = mean(
                        cfmg_business_plan,
                        na.rm = TRUE),
                `User groups` = mean(
                        cfmg_user_groups,
                        na.rm = TRUE)
        ) |>
        mutate(across(where(is.numeric), ~ifelse(is.nan(.x), NA_real_, .x))) |>
        gt() |>
        fmt_percent(columns = -province, decimals = 0) |>
        tab_header(title = "Business processes in the management of CFMGs",
                   subtitle = "Values represent the % of CFMGs with each business process in place by province. n = 75 CFMGs.") |>
        opt_align_table_header(align = "left") |>
        cols_align(align = "center", columns = -province) |>
        cols_align(align = "left", columns = province) |>
        tab_style(
          style     = cell_text(align = "center"),
          locations = cells_column_labels(columns = province)
        ) |>
        sub_missing()


```

As is evident from the table, these business management processes were even more erratically completed than the legal and governance processes. As substantial proportion of CFMGs did not have workplans, budgets or a business plan. More surprising, many did not have any recognised user groups, which is critical to managing forest resources.

## External support

All CFMGs are supported by the Forestry Deparment, which has a statutory duty to implement the community forest policy and regulations. Nonetheless, the frequency of visits by District Forest Officers (DFOs) varied widely and in many cases they rarely or never visited.

```{r}
#| label: fig-dfo_visits
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Frequency of visits by District Forest Officers (DFO) by province. The length of the bars represents the proportion of CFMGs with a specified frequency of visits. The numbers superimposed on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    dfo = case_match(dfo_visits,
      "never"        ~ "Never",
      "rarely"       ~ "Rarely",
      "occasionally" ~ "Occasionally",
      "often"        ~ "Often",
      "very_often"   ~ "Very often",
      .default       = NA
    ),
    dfo = factor(dfo, levels = c("Very often", "Often", "Occasionally", "Rarely", "Never"))
  ) |>
  drop_na(dfo) |>
  count(province, dfo) |>
  group_by(province) |>
  mutate(
    prop       = n / sum(n),
    mean_score = sum(as.numeric(dfo) * prop)
  ) |>
  ungroup() |>
  ggplot(aes(y = reorder(province, mean_score), x = prop, fill = dfo)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.08, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      "Never"        = "#C0392B",
      "Rarely"       = "#E07B54",
      "Occasionally" = "#F0C040",
      "Often"        = "#76B7A0",
      "Very often"   = "#2A9D8F"
    )
  ) +
        guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Many CFMGs also operate in partnership with an NGO, such as a conservation organisation, or a private enterprise, such as a carbon trader (@fig-support).

```{r}
#| label: fig-support
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "CFMGs partnered with an external organisation, such as a conservation NGO or carbon enterprise, by province. The length of the bars represents the number of CFMGs either partnered (Yes) or not partnered (No) with an external organisation. The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(cfmg_governance_yn_institutions_support_cfmg, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

The table below (@tbl-support_organisations) lists the partners organisations and the number of CFMGs they support. It is important to note that this list does not cover all organisations supporting CFMGs or all the CFMGs that are supported by a named organisation, but only those in our sample of 75 CFMGs

```{r}
#| label: tbl-support_organisations
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Organisations partnering with CFMGs"

cfmg |> filter(interview_type == "cfmg_executive") |>
        mutate(
               `CFMG partner` = cfmg_governance_support_institutions_quests_supportinstitutionsdetails_1__institution_partner_name,
               `CFMG partner` = case_match(
                       `CFMG partner`,
                       "EP-JSLP" ~ "Forestry Department",
                       "Community Climate Solutions Zambia" ~ "Community Climate Solutions",
                       .default = `CFMG partner`
               )
               ) |>
        count(`CFMG partner`) |>
        gt() |>
        tab_header(title = "Partner organisations",
                   subtitle = "n = number of CFMGs partnered with each organisation.") |>
        opt_align_table_header(align = "left") |>
        cols_align(align = "center", columns = -`CFMG partner`) |>
        cols_align(align = "left", columns = `CFMG partner`) |>
        tab_style(
          style     = cell_text(align = "center"),
          locations = cells_column_labels(columns = `CFMG partner`)
        ) |>
        sub_missing()
```

The figure below (@fig-partner_visits) indicates the frequence of visits to CFMGs by partner organiastions. The respondents did not state a specified number of visits per year, but from the context we assume the frequently indicates monthly or every other month, 1-4 times a year, while never means that they had not visited in over a year.

```{r}
#| label: fig-partner_visits
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Frequency of visits by the supporting organisation. The length of the bars represents the proportion of CFMGs with a specified frequency of visits. The numbers superimposed on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
          `Support freq` = cfmg_governance_support_institutions_quests_supportinstitutionsdetails_1__times_institutions_visit_cfmg,
          `Support freq` = case_match(`Support freq`,
                                         "never_0_times"        ~ "Never",
                                         "twice_2_times" ~ "Occasionally",
                                         "thrice_3_times" ~ "Occasionally",
                                         "once_1_time" ~ "Occasionally",
                                         "more_than_5_times" ~ "Occasionally",
                                         "more_than_3_times" ~ "Occasionally",
                                         "More than 3 times" ~ "Occasionally",
                                         "More 3" ~ "Occasionally",
                                         "Dependent on their activities" ~ "Occasionally",
                                         "8_times" ~ "Occasionally",
                                         "4_times" ~ "Occasionally",
                                         "12_time (at_least_every_month)" ~ "Frequently",
                                         "12_times (every_month)" ~ "Frequently",
                                         "36_times" ~ "Frequently",
                                         "COMACO 15 to 20 times per year" ~ "Frequently",
                                         "Very often" ~ "Frequently",
                                         "Very often/regularly. They have a Field Officer living within the community" ~ "Frequently",
                                         "Very_often" ~ "Frequently",
                                         "more_than_12_times" ~ "Frequently",
                                         "several/often_times" ~ "Frequently",
                                         "very_often" ~ "Frequently",
                                         "very_often (60_meetings_per_quater)" ~ "Frequently",
                                         "very_often (they_have_a_lot_of_activities_with_the_cfmg)" ~ "Frequently",
                                         "very_often (they_have_an_officer_within_the_community)" ~ "Frequently",
                                         .default = NA),
    `Support freq` = factor(`Support freq`, levels = c("Frequently", "Occasionally", "Never")),
    `CFMG partner` = cfmg_governance_support_institutions_quests_supportinstitutionsdetails_1__institution_partner_name,
               `CFMG partner` = case_match(
                       `CFMG partner`,
                       "EP-JSLP" ~ "Forestry Department",
                       "Community Climate Solutions Zambia" ~ "Community Climate Solutions",
                       .default = `CFMG partner`
               )
  ) |>
  drop_na(`Support freq`) |>
        drop_na(`CFMG partner`) |>
  count(`CFMG partner`, `Support freq`) |>
  group_by(`CFMG partner`) |>
  mutate(
    prop       = n / sum(n),
    mean_score = sum(as.numeric(`Support freq`) * prop)
  ) |>
  ungroup() |>
  ggplot(aes(y = reorder(`CFMG partner`, mean_score), x = prop, fill = `Support freq`)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.08, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      "Never"        = "#C0392B",
      "Occasionally" = "#F0C040",
      "Frequently"   = "#2A9D8F"
    )
  ) +
        guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

## Interest in expanding the community forest

Next we asked interviewees whether they were interested in expanding the area of their community forest. A "yes" answer can be taken as an endorsement of community forest management (CFM). On the other hand, a negative response might reflect dissatisfaction with the CFMG management or the benefits CFM has brought. However, a negative response could also reflect other reasons, such as a lack of sufficient addition forest area in the community. We compare the responses from different FGD groups.

```{r}
#| label: fig-interest_in_expanding_FGDs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Interest in expanding the community forest by FGD type. The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    expand_cfma = substr(cfmg_governance_interest_expanding_cfma, 1, 1),
    expand_cfma = case_match(
      expand_cfma,
      "y" ~ "Yes",
      "n" ~ "No",
      "d" ~ "Don't know"
    ),
    expand_cfma = factor(expand_cfma, levels = c("Yes", "Don't know", "No")),
    interview_type = case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    )
  ) |>
  filter(!is.na(expand_cfma)) |>
  count(interview_type, expand_cfma) |>
  group_by(interview_type) |>
  mutate(
    prop       = n / sum(n),
    mean_score = sum(as.numeric(expand_cfma) * prop)
  ) |>
  ungroup() |>
  ggplot(aes(y = reorder(interview_type, -mean_score), x = prop, fill = expand_cfma)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "No"         = "#C0392B",
      "Don't know" = "#F0C040",
      "Yes"        = "#2A9D8F"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen in @fig-interest_in_expanding_FGDs, the CFMG executives were enthusiastic to expand the CF in most cases. However, among the community this number dropped to 41 for female groups, 37 for male groups and to just 23 for youth groups.

The difference in attitudes between CFMG Executive and Youth FGD groups was significant, but the pairwise comparisons among other FGD groups were not significant. These results might suggest that community forest management is not benefiting youth to the same degree as other groups, or perhaps that they bear higher opportunity costs. Our analyses in other sections will help elucidate the reasons.

::: {.callout-note collapse="true"}
### Interest in expanding community forest - FGDs - Model methods

To examine the differences among FGD types, we employed a binomial GLMM (i.e. yes = want to expand, no = do not want to expand), with random intercepts for province and CFMG nested within province. We used FDG type as the factor. We examined the significance of pairwise differences among FDG types using Tukey HSD.
:::

```{r}
#| label: model_expand_cfma_fgd
#| include: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |> filter(interview_type != "individual_interview") |>
        mutate(
                expand_cfma = substr(
                        cfmg_governance_interest_expanding_cfma, 1,1),
                expand_cfma = case_match(
                        expand_cfma,
                        "y" ~ "Yes",
                        "n" ~ "No",
                        "d" ~ "Don't know"
                        ),
                expand_cfma = factor(expand_cfma,
                         levels = c("Yes", "No", "Don't know")
                         )
        ) |>
        mutate(
                expand_cfma = if_else(expand_cfma == "Don't know", NA, expand_cfma)
        )

M1 <- glmer(expand_cfma ~
                    interview_type +
                    (1 | province/cfmg_name),
            family = binomial(link = "logit"),
            data = test.data
)

emm.M1 <- emmeans(M1, specs = ~ interview_type)
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: M1_expand_cfma_FDG
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

summary(M1)

```

```{r}
#| label: pairwise_expand_cfma_FDG
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

summary(pairwise.M1)

```

We also compared the responses from the individual interviewees (@tbl-interest_in_expanding_individual). The traditional leaders and community members more often reported that they did not want to expand the CF than other groups. However, the only pairwise comparisons that were significant were NGO representative with traditional leader and NGO representative with community member.

```{r}
#| label: tbl-interest_in_expanding_individual
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Interest in expanding community forest by FDG type."

cfmg |> filter(interview_type == "individual_interview") |>
  mutate(
    expand_cfma = substr(cfmg_governance_interest_expanding_cfma, 1, 1),
    expand_cfma = case_match(expand_cfma, "y"~"Yes", "n"~"No", "d"~"Don't know"),
    expand_cfma = factor(expand_cfma, levels = c("Yes", "No", "Don't know")),
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_exec_member" ~ "CFMG executive member",
      "cfmg_member"      ~ "Community member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "CFMG executive member", "Community member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    ))
  ) |>
  filter(!is.na(expand_cfma)) |>
  count(interviewee_role, expand_cfma, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, expand_cfma, prop) |>
  pivot_wider(names_from = expand_cfma, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Yes, No, `Don't know`), decimals = 0) |>
  tab_header(title = "Interest in expanding the community forest area by interviewee role",
             subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses.") |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()

```

::: {.callout-note collapse="true"}
### Interest in expanding community forest - Individual interviewees - Model methods

To examine differences among individual interviewees in their stated interest in expanding their community forest, we employed a binomial GLMM (i.e. yes = want to expand, no = do not want to expand) and included interviewee role, interviewee gender and interviewee age as factors. We specified random intercepts for province and cfmg nested within province. To examine the pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: test_data_expand_cfma_individual
#| echo: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |> filter(interview_type == "individual_interview") |>
        mutate(
                expand_cfma = substr(
                        cfmg_governance_interest_expanding_cfma, 1,1),
                expand_cfma = case_match(
                        expand_cfma,
                        "y" ~ "Yes",
                        "n" ~ "No",
                        "d" ~ "Don't know"
                        ),
                expand_cfma = factor(expand_cfma,
                         levels = c("Yes", "No", "Don't know")
                         )
        ) |>
        mutate(
                expand_cfma = if_else(expand_cfma == "Don't know", NA, expand_cfma),
                interviewee_gender = demo_information_interviewee_gender,
                interviewee_age = demo_information_interviewee_age
        )

M2 <- glmer(expand_cfma ~
                    interviewee_role +
                    interviewee_gender +
                    interviewee_age +
                    (1 | province/cfmg_name),
            family = binomial(link = "logit"),
            data = test.data
)

emm.M2 <- emmeans(M2, specs = ~ interviewee_role)
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")

```

```{r}
#| label: M2_expand_cfma_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

summary(M2)

```

```{r}
#| label: pairwise_expand_cfma_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

summary(pairwise.M2)

```

## Defined user groups

Among those CFMGs that had recognised user groups, they were involved in managing the following resources (@fig-gov_user_groups).

```{r}
#| label: fig-gov_user_groups
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Forest resources managed by user groups within CFMGs, as reported by CFMG executives. Word size is proportional to the number of CFMGs reporting a user-group for that resource type."
#| fig-width: 10
#| fig-height: 7

prefix <- "cfmg_governance_forest_resources_managed_usergroups_"

resource_labels <- c(
  "timber"           = "Timber",
  "poles"            = "Poles",
  "firewood"         = "Firewood",
  "charcoal"         = "Charcoal",
  "honey"            = "Honey",
  "mushrooms"        = "Mushrooms",
  "wild_fruits"      = "Wild fruit",
  "caterpillars"     = "Caterpillars",
  "chikanda"         = "Chikanda",
  "munkoyo"          = "Munkoyo",
  "lusala_busala"    = "Lusala busala",
  "mungongo"         = "Mungongo",
  "devils_claw"      = "Devil's claw",
  "fish"             = "Fish",
  "pupwe"            = "Pupwe",
  "brooms"           = "Brooms",
  "bamboo"           = "Bamboo",
  "medicinal_plants" = "Medicinal plants",
  "bushmeat"         = "Wild meat",
  "resins_gums"      = "Resins",
  "thatching_grass"  = "Thatching",
  "fibers"           = "Fibres",
  "eco_tourism"      = "Eco-tourism"
)

full_var_names <- paste0(prefix, names(resource_labels))

wc_data <- cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(full_var_names)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = resource_labels[sub(prefix, "", variable, fixed = TRUE)]) |>
  filter(freq > 0)

set.seed(42)
ggplot(wc_data, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 30)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

Evidently, people are harvesting a wide variety of products from CFs, although honey and mushrooms appear especially important. It is not clear from the responses, if these user-groups are defined within the CF management plan or whether these are informal groups that are nevertheless recognised by the CFMG.

## Discussion

In this chapter we assessed CFMG governance across five domains: legal processes, governance processes, business processes, external support, and community buy-in as reflected by interest in expanding the CF.

The most consistent pattern across all three process domains — legal, governance, and business — is that foundational steps have been taken by most CFMGs, but the more demanding, ongoing requirements are often incomplete. Early legal steps, such as obtaining a recognition letter and signing a forest management agreement, have been completed by the large majority of CFMGs. But later and more technically involved documents — the registration certificate, forest-use permit, and benefit-sharing agreement — are absent in a substantial proportion. Similarly in governance, most CFMGs have held elections and have term limits in place, but basic financial management infrastructure — bank accounts, receipt books, financial and annual reporting — is frequently missing. Business processes are the weakest area of all: workplans, budgets and business plans are not in place in many CFMGs. This is a critical gap. Without these planning and financial management tools, CFMGs cannot systematically manage their resources, are poorly positioned to account for and redistribute income — including carbon payments — and will struggle to demonstrate compliance to the Forestry Department or to external partners.

The fact that most CFMGs are overdue for elections is particularly concerning. Timely elections are a basic requirement of the Bylaws & Constitution, and overdue elections undermine the democratic legitimacy of executive committees and could contribute to the perception of elite capture noted elsewhere in this report. Ensuring that CFMGs run elections on schedule should be a priority for the Forestry Department and for supporting NGOs.

External support from the Forestry Department is insufficient. Most CFMGs are rarely or never visited by DFOs, despite the statutory duty of the Forestry Department to implement the community forest policy. This gap is partly compensated by NGOs and carbon trading enterprises, many of which visit their partner CFMGs more frequently. However, this support is not universal: a significant proportion of CFMGs in our sample had no external partner at all. Given the governance deficits identified above, regular external support for capacity building and oversight is not a luxury but a necessity. The current model, in which CFM is effectively co-governed by the private sector rather than the state, may be adequate for CFMGs with active and capable partners but leaves others without meaningful support.

Interest in expanding the CF is highest among CFMG executives and drops markedly among community members, particularly youth. Among individual interviewees, NGO representatives are significantly more enthusiastic about expansion than traditional leaders or community members. This divergence — consistent with the equity chapter findings — suggests that the leadership and external partner class has a more optimistic view of CFM outcomes than the broader community. Community-level dissatisfaction, if unaddressed, poses a long-term risk to the legitimacy and sustainability of CFM.

In conclusion, CFMG governance is still maturing. The legal framework for CFM is largely in place, but institutional and business management processes lag behind. Strengthening financial management, ensuring timely elections, expanding DFO engagement, and closing the gap between executive and community perspectives on CFM will be essential for CFM to fulfil its potential in Zambia.
