---
title: "Governance"
format:
  html:
    css: custom.css
  pdf:
    include-in-header:
      text: |
        \pagenumbering{arabic}
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| #| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ggwordcloud)

load("cfmg.RData")
```

## Summary

The implementation of CFM processes is patchy. The legal processes required for establishment of a CFA were mostly complete across our sample, but later processes such as the registration certificate, forest-use permits and benefit-sharing agreement have not been completed for a substantial number of CFAs. While most CFMGs have held elections, have term-limits for executive officers and hold meetings with reasonable frequency, a large proportion have not opened a bank account, do not have a receipt book and have not completed annual financial and technical reports. Most CFMGs are overdue elections and hence are not compliant with their By-laws & Constitutions. This is particularly worrying as, in some cases, it may reflect an attempt by CFMG executives to retain control for personal benefit. Business processes are even less well implemented. Many do not have a business plan, annual budgets or workplans. Many also do not have defined user groups or issue permits. Most CFMGs are only rarely visited by district forestry staff, but most are supported by conservation NGOs or carbon developers, who do visit more frequently. The frequency of visits by district forestry staff was significantly associated with a higher governance index score. Interest in expanding the CFA was highest among CFMG executives and lowest among community youth, while among individual interview respondents, private enterprise representatives indicated they would like to expand the CFA significantly more often than traditional leaders (usually headmen or headwomen) and community members. The governance index was a highly significant indicator of community interest in expanding the CFA, an indicator of satisfaction with CFM. Thus, conversely, poor governance constitutes a risk to the sustainability of CFM in Zambia.

::: {.callout-note icon="false"}
## Key conclusions

CFMG governance is patchy. There are some areas of serious concern. The legal framework for CFM is in place across most CFMGs, but institutional and business management processes lag behind. Many CFMGs are not in compliance with their own by-laws and constitutions, with respect to annual reporting and term limits for executive officers. Oversight of CFM is a statutory duty of the FD, but visits from district forestry office staff are often rare to non-existent. Higher visit frequency from district forestry staff was a significant predictor of better governance. Improving governance, primarily through strengthening financial management and CFMG democratic processes by expanding district forestry staff engagement, will be essential for CFM to fulfil its potential in Zambia.
:::

## Introduction

In this chapter, we consider questions concerning the governance and management of CFMGs. As our questions pertain to the running of the CFMGs, our analysis focuses on the responses of the CFMG executives. However, the responses from other FGDs and from the individual interviews are included for comparison where appropriate.

We examine the hypothesis that CFMGs with stronger governance — completing legal requirements, maintaining transparent financial records, holding regular elections, and making decisions transparently — would be more successful at protecting their forests and generating economic benefits for their communities, thereby generating greater community support for CFM. We test this by examining the completion of governance processes, comparing these results across CFMGs, and interrogating community groups as to whether they would support expansion of the CFA, as an indicator of satisfaction with CFM. In later chapters, we assess whether better governance predicts better outcomes for forest management and economic development.

## CFMG legal processes

We asked CFMG executives what required legal processes had been completed, including the recognition letter, forest management agreement, forest management plan, community forest registration certificate, forest user-rights permit, CFMG by-laws & constitution, and benefit sharing agreement (@tbl-legal_processes).

```{r}
#| label: tbl-legal_processes
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "CFMG completed legal processes by province."

cfmg |> filter(interview_type == "cfmg_executive") |>
        group_by(province) |>
        summarise(
                `CFMG recongnition` = mean(
                        cfmg_governance_legal_documents_recognition_letter,
                        na.rm = TRUE),
                `CFA agreement` = mean(
                        cfmg_governance_legal_documents_community_forest_management_agreement,
                        na.rm = TRUE),
                `Management plan` = mean(
                        cfmg_governance_legal_documents_forest_management_plan,
                        na.rm = TRUE),
                `Registration certificate` = mean(
                        cfmg_governance_legal_documents_community_forest_registration_certificate,
                        na.rm = TRUE),
                `User-rights permit` = mean(
                        cfmg_governance_legal_documents_forest_user_rights_license_permit,
                        na.rm = TRUE),
                `Bylaws & constitution` = mean(
                        cfmg_governance_legal_documents_bylaws_and_constitution,
                        na.rm = TRUE),
                `Benefit sharing agreement` = mean(
                        cfmg_governance_legal_documents_benefit_sharing_agreement,
                        na.rm = TRUE)
        ) |>
        gt() |>
        fmt_percent(columns = -province, decimals = 0) |>
        tab_header(title = "CFMG legal processes",
                   subtitle = "Values represent the % of CFMGs with each legal document by province. n = 75 CFMGs.") |>
        opt_align_table_header(align = "left") |>
        cols_align(align = "center", columns = -province) |>
        cols_align(align = "left", columns = province) |>
        tab_style(
          style     = cell_text(align = "center"),
          locations = cells_column_labels(columns = province)
        ) |>
        sub_missing()


```

::: return-to-summary
[← Return to Summary](Summary.html#governance)
:::

Lusaka Province was the only province in which all processes had been completed across all CFMGs, while Eastern Province had the highest proportion of incomplete legal processes. Overall, completion of legal processes was quite patchy. Some processes, such as the Recognition letter, CFA Agreement, Management Plan and By-laws & Constitution, seem to have been completed for most CFMGs, but the Registration Certificate, User-rights Permit and Benefit-sharing Agreement were less often completed. This assessment is based on the responses of the CFMG executives, who should be well informed. Nevertheless, it may be that more processes have been completed, especially where the CFMG is being supported by external partners, such as NGOs or carbon developers.

## CFMG governance processes

Through the survey, we asked CFMG executives what governance processes were being implemented, such as holding CFMG elections, issuing permits and annual reporting (@tbl-governance_processes).

```{r}
#| label: tbl-governance_processes
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "CFMG governance processes conducted by province"

cfmg |> filter(interview_type == "cfmg_executive") |>
        mutate(
               cfmg_elections = case_match(
                       cfmg_governance_cfmg_held_election,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_term_limits = case_match(
                       cfmg_governance_term_limits_exist,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_HFO = case_match(
                       cfmg_governance_yn_employ_honorary_forest_officers,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_receipt_book = case_match(
                       cfmg_governance_receipt_book_maintained,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_bank_acc = case_match(
                       cfmg_governance_bank_account,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_financial_rep = case_match(
                       cfmg_governance_financial_reports_to_community,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_annual_rep = case_match(
                       cfmg_governance_annual_report_to_community,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_),
               cfmg_permits = case_match(
                       cfmg_governance_issue_licenses_permits,
                       "yes" ~ 1,
                       "no" ~ 0,
                       .default = NA_integer_)

        ) |>
        group_by(province) |>
        summarise(
                `CFMG elections` = mean(
                        cfmg_elections,
                        na.rm = TRUE),
                `CFMG term-limits` = mean(
                        cfmg_term_limits,
                        na.rm = TRUE),
                `Receipt book` = mean(
                        cfmg_receipt_book,
                        na.rm = TRUE),
                `Bank acc` = mean(
                        cfmg_bank_acc,
                        na.rm = TRUE),
                `Permits issued` = mean(
                        cfmg_permits,
                        na.rm = TRUE),
                `Financial reporting` = mean(
                        cfmg_financial_rep,
                        na.rm = TRUE),
                `Annual reporting` = mean(
                        cfmg_annual_rep,
                        na.rm = TRUE),
                `HFOs appointed` = mean(
                        cfmg_HFO,
                        na.rm = TRUE)
        ) |>
        gt() |>
        fmt_percent(columns = -province, decimals = 0) |>
        tab_header(title = "CFMG governance processes",
                   subtitle = "Values represent the % of CFMGs completing/conducting each governance process by province. n = 75 CFMGs.") |>
        opt_align_table_header(align = "left") |>
        cols_align(align = "center", columns = -province) |>
        cols_align(align = "left", columns = province) |>
        tab_style(
          style     = cell_text(align = "center"),
          locations = cells_column_labels(columns = province)
        ) |>
        sub_missing()


```

::: return-to-summary
[← Return to Summary](Summary.html#governance)
:::

While most CFMGs have held elections, have term limits for executive officers and employ honorary forest officers (HFOs), other governance processes were not as consistently implemented. Most CFMGs have not opened bank accounts, do not maintain a receipt book, or issue forest-use permits. Financial and annual reporting to community members was also apparently rarely implemented.

It appears that most CFMGs have not held elections since they were formed, up to five years ago (@fig-last_election).

```{r}
#| label: fig-last_election
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Number of years since the last CFMG executive election. The length of each bar segment represents the proportion of CFMGs in that category. The numbers superimposed on the bars represent the actual number of CFMGs in each category."
#| fig-height: 2.5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    election = case_match(
      cfmg_governance_last_cfmg_election_when,
      "Less than a month"     ~ "<1 year ago",
      "less_than_1_year_ago"  ~ "<1 year ago",
      "1_2_years_ago"         ~ "1 to 2 years ago",
      "2 years"               ~ "1 to 2 years ago",
      "2 years ago"           ~ "1 to 2 years ago",
      "3_5_years_ago"         ~ "3 to 5 years ago",
      "3 years"               ~ "3 to 5 years ago",
      "3 years ago"           ~ "3 to 5 years ago",
      "4 years ago"           ~ "3 to 5 years ago",
      "more_than_5_years_ago" ~ ">5 years ago",
      .default = NA
    ),
    election = factor(election,
      levels = c(">5 years ago", "3 to 5 years ago", "1 to 2 years ago", "<1 year ago"))
  ) |>
  drop_na(election) |>
  count(election) |>
  mutate(prop = n / sum(n), group = "All CFMGs") |>
  ggplot(aes(y = group, x = prop, fill = election)) +
  geom_col(width = 0.4, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.06, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(add = c(0, 0.02)),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      "<1 year ago"      = "#2A9D8F",
      "1 to 2 years ago" = "#76B7A0",
      "3 to 5 years ago" = "#E07B54",
      ">5 years ago"     = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  scale_y_discrete(expand = expansion(add = c(0.3, 0.3))) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(2, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#governance)
:::

Forty-three out of 75 CFMGs have not held elections for over 3 yrs, although term limits for executive officers are typically 2-3 years (@fig-tenure). This indicates that most CFMGs are not in compliance with their By-laws & Constitutions.

```{r}
#| label: fig-tenure
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Tenure of CFMG executive positions. The length of each bar segment represents the proportion of CFMGs with that tenure length. The numbers superimposed on the bars represent the actual number of CFMGs in each category."
#| fig-height: 2.5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    tenure = case_match(
      cfmg_governance_cfmg_postions_tenure,
      "2_years"  ~ "2 years",
      "3_years"  ~ "3 years",
      "5 time"   ~ "5 years",
      "5years"   ~ "5 years",
      "6 years"  ~ ">5 years",
      "7 years"  ~ ">5 years",
      "dont_know" ~ NA,
      .default   = NA
    ),
    tenure = factor(tenure, levels = c(">5 years", "5 years", "3 years", "2 years"))
  ) |>
  drop_na(tenure) |>
  count(tenure) |>
  mutate(prop = n / sum(n), group = "All CFMGs") |>
  ggplot(aes(y = group, x = prop, fill = tenure)) +
  geom_col(width = 0.4, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.06, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(add = c(0, 0.02)),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      "2 years"  = "#2A9D8F",
      "3 years"  = "#76B7A0",
      "5 years"  = "#E07B54",
      ">5 years" = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  scale_y_discrete(expand = expansion(add = c(0.3, 0.3))) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(2, 25, 10, 10)
  )
```

In contrast, most CFMGs do appear to be holding meetings regularly (@fig-meetings). However, a small proportion of CFMGs holds meetings only rarely or never.

```{r}
#| label: fig-meetings
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Number of CFMG meetings per year. The length of each bar segment represents the proportion of CFMGs with the specified meeting frequency. The numbers superimposed on the bars represent the actual number of CFMGs in each category."
#| fig-height: 2.5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    meetings = case_match(
      cfmg_governance_cfmg_meetings,
      "Bi monthly"                                                            ~ ">4 yr-1",
      "monthly"                                                               ~ ">4 yr-1",
      "Quarterly"                                                             ~ "4 yr-1",
      "quarterly"                                                             ~ "4 yr-1",
      "yearly"                                                                ~ "1 yr-1",
      "only_when_needed_adhoc_meetings "                                      ~ "rarely",
      "Due to lack of support the Executive does not meet often"              ~ "rarely",
      "Quarterly but the executive has not had meetings in the last 2 years"  ~ "never",
      "Quarterly but they don't meet"                                         ~ "never",
      .default = NA
    ),
    meetings = factor(meetings,
      levels = c("never", "rarely", "1 yr-1", "4 yr-1", ">4 yr-1"))
  ) |>
  drop_na(meetings) |>
  count(meetings) |>
  mutate(prop = n / sum(n), group = "All CFMGs") |>
  ggplot(aes(y = group, x = prop, fill = meetings)) +
  geom_col(width = 0.4, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.06, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(add = c(0, 0.02)),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      ">4 yr-1" = "#2A9D8F",
      "4 yr-1"  = "#76B7A0",
      "1 yr-1"  = "#F0C040",
      "rarely"  = "#E07B54",
      "never"   = "#C0392B"
    ),
    labels = c(
      ">4 yr-1" = expression(">4 yr"^{-1}),
      "4 yr-1"  = expression("4 yr"^{-1}),
      "1 yr-1"  = expression("1 yr"^{-1}),
      "rarely"  = "rarely",
      "never"   = "never"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  scale_y_discrete(expand = expansion(add = c(0.3, 0.3))) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_blank(),
    axis.ticks.y       = element_blank(),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(2, 25, 10, 10)
  )
```

## CFMG business processes

As with the legal processes, we asked CFMG executives about the state of their business processes. Specifically, we asked if they had developed annual workplans and budgets, and whether these plans were being implemented. We also asked if they had developed a business plan and if any forest user groups had been established (@tbl-business_processes).

```{r}
#| label: tbl-business_processes
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "CFMG completed business process by province."

cfmg |> filter(interview_type == "cfmg_executive") |>
        mutate(
               cfmg_workplan = case_match(
                       cfmg_governance_cfmg_workplan,
                       "yes" ~ 1,
                       "no" ~ 0,
                       "don't_know" ~ NA),
               cfmg_budget = case_match(
                       cfmg_governance_cfmg_budget,
                       "yes" ~ 1,
                       "no" ~ 0,
                       "don't_know" ~ NA),
               cfmg_business_plan = case_match(
                       cfmg_governance_cfma_business_plan,
                       "yes" ~ 1,
                       "no" ~ 0,
                       "don't_know" ~ NA),
               cfmg_user_groups = case_match(
                       cfmg_governance_any_user_groups,
                       "yes" ~ 1,
                       "no" ~ 0,
                       "don't_know" ~ NA)
        ) |>
        group_by(province) |>
        summarise(
                `CFMG workplan` = mean(
                        cfmg_workplan,
                        na.rm = TRUE),
                `CFMG budget` = mean(
                        cfmg_budget,
                        na.rm = TRUE),
                `Implementing plan` = mean(
                        cfmg_governance_implementing_plans_cfmg_executive,
                        na.rm = TRUE),
                `Business plan` = mean(
                        cfmg_business_plan,
                        na.rm = TRUE),
                `User groups` = mean(
                        cfmg_user_groups,
                        na.rm = TRUE)
        ) |>
        mutate(across(where(is.numeric), ~ifelse(is.nan(.x), NA_real_, .x))) |>
        gt() |>
        fmt_percent(columns = -province, decimals = 0) |>
        tab_header(title = "Business processes in the management of CFMGs",
                   subtitle = "Values represent the % of CFMGs with each business process in place by province. n = 75 CFMGs.") |>
        opt_align_table_header(align = "left") |>
        cols_align(align = "center", columns = -province) |>
        cols_align(align = "left", columns = province) |>
        tab_style(
          style     = cell_text(align = "center"),
          locations = cells_column_labels(columns = province)
        ) |>
        sub_missing()


```

::: return-to-summary
[← Return to Summary](Summary.html#governance)
:::

As is evident from the table, these business management processes were even more erratically completed than the legal and governance processes. A substantial proportion of CFMGs did not have workplans, budgets or a business plan. More surprising, many did not have any recognised user groups, which is critical to managing forest resources.

## Governance index

To summarise CFMG governance performance, we constructed a governance index (0–10) based on the presence or absence of ten selected features reported by CFMG executives: receipt book, bank account, financial reporting, annual reporting, issuing permits, HFOs appointed, workplan, budget, business plan, and term limits. We then tested whether two aspects of external support — the frequency of district forest office staff visits and the presence of a partner organisation — were associated with higher governance scores (@fig-gov_index; [Appendix III](Appendix_III-Surveyed_CFAs.qmd)).

```{r}
#| label: fig-gov_index
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Distribution of CFMG governance index scores (0–10). Each point represents one CFMG. The index counts the number of ten governance features reported as present by the CFMG executive FGD."
#| fig-height: 3

load("cfmg.RData")

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  distinct(cfmg_name, .keep_all = TRUE) |>
  drop_na(gov_index) |>
  ggplot(aes(x = gov_index)) +
  geom_histogram(binwidth = 1, fill = "#2A9D8F", colour = "white", linewidth = 0.4) +
  scale_x_continuous(breaks = 0:10, name = "Governance index (out of 10)") +
  scale_y_continuous(name = "Number of CFMGs") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5)
  )
```

```{r}
#| label: data_gov_index_model
#| include: false
#| warning: false
#| message: false
#| error:   false

exec_cfmg <- cfmg |>
  filter(interview_type == "cfmg_executive") |>
  distinct(cfmg_name, .keep_all = TRUE) |>
  mutate(
    dfo_visits_num = case_match(
      dfo_visits,
      "never"        ~ 1,
      "rarely"       ~ 2,
      "occasionally" ~ 3,
      "often"        ~ 4,
      "very_often"   ~ 5,
      .default       = NA_real_
    ),
    external_partner = case_match(
      cfmg_governance_yn_institutions_support_cfmg,
      "yes" ~ 1,
      "no"  ~ 0,
      .default = NA_real_
    )
  )
```

```{r}
#| label: model_gov_index
#| include: false
#| warning: false
#| message: false
#| error:   false

M_gov <- lmer(
  gov_index ~ dfo_visits_num + external_partner + (1 | province),
  data = exec_cfmg,
  REML = TRUE
)
```

## External support

All CFMGs are supported by FD, which has a statutory duty to implement the CFM policy and regulations. Nonetheless, the frequency of visits by district forestry staff varied widely and in many cases they rarely or never visited.

```{r}
#| label: fig-dfo_visits
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Frequency of visits by District Forest Officers (DFO) or other district forestry staff by province. The length of the bars represents the proportion of CFMGs with a specified frequency of visits. The numbers superimposed on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    dfo = case_match(dfo_visits,
      "never"        ~ "Never",
      "rarely"       ~ "Rarely",
      "occasionally" ~ "Occasionally",
      "often"        ~ "Often",
      "very_often"   ~ "Very often",
      .default       = NA
    ),
    dfo = factor(dfo, levels = c("Very often", "Often", "Occasionally", "Rarely", "Never"))
  ) |>
  drop_na(dfo) |>
  count(province, dfo) |>
  group_by(province) |>
  mutate(
    prop       = n / sum(n),
    mean_score = sum(as.numeric(dfo) * prop)
  ) |>
  ungroup() |>
  ggplot(aes(y = reorder(province, mean_score), x = prop, fill = dfo)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.08, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      "Never"        = "#C0392B",
      "Rarely"       = "#E07B54",
      "Occasionally" = "#F0C040",
      "Often"        = "#76B7A0",
      "Very often"   = "#2A9D8F"
    )
  ) +
        guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#governance)
:::

Many CFMGs also operate in partnership with an NGO, such as a conservation organisation, or a private enterprise, such as a carbon developer (@fig-support).

```{r}
#| label: fig-support
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "CFMGs partnered with an external organisation, such as a conservation NGO or carbon developer, by province. The length of the bars represents the number of CFMGs either partnered (Yes) or not partnered (No) with an external organisation. The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(cfmg_governance_yn_institutions_support_cfmg, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#governance)
:::

The table below (@tbl-support_organisations) lists the partners organisations and the number of CFMGs they support. It is important to note that this list does not cover all organisations supporting CFMGs or all the CFMGs that are supported by a named organisation, but only those in our sample of 75 CFMGs. The FD is named as a partner in situations where the district forestry office was responsible for setting up the CFMG and other partners are not involved.

```{r}
#| label: tbl-support_organisations
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Organisations partnering with CFMGs"

cfmg |> filter(interview_type == "cfmg_executive") |>
        mutate(
               `CFMG partner` = cfmg_governance_support_institutions_quests_supportinstitutionsdetails_1__institution_partner_name,
               `CFMG partner` = case_match(
                       `CFMG partner`,
                       "EP-JSLP" ~ "Forestry Department",
                       "Community Climate Solutions Zambia" ~ "Community Climate Solutions",
                       .default = `CFMG partner`
               )
               ) |>
        count(`CFMG partner`) |>
        gt() |>
        tab_header(title = "Partner organisations",
                   subtitle = "n = number of CFMGs partnered with each organisation.") |>
        opt_align_table_header(align = "left") |>
        cols_align(align = "center", columns = -`CFMG partner`) |>
        cols_align(align = "left", columns = `CFMG partner`) |>
        tab_style(
          style     = cell_text(align = "center"),
          locations = cells_column_labels(columns = `CFMG partner`)
        ) |>
        sub_missing()
```

The figure below (@fig-partner_visits) indicates the frequency of visits to CFMGs by partner organisations. The respondents did not state a specified number of visits per year, but from the context we can assume that frequently indicates monthly or bimonthly, occasionally indicates 1-4 times a year, while never means that they had not visited in over a year.

```{r}
#| label: fig-partner_visits
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Frequency of visits by the supporting organisation. The length of the bars represents the proportion of CFMGs with a specified frequency of visits. The numbers superimposed on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
          `Support freq` = cfmg_governance_support_institutions_quests_supportinstitutionsdetails_1__times_institutions_visit_cfmg,
          `Support freq` = case_match(`Support freq`,
                                         "never_0_times"        ~ "Never",
                                         "twice_2_times" ~ "Occasionally",
                                         "thrice_3_times" ~ "Occasionally",
                                         "once_1_time" ~ "Occasionally",
                                         "more_than_5_times" ~ "Occasionally",
                                         "more_than_3_times" ~ "Occasionally",
                                         "More than 3 times" ~ "Occasionally",
                                         "More 3" ~ "Occasionally",
                                         "Dependent on their activities" ~ "Occasionally",
                                         "8_times" ~ "Occasionally",
                                         "4_times" ~ "Occasionally",
                                         "12_time (at_least_every_month)" ~ "Frequently",
                                         "12_times (every_month)" ~ "Frequently",
                                         "36_times" ~ "Frequently",
                                         "COMACO 15 to 20 times per year" ~ "Frequently",
                                         "Very often" ~ "Frequently",
                                         "Very often/regularly. They have a Field Officer living within the community" ~ "Frequently",
                                         "Very_often" ~ "Frequently",
                                         "more_than_12_times" ~ "Frequently",
                                         "several/often_times" ~ "Frequently",
                                         "very_often" ~ "Frequently",
                                         "very_often (60_meetings_per_quater)" ~ "Frequently",
                                         "very_often (they_have_a_lot_of_activities_with_the_cfmg)" ~ "Frequently",
                                         "very_often (they_have_an_officer_within_the_community)" ~ "Frequently",
                                         .default = NA),
    `Support freq` = factor(`Support freq`, levels = c("Frequently", "Occasionally", "Never")),
    `CFMG partner` = cfmg_governance_support_institutions_quests_supportinstitutionsdetails_1__institution_partner_name,
               `CFMG partner` = case_match(
                       `CFMG partner`,
                       "EP-JSLP" ~ "Forestry Department",
                       "Community Climate Solutions Zambia" ~ "Community Climate Solutions",
                       .default = `CFMG partner`
               )
  ) |>
  drop_na(`Support freq`) |>
        drop_na(`CFMG partner`) |>
  count(`CFMG partner`, `Support freq`) |>
  group_by(`CFMG partner`) |>
  mutate(
    prop       = n / sum(n),
    mean_score = sum(as.numeric(`Support freq`) * prop)
  ) |>
  ungroup() |>
  ggplot(aes(y = reorder(`CFMG partner`, mean_score), x = prop, fill = `Support freq`)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.08, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of CFMGs"
  ) +
  scale_fill_manual(
    values = c(
      "Never"        = "#C0392B",
      "Occasionally" = "#F0C040",
      "Frequently"   = "#2A9D8F"
    )
  ) +
        guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We modelled the governance index against district forest office staff visit frequency and presence of an external partner to determine which aspects of external support are associated with higher governance scores (@tbl-gov_index_model, @fig-gov_index_predictors). As the table shows, the frequency of visits from district forestry staff was a highly significant predictor of governance index score. Thus, where district forestry staff visited more frequently CFMGs were better governed.

```{r}
#| label: tbl-gov_index_model
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Linear mixed model of the CFMG governance index."

re_var <- as.data.frame(VarCorr(M_gov))
prov_var  <- round(re_var$vcov[re_var$grp == "province"], 3)
resid_var <- round(re_var$vcov[re_var$grp == "Residual"], 3)
r2 <- performance::r2_nakagawa(M_gov)

tidy(M_gov, effects = "fixed") |>
  mutate(
    term = case_match(term,
      "(Intercept)"      ~ "Intercept",
      "dfo_visits_num"   ~ "DFO visit frequency (1–5)",
      "external_partner" ~ "External partner (yes)"
    )
  ) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(
    term      = "Factor",
    estimate  = "Estimate",
    std.error = "SE",
    p.value   = md("*P* value")
  ) |>
  fmt_number(columns = c(estimate, std.error), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Predictors of CFMG governance index",
    subtitle = glue::glue(
      "Province random intercept variance = {prov_var}; ",
      "residual variance = {resid_var}. ",
      "Marginal R² = {round(r2$R2_marginal, 3)}; ",
      "conditional R² = {round(r2$R2_conditional, 3)}. ",
      "n = {nobs(M_gov)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -term) |>
  cols_align(align = "left", columns = term) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = term)
  ) |>
  sub_missing()
```

```{r}
#| label: fig-gov_index_predictors
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "CFMG governance index in relation to (a) DFO visit frequency and (b) presence of an external partner organisation. Points are individual CFMGs, jittered slightly. Boxes show median and interquartile range."
#| fig-height: 4
#| fig-width: 9

p_dfo <- exec_cfmg |>
  drop_na(dfo_visits_num, gov_index) |>
  mutate(dfo_label = case_match(
    dfo_visits_num,
    1 ~ "Never", 2 ~ "Rarely", 3 ~ "Occasionally", 4 ~ "Often", 5 ~ "Very often"
  ),
  dfo_label = factor(dfo_label,
    levels = c("Never", "Rarely", "Occasionally", "Often", "Very often"))) |>
  ggplot(aes(x = dfo_label, y = gov_index)) +
  geom_boxplot(fill = "#76B7A0", colour = "grey30", outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, height = 0.2, colour = "#1A6B63", alpha = 0.6, size = 1.8) +
  scale_y_continuous(name = "Governance index", limits = c(0, 10), breaks = seq(0, 10, 2)) +
  scale_x_discrete(name = "DFO visit frequency") +
  labs(tag = "(a)") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5)
  )

p_partner <- exec_cfmg |>
  drop_na(external_partner, gov_index) |>
  mutate(partner_label = if_else(external_partner == 1, "Yes", "No")) |>
  ggplot(aes(x = partner_label, y = gov_index)) +
  geom_boxplot(fill = "#76B7A0", colour = "grey30", outlier.shape = NA, width = 0.4) +
  geom_jitter(width = 0.12, height = 0.2, colour = "#1A6B63", alpha = 0.6, size = 1.8) +
  scale_y_continuous(name = NULL, limits = c(0, 10), breaks = seq(0, 10, 2)) +
  scale_x_discrete(name = "External partner organisation") +
  labs(tag = "(b)") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5)
  )

gridExtra::grid.arrange(p_dfo, p_partner, ncol = 2, widths = c(2, 1))
```

## Interest in expanding the community forest

Next we asked interviewees whether they were interested in expanding the area of their CFA. A "yes" answer can be taken as an endorsement of CFM. On the other hand, a negative response might reflect dissatisfaction with the CFMG management or the benefits CFM has brought. However, a negative response could also reflect other reasons, such as a lack of sufficient additional forest area in the community. We compare the responses from different FGD groups.

```{r}
#| label: fig-interest_in_expanding_FGDs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Interest in expanding the community forest by FGD type. The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    expand_cfma = substr(cfmg_governance_interest_expanding_cfma, 1, 1),
    expand_cfma = case_match(
      expand_cfma,
      "y" ~ "Yes",
      "n" ~ "No",
      "d" ~ "Don't know"
    ),
    expand_cfma = factor(expand_cfma, levels = c("Yes", "Don't know", "No")),
    interview_type = case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    )
  ) |>
  filter(!is.na(expand_cfma)) |>
  count(interview_type, expand_cfma) |>
  group_by(interview_type) |>
  mutate(
    prop       = n / sum(n),
    mean_score = sum(as.numeric(expand_cfma) * prop)
  ) |>
  ungroup() |>
  ggplot(aes(y = reorder(interview_type, -mean_score), x = prop, fill = expand_cfma)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "No"         = "#C0392B",
      "Don't know" = "#F0C040",
      "Yes"        = "#2A9D8F"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#governance)
:::

As can be seen in @fig-interest_in_expanding_FGDs, the CFMG executives were enthusiastic to expand the CFA in most cases. However, among the community this number dropped to 41 for female groups, 37 for male groups and to just 23 for youth groups.

The difference in attitudes between CFMG executive and youth FGD groups was significant, but the pairwise comparisons among other FGD groups were not significant. These results might suggest that CFM is not benefiting youth to the same degree as other groups, or perhaps that they bear higher opportunity costs. Our analyses in other sections will help elucidate the reasons.

```{r}
#| label: model_expand_cfma_fgd
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |> filter(interview_type != "individual_interview") |>
        mutate(
                expand_cfma = substr(
                        cfmg_governance_interest_expanding_cfma, 1,1),
                expand_cfma = case_match(
                        expand_cfma,
                        "y" ~ "Yes",
                        "n" ~ "No",
                        "d" ~ "Don't know"
                        ),
                expand_cfma = factor(expand_cfma,
                         levels = c("Yes", "No", "Don't know")
                         )
        ) |>
        mutate(
                expand_cfma = if_else(expand_cfma == "Don't know", NA, expand_cfma)
        )

M1 <- glmer(expand_cfma ~
                    interview_type +
                    gov_index +
                    (1 | cfmg_name),
            family = binomial(link = "logit"),
            data = test.data
)

emm.M1 <- emmeans(M1, specs = ~ interview_type)
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: tbl-M1_expand_cfma_FGD
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Binomial GLMM of interest in expanding the community forest — FGD groups."

re_M1     <- as.data.frame(VarCorr(M1))
cfmg_var  <- round(re_M1$vcov[re_M1$grp == "cfmg_name"], 3)
r2_M1     <- performance::r2_nakagawa(M1)

tidy(M1, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"               ~ "Intercept (Executive FGD)",
    "interview_typecfmg_female" ~ "Female FGD",
    "interview_typecfmg_male"   ~ "Male FGD",
    "interview_typecfmg_youth"  ~ "Youth FGD",
    "gov_index"                 ~ "Governance index"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(
    term      = "Factor",
    estimate  = "Estimate",
    std.error = "SE",
    p.value   = md("*P* value")
  ) |>
  fmt_number(columns = c(estimate, std.error), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Interest in expanding the community forest — FGD groups",
    subtitle = glue::glue(
      "Estimates are log-odds; reference level = Executive FGD. ",
      "CFMG random intercept variance = {cfmg_var}. ",
      "Marginal R² = {round(r2_M1$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M1$R2_conditional, 3)}. ",
      "n = {nobs(M1)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -term) |>
  cols_align(align = "left", columns = term) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = term)
  ) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_expand_cfma_FGD
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Tukey-adjusted pairwise comparisons for FGD type — interest in expanding the community forest."

as.data.frame(pairwise.M1) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(
    contrast  = "Contrast",
    estimate  = "Estimate",
    SE        = "SE",
    z.ratio   = "z ratio",
    p.value   = md("*P* value")
  ) |>
  fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Pairwise comparisons — FGD type",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -contrast) |>
  cols_align(align = "left", columns = contrast) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = contrast)
  ) |>
  sub_missing()
```

We also compared the responses from the individual interviewees (@tbl-interest_in_expanding_individual). The traditional leaders and community members more often reported that they did not want to expand the CFA than other groups. However, the only pairwise comparisons that were significant were private enterprise with traditional leader and private enterprise with community member.

```{r}
#| label: tbl-interest_in_expanding_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Interest in expanding community forest by FGD type."

cfmg |> filter(interview_type == "individual_interview") |>
  mutate(
    expand_cfma = substr(cfmg_governance_interest_expanding_cfma, 1, 1),
    expand_cfma = case_match(expand_cfma, "y"~"Yes", "n"~"No", "d"~"Don't know"),
    expand_cfma = factor(expand_cfma, levels = c("Yes", "No", "Don't know")),
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_exec_member" ~ "CFMG executive member",
      "cfmg_member"      ~ "Community member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "Private enterprise rep"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "CFMG executive member", "Community member", "Traditional leader",
      "District Forestry Officer", "Private enterprise rep"
    ))
  ) |>
  filter(!is.na(expand_cfma)) |>
  count(interviewee_role, expand_cfma, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, expand_cfma, prop) |>
  pivot_wider(names_from = expand_cfma, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Yes, No, `Don't know`), decimals = 0) |>
  tab_header(title = "Interest in expanding the community forest area by interviewee role",
             subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses.") |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()

```

```{r}
#| label: test_data_expand_cfma_individual
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |> filter(interview_type == "individual_interview") |>
        mutate(
                expand_cfma = substr(
                        cfmg_governance_interest_expanding_cfma, 1,1),
                expand_cfma = case_match(
                        expand_cfma,
                        "y" ~ "Yes",
                        "n" ~ "No",
                        "d" ~ "Don't know"
                        ),
                expand_cfma = factor(expand_cfma,
                         levels = c("Yes", "No", "Don't know")
                         )
        ) |>
        mutate(
                expand_cfma = if_else(expand_cfma == "Don't know", NA, expand_cfma),
                interviewee_gender = demo_information_interviewee_gender,
                interviewee_age = demo_information_interviewee_age
        )

M2 <- glmer(expand_cfma ~
                    interviewee_role +
                    interviewee_gender +
                    interviewee_age +
                    (1 | province/cfmg_name),
            family = binomial(link = "logit"),
            data = test.data
)

emm.M2 <- emmeans(M2, specs = ~ interviewee_role)
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")

```

```{r}
#| label: tbl-M2_expand_cfma_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Binomial GLMM of interest in expanding the community forest — individual interviewees."

re_M2     <- as.data.frame(VarCorr(M2))
cfmg_var2 <- round(re_M2$vcov[re_M2$grp == "cfmg_name:province"], 3)
prov_var2 <- round(re_M2$vcov[re_M2$grp == "province"], 3)
r2_M2     <- suppressWarnings(suppressMessages(performance::r2_nakagawa(M2)))

tidy(M2, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"                         ~ "Intercept (Community member)",
    "interviewee_rolecfmg_exec_member"    ~ "CFMG executive member",
    "interviewee_roletrad_leader"         ~ "Traditional leader",
    "interviewee_roledfo"                 ~ "District Forestry Officer",
    "interviewee_roleNGO_rep"             ~ "Private enterprise rep",
    "interviewee_gendermale"              ~ "Gender (male)",
    "interviewee_age"                     ~ "Age"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(
    term      = "Factor",
    estimate  = "Estimate",
    std.error = "SE",
    p.value   = md("*P* value")
  ) |>
  fmt_number(columns = c(estimate, std.error), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Interest in expanding the community forest — individual interviewees",
    subtitle = glue::glue(
      "Estimates are log-odds; reference level = Community member. ",
      "Province variance = {prov_var2}; CFMG:province variance = {cfmg_var2}. ",
      "Marginal R² = {round(r2_M2$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M2$R2_conditional, 3)}. ",
      "n = {nobs(M2)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -term) |>
  cols_align(align = "left", columns = term) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = term)
  ) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_expand_cfma_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Tukey-adjusted pairwise comparisons for interviewee role — interest in expanding the community forest."

as.data.frame(pairwise.M2) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(
    contrast  = "Contrast",
    estimate  = "Estimate",
    SE        = "SE",
    z.ratio   = "z ratio",
    p.value   = md("*P* value")
  ) |>
  fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Pairwise comparisons — interviewee role",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -contrast) |>
  cols_align(align = "left", columns = contrast) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = contrast)
  ) |>
  sub_missing()
```

## Defined user groups

Among those CFMGs that had recognised user groups, they were involved in managing the following resources (@fig-gov_user_groups).

```{r}
#| label: fig-gov_user_groups
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Forest resources managed by user groups within CFMGs, as reported by CFMG executives. Word size is proportional to the number of CFMGs reporting a user-group for that resource type."
#| fig-width: 10
#| fig-height: 7

prefix <- "cfmg_governance_forest_resources_managed_usergroups_"

resource_labels <- c(
  "timber"           = "Timber",
  "poles"            = "Poles",
  "firewood"         = "Firewood",
  "charcoal"         = "Charcoal",
  "honey"            = "Honey",
  "mushrooms"        = "Mushrooms",
  "wild_fruits"      = "Wild fruit",
  "caterpillars"     = "Caterpillars",
  "chikanda"         = "Chikanda",
  "munkoyo"          = "Munkoyo",
  "lusala_busala"    = "Lusala busala",
  "mungongo"         = "Mungongo",
  "devils_claw"      = "Devil's claw",
  "fish"             = "Fish",
  "pupwe"            = "Pupwe",
  "brooms"           = "Brooms",
  "bamboo"           = "Bamboo",
  "medicinal_plants" = "Medicinal plants",
  "bushmeat"         = "Wild meat",
  "resins_gums"      = "Resins",
  "thatching_grass"  = "Thatching",
  "fibers"           = "Fibres",
  "eco_tourism"      = "Eco-tourism"
)

full_var_names <- paste0(prefix, names(resource_labels))

wc_data <- cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(full_var_names)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = resource_labels[sub(prefix, "", variable, fixed = TRUE)]) |>
  filter(freq > 0)

set.seed(42)
ggplot(wc_data, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 30)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

Evidently, people are harvesting a wide variety of products from CFAs, although honey and mushrooms appear especially important. It is not clear from the responses, if these user-groups are defined within the CFA management plan or whether these are informal groups that are nevertheless recognised by the CFMG.

## Discussion

In this chapter we assessed CFMG governance across five domains: legal processes, governance processes, business processes, external support, and community buy-in as reflected by interest in expanding the CFA and the activities of user-groups.

The most consistent pattern across all three process domains — legal, governance, and business — is that foundational steps have been taken by most CFMGs, but the more demanding, ongoing requirements are often incomplete. Early legal steps, such as obtaining a recognition letter and signing a forest management agreement, have been completed by most CFMGs, but later and more involved documents — the registration certificate, forest-use permit, and benefit-sharing agreement — are absent in a substantial proportion. Similarly with respect to governance processes, most CFMGs have held elections and have term limits in place, but basic management actions — opening bank accounts, operating receipt books, and financial and annual reporting — are frequently missing. Business processes are the weakest area of all. Annual workplans, budgets and business plans are rarely in place. Without these planning and financial management tools, CFMGs cannot sustainably manage their resources, are poorly positioned to account for and redistribute income — including carbon payments — and will struggle to demonstrate compliance with the forest management license or to agreements with external partners.

The fact that most CFMGs are overdue for elections is particularly concerning. Timely elections are a basic requirement of the CFMG By-laws & Constitution, and overdue elections undermine the democratic legitimacy of the CFMG executives and could contribute to the perception of elite capture noted elsewhere in this report. These perceptions may be very valid in some cases. The enumerators noted that while some CFMGs had not held elections because they were no longer active, the CFMG executives of others stated that the CFMG had yet to receive carbon income and they were unwilling to hold elections, because they anticipated personal benefits from being a CFMG executive member once the carbon income started to flow. Still others had received carbon income and were unwilling to give up their positions because they do receive some kind of personal benefit. It is essential that this kind of corruption is stamped out, as it may threaten the legitimacy of CFM in the eyes of the community and potentially undo gains made in SFM and sustainable development. Ensuring that CFMGs hold elections on schedule and fairly should be a priority for the FD and for supporting partners.

External support from the FD is insufficient. Most CFMGs are rarely or never visited by district forestry staff, despite the statutory duty of the FD to implement the CFM policy. This gap is partly compensated by conservation NGOs and carbon developers, many of which visit their partner CFMGs more frequently. However, this support is not universal: a significant proportion of CFMGs in our sample had no external partner. Given the governance deficits identified above, regular external support from CFMG partners for capacity building and oversight is clearly a necessity. The current model, in which CFM is effectively co-governed by the community and the private sector rather than the state, may be adequate for CFMGs with active and capable partners, but leaves others without meaningful support. It is noteworthy that, according to the analysis of our Governance Index, frequency of visits from district forestry staff was a highly significant predictor of governance performance.

Interest in expanding the CFA is highest among CFMG executives and drops markedly among community members, particularly youth. Among individual interviewees, Private enterprise reps are significantly more enthusiastic about expansion than traditional leaders or community members. This divergence — consistent with the equity chapter findings — suggests that the CFMG leadership and external partner stakeholders have a more optimistic view of CFM outcomes than the broader community. Furthermore, we found that our governance index was a highly significant predictor of community interest in expanding the CFA, which we take as a proxy indicator for satisfaction with CFM. Thus, as we hypothesised, when CFMGs are better governed there is greater community support for CFM. Thus, conversely, poor governance could lead to increased levels of community dissatisfaction with CFM and, if not addressed, could pose a long-term risk to the legitimacy and sustainability of CFM in Zambia.
