---
title: "Economic Transformation"
author: "Rhett D Harrison"
format:
        html:
                css: custom.css
        pdf: default
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ordinal)
library(tidytext)
library(ggwordcloud)
library(ggbreak)
library(patchwork)

load("cfmg.RData")
```

## Summary

Most FGD groups reported that CFM had not improved access to natural resources for either household use or sale, though approximately one third indicated that access had increased. Community youth were significantly less likely than CFMG Executive FGDs to report improved access for household use. Individual interviewees showed a similar pattern, with DFOs significantly more likely than CFMG executives or traditional leaders to consider that CFM had improved household resource access.

Employment outcomes were more positive. Approximately half of FGD groups agreed that CFM had generated new employment opportunities, with a marked contrast between community youth (who were most likely to agree) and community female FGDs (who were most likely to disagree). Across the surveyed CFMGs, reported employment increased substantially after CFM establishment, though there was considerable variation among provinces (means ranging from 3 in Western Province to 21 in Lusaka Province). Employment was strongly male-biased and the dominant employment category was HFO or Community Scout.

Income earning opportunities from CFM remain limited for most communities. Around one third of CFMGs had received carbon finance payments, and awareness of carbon credit agreements was generally high, though youth were significantly less informed than CFMG Executives. Carbon trader support was primarily directed at forest management and agricultural extension. Natural resource value chains — particularly honey and mushrooms — dominated both realised and aspirational income sources. Most forest products are sold only at local markets and private enterprise partnerships were uncommon, with the vast majority involving honey off-takers. Communities identified timber as the most important unrealised commercial opportunity, despite none of the surveyed CFMGs being involved in timber production.

Opportunity costs of CFM were recognised by only about one third of FGD groups. Unexpectedly, non-timber forest product (NTFP) harvesting was frequently cited as a precluded activity, suggesting possible misunderstandings around forest use rules in some communities.

It is evident from the survey that most CFMG business plans are overly reliant on carbon and potentially valuable NTFP value chains remain under-exploited.

## Introduction
In this chapter, we consider the effect of CFM management on indicators of economic transformation.

### Access to natural resources
We asked FGD groups whether CFM had improved access to natural resources for household use or for sale (@fig-et_household_resources, @fig-et_resources_for_sale).

```{r}
#| label: fig-et_household_resources
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Has CFM improved access to natural resources for household use? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = case_match(
      as.character(economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials),
      "1" ~ "Yes",
      "0" ~ "No"
    ),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

```{r}
#| label: fig-et_resources_for_sale
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Has CFM improved access to natural resources for sale? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = case_match(
      as.character(economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_sale_e_g_timber_honey_mushrooms),
      "1" ~ "Yes",
      "0" ~ "No"
    ),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Responses were similar for both household use and sale. Most FGD groups said that there was no increase in access to natural resources for household use or for sale. However, approximately one third of the groups indicated that access to natural resources has increased under CFM. Community Youth were significantly more likely to say there had not been any improvement in access to natural resources for household use, as compared to CFMG Executive FGDs. None of the other pairwise comparisons were significant.

::: {.callout-note collapse="true"}
### Access to natural resources - FGDs - Model methods
To examine differences among FGD types in whether they reported improved access to natural resources as an economic benefit of CFM, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We ran separate models for household use and resources for sale. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_et_fgd_household
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    hh = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(hh) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M1 <- glmer(hh ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M1      <- emmeans(M1, specs = ~ interview_type, type = "response")
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: M1_et_fgd_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Household resources – FGD model results (click to expand)"

summary(M1)
```

```{r}
#| label: pairwise_et_fgd_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Household resources – FGD pairwise comparisons (click to expand)"

summary(pairwise.M1)
```

```{r}
#| label: model_et_fgd_sale
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    sale = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_sale_e_g_timber_honey_mushrooms,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(sale) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M2 <- glmer(sale ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M2      <- emmeans(M2, specs = ~ interview_type, type = "response")
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")
```

```{r}
#| label: M2_et_fgd_sale
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Resources for sale – FGD model results (click to expand)"

summary(M2)
```

```{r}
#| label: pairwise_et_fgd_sale
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Resources for sale – FGD pairwise comparisons (click to expand)"

summary(pairwise.M2)
```

Individual interviewees were also asked whether CFM had improved access to natural resources for household use or for sale. @tbl-et_individual_access shows the percentage reporting each by interviewee role. Again a majority of respondents indicated that they did not think that CFM had improved access to natural resources for either household use or sale. DFOs were significantly more likely to suggest that CFM had improved access to natural resources for household use than CFMG executives or traditional leaders. None of the other pairwise comparisons were significant.

```{r}
#| label: tbl-et_individual_access
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting improved access to natural resources as an economic benefit of CFM."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    hh   = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials,
    sale = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_sale_e_g_timber_honey_mushrooms
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    ))
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n         = n(),
    Household = mean(hh,   na.rm = TRUE),
    Sale      = mean(sale, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Household        = "Household use",
    Sale             = "For sale"
  ) |>
  fmt_percent(
    columns  = c(Household, Sale),
    decimals = 0
  ) |>
  tab_header(
    title    = "Improved access to natural resources from CFM as reported by individual interviewees",
    subtitle = "Values represent the % of responses agreeing that CFM has improved assess to resources for household use or sale. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Household, Sale)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Access to natural resources - Individual interviewees - Model methods
To examine differences among individual interviewee roles in whether they reported improved access to natural resources as an economic benefit of CFM, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. We ran separate models for household use and resources for sale. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_et_individual_household
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    hh                 = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials,
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(hh) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M4 <- glmer(hh ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M4      <- emmeans(M4, specs = ~ interviewee_role, type = "response")
pairwise.M4 <- pairs(emm.M4, adjust = "Tukey")
```

```{r}
#| label: M4_et_individual_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Household resources – Individual model results (click to expand)"

summary(M4)
```

```{r}
#| label: pairwise_et_individual_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Household resources – Individual pairwise comparisons (click to expand)"

summary(pairwise.M4)
```

```{r}
#| label: model_et_individual_sale
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    sale               = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_sale_e_g_timber_honey_mushrooms,
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(sale) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M5 <- glmer(sale ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M5      <- emmeans(M5, specs = ~ interviewee_role, type = "response")
pairwise.M5 <- pairs(emm.M5, adjust = "Tukey")
```

```{r}
#| label: M5_et_individual_sale
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Resources for sale – Individual model results (click to expand)"

summary(M5)
```

```{r}
#| label: pairwise_et_individual_sale
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Resources for sale – Individual pairwise comparisons (click to expand)"

summary(pairwise.M5)
```

### Employment

We asked FGD groups whether CFM had generated new employment opportunities (@fig-et_employment).

```{r}
#| label: fig-et_employment
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Responses to whether CFM has generated employment opportunities. The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = economic_transformation_yn_formal_employment_opportunities_created_cfm,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen from the figure, approximately half of the FGD groups agreed that CFM had increased employment opportunities. Interestingly, community youth were especially likely to agreed that CFM had created employment opportunities. In contrast, community female FGD groups were much more likely to say that CFM had not increased employment opportunities. All pairwise comparisons were highly significant.

::: {.callout-note collapse="true"}
### Employment - FGDs - Model methods
To examine differences among FGD types in whether they reported CFM had generated employment opportunities, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_et_fgd_employment
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = economic_transformation_yn_formal_employment_opportunities_created_cfm,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M3 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M3      <- emmeans(M3, specs = ~ interview_type, type = "response")
pairwise.M3 <- pairs(emm.M3, adjust = "Tukey")
```

```{r}
#| label: M3_et_fgd_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Employment – FGD model results (click to expand)"

summary(M3)
```

```{r}
#| label: pairwise_et_fgd_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Employment – FGD pairwise comparisons (click to expand)"

summary(pairwise.M3)
```

Individual interviewees were also asked whether CFM had generated employment opportunities (@tbl-et_individual_employment). Interestingly, community respondents were much more likely to agree that CFM had created employment opportunities than NGO representatives or DFOs. The differences between DFOs and CFMG executives and traditional leaders were significant, but none of the other pairwise comparisons were significant.

```{r}
#| label: tbl-et_individual_employment
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that CFM has generated employment opportunities by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_yn_formal_employment_opportunities_created_cfm
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
    
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n          = n(),
    Employment = mean(empl, na.rm = TRUE),
    .groups    = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Employment       = "Agree CFM has generated employment opportunities"
  ) |>
  fmt_percent(
    columns  = Employment,
    decimals = 0
  ) |>
  tab_header(
    title    = "Employment opportunities reported by individual interviewees",
    subtitle = "Values represent the % of responses agreeing that CFM has generated employment opportunities. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Employment)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Employment - Individual interviewees - Model methods
To examine differences among individual interviewee roles in whether they reported CFM had generated employment opportunities, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_et_individual_employment
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = economic_transformation_yn_formal_employment_opportunities_created_cfm,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M6 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M6      <- emmeans(M6, specs = ~ interviewee_role, type = "response")
pairwise.M6 <- pairs(emm.M6, adjust = "Tukey")
```

```{r}
#| label: M6_et_individual_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Employment – Individual model results (click to expand)"

summary(M6)
```

```{r}
#| label: pairwise_et_individual_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Employment – Individual pairwise comparisons (click to expand)"

summary(pairwise.M6)
```


#### Changes in numbers employed
We asked CFMG Executive FGDs to estimate the number of people employed before and after CFM (@fig-et_num_employed). The mean number of people employed per CFMG varied from three in Western Province to 21 in Lusaka Province. The figure also indicates that the numbers of people employed in forest management had substantially increased.

```{r}
#| label: fig-et_num_employed
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Mean number of people per CFMG employed before and after CFM by province, as reported by CFMG Executives."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    province,
    "Before CFM" = economic_transformation_num_employed_before_cfmg,
    "After CFM"  = economic_transformation_num_employed_after_cfmg
  ) |>
  pivot_longer(-province, names_to = "timing", values_to = "n_employed") |>
  filter(!is.na(n_employed)) |>
  mutate(
    timing     = factor(timing, levels = c("Before CFM", "After CFM")),
    n_employed = if_else(province == "Western" & timing == "After CFM",
                         n_employed / 2, n_employed)
  ) |>
  group_by(province, timing) |>
  summarise(mean_employed = mean(n_employed, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(y = reorder(province, mean_employed), x = mean_employed, fill = timing)) +
  geom_col(
    position = position_dodge(width = 0.7), width = 0.6,
    colour = "white", linewidth = 0.3
  ) +
  geom_text(
    aes(label = round(mean_employed, 0)),
    position = position_dodge(width = 0.7),
    hjust = -0.2, size = 3.5, colour = "grey25", fontface = "bold"
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.18)),
    name   = "Mean number employed per CFMG"
  ) +
  coord_cartesian(clip = "off") +
  scale_fill_manual(
    values = c("Before CFM" = "#E07B54", "After CFM" = "#2A9D8F")
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.4),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

#### Full-time senior employment by gender and province
CFMG Executive FGDs also reported on the number of senior males and females in full-time employment in the CFMG (@fig-et_fulltime_gender). Clearly, there are many more senior men in full-time employment than women. According to the response across the 75 CFMGs surveyed, 278 senior men were employed as compared to 56 women.

```{r}
#| label: fig-et_fulltime_gender
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Number of males and females employed full-time in CFMGs, by province, as reported by executive FGDs."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    province,
    "Male"   = economic_transformation_formal_employment_quests_formalemploymentdetails_1__males_employed_cfmg_full_time,
    "Female" = economic_transformation_formal_employment_quests_formalemploymentdetails_1__females_employed_cfmg_full_time
  ) |>
  pivot_longer(-province, names_to = "gender", values_to = "n_employed") |>
  filter(!is.na(n_employed)) |>
  mutate(gender = factor(gender, levels = c("Male", "Female"))) |>
  group_by(province, gender) |>
  summarise(total = sum(n_employed, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(y = reorder(province, total), x = total, fill = gender)) +
  geom_col(
    position = position_dodge(width = 0.7), width = 0.6,
    colour = "white", linewidth = 0.3
  ) +
  geom_text(
    aes(label = total),
    position = position_dodge(width = 0.7),
    hjust = -0.2, size = 3.5, colour = "grey25", fontface = "bold"
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.18)),
    name   = "Number employed full-time"
  ) +
  coord_cartesian(clip = "off") +
  scale_fill_manual(
    values = c("Male" = "#2A9D8F", "Female" = "#E07B54")
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.4),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

CFMG Executive FGDs also reported the number of male and female youth in full-time employment in the CFMG (@fig-et_fulltime_youth). Across the sample of 75 CFMGs, 207 male youths were employed comapred to 48 female youths.

```{r}
#| label: fig-et_fulltime_youth
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Number of male and female youth employed full-time in CFMGs, by province, as reported by executive FGDs."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    province,
    "Male youth"   = economic_transformation_formal_employment_quests_formalemploymentdetails_1__male_youth_employed_cfmg_full_time,
    "Female youth" = economic_transformation_formal_employment_quests_formalemploymentdetails_1__female_youth_employed_cfmg_full_time
  ) |>
  pivot_longer(-province, names_to = "gender", values_to = "n_employed") |>
  filter(!is.na(n_employed)) |>
  mutate(gender = factor(gender, levels = c("Male youth", "Female youth"))) |>
  group_by(province, gender) |>
  summarise(total = sum(n_employed, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(y = reorder(province, total), x = total, fill = gender)) +
  geom_col(
    position = position_dodge(width = 0.7), width = 0.6,
    colour = "white", linewidth = 0.3
  ) +
  geom_text(
    aes(label = total),
    position = position_dodge(width = 0.7),
    hjust = -0.2, size = 3.5, colour = "grey25", fontface = "bold"
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.18)),
    name   = "Number of youth employed full-time"
  ) +
  coord_cartesian(clip = "off") +
  scale_fill_manual(
    values = c("Male youth" = "#2A9D8F", "Female youth" = "#E07B54")
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.4),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```


#### Types of formal employment opportunities
Community female, male, and youth FGDs identified the types of formal employment opportunities generated by CFM (@fig-et_employment_types). Clearly, by far the largest type of job created is Honorary Forest Office or Community Scout.

```{r}
#| label: fig-et_employment_types
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of formal employment opportunities reported by female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting that job type."
#| fig-width: 10
#| fig-height: 7

prefix <- "economic_transformation_formal_employment_opportunities_from_cfm_"

job_labels <- c(
  "forest_managers"                                          = "Forest managers",
  "forest_rangers"                                           = "Forest rangers",
  "community_scouts"                                         = "Community scouts",
  "hfos"                                                     = "HFOs",
  "monitoring_and_evaluation_officers"                       = "M&E officers",
  "administrative_staff"                                     = "Admin staff",
  "program_coordinators"                                     = "Program coordinators",
  "training_coordinators_facilitators"                       = "Training coordinators",
  "agroforestry_extension_officers"                          = "Agroforestry extension",
  "nursery_managers"                                         = "Nursery managers",
  "nursery_workers"                                          = "Nursery workers",
  "technical_support_roles_e_g_soil_scientists_agronomists"  = "Technical support",
  "eco_tourism_operators_e_g_tour_guides_lodge_staff"        = "Eco-tourism operators",
  "cultural_heritage_and_craft_trainers"                     = "Cultural heritage & craft",
  "craftsmen_and_woodcarvers"                                = "Craftsmen & woodcarvers",
  "processing_plant_staff_e_g_honey_timber_charcoal"         = "Processing natural resources",
  "marketing_and_sales_agents"                               = "Marketing & sales",
  "researchers"                                              = "Researchers",
  "data_analysts"                                            = "Data analysts",
  "surveyors"                                                = "Surveyors",
  "field_data_collectors"                                    = "Field data collectors",
  "biodiversity_monitors"                                    = "Biodiversity monitors",
  "forest_extension_workers"                                 = "Forest extension workers",
  "forest_protection_and_fire_mnagement"                     = "Forest protection & fire",
  "enterprise_and_processing_roles"                          = "Enterprise",
  "training_and_facilitation"                                = "Training & facilitation",
  "security_services"                                        = "Security services",
  "livestock_and_resource_management"                        = "Livestock & resource mgmt",
  "interns_trainees"                                         = "Interns & trainees",
  "technical_extension_roles"                                = "Technical extension"
)

full_var_names <- paste0(prefix, names(job_labels))

binary_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_var_names)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = job_labels[sub(prefix, "", variable)]) |>
  filter(freq > 0) |>
  select(word, freq)

comment_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(text = economic_transformation_formal_employment_opportunities_from_cfm_comment) |>
  filter(!is.na(text), trimws(text) != "") |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  mutate(word = case_match(word,
    "hammer" ~ "Hammer mill",
    "mill"   ~ "Hammer mill",
    "cows"   ~ "Cattle",
    .default = word
  )) |>
  count(word, name = "freq")

wc_data <- bind_rows(binary_freq, comment_freq) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(wc_data, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 30)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

### Income earning opportunities
#### Carbon finance
We asked the CFMG Executive FGDs whether their CFMG had received income from carbon finance (@fig-et_carbon_finance). According to the responses, around one third of CFs have received carbon finance. Later responses suggest this is an under-representation for the number of CFMGs involved in carbon finance and probably reflects only those CFMGs that have actually started to receive payouts from carbon.

```{r}
#| label: fig-et_carbon_finance
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "CFMG involvement in carbon finance by province, as reported by CFMG executives."
#| fig-width: 8
#| fig-height: 6

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  transmute(
    province,
    response = case_when(
      economic_transformation_cfma_involved_carbon_finance == "yes"                          ~ "Yes",
      str_starts(economic_transformation_cfma_involved_carbon_finance, "Yes")                ~ "Yes",
      economic_transformation_cfma_involved_carbon_finance == "no"                           ~ "No",
      str_starts(economic_transformation_cfma_involved_carbon_finance, "No")                 ~ "No",
      economic_transformation_cfma_involved_carbon_finance %in% c(
        "Not yet", "At the moment_they are not",
        "The cfmg is yet to enter into partnership with the carbon credit organization",
        "The CFMG has not signed form 7 and MOU. They had to cancel the one they had signed earlier due to disagreements in the benefits sharing mechanism for income from carbon credits",
        "Has not signed the agreement with CCS yet due to dispute  over percentages in the benefit sharing mechanism"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(response = factor(response, levels = c("Yes", "No"))) |>
  count(province, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 1.5),
                     name   = "Number of CFMGs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

FGD groups were also asked whether the community was aware of the carbon credit agreement (@fig-et_carbon_awareness). While most CFMG Executive groups stated that the community was aware of the carbon finance agreement, interestingly two stated that they were not, which is a little surprising. Most community female and male groups indicated they were aware of the carbon finance agreement, but more than half of the community youth groups stated they were not aware of the carbon finance agreement. Community youth were significantly less likely to know about the carbon finance agreement than CFMG Executives, but other pairwise comparisons were not significant.

```{r}
#| label: fig-et_carbon_awareness
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Community awareness of the carbon credit agreement by FGD group."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    group = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    ),
    response = case_when(
      economic_transformation_community_awareness_carbon_credit_agreement == "yes" ~ "Yes",
      economic_transformation_community_awareness_carbon_credit_agreement == "no"  ~ "No",
      economic_transformation_community_awareness_carbon_credit_agreement %in% c(
        "Not aware of the agreement of the Juristiction project we just know what's involved",
        "The agreement is not clear we don't know where they sale the carbon and we don't know how much they get from the money",
        "So far we have just heard the chief signed"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(response = factor(response, levels = c("Yes", "No"))) |>
  count(group, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = group, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 3),
                     name   = "Number of FGDs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Carbon finance awareness – FGDs – Model methods
To examine differences among FGD types in whether they reported community awareness of the carbon credit agreement, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_et_fgd_carbon_awareness
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data.ca <- cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    province, cfmg_name,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    ),
    empl = case_when(
      economic_transformation_community_awareness_carbon_credit_agreement == "yes" ~ 1L,
      economic_transformation_community_awareness_carbon_credit_agreement == "no"  ~ 0L,
      economic_transformation_community_awareness_carbon_credit_agreement %in% c(
        "Not aware of the agreement of the Juristiction project we just know what's involved",
        "The agreement is not clear we don't know where they sale the carbon and we don't know how much they get from the money",
        "So far we have just heard the chief signed"
      ) ~ 0L,
      TRUE ~ NA_integer_
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M11 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data.ca, family = binomial)

emm.M11      <- emmeans(M11, specs = ~ interview_type, type = "response")
pairwise.M11 <- pairs(emm.M11, adjust = "Tukey")
```

```{r}
#| label: M11_et_fgd_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Carbon finance awareness – FGD model results (click to expand)"

summary(M11)
```

```{r}
#| label: pairwise_et_fgd_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Carbon finance awareness – FGD pairwise comparisons (click to expand)"

summary(pairwise.M11)
```

Individual interviewees were also asked whether the community was aware of the carbon credit agreement (@tbl-et_individual_carbon_awareness). Most respondents indicated that the community was aware, but the percentage of positive responses varied across interviewee roles. Interestingly, the NGO represenatatives (a groups that also included carbon enterprise representatives), overwhelmingly (~95%) suggested the community was informed, but the only 70% of DFOs thought the communities were informed. However, none of these differences were significant.

```{r}
#| label: tbl-et_individual_carbon_awareness
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting community awareness of the carbon credit agreement, by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_community_awareness_carbon_credit_agreement
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no"  ~ 0,
                      .default = NA)
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n         = n(),
    Awareness = mean(empl, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Awareness        = "Aware of carbon credit agreement"
  ) |>
  fmt_percent(
    columns  = Awareness,
    decimals = 0
  ) |>
  tab_header(
    title    = "Community awareness of the carbon credit agreement, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Awareness)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Carbon finance awareness – Individual interviewees – Model methods
To examine differences among individual interviewee roles in whether they reported community awareness of the carbon credit agreement, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_et_individual_carbon_awareness
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data.ca.ind <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl = case_match(
      economic_transformation_community_awareness_carbon_credit_agreement,
      "yes" ~ 1,
      "no"  ~ 0,
      .default = NA
    ),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M12 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data.ca.ind, family = binomial)

emm.M12      <- emmeans(M12, specs = ~ interviewee_role, type = "response")
pairwise.M12 <- pairs(emm.M12, adjust = "Tukey")
```

```{r}
#| label: M12_et_individual_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Carbon finance awareness – Individual model results (click to expand)"

summary(M12)
```

```{r}
#| label: pairwise_et_individual_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Carbon finance awareness – Individual pairwise comparisons (click to expand)"

summary(pairwise.M12)
```


#### Carbon trader support
The types of CFM activities supported by the carbon trader are shown in @fig-et_trader_support. As can be seen support focused on aspects of forest management or agricultural extension. Other potential elements of support, such as NTFPs, infrastructure development or market access, featured substantially less.

```{r}
#| label: fig-et_trader_support
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of CFM activities supported by the carbon trader."
#| fig-width: 8
#| fig-height: 5

prefix_ts <- "economic_transformation_cfm_activities_supported_by_carbon_trader_"

ts_labels <- c(
  "patrolling_and_monitoring"                        = "Patrolling",
  "fire_management"                                  = "Fire management",
  "biodiversity_conservation"                        = "Biodiversity",
  "agroforestry"                                     = "Agroforestry",
  "climate_smart_farming"                            = "Climate-smart farming",
  "non_timber_forest_products_ntfps"                 = "NTFPs",
  "eco_tourism"                                      = "Eco-tourism",
  "tree_planting"                                    = "Tree planting",
  "training_and_capacity_building"                   = "Training",
  "carbon_stock_tracking"                            = "Carbon tracking",
  "market"                                           = "Market access",
  "alternative_livelihood_activities"                = "Livelihoods",
  "infractrure"                                      = "Infrastructure",
  "social_gatherings"                                = "Social gatherings",
  "employment"                                       = "Employment",
  "community_sensitization_awareness_engangement"    = "Sensitisation"
)

full_ts_vars <- paste0(prefix_ts, names(ts_labels))

ts_freq <- cfmg |>
  select(all_of(full_ts_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = ts_labels[sub(prefix_ts, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(ts_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 18)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

#### Other income earning opportunities
FGD groups also reported whether CFM had generated income earning opportunities through specific interventions, including market linkages, alternative livelihood interventions, conservation revenue, forest product processing and ecotourism (@fig-et_income_opportunities). Although some community FGDs indicated improvement in income earning opportunities under some categories, especially conservation revenue and forest product processing, most indicated that CFM had not generated new income earning opportunities in these areas.

```{r}
#| label: fig-et_income_opportunities
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Income earning opportunities generated by CFM, as reported by female, male, and youth FGDs. The length of the bars represents the number of FGDs reporting a benefit (Yes) or no benefit (No) for each income source."
#| fig-width: 7
#| fig-height: 5

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  mutate(interview_type = case_match(
    interview_type,
    "cfmg_female" ~ "Female",
    "cfmg_male"   ~ "Male",
    "cfmg_youth"  ~ "Youth"
  )) |>
  select(
    interview_type,
    "Ecotourism income"         = economic_transformation_economic_benefits_from_cfm_income_from_ecotourism_e_g_guided_tours_camping_fees,
    "Forest product processing" = economic_transformation_economic_benefits_from_cfm_income_from_forest_product_processing_e_g_honey_bamboo,
    "Conservation revenue"      = economic_transformation_economic_benefits_from_cfm_revenue_from_forest_conservation_programs_e_g_carbon_credits,
    "Alternative livelihoods"   = `economic_transformation_economic_benefits_from_cfm_alternative-livelihoods_local_business_development_e_g_forest_based_crafts_small_scale_industries`,
    "Market linkages"           = economic_transformation_economic_benefits_from_cfm_market_linkages_and_trade_opportunities_for_forest_products
  ) |>
  pivot_longer(-interview_type, names_to = "income_source", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    income_source = factor(income_source, levels = c(
      "Ecotourism income", "Forest product processing", "Conservation revenue",
      "Alternative livelihoods", "Market linkages"
    )),
    interview_type = factor(interview_type, levels = c("Female", "Male", "Youth"))
  ) |>
  count(interview_type, income_source, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  (\(d) ggplot(d, aes(y = income_source, x = n_signed, fill = response)) +
    geom_blank(data = tibble(
      interview_type = factor(
        c("Female", "Female", "Male", "Male", "Youth", "Youth"),
        levels = c("Female", "Male", "Youth")
      ),
      income_source = factor(
        rep("Ecotourism income", 6),
        levels = c("Ecotourism income", "Forest product processing",
                   "Conservation revenue", "Alternative livelihoods", "Market linkages")
      ),
      n_signed = c(-60, 0, -60, 10, -60, 10),
      response = factor(c("No", "Yes", "No", "Yes", "No", "Yes"), levels = c("No", "Yes"))
    )) +
    geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
    geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
    geom_text(
      aes(label  = n,
          hjust  = if_else(response == "No", -0.4, -0.4),
          colour = response),
      size = 3, fontface = "bold"
    ) +
    scale_colour_manual(
      values = c("No" = "white", "Yes" = "grey25"),
      guide  = "none"
    ) +
    scale_x_continuous(
      labels = ~abs(.x),
      breaks = seq(-60, 10, by = 20),
      expand = expansion(add = 0),
      name   = "Number of FGDs"
    ) +
    coord_cartesian(clip = "off") +
    scale_y_discrete(expand = expansion(add = 0.5)) +
    scale_fill_manual(
      values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
      guide  = guide_legend(reverse = FALSE)
    ) +
    facet_wrap(~ interview_type, scales = "free_x") +
    labs(y = NULL, fill = NULL) +
    theme_minimal(base_size = 18) +
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
      axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
      axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
      axis.title.x       = element_text(size = rel(1), colour = "grey40"),
      legend.position    = "bottom",
      legend.text        = element_text(size = rel(1)),
      strip.text         = element_text(size = rel(1.14), face = "bold"),
      plot.margin        = margin(10, 25, 10, 10)
    )
  )()
```

However, community female, male, and youth FGD groups did identified other additional sources of income generated by CFM (@fig-et_income_sources). Natural resource based value chains featured prominently among these, but interestingly community business was also mentioned.

```{r}
#| label: fig-et_income_sources
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Additional sources of income from CFM reported by female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each income source."
#| fig-width: 10
#| fig-height: 7

prefix_income <- "economic_transformatioNAdditional_sources_of_income_additional_income_sources_from_cfm_"

income_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(starts_with(prefix_income)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(
    word = variable |>
      sub(prefix_income, "", x = _, fixed = TRUE) |>
      gsub("_", " ", x = _) |>
      str_to_title(),
    word = case_match(word,
      "Community Businesses And Enterprises" ~ "Community business",
      "Wildlife Based And Safari Income"  ~ "Wildlife based",
      .default = word
    )
  ) |>
  filter(
    freq > 0,
    !str_starts(word, "Rank"),
    !word %in% c("None", "Dont Know")
  ) |>
  select(word, freq)

set.seed(42)
ggplot(income_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

To better understand the importance of different income earning opportunities we asked the community female, male, and youth FGD groups to rank their top three additional sources of income from CFM (@fig-et_income_ranks). Interestingly, Among the 14 value chains identified, 12 were mentioned as 1st rank at least once. This immediately points to the diversity of forest products being harvested from CFs across the country, but also to the fact that for most CFs only one product features strongly as an addition source of income. Among these, mushrooms and honey are clearly the most frequently mentioned. 

```{r}
#| label: fig-et_income_ranks
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Ranking of additional income sources from CFM by female, male, and youth FGDs. Bars show the number of FGDs assigning each rank. Only income sources receiving at least one top-three ranking are shown."
#| fig-width: 9
#| fig-height: 7

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(matches("Additional_sources_of_income.*rank_")) |>
  pivot_longer(everything(), names_to = "variable", values_to = "rank_val") |>
  mutate(
    rank_val = as.numeric(rank_val),
    source   = sub(".*rank_", "", variable) |>
      gsub("_", " ", x = _) |>
      str_to_title(),
    source = case_match(source,
      "Wildlife Based And Safari Income"  ~ "Wildlife Based",
      "Community Businesses And Enterprises" ~ "Community Business",
      .default = source
    )
  ) |>
  filter(rank_val %in% c(1, 2, 3), source != "Carbon Credits") |>
  mutate(rank = factor(paste0("Rank ", rank_val), levels = c("Rank 3", "Rank 2", "Rank 1"))) |>
  count(source, rank) |>
  group_by(source) |>
  mutate(total = sum(n)) |>
  ungroup() |>
  ggplot(aes(y = reorder(source, total), x = n, fill = rank)) +
  geom_col(width = 0.7, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_fill_manual(
    values = c(
      "Rank 1" = "#264653",
      "Rank 2" = "#E9C46A",
      "Rank 3" = "#E07B54"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.4),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

CFMG Executive FGDs were asked the estimated annual income generated through CFM (@fig-et_annual_income). Unfortunately, there was a lot of inconsistency in the responses to this question. Most FGD groups did not include income from carbon credit sales, but some did. The figure below includes only those that did not include the carbon credit income to facilitate comparison. 

```{r}
#| label: fig-et_annual_income
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Mean estimated annual income from CFM by province, as reported by executive FGDs."
#| fig-width: 9
#| fig-height: 5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    province,
    income = economic_transformation_estimated_annual_income_from_cfm
  ) |>
  filter(!is.na(income), income <= 250000) |>
  mutate(province = reorder(province, income, FUN = median)) |>
  ggplot(aes(y = province, x = income)) +
  geom_boxplot(
    width = 0.5, fill = "grey92", colour = "grey40",
    outlier.shape = NA
  ) +
  geom_jitter(
    height = 0.15, width = 0,
    size = 2, alpha = 0.7, colour = "#2A9D8F"
  ) +
  stat_summary(
    fun = mean, geom = "point",
    colour = "red", size = 3, shape = 16
  ) +
  scale_x_continuous(
    labels = label_comma(),
    expand = expansion(mult = c(0.02, 0.1)),
    name   = "Estimated annual income from CFM"
  ) +
  labs(y = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.4),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Female, male, and youth FGDs also reported which forest products were harvested for household use (@fig-et_household_products). Mushrooms again featured prominently, but other natural resources are also clearly important, such as wild fruits, medicinal plants, thatching grass and caterpillars.

```{r}
#| label: fig-et_household_products
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Forest products harvested for household use reported by female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each product."
#| fig-width: 10
#| fig-height: 7

prefix_hh <- "economic_transformation_products_harvested_household_use_"

hh_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(starts_with(prefix_hh)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(
    word = variable |>
      sub(prefix_hh, "", x = _, fixed = TRUE) |>
      gsub("_", " ", x = _) |>
      str_to_title()
  ) |>
  filter(
    freq > 0,
    !word %in% c("Carbon Credits", "Dont Know", "None")
  ) |>
  select(word, freq)

set.seed(42)
ggplot(hh_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

CFMG Executive FGDs were asked whether the CFMG was collaborating with any private enterprises (@fig-et_collaboration). The answers indicate that only a small number of CFMGs were collaborating with a private enterprise.

```{r}
#| label: fig-et_collaboration
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "CFMG collaboration with private enterprise as reported by executive FGDs by province."
#| fig-width: 8
#| fig-height: 6

collab_theme <- list(
  theme_minimal(base_size = 18),
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.title         = element_text(size = rel(1.14), face = "bold"),
    plot.margin        = margin(10, 25, 10, 10)
  )
)

# Plot 1: private enterprise collaboration (Yes / No)
p1 <- cfmg |>
  filter(interview_type == "cfmg_executive") |>
  transmute(
    province,
    response = case_match(
      economic_transformation_cfmg_private_enterprise_collaboration,
      "yes" ~ "Yes", "no" ~ "No"
    )
  ) |>
  drop_na(response) |>
  mutate(response = factor(response, levels = c("Yes", "No"))) |>
  count(province, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 3.5, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"), guide = "none") +
  scale_x_continuous(labels = ~abs(.x), expand = expansion(add = 1),
                     name = "Number of CFMGs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F")) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL,
       title = "Collaboration with private enterprise") +
  collab_theme

p1
```

The value chains that CFMGs reported working with through private enterprise partnerships are shown in @fig-et_value_chains. Sixteen out of 17 involved partnerships with honey off-takers and just one involved mushrooms.

```{r}
#| label: fig-et_value_chains
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Value chains that CFMGs collaborate with through private enterprise partnerships, as reported by executive FGDs."
#| fig-width: 8
#| fig-height: 5

prefix_vc <- "economic_transformation_private_enterprises_cfmg_works_with_"

vc_labels <- c(
  "timber_companies"               = "Timber",
  "sawmills"                       = "Sawmills",
  "furniture_manufacturers"        = "Furniture",
  "wood_processing_enterprises"    = "Wood processing",
  "eco_tourism_operators"          = "Eco-tourism",
  "mushroom_processors_traders"    = "Mushrooms",
  "wild_fruit_processors_traders"  = "Wild fruit",
  "insect_processors_traders"      = "Insects",
  "chikanda_processors_traders"    = "Chikanda",
 # "carbon_credit_traders"          = "Carbon credits",
  "beekeeping_and_honey_producers" = "Honey",
  "charcoal"                       = "Charcoal",
  "botanicals_harvesting"          = "Botanicals"
)

full_vc_vars <- paste0(prefix_vc, names(vc_labels))

vc_freq <- cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(full_vc_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(label = vc_labels[sub(prefix_vc, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(label, freq)

ggplot(vc_freq, aes(x = freq, y = reorder(label, freq))) +
  geom_col(fill = "#2A9D8F", width = 0.6) +
  geom_text(aes(label = freq), hjust = -0.3, size = 4.5, colour = "grey30") +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of CFMGs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We asked the community female, male, and youth FGD groups to identify economic opportunities that remained unrealised within their CFs (@fig-et_unrealised). Interestingly, although a wide range of different natural resource related value chains were mentioned, timber was mention the most often. This is noteworthy as none of the CFMGs we surveyed are involved in timber production.

```{r}
#| label: fig-et_unrealised
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Unrealised economic opportunities identified by FGD groups."
#| fig-width: 8
#| fig-height: 5

prefix_ur <- "economic_transformation_economic_oppportunities_unrealized_economic_opportunities_"

ur_labels <- c(
  "timber"                    = "Timber",
  "firewood"                  = "Firewood",
  "charcoal"                  = "Charcoal",
  "honey"                     = "Honey",
  "mushrooms"                 = "Mushrooms",
  "wild_fruits"               = "Wild fruits",
  "caterpillars"              = "Caterpillars",
  "chikanda"                  = "Chikanda",
  "munkoyo"                   = "Munkoyo",
  "lusala_busala"             = "Lusala/Busala",
  "mungongo"                  = "Mungongo",
  "water"                     = "Water",
  "devils_claw"               = "Devil's claw",
  "fish"                      = "Fish",
  "wildlife"                  = "Wildlife",
  "minerals_mining"           = "Minerals",
  "clay"                      = "Clay",
  "sand"                      = "Sand",
  "grass"                     = "Grass",
  "livestock"                 = "Livestock",
  "crops"                     = "Crops",
  "tourism_education_tours"   = "Tourism",
  "permitting"                = "Permitting",
  "infrastruture_utilities"   = "Infrastructure",
  "bamboo"                    = "Bamboo",
  "medicinal_plants"          = "Medicinal plants",
  "bushmeat"                  = "Bushmeat",
  "resins_gums"               = "Resins",
  "thatching_grass"           = "Thatching grass",
  "fibers"                    = "Fibres",
  "eco_tourism"               = "Eco-tourism"
)

full_ur_vars <- paste0(prefix_ur, names(ur_labels))

ur_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_ur_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = ur_labels[sub(prefix_ur, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(ur_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

The ranking of these unrealised opportunities is shown in @fig-et_unrealised_ranks. Again a wide range of resources are mentioned, but the top three are timber, honey and mushrooms. Again, tt is interesting that community groups consider their timber resources to have high unrealised potential for commercial development.

```{r}
#| label: fig-et_unrealised_ranks
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Ranking of unrealised economic opportunities by FGD groups. Bars show the frequency of rank 1 (highest priority), rank 2, and rank 3 responses."
#| fig-width: 9
#| fig-height: 7

prefix_ur_r <- "economic_transformation_economic_oppportunities_rank_"

ur_rank_labels <- c(
  "timber"                    = "Timber",
  "firewood"                  = "Firewood",
  "charcoal"                  = "Charcoal",
  "honey"                     = "Honey",
  "mushrooms"                 = "Mushrooms",
  "wild_fruits"               = "Wild fruits",
  "caterpillars"              = "Caterpillars",
  "chikanda"                  = "Chikanda",
  "munkoyo"                   = "Munkoyo",
  "lusala_busala"             = "Lusala/Busala",
  "mungongo"                  = "Mungongo",
  "water"                     = "Water",
  "devils_claw"               = "Devil's claw",
  "fish"                      = "Fish",
  "wildlife"                  = "Wildlife",
  "minerals_mining"           = "Minerals",
  "clay"                      = "Clay",
  "sand"                      = "Sand",
  "grass"                     = "Grass",
  "livestock"                 = "Livestock",
  "crops"                     = "Crops",
  "tourism_education_tours"   = "Tourism",
  "permitting"                = "Permitting",
  "bamboo"                    = "Bamboo",
  "medicinal_plants"          = "Medicinal plants",
  "bushmeat"                  = "Bushmeat",
  "resins_gums"               = "Resins",
  "thatching_grass"           = "Thatching grass",
  "fibers"                    = "Fibres",
  "eco_tourism"               = "Eco-tourism"
)

full_ur_rank_vars <- paste0(prefix_ur_r, names(ur_rank_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_ur_rank_vars)) |>
  pivot_longer(everything(), names_to = "variable", values_to = "rank_val") |>
  mutate(
    rank_val = as.numeric(rank_val),
    source   = ur_rank_labels[sub(prefix_ur_r, "", variable, fixed = TRUE)]
  ) |>
  filter(rank_val %in% c(1, 2, 3)) |>
  mutate(rank = factor(paste0("Rank ", rank_val), levels = c("Rank 3", "Rank 2", "Rank 1"))) |>
  count(source, rank) |>
  group_by(source) |>
  mutate(total = sum(n)) |>
  ungroup() |>
  ggplot(aes(y = reorder(source, total), x = n, fill = rank)) +
  geom_col(width = 0.7, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position  = position_stack(vjust = 0.5),
    size      = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c("Rank 1" = "#264653", "Rank 2" = "#E9C46A", "Rank 3" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of responses"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 15, 10, 10)
  )
```

FGD groups also identified potential value-added opportunities from CFM resources (@fig-et_value_added). The overlap with the previous two questions is substantial, but it is interesting that the communities specifically identify honey processing and drying mushrooms, as well as timber processing. This suggests that the communities are aware that there value addition opportunities that they are currently not realising.

```{r}
#| label: fig-et_value_added
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Potential value-added opportunities identified by FGD groups."
#| fig-width: 8
#| fig-height: 5

prefix_va <- "economic_transformation_potential_value_added_opportunities_"

va_labels <- c(
  "timber_processing_e_g_furniture_carpentry_woodcrafts"                                                    = "Timber processing",
  "charcoal_briquetting_eco_friendly_alternatives_to_traditional_charcoal"                                  = "Charcoal briquettes",
  "honey_processing_and_packaging_e_g_branded_honey_beeswax_products"                                      = "Honey processing",
  "mushroom_drying_and_packaging_for_extended_shelf_life_and_market_access"                                 = "Dried mushrooms",
  "wild_fruit_processing_e_g_juices_jams_dried_fruits"                                                     = "Wild fruits",
  "caterpillar_preservation_and_packaging_e_g_dried_or_canned_for_sale"                                    = "Caterpillars",
  "chikanda"                                                                                                 = "Chikanda",
  "munkoyo"                                                                                                  = "Munkoyo",
  "lusala_busala"                                                                                            = "Lusala/Busala",
  "mungongo"                                                                                                 = "Mungongo",
  "devils_claw"                                                                                              = "Devil's claw",
  "palm_leaves"                                                                                              = "Palm leaves",
  "grass"                                                                                                    = "Grass",
  "fish"                                                                                                     = "Fish",
  "minerals"                                                                                                 = "Minerals",
  "bamboo_products_e_g_furniture_handicrafts_construction_materials"                                        = "Bamboo products",
  "medicinal_plant_extraction_e_g_herbal_teas_essential_oils_medicinal_creams"                              = "Medicinal plants",
  "handicrafts_and_woven_products_e_g_baskets_mats_textiles"                                                = "Handicrafts",
  "resin_and_gum_processing_e_g_adhesives_fragrances"                                                       = "Resins",
  "ecotourism_development_e_g_eco_lodges_cultural_experiences_tour_guiding"                                 = "Eco-tourism",
  "agroforestry_based_businesses_e_g_tree_nurseries_sustainable_agriculture"                                = "Agroforestry"
)

full_va_vars <- paste0(prefix_va, names(va_labels))

va_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_va_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = va_labels[sub(prefix_va, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(va_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

We asked the CFMG Executive FGDs to report which primary markets CFM products are sold into (@fig-et_markets). Evidently, most products are sold only at the local market level, with some also being transported to national markets.

```{r}
#| label: fig-et_markets
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Primary markets for CFM products, as reported by executive FGDs."
#| fig-width: 7
#| fig-height: 4

market_labels <- c(
  "economic_transformation_cfm_products_primary_market_local_markets"    = "Local",
  "economic_transformation_cfm_products_primary_market_national_markets" = "National",
  "economic_transformation_cfm_products_primary_market_regional_markets" = "Regional",
  "economic_transformation_cfm_products_primary_market_eu_markets"       = "EU"
)

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(names(market_labels))) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n") |>
  mutate(market = factor(market_labels[variable],
                         levels = c("Local", "National", "Regional", "EU"))) |>
  ggplot(aes(x = n, y = market)) +
  geom_col(fill = "#2A9D8F", width = 0.55) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, colour = "grey30") +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of CFMGs"
  ) +
  scale_y_discrete(limits = rev) +
  labs(y = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Opportunity costs
#### Precluded activities

The community female, male, and youth FGD groups identified economic activities that are precluded under CFM regulations (@fig-et_precluded). While many of these are self evident, the fact that NTFP harvesting is listed quite prominently is surprising. Whether this reflects a misunderstanding on the part of the communities or indeed that they are being told they cannot collect NTFPs is not clear.

```{r}
#| label: fig-et_precluded
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Economic activities precluded by CFM, as reported by FGD groups."
#| fig-width: 8
#| fig-height: 5

prefix_p <- "economic_transformation_economic_activities_precluded_cfm_"

precluded_labels <- c(
  "logging"                              = "Logging",
  "timber_harvesting"                    = "Timber harvesting",
  "hunting"                              = "Hunting",
  "charcoal_production"                  = "Charcoal",
  "commercial_mining"                    = "Mining",
  "farming"                              = "Farming",
  "livestock_grazing"                    = "Livestock grazing",
  "land_conversion"                      = "Land conversion",
  "introduction_of_invasive_species"     = "Invasive species",
  "ntfp_harvesting_of_forest_products"   = "NTFP harvesting",
  "fishing"                              = "Fishing",
  "beekeping_tradtional"                 = "Beekeeping",
  "fire_based_livelihood_activities"     = "Fire livelihoods"
)

full_p_vars <- paste0(prefix_p, names(precluded_labels))

p_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_p_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = precluded_labels[sub(prefix_p, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(p_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#F4A582", high = "#8B0000") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```


#### Other opportunity costs
FGD groups were asked whether CFM entails opportunity costs for the community (@fig-et_opp_costs). Interestingly, many of the FGDs did not recognise any opportunity costs. However, about a third of the FGDs did. The proportion of FGDs recognising opportunity costs was highest among community male FGDs and lowest among community female FGDs. However, none of the pairwise comparisons were significant.

```{r}
#| label: fig-et_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFM entails opportunity costs for the community, by FGD group."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    group = case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    response = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"          ~ "Yes",
      economic_transformation_opportunity_costs_of_cfm == "no"           ~ "No",
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ "No",
      economic_transformation_opportunity_costs_of_cfm %in% c(
        "Community forest situated very far from community",
        "The Major livelihood for this community is fishing. They don't depend so much on the forest. They can still get materials for construction in the open areas"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(
    group    = factor(group, levels = c("Youth", "Male", "Female", "Executive")),
    response = factor(response, levels = c("Yes", "No"))
  ) |>
  count(group, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = group, x = n_signed, fill = response)) +
  geom_col(width = 0.6, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 3),
                     name   = "Number of FGDs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Opportunity costs – FGDs – Model methods
To examine differences among FGD types in whether they reported that CFM entails opportunity costs for the community, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_et_fgd_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data.opp <- cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    province, cfmg_name,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    ),
    empl = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"           ~ 1L,
      economic_transformation_opportunity_costs_of_cfm == "no"            ~ 0L,
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ 0L,
      economic_transformation_opportunity_costs_of_cfm %in% c(
        "Community forest situated very far from community",
        "The Major livelihood for this community is fishing. They don't depend so much on the forest. They can still get materials for construction in the open areas"
      ) ~ 0L,
      TRUE ~ NA_integer_
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M9 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data.opp, family = binomial)

emm.M9      <- emmeans(M9, specs = ~ interview_type, type = "response")
pairwise.M9 <- pairs(emm.M9, adjust = "Tukey")
```

```{r}
#| label: M9_et_fgd_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Opportunity costs – FGD model results (click to expand)"

summary(M9)
```

```{r}
#| label: pairwise_et_fgd_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Opportunity costs – FGD pairwise comparisons (click to expand)"

summary(pairwise.M9)
```

Individual interviewees were also asked whether CFM entails opportunity costs. @tbl-et_individual_opp_costs shows the percentage reporting this broken down by interviewee role. Interestingly, the community traditional leaders and community members recognise opportunity costs less of the time than other individual interviewee types, although these differences were not significant.

```{r}
#| label: tbl-et_individual_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that CFM entails opportunity costs for the community, by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_opportunity_costs_of_cfm
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_when(
      empl == "yes"           ~ 1,
      empl == "no"            ~ 0,
      str_starts(empl, "No_") ~ 0,
      TRUE                    ~ NA_real_
    )
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n             = n(),
    Opp_costs     = mean(empl, na.rm = TRUE),
    .groups       = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Opp_costs        = "Report opportunity costs"
  ) |>
  fmt_percent(
    columns  = Opp_costs,
    decimals = 0
  ) |>
  tab_header(
    title    = "Opportunity costs of CFM, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses. A positive response indicates that the respondent considers CFM to entail opportunity costs for the community."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Opp_costs)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Opportunity costs – Individual interviewees – Model methods
To examine differences among individual interviewee roles in whether they reported that CFM entails opportunity costs for the community, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_et_individual_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data.opp.ind <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"           ~ 1,
      economic_transformation_opportunity_costs_of_cfm == "no"            ~ 0,
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ 0,
      TRUE                                                                 ~ NA_real_
    ),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M10 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
               (1 | province) + (1 | province_cfmg),
             data = test.data.opp.ind, family = binomial)

emm.M10      <- emmeans(M10, specs = ~ interviewee_role, type = "response")
pairwise.M10 <- pairs(emm.M10, adjust = "Tukey")
```

```{r}
#| label: M10_et_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Opportunity costs – Individual model results (click to expand)"

summary(M10)
```

```{r}
#| label: pairwise_et_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Opportunity costs – Individual pairwise comparisons (click to expand)"

summary(pairwise.M10)
```

The types of opportunity costs identified are shown in @fig-et_opp_cost_types. Mentions of reduced access, land restrictions and land use conflict are by some margin the most frequently identified opportunity costs. However, it is also interesting that time costs and diverting community resources are also quite prevalent. CFM requires that community members meet which entails a time cost and community resources are diverted to managing the forest that might otherwise be spent on other activities, such as agriculture.

```{r}
#| label: fig-et_opp_cost_types
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of opportunity costs experienced under CFM."
#| fig-width: 8
#| fig-height: 5

prefix_oc <- "economic_transformation_main_opportunity_costs_experienced_"

oc_labels <- c(
  "reduced_access_to_forest_resources_e_g_firewood_building_materials"    = "Reduced access",
  "time_spent_on_cfm_activities_instead_of_income_generating_work"        = "Time cost",
  "conflict_with_other_land_resource_uses"                                = "Land use conflict",
  "restrictions_on_land_use_for_farming_or_other_activities"              = "Land restrictions",
  "economic_costs_from_diverting_community_resources_to_cfm"              = "Diverting resources"
)

full_oc_vars <- paste0(prefix_oc, names(oc_labels))

binary_oc <- cfmg |>
  select(all_of(full_oc_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = oc_labels[sub(prefix_oc, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

comment_oc <- cfmg |>
  select(text = economic_transformation_main_opportunity_costs_experienced_comment) |>
  filter(!is.na(text), trimws(text) != "") |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  mutate(word = case_match(
    word,
    c("charcoal", "production", "prohibition", "stopped",
      "restriction", "restrictions") ~ "Charcoal",
    "chitemene"                        ~ "Chitemene",
    "grazing"                          ~ "Grazing",
    "timber"                           ~ "Timber",
    "logging"                          ~ "Logging",
    "fruits"                           ~ "Fruits",
    "medicine"                         ~ "Medicinal plants",
    .default                           = word
  )) |>
  filter(!word %in% c("reduced", "involved", "traditional",
                       "interference", "access", "source",
                       "villages", "moment")) |>
  count(word, name = "freq")

wc_oc <- bind_rows(binary_oc, comment_oc) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(wc_oc, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 18)) +
  scale_colour_gradient(low = "#F4A582", high = "#8B0000") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

## Discussion

Through the FGDs and the individual interviews, we investigated a diverse set of indicators of economic transformation. We covered access to natural resources for household use and sale, employment, income earning opportunities, including carbon finance and other sources of income, and opportunity costs.

Most FGD groups and individual interviewees indicated that CFM had not increased access to natural resources for either household use or sale. However, about a third of respondents considered that access to resources had improved under CFM. Given that the communities were involved in harvesting forest products before the instigation of CFM, it is not surprising that most respondents do not recognise any change. However, around one third of respondent groups and individuals felt that access had improved under CFM. This presumably reflects the formalisation of community rights to access forest resources.

A larger proportion of FGD groups considered that CFM had had a positive effect on full-time employment. However, the differences among different FGD types were highly significant. More community female groups suggested there was no change in employment and more community youth groups felt full-time formal employment opportunities had increased. We also showed that the numbers employed included many more men than women and, given that by far the most important employment roles were HFOs and Community Scouts, it is likely they also included a disproportionate number of youth, which would explain these differences in perception. Moreover, while CFM does appear to have increased formal employment, the numbers are quite paltry. However, this may reflect the fact that employment in forest product harvesting and processing is rarely formal or full-time. In short, while CMF does appear to have resulted in some new employment opportunities, these have been largely limited to HFO and Community Scout roles and have not benefited many women.

A large proportion of CFs in Zambia were formed on the promise of carbon dollars. However, only approximately one third of the CFs in our sample appeared to have benefited from carbon income at the time of the survey. The fact that a number of the FDG groups, especially youth groups, were unaware of the carbon agreement is concerning. It suggests that the carbon trading companies may have failed to conduct a proper FPIC process involving the entire community and that the CFMG Executives may not be communicating properly with their communities. Nevertheless, many carbon traders are clearly supporting communities. Most of the roles concern forest protection and agricultural extension for sustainable intensification, for example, climate smart agriculture. Both of these may be considered as essential to a carbon business plan. NTFPs and enterprise development were less often mentioned as a part of the carbon trader activities. This is perhaps an oversight on their part, because enhancing forest product value chains would increase the value of forest conservation for local communities.

The above results are backed up by the response to questions concerning other income earning opportunities. It seems CFM has not led to increased income from market linkages, alternative livelihoods, conservation revenue, forest product processing or ecotourism except for in a very few cases. Nevertheless, a wide variety of forest products clearly contribute to household use and some additional income. Although honey and mushrooms were most often identified as the most important source of additional income, another 10 NTFPs were mention as the most important at least once. However, these additional sources of income were identified in less than half of the CFs and only 13 identified a second rank. Thus, forest products are rarely being exploited as a source of additional income and if they are it is usually only one. Only a small number of CFMGs are partnering with an enterprise, such as an off-taker, and among those that do it almost exclusively concerns honey. Nonetheless, the communities identified a wide diversity of forest products that could potentially be harvested and processed to provide an additional source of income. Prominent among these were timber and wild fruit, as well as mushrooms and honey. In some cases timber harvesting may conflict with a carbon business plans, but not if plantations were established. The potential of CFM to support forest product enterprises is clearly massively under appreciated. 

Unfortunately, the information on incomes from CFM were not very informative. Some clearly included income from carbon while most did not and it seems that the CFMG Executives have little understanding of the value of forest products being harvested from their CFs. To get a more accurate picture of these values a more in depth PEN type survey is recommended.

The local communities identified a number of opportunity costs to CFM. Most of these pertained to activities that are precluded, such as those involving forest clearance. However, many groups also identified other opportunity costs, such as time costs and diverted resources. Recognition of opportunity costs is important to understand the cost-benefits of CFM for communities.

Overall, a positive picture of economic transformation emerges from the responses of the FGDs and individual interviews. While changes in indicators, such as employment or additional income from forest products, many not appear substantial, they are in a positive direction. Most CFMGs will also receive carbon income, even if this has only benefited around one third of CFMGs so far. However, what is also evident is that CFMG business plans are over-reliant on just carbon or sometimes carbon and honey. Substantial opportunities for developing much more diversfied business plans exist. These could include a range of different NTFPs and potentially also timber. In addition to simply increasing income earning opportunities, diversified business plans would also provide a buffer against fluctuating commodity prices and because different products are harvested at different times of the year, they can help stabilise income throughout the year as well.