---
title: "Equity and Inclusion"
author: "Rhett D Harrison"
format:
        html:
                css: custom.css
        pdf: default
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ordinal)
library(tidytext)
library(ggwordcloud)
library(patchwork)

load("cfmg.RData")
```

## Summary

The overall picture for equity and inclusion is broadly positive, though with important caveats. A substantial majority of FGD groups reported that the benefits of CFM are being shared equitably and that women benefit from CFM without disproportionately suffering opportunity costs. Notably, a majority of community female FGD groups themselves agreed that women have benefited from CFM. Carbon benefit-sharing agreements between carbon companies and communities are dominated by revenue sharing, while employment and livelihood support components are less commonly included. Responsibility for carbon income sharing decisions within communities was roughly equally split between instances where the CFMG executive makes the decisions and those where the whole CFMG is involved. Importantly, community female and youth FGD groups were significantly more likely than executive and male FGD groups to report inequitable benefit sharing, inadequate participation of women in decision making, and a lack of transparency in benefit sharing. In some CFMGs, community youth groups also identified issues with elite capture and marginalisation of youth. This common divergence in perspective between community members and CFMG executives is a concerning pattern suggesting that a minority of CFMGs have real equity problems that are not visible from the executive's perspective. Poor transparency in benefit sharing and the non-existence of grievance and redress mechanisms were also noted for many CFMGs. Without functioning formal channels through which equity concerns can be raised and addressed, inequitable outcomes risk going undetected and uncorrected. Strengthening these governance mechanisms is essential if CFM is to deliver equitable outcomes for all community members, including women, youth and other vulnerable groups.

## Introduction

In this chapter, we consider the equitable sharing of benefits arising from community forest management (CFM) and what the impacts of CFM are on vulnerable groups.

## Benefit sharing

We asked FGD groups whether the benefits of CFM were equitably distributed (@fig-bs_equity). On the one hand, a substantial majority of responses across the FGD groups asserted that benefits were being shared equitably. However, the proportion of community female and community youth groups that felt that benefits were not being shared equitably was significantly higher than for the CFMG executive and community male FGD groups. Moreover, even among the CFMG executive and community male groups, there were some that agreeded that benefits were not being shared equitably.

```{r}
#| label: fig-bs_equity
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Are the benefits of CFM are equitably shared? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = economic_transformation_benefits_equity_across_groups,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Equity - FGDs - Model methods

To examine differences among FGD types in whether they agreed the benefits of CFM were equitably shared, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_equity
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = economic_transformation_benefits_equity_across_groups,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M1 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M1      <- emmeans(M1, specs = ~ interview_type, type = "response")
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: M1_bs_fgd_equity
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Equity – FGD model results (click to expand)"

summary(M1)
```

```{r}
#| label: pairwise_bs_fgd_equity
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Equity – FGD pairwise comparisons (click to expand)"

summary(pairwise.M1)
```

Individual interviewees were also asked whether the benefits of CFM were equitably shared. @tbl-bs_individual_equity shows the percentage reporting this broken down by interviewee role. Again a substantial majority of respondents (83% - 92%) thought that benefits were being shared equitably, and the differences among respondents were not significant. However, there was still minority of respondents that did not think benefits were being equitably shared. This suggests that there are some CFs where the benefit sharing arrangements are not optimal or appropriate arrangements are in place but not being followed.

```{r}
#| label: tbl-bs_individual_equity
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that the benefits of CFM are equitably shared by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_benefits_equity_across_groups
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
    
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n          = n(),
    Equity = mean(empl, na.rm = TRUE),
    .groups    = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Equity       = "Agree benefits are equitably shared"
  ) |>
  fmt_percent(
    columns  = Equity,
    decimals = 0
  ) |>
  tab_header(
    title    = "Equitable sharing of benefits, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses. A positive response indicates that benefits of CFM are equitably shared."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Equity)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Equity - Individual interviewees - Model methods

To examine differences among individual interviewee roles in whether they agreed the benefits of CFM were equitably shared, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_equity
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = economic_transformation_benefits_equity_across_groups,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M2 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M2      <- emmeans(M2, specs = ~ interviewee_role, type = "response")
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")
```

```{r}
#| label: M2_bs_individual_equity
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Equitable benefit sharing – Individual model results (click to expand)"

summary(M2)
```

```{r}
#| label: pairwise_bs_individual_equity
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Equitable benefit sharing – Individual pairwise comparisons (click to expand)"

summary(pairwise.M2)
```

## Carbon income benefit sharing

### Company-community arrangements

CFMG executive FGDs reported on the components were included in their carbon finance benefit-sharing arrangement with the carbon trading enterprise (@fig-bs_benefit_sharing). Clearly the agreements are dominated by revenue sharing, with relatively few also containing agreements on employment, infrastructure development and livelihood support.

```{r}
#| label: fig-bs_benefit_sharing
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Components included in the carbon finance benefit-sharing arrangement between the carbon company and the community as reported by CFMG executive FGDs. The sample includes only those CFMGs that stated they had an agreement with a carbon trading company."
#| fig-width: 9
#| fig-height: 4

prefix_bs <- "economic_transformation_benefit_sharing_arrangement_company_community_"

bs_labels <- c(
  "revenue_sharing_percentage_of_carbon_finance_proceeds_given_to_the_community"              = "Revenue sharing",
  "employment_opportunities_for_community_members"                                             = "Employment",
  "investment_in_community_infrastructure_e_g_schools_clinics_roads"                          = "Infrastructure investment",
  "support_for_sustainable_livelihoods_e_g_alternative_income_generating_activities_training_programs" = "Livelihood support"
)

full_bs_vars <- paste0(prefix_bs, names(bs_labels))

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(full_bs_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n") |>
  mutate(label = factor(bs_labels[sub(prefix_bs, "", variable, fixed = TRUE)],
                        levels = rev(bs_labels))) |>
  ggplot(aes(x = n, y = label)) +
  geom_col(fill = "#2A9D8F", width = 0.55) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, colour = "grey30") +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of CFMGs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Within-community sharing

We also asked FGD groups about the carbon income sharing arrangements within the community. We first asked who is responsible for the decision-making (@fig-bs_benefit_sharing_decision) and then we asked on what principles the benefit sharing was based (@fig-bs_benefit_sharing_principles). As can be seen there is broad agreement among the FGD groups on who makes the decision about the carbon benefit sharing arrangements. In approximately half the CFs the CFMG executive is responsible for the decision, while in the other half the entire CFMG is responsible.

```{r}
#| label: fig-bs_benefit_sharing_decision
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Who decides how carbon income is shared within the community by FGD group. The length of the segments is equal to the number of times that option was mentioned by the FGD group. Note: for 12 CFs the benefit-sharing arrangement is defined in the CFMG constitution."
#| fig-width: 10
#| fig-height: 5

prefix_bsc <- "economic_transformation_benefit_sharing_arrangement_carbon_income_"

decision_labels <- c(
  "decided_by_the_whole_cfmg"     = "Whole CFMG",
  "decided_by_the_cfmg_executive" = "Executive decision"
)

full_decision_vars <- paste0(prefix_bsc, names(decision_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female", "Executive")
  )) |>
  select(group, all_of(full_decision_vars)) |>
  pivot_longer(all_of(full_decision_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    decision_labels[sub(prefix_bsc, "", variable, fixed = TRUE)],
    levels = decision_labels
  )) |>
  count(group, mechanism) |>
  ggplot(aes(y = group, x = n, fill = mechanism)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Whole CFMG"         = "#2A9D8F",
      "Executive decision" = "#E07B54"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(0.85)),
    plot.margin        = margin(10, 15, 10, 10)
  )
```

```{r}
#| label: fig-bs_benefit_sharing_principles
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Principles governing carbon income sharing within the community by FGD group. The length of the segments is equal to the number of times that option was mentioned by the FGD group. Note: for 12 CFs the benefit-sharing arrangement is defined in the CFMG constitution."
#| fig-width: 10
#| fig-height: 5

prefix_bsc <- "economic_transformation_benefit_sharing_arrangement_carbon_income_"

principle_labels <- c(
  "equally_among_all_members_regardless_of_gender_or_group"                            = "Equal shares",
  "based_on_contributions_or_roles_in_cfmg_activities"                                 = "By contribution",
  "prioritized_for_marginalized_groups_e_g_women_youth_aged_persons_with_disabilities" = "Marginalised groups"
)

full_principle_vars <- paste0(prefix_bsc, names(principle_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female", "Executive")
  )) |>
  select(group, all_of(full_principle_vars)) |>
  pivot_longer(all_of(full_principle_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    principle_labels[sub(prefix_bsc, "", variable, fixed = TRUE)],
    levels = principle_labels
  )) |>
  count(group, mechanism) |>
  ggplot(aes(y = group, x = n, fill = mechanism)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Equal shares"        = "#2A9D8F",
      "By contribution"     = "#57B8B0",
      "Marginalised groups" = "#E9C46A"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(0.85)),
    plot.margin        = margin(10, 15, 10, 10)
  )
```

According to the responses of the CFMG executives, seven CFs shared benefits according to members contribution to CFM while six shared them equally among the community households. Only two said that consideration of marginalised or vulnerable groups was a consideration. The other FGD groups were broadly similar in their responses and differences may be due to their knowledge of the arrangements, rather than actual differences in the arrangements.

### External determination of benefit sharing

We also asked FGD groups whether stakeholders outside the community were also involved in determining the benefit-sharing arrangement (@fig-bs_benefit_sharing_external).

```{r}
#| label: fig-bs_benefit_sharing_external
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Who determines the benefit-sharing arrangement outside the community, by FGD group."
#| fig-width: 10
#| fig-height: 5

prefix_bse <- "economic_transformation_benefit_sharing_mechanism_determined_by_"

bse_labels <- c(
  "carbon_trader"                        = "Carbon trader",
  "the_forestry_department_fd"           = "Forestry Dept",
  "an_ngo_or_external_organization"      = "NGO",
  "a_multiple_stakeholder_joint_decision"= "Joint decision",
  "traditional_leaders"                  = "Traditional leaders"
)

full_bse_vars <- paste0(prefix_bse, names(bse_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female", "Executive")
  )) |>
  select(group, all_of(full_bse_vars)) |>
  pivot_longer(all_of(full_bse_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    bse_labels[sub(prefix_bse, "", variable, fixed = TRUE)],
    levels = bse_labels
  )) |>
  count(group, mechanism) |>
  ggplot(aes(y = group, x = n, fill = mechanism)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Carbon trader"      = "#2A9D8F",
      "Forestry Dept"      = "#57B8B0",
      "NGO"                = "#E9C46A",
      "Joint decision"     = "#F4A261",
      "Traditional leaders"= "#E07B54"
    ),
    guide = guide_legend(reverse = TRUE, nrow = 2, byrow = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(0.85)),
    plot.margin        = margin(10, 15, 10, 10)
  )
```

## Gender equity

We asked FGD groups whether women benefit from CFM (@fig-bs_equity_women_benefit).

```{r}
#| label: fig-bs_equity_women_benefit
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Do women benefit from CFM? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = equity_and_inclusion_women_benefit_from_cfm,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Women benefit from CFM - FGDs - Model methods

To examine differences among FGD types in whether they agreed that women benefit from CFM, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_women_benefit
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = equity_and_inclusion_women_benefit_from_cfm,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M3 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M3      <- emmeans(M3, specs = ~ interview_type, type = "response")
pairwise.M3 <- pairs(emm.M3, adjust = "Tukey")
```

```{r}
#| label: M3_bs_fgd_women_benefit
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women benefit from CFM – FGD model results (click to expand)"

summary(M3)
```

```{r}
#| label: pairwise_bs_fgd_women_benefit
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women benefit from CFM – FGD pairwise comparisons (click to expand)"

summary(pairwise.M3)
```

Individual interviewees were also asked whether women benefit from CFM. @tbl-bs_individual_women_benefit shows the percentage reporting this broken down by interviewee role.

```{r}
#| label: tbl-bs_individual_women_benefit
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that women benefit from CFM by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = equity_and_inclusion_women_benefit_from_cfm
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n             = n(),
    Women_benefit = mean(empl, na.rm = TRUE),
    .groups       = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Women_benefit    = "Agree women benefit from CFM"
  ) |>
  fmt_percent(
    columns  = Women_benefit,
    decimals = 0
  ) |>
  tab_header(
    title    = "Women benefit from CFM, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses. A positive response indicates that women benefit from CFM."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Women_benefit)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Women benefit from CFM - Individual interviewees - Model methods

To examine differences among individual interviewee roles in whether they agreed that women benefit from CFM, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_women_benefit
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = equity_and_inclusion_women_benefit_from_cfm,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M4 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M4      <- emmeans(M4, specs = ~ interviewee_role, type = "response")
pairwise.M4 <- pairs(emm.M4, adjust = "Tukey")
```

```{r}
#| label: M4_bs_individual_women_benefit
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women benefit from CFM – Individual model results (click to expand)"

summary(M4)
```

```{r}
#| label: pairwise_bs_individual_women_benefit
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women benefit from CFM – Individual pairwise comparisons (click to expand)"

summary(pairwise.M4)
```

### Women disproportionately suffer opportunity costs

We asked FGD groups whether women disproportionately suffer the opportunity costs of CFM (@fig-bs_opp_costs).

```{r}
#| label: fig-bs_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Do women disproportionately suffer the opportunity costs of CFM? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = equity_and_inclusion_women_disproportionally_suffer_opportunity_costs,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Women opportunity costs - FGDs - Model methods

To examine differences among FGD types in whether they agreed that women disproportionately suffer the opportunity costs of CFM, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = equity_and_inclusion_women_disproportionally_suffer_opportunity_costs,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M5 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M5      <- emmeans(M5, specs = ~ interview_type, type = "response")
pairwise.M5 <- pairs(emm.M5, adjust = "Tukey")
```

```{r}
#| label: M5_bs_fgd_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women opportunity costs – FGD model results (click to expand)"

summary(M5)
```

```{r}
#| label: pairwise_bs_fgd_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women opportunity costs – FGD pairwise comparisons (click to expand)"

summary(pairwise.M5)
```

Individual interviewees were also asked whether women disproportionately suffer the opportunity costs of CFM. @tbl-bs_individual_opp_costs shows the percentage reporting this broken down by interviewee role.

```{r}
#| label: tbl-bs_individual_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that women disproportionately suffer the opportunity costs of CFM by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = equity_and_inclusion_women_disproportionally_suffer_opportunity_costs
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n         = n(),
    Opp_costs = mean(empl, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Opp_costs        = "Agree women disproportionately suffer opportunity costs"
  ) |>
  fmt_percent(
    columns  = Opp_costs,
    decimals = 0
  ) |>
  tab_header(
    title    = "Women disproportionately suffer opportunity costs, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Opp_costs)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Women opportunity costs - Individual interviewees - Model methods

To examine differences among individual interviewee roles in whether they agreed that women disproportionately suffer the opportunity costs of CFM, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = equity_and_inclusion_women_disproportionally_suffer_opportunity_costs,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M6 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M6      <- emmeans(M6, specs = ~ interviewee_role, type = "response")
pairwise.M6 <- pairs(emm.M6, adjust = "Tukey")
```

```{r}
#| label: M6_bs_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women opportunity costs – Individual model results (click to expand)"

summary(M6)
```

```{r}
#| label: pairwise_bs_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women opportunity costs – Individual pairwise comparisons (click to expand)"

summary(pairwise.M6)
```

### Women participate in decision making

We asked FGD groups whether women participate in CFM decision making (@fig-bs_decision_making).

```{r}
#| label: fig-bs_decision_making
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Do women participate in CFM decision making? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = equity_and_inclusion_women_participate_in_decision_making,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Women decision making - FGDs - Model methods

To examine differences among FGD types in whether they agreed that women participate in CFM decision making, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_decision_making
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = equity_and_inclusion_women_participate_in_decision_making,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M7 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M7      <- emmeans(M7, specs = ~ interview_type, type = "response")
pairwise.M7 <- pairs(emm.M7, adjust = "Tukey")
```

```{r}
#| label: M7_bs_fgd_decision_making
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women decision making – FGD model results (click to expand)"

summary(M7)
```

```{r}
#| label: pairwise_bs_fgd_decision_making
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women decision making – FGD pairwise comparisons (click to expand)"

summary(pairwise.M7)
```

Individual interviewees were also asked whether women participate in CFM decision making. @tbl-bs_individual_decision_making shows the percentage reporting this broken down by interviewee role.

```{r}
#| label: tbl-bs_individual_decision_making
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that women participate in CFM decision making by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = equity_and_inclusion_women_participate_in_decision_making
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n               = n(),
    Decision_making = mean(empl, na.rm = TRUE),
    .groups         = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Decision_making  = "Agree women participate in decision making"
  ) |>
  fmt_percent(
    columns  = Decision_making,
    decimals = 0
  ) |>
  tab_header(
    title    = "Women participate in CFM decision making, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Decision_making)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Women decision making - Individual interviewees - Model methods

To examine differences among individual interviewee roles in whether they agreed that women participate in CFM decision making, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_decision_making
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = equity_and_inclusion_women_participate_in_decision_making,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M8 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M8      <- emmeans(M8, specs = ~ interviewee_role, type = "response")
pairwise.M8 <- pairs(emm.M8, adjust = "Tukey")
```

```{r}
#| label: M8_bs_individual_decision_making
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women decision making – Individual model results (click to expand)"

summary(M8)
```

```{r}
#| label: pairwise_bs_individual_decision_making
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Women decision making – Individual pairwise comparisons (click to expand)"

summary(pairwise.M8)
```

### Transparency in benefit sharing

We asked FGD groups whether there is transparency in CFM benefit sharing (@fig-bs_transparency).

```{r}
#| label: fig-bs_transparency
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Is there transparency in CFM benefit sharing? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = equity_and_inclusion_transparency_in_benefit_sharing,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Transparency in benefit sharing - FGDs - Model methods

To examine differences among FGD types in whether they agreed there is transparency in CFM benefit sharing, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_transparency
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = equity_and_inclusion_transparency_in_benefit_sharing,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M9 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M9      <- emmeans(M9, specs = ~ interview_type, type = "response")
pairwise.M9 <- pairs(emm.M9, adjust = "Tukey")
```

```{r}
#| label: M9_bs_fgd_transparency
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Transparency in benefit sharing – FGD model results (click to expand)"

summary(M9)
```

```{r}
#| label: pairwise_bs_fgd_transparency
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Transparency in benefit sharing – FGD pairwise comparisons (click to expand)"

summary(pairwise.M9)
```

Individual interviewees were also asked whether there is transparency in CFM benefit sharing. @tbl-bs_individual_transparency shows the percentage reporting this broken down by interviewee role.

```{r}
#| label: tbl-bs_individual_transparency
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that there is transparency in CFM benefit sharing by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = equity_and_inclusion_transparency_in_benefit_sharing
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n            = n(),
    Transparency = mean(empl, na.rm = TRUE),
    .groups      = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Transparency     = "Agree there is transparency in benefit sharing"
  ) |>
  fmt_percent(
    columns  = Transparency,
    decimals = 0
  ) |>
  tab_header(
    title    = "Transparency in CFM benefit sharing, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Transparency)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Transparency in benefit sharing - Individual interviewees - Model methods

To examine differences among individual interviewee roles in whether they agreed there is transparency in CFM benefit sharing, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_transparency
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = equity_and_inclusion_transparency_in_benefit_sharing,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M10 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
               (1 | province) + (1 | province_cfmg),
             data = test.data, family = binomial)

emm.M10      <- emmeans(M10, specs = ~ interviewee_role, type = "response")
pairwise.M10 <- pairs(emm.M10, adjust = "Tukey")
```

```{r}
#| label: M10_bs_individual_transparency
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Transparency in benefit sharing – Individual model results (click to expand)"

summary(M10)
```

```{r}
#| label: pairwise_bs_individual_transparency
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Transparency in benefit sharing – Individual pairwise comparisons (click to expand)"

summary(pairwise.M10)
```


We examined which mechanisms are being used to achieve transparency in benefit sharing among the community female, male, and youth FGDs (@fig-bs_transparency_mechanisms).

```{r}
#| label: fig-bs_transparency_mechanisms
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Mechanisms used to achieve transparency in CFM benefit sharing, as reported by community female, male, and youth FGDs. Bars show the number of FGDs reporting each mechanism."
#| fig-width: 10
#| fig-height: 8

prefix_trans <- "equity_and_inclusion_mechanisms_for_transparency_fairness_"

trans_labels <- c(
  "community_meetings_to_discuss_benefit_sharing_decisions"                                                    = "Community meetings",
  "written_agreements_or_guidelines_outlining_benefit_distribution_rules"                                     = "Written agreements",
  "memos"                                                                                                      = "Memos",
  "financial_reports_or_audits_shared_with_community_members"                                                 = "Financial reports/audits",
  "inclusive_representation_of_women_youth_and_marginalized_groups_in_decision_making"                        = "Inclusive representation",
  "external_oversight_by_government_agencies_ngos_or_independent_bodies"                                     = "External oversight",
  "grievance_and_dispute_resolution_mechanisms_for_complaints_or_conflicts"                                   = "Grievance mechanisms",
  "needs_assessment"                                                                                           = "Needs assessment"
)

full_trans_vars <- paste0(prefix_trans, names(trans_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_female" ~ "Female",
      "cfmg_male"   ~ "Male",
      "cfmg_youth"  ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female")
  )) |>
  select(group, all_of(full_trans_vars)) |>
  pivot_longer(all_of(full_trans_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    trans_labels[sub(prefix_trans, "", variable, fixed = TRUE)],
    levels = rev(trans_labels)
  )) |>
  count(group, mechanism) |>
  complete(group, mechanism, fill = list(n = 0)) |>
  ggplot(aes(y = mechanism, x = n, fill = group)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65, colour = "white") +
  geom_text(
    aes(label = ifelse(n > 0, n, "")),
    position = position_dodge(width = 0.75),
    hjust = -0.3, size = 3.5, colour = "grey30"
  ) +
  scale_fill_manual(
    values = c(
      "Female" = "#2A9D8F",
      "Male"   = "#57B8B0",
      "Youth"  = "#E9C46A"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(labels = \(x) str_wrap(x, width = 25), expand = expansion(add = 0.6)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Grievance and Redress

We asked FGD groups whether their CFMG has a grievance and redress mechanism in place (@fig-bs_grievance).

```{r}
#| label: fig-bs_grievance
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Does the CFMG have a grievance and redress mechanism? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = equity_and_inclusion_grievance_redress_mechanism,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Grievance and redress - FGDs - Model methods

To examine differences among FGD types in whether they reported having a grievance and redress mechanism, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_grievance
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = equity_and_inclusion_grievance_redress_mechanism,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M11 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
             data = test.data, family = binomial)

emm.M11      <- emmeans(M11, specs = ~ interview_type, type = "response")
pairwise.M11 <- pairs(emm.M11, adjust = "Tukey")
```

```{r}
#| label: M11_bs_fgd_grievance
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Grievance and redress – FGD model results (click to expand)"

summary(M11)
```

```{r}
#| label: pairwise_bs_fgd_grievance
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Grievance and redress – FGD pairwise comparisons (click to expand)"

summary(pairwise.M11)
```

Individual interviewees were also asked whether their CFMG has a grievance and redress mechanism. @tbl-bs_individual_grievance shows the percentage reporting this broken down by interviewee role.

```{r}
#| label: tbl-bs_individual_grievance
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that the CFMG has a grievance and redress mechanism by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = equity_and_inclusion_grievance_redress_mechanism
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n         = n(),
    Grievance = mean(empl, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Grievance        = "Grievance and redress mechanism"
  ) |>
  fmt_percent(
    columns  = Grievance,
    decimals = 0
  ) |>
  tab_header(
    title    = "Grievance and redress mechanism, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Grievance)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Grievance and redress - Individual interviewees - Model methods

To examine differences among individual interviewee roles in whether they reported the CFMG has a grievance and redress mechanism, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_grievance
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = equity_and_inclusion_grievance_redress_mechanism,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M12 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
               (1 | province) + (1 | province_cfmg),
             data = test.data, family = binomial)

emm.M12      <- emmeans(M12, specs = ~ interviewee_role, type = "response")
pairwise.M12 <- pairs(emm.M12, adjust = "Tukey")
```

```{r}
#| label: M12_bs_individual_grievance
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Grievance and redress – Individual model results (click to expand)"

summary(M12)
```

```{r}
#| label: pairwise_bs_individual_grievance
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Grievance and redress – Individual pairwise comparisons (click to expand)"

summary(pairwise.M12)
```

We also examined which mechanisms are being used to achieve grievance and redress among the community female, male, and youth FGDs (@fig-bs_grievance_mechanisms).

```{r}
#| label: fig-bs_grievance_mechanisms
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Mechanisms used for grievance and redress in CFMGs, as reported by community female, male, and youth FGDs. Bars show the number of FGDs reporting each mechanism."
#| fig-width: 10
#| fig-height: 8

prefix_griev <- "equity_and_inclusion_grievance_redress_mechanisms_"


griev_labels <- c(
  "community_meetings_or_forums_for_conflict_resolution"  = "Community meetings/forums",
  "mediation_by_traditional_leaders"                      = "Mediation by traditional leaders",
  "formal_grievance_committees_within_the_cfm_group"      = "Formal grievance committees",
  "legal_channels_local_courts_government_agencies"       = "Legal channels",
  "external_mediation_by_ngos_or_third_parties"           = "External mediation (NGOs)",
  "anonymous_reporting_mechanisms_suggestion_boxes"        = "Anonymous reporting",
  "stipulated_in_constitution"                            = "Stipulated in constitution",
  "stipulated_in_redress_document"                        = "Stipulated in redress document"
)

full_griev_vars <- paste0(prefix_griev, names(griev_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_female" ~ "Female",
      "cfmg_male"   ~ "Male",
      "cfmg_youth"  ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female")
  )) |>
  select(group, all_of(full_griev_vars)) |>
  pivot_longer(all_of(full_griev_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    griev_labels[sub(prefix_griev, "", variable, fixed = TRUE)],
    levels = rev(griev_labels)
  )) |>
  count(group, mechanism) |>
  complete(group, mechanism, fill = list(n = 0)) |>
  ggplot(aes(y = mechanism, x = n, fill = group)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65, colour = "white") +
  geom_text(
    aes(label = ifelse(n > 0, n, "")),
    position = position_dodge(width = 0.75),
    hjust = -0.3, size = 3.5, colour = "grey30"
  ) +
  scale_fill_manual(
    values = c(
      "Female" = "#2A9D8F",
      "Male"   = "#57B8B0",
      "Youth"  = "#E9C46A"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(labels = \(x) str_wrap(x, width = 25), expand = expansion(add = 0.6)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Impact on poverty

We asked the community female, male, and youth FGDs to identify the ways in which CFM has had an impact on poverty in their communities (@fig-bs_poverty_impact).

```{r}
#| label: fig-bs_poverty_impact
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Reported impacts of CFM on poverty, as identified by community female, male, and youth FGDs. Bars show the number of FGDs reporting each impact."
#| fig-width: 10
#| fig-height: 8

prefix_pov <- "equity_and_inclusion_impact_on_poverty_"

pov_labels <- c(
  "increase_in_household_incomes"                                                    = "Increase in household incomes",
  "improved_nutrition_and_food_security"                                             = "Improved nutrition and food security",
  "improved_water_security"                                                          = "Improved water security",
  "higher_school_attendance_rates"                                                   = "Higher school attendance rates",
  "improved_access_to_healthcare"                                                    = "Improved access to healthcare",
  "increase_in_household_assets_livestock_tools_housing_improvements"               = "Increase in household assets",
  "creation_of_new_income_generating_activities_eco_tourism_handicrafts_ntfp_trade" = "New income-generating activities",
  "reduction_in_economic_inequality_within_the_community"                           = "Reduction in economic inequality"
)

full_pov_vars <- paste0(prefix_pov, names(pov_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_female" ~ "Female",
      "cfmg_male"   ~ "Male",
      "cfmg_youth"  ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female")
  )) |>
  select(group, all_of(full_pov_vars)) |>
  pivot_longer(all_of(full_pov_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(impact = factor(
    pov_labels[sub(prefix_pov, "", variable, fixed = TRUE)],
    levels = rev(pov_labels)
  )) |>
  count(group, impact) |>
  complete(group, impact, fill = list(n = 0)) |>
  ggplot(aes(y = impact, x = n, fill = group)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65, colour = "white") +
  geom_text(
    aes(label = ifelse(n > 0, n, "")),
    position = position_dodge(width = 0.75),
    hjust = -0.3, size = 3.5, colour = "grey30"
  ) +
  scale_fill_manual(
    values = c(
      "Female" = "#2A9D8F",
      "Male"   = "#57B8B0",
      "Youth"  = "#E9C46A"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(labels = \(x) str_wrap(x, width = 25), expand = expansion(add = 0.6)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Barriers to poorer community members

We asked the community female, male, and youth FGDs to identify the barriers that prevent poorer community members from benefiting from CFM (@fig-bs_poverty_barriers).

```{r}
#| label: fig-bs_poverty_barriers
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Barriers preventing poorer community members from benefiting from CFM, as reported by community female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each barrier."
#| fig-width: 10
#| fig-height: 7

prefix_b <- "equity_and_inclusion_challenges_for_poorer_community_members_"

barrier_labels <- c(
  "lack_or_limited_cfm_benefits"                                                         = "Limited CFM benefits",
  "limited_access_to_information_about_cfm_opportunities_and_benefits"                   = "Limited information",
  "lack_of_financial_and_other_resources_to_participate_in_cfm_activities"              = "Lack of resources",
  "exclusion_from_decision_making_processes_within_the_cfm_group"                       = "Excluded from decisions",
  "unequal_distribution_of_benefits_among_community_members"                             = "Unequal benefit distribution",
  "limited_access_to_training_or_capacity_building_opportunities"                       = "Limited training",
  "high_opportunity_costs_time_labor_or_resources_lost_by_participating_in_cfm"        = "High opportunity costs",
  "social_or_cultural_barriers_discrimination_based_on_gender_age_or_status"          = "Social/cultural barriers",
  "physical_distance_from_cfm_activities_or_resources"                                  = "Physical distance",
  "lack_of_trust_in_cfm_leadership_management_or_benefit_sharing_mechanisms"          = "No trust in leadership",
  "lack_of_interest_in_cfm_activities"                                                  = "Lack of interest",
  "cfma_access_restrictions"                                                             = "CFMA access restrictions",
  "influence_by_stronger_indiv"                                                          = "Stronger individual influence",
  "human_wildlife_conflicts"                                                             = "Human-wildlife conflicts",
  "selfishness"                                                                          = "Selfishness",
  "interference_from_govt_and_traditional_leadership"                                   = "Govt/traditional leader interference"
)

full_b_vars <- paste0(prefix_b, names(barrier_labels))

# Frequency counts from binary checkbox variables
barrier_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_b_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = barrier_labels[sub(prefix_b, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

# Word frequencies from open-ended comments
comment_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(comment = equity_and_inclusion_challenges_for_poorer_community_members_comment) |>
  drop_na(comment) |>
  unnest_tokens(word, comment) |>
  anti_join(stop_words, by = "word") |>
  filter(nchar(word) > 2, !word %in% c(
    "cfm", "cfmg", "forest", "limited", "lack",
    "youths", "visible", "accessing", "people", "poorer", "community", "coming"
  )) |>
  count(word, name = "freq") |>
  filter(freq > 1)

# Combine checkbox labels and comment words
barriers_all <- bind_rows(barrier_freq, comment_freq) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(barriers_all, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#F4A582", high = "#8B0000") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

### Addressing challenges

We asked the community female, male, and youth FGDs to suggest solutions to address the barriers facing poorer community members in benefiting from CFM (@fig-bs_solutions).

```{r}
#| label: fig-bs_solutions
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Suggested solutions to address barriers preventing poorer community members from benefiting from CFM, as reported by community female, male, and youth FGDs. Bars show the number of FGDs reporting each solution."
#| fig-width: 10
#| fig-height: 6

prefix_sol <- "equity_and_inclusioNAddressing_challenges_"

sol_labels <- c(
  "provision_of_financial_support_or_subsidies_to_poorer_households"      = "Financial support/subsidies",
  "creation_of_targeted_programs_for_marginalized_groups"                  = "Targeted programs for marginalised groups",
  "community_meetings_or_forums_to_discuss_and_resolve_issues"             = "Community meetings/forums",
  "adjustments_to_benefit_sharing_arrangements_to_ensure_fairness"         = "Adjustments to benefit sharing",
  "training_and_capacity_building_for_disadvantaged_groups"                = "Training and capacity building",
  "leadership_and_governance"                                               = "Leadership and governance"
)

full_sol_vars <- paste0(prefix_sol, names(sol_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_female" ~ "Female",
      "cfmg_male"   ~ "Male",
      "cfmg_youth"  ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female")
  )) |>
  select(group, all_of(full_sol_vars)) |>
  pivot_longer(all_of(full_sol_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(solution = factor(
    sol_labels[sub(prefix_sol, "", variable, fixed = TRUE)],
    levels = rev(sol_labels)
  )) |>
  count(group, solution) |>
  complete(group, solution, fill = list(n = 0)) |>
  ggplot(aes(y = solution, x = n, fill = group)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65, colour = "white") +
  geom_text(
    aes(label = ifelse(n > 0, n, "")),
    position = position_dodge(width = 0.75),
    hjust = -0.3, size = 3.5, colour = "grey30"
  ) +
  scale_fill_manual(
    values = c(
      "Female" = "#2A9D8F",
      "Male"   = "#57B8B0",
      "Youth"  = "#E9C46A"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(labels = \(x) str_wrap(x, width = 25), expand = expansion(add = 0.6)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

## Discussion

Through the FGDs and individual interviews, we investigated a wide range of equity and inclusion indicators, covering benefit sharing, gender equity, transparency, grievance and redress, and poverty impacts.

The most striking and consistent finding is a divergence in perspective by FGD group. Across virtually every indicator — equitable benefit sharing, women benefiting from CFM, women suffering disproportionate opportunity costs, women's participation in decision making, and transparency in benefit sharing — community female and youth groups were significantly more likely than CFMG executive and male groups to report negative outcomes. Given that CFMG executives and community males may typically be better informed about formal governance arrangements, it is possible that this divergence reflects an information asymmetry rather than disagreement about conditions. However, it may also reflect that the benefits and costs of CFM are being distributed unevenly, in ways that are visible to female and youth community members but not acknowledged — or perhaps not noticed — by those in the CFMG executives. It is also possible that these differences in perspective reflect elite capture or corruption.

Notwithstanding these these issues, the overall picture is more positive than might have been expected. A clear majority of all FGD groups — including community female groups — reported that benefits are being shared equitably and that women benefit from CFM. Most female groups also stated that women do not disproportionately suffer opportunity costs and that women participate in CFM decision making. This is somewhat surprising, given that natural resource management decisions in Zambia are traditionally the purview of male community members. It may be that the formal governance structures mandated by CFM — elections, executive positions with female representation, and documented benefit-sharing arrangements — are providing women with a degree of recognition and voice that they would not otherwise have had.

Carbon benefit-sharing agreements between companies and communities are almost entirely structured around revenue sharing. Employment, infrastructure investment and livelihood support are less commonly included in these agreements. While revenue sharing provides a direct financial return, it means communities that have not yet received a carbon payment have little to show for CFM. Diversifying the components of company-community agreements could broaden both the distribution of benefits and community buy-in.

The reported poverty impacts of CFM — primarily increased household incomes and new income-generating activities — are broadly consistent with the economic transformation chapter findings. However, barriers to poorer community members benefiting remain, including limited access to resources, capital and markets. The solutions communities themselves proposed — financial support, targeted programmes for marginalised groups, and better governance — align closely with the governance and economic transformation recommendations identified elsewhere in this report.

In conclusion, the equity and inclusion outcomes of CFM in Zambia are broadly positive but uneven. The divergence between female and youth perspectives and those of CFMG executives is the most important finding of this chapter. It signals that equity problems in a subset of CFMGs are likely more widespread than executive-level data alone would suggest, and that strengthening transparency, grievance and redress mechanisms and community-wide engagement is essential to ensuring CFM benefits reach all community members equitably.
