---
title: "CFMG Equitable Benefit Sharing"
author: "Rhett D Harrison"
format:
        html:
                css: custom.css
        pdf: default
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ordinal)
library(tidytext)
library(ggwordcloud)
library(patchwork)

load("cfmg.RData")
```

## Introduction

Here we consider the equitable sharing of benefits arising from community forest management (CFM) among CFMG members.

## Benefit sharing
We asked FGD groups whether the benefits of CFM were equitably distributed (@fig-bs_equity).

```{r}
#| label: fig-bs_equity
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Responses to whether the benefits of CFM are equitably shared. The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = economic_transformation_benefits_equity_across_groups,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Equity - FGDs - Model methods
To examine differences among FGD types in whether they agreed the benefits of CFM were equitably shared, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CFMG nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_equity
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = economic_transformation_benefits_equity_across_groups,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M1 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M1      <- emmeans(M1, specs = ~ interview_type, type = "response")
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: M1_bs_fgd_equity
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Equity – FGD model results (click to expand)"

summary(M1)
```

```{r}
#| label: pairwise_bs_fgd_equity
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Equity – FGD pairwise comparisons (click to expand)"

summary(pairwise.M1)
```

Individual interviewees were also asked whether the benefits of CFM were equitably shared. @tbl-bs_individual_equity shows the percentage reporting this, broken down by interviewee role.

```{r}
#| label: tbl-bs_individual_equity
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that the benefits of CFM are equitably shared, by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_benefits_equity_across_groups
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
    
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n          = n(),
    Equity = mean(empl, na.rm = TRUE),
    .groups    = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Equity       = "Agree benefits are equitably shared"
  ) |>
  fmt_percent(
    columns  = Equity,
    decimals = 0
  ) |>
  tab_header(
    title    = "Equitable sharing of benefits, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses. A positive response indicates that benefits of CFM are equitably shared."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Equity)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Equity - Individual interviewees - Model methods
To examine differences among individual interviewee roles in whether they agreed the benefits of CFM were equitably shared, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CFMG nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_equity
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = economic_transformation_benefits_equity_across_groups,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M2 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M2      <- emmeans(M2, specs = ~ interviewee_role, type = "response")
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")
```

```{r}
#| label: M2_bs_individual_equity
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Equity – Individual model results (click to expand)"

summary(M2)
```

```{r}
#| label: pairwise_bs_individual_equity
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Equity – Individual pairwise comparisons (click to expand)"

summary(pairwise.M2)
```

## Carbon income benefit sharing
### Company-community arrangements
Executive FGDs reported which elements were included in their carbon finance benefit-sharing arrangement (@fig-bs_benefit_sharing).

```{r}
#| label: fig-bs_benefit_sharing
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Elements included in the carbon finance benefit-sharing arrangement between company and community, as reported by executive FGDs."
#| fig-width: 9
#| fig-height: 4

prefix_bs <- "economic_transformation_benefit_sharing_arrangement_company_community_"

bs_labels <- c(
  "revenue_sharing_percentage_of_carbon_finance_proceeds_given_to_the_community"              = "Revenue sharing",
  "employment_opportunities_for_community_members"                                             = "Employment",
  "investment_in_community_infrastructure_e_g_schools_clinics_roads"                          = "Infrastructure investment",
  "support_for_sustainable_livelihoods_e_g_alternative_income_generating_activities_training_programs" = "Livelihood support"
)

full_bs_vars <- paste0(prefix_bs, names(bs_labels))

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(full_bs_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n") |>
  mutate(label = factor(bs_labels[sub(prefix_bs, "", variable, fixed = TRUE)],
                        levels = rev(bs_labels))) |>
  ggplot(aes(x = n, y = label)) +
  geom_col(fill = "#2A9D8F", width = 0.55) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, colour = "grey30") +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of CFMGs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Within-community sharing
FGD groups reported how carbon income is shared within the community (@fig-bs_benefit_sharing_community).

```{r}
#| label: fig-bs_benefit_sharing_community
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Benefit-sharing arrangements for carbon income within the community, by FGD group. Note: for 12 executive FGDs, the benefit-sharing arrangement is enshrined in the CFMG constitution; this is not shown in the figure."
#| fig-width: 10
#| fig-height: 5

prefix_bsc <- "economic_transformation_benefit_sharing_arrangement_carbon_income_"

bsc_labels <- c(
  "equally_among_all_members_regardless_of_gender_or_group"                                = "Equal shares",
  "based_on_contributions_or_roles_in_cfmg_activities"                                     = "By contribution",
  "prioritized_for_marginalized_groups_e_g_women_youth_aged_persons_with_disabilities"     = "Marginalised groups",
  "decided_by_the_cfmg_executive"                                                           = "Executive decision",
  "decided_by_the_whole_cfmg"                                                               = "Whole CFMG"
)

full_bsc_vars <- paste0(prefix_bsc, names(bsc_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female", "Executive")
  )) |>
  select(group, all_of(full_bsc_vars)) |>
  pivot_longer(all_of(full_bsc_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    bsc_labels[sub(prefix_bsc, "", variable, fixed = TRUE)],
    levels = bsc_labels
  )) |>
  count(group, mechanism) |>
  ggplot(aes(y = group, x = n, fill = mechanism)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Equal shares"       = "#2A9D8F",
      "By contribution"    = "#57B8B0",
      "Marginalised groups"= "#E9C46A",
      "Executive decision" = "#F4A261",
      "Whole CFMG"         = "#E07B54"
    ),
    guide = guide_legend(reverse = TRUE, nrow = 2, byrow = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(0.85)),
    plot.margin        = margin(10, 15, 10, 10)
  )
```

### External determination of benefit sharing
FGD groups also reported who determines the benefit-sharing arrangement outside the community (@fig-bs_benefit_sharing_external).

```{r}
#| label: fig-bs_benefit_sharing_external
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Who determines the benefit-sharing arrangement outside the community, by FGD group."
#| fig-width: 10
#| fig-height: 5

prefix_bse <- "economic_transformation_benefit_sharing_mechanism_determined_by_"

bse_labels <- c(
  "the_community"                        = "Community",
  "carbon_trader"                        = "Carbon trader",
  "the_forestry_department_fd"           = "Forestry Dept",
  "an_ngo_or_external_organization"      = "NGO",
  "a_multiple_stakeholder_joint_decision"= "Joint decision",
  "traditional_leaders"                  = "Traditional leaders"
)

full_bse_vars <- paste0(prefix_bse, names(bse_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female", "Executive")
  )) |>
  select(group, all_of(full_bse_vars)) |>
  pivot_longer(all_of(full_bse_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    bse_labels[sub(prefix_bse, "", variable, fixed = TRUE)],
    levels = bse_labels
  )) |>
  count(group, mechanism) |>
  ggplot(aes(y = group, x = n, fill = mechanism)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Community"          = "#264653",
      "Carbon trader"      = "#2A9D8F",
      "Forestry Dept"      = "#57B8B0",
      "NGO"                = "#E9C46A",
      "Joint decision"     = "#F4A261",
      "Traditional leaders"= "#E07B54"
    ),
    guide = guide_legend(reverse = TRUE, nrow = 2, byrow = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(0.85)),
    plot.margin        = margin(10, 15, 10, 10)
  )
```


## Opportunity costs
FGD groups were asked whether CFM entails opportunity costs for the community (@fig-bs_opp_costs).

```{r}
#| label: fig-bs_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFM entails opportunity costs for the community, by FGD group."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    group = case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    response = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"          ~ "Yes",
      economic_transformation_opportunity_costs_of_cfm == "no"           ~ "No",
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ "No",
      economic_transformation_opportunity_costs_of_cfm %in% c(
        "Community forest situated very far from community",
        "The Major livelihood for this community is fishing. They don't depend so much on the forest. They can still get materials for construction in the open areas"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(
    group    = factor(group, levels = c("Youth", "Male", "Female", "Executive")),
    response = factor(response, levels = c("Yes", "No"))
  ) |>
  count(group, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = group, x = n_signed, fill = response)) +
  geom_col(width = 0.6, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 3),
                     name   = "Number of FGDs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Opportunity costs – FGDs – Model methods
To examine differences among FGD types in whether they reported that CFM entails opportunity costs for the community, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CFMG nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data.opp <- cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    province, cfmg_name,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    ),
    empl = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"           ~ 1L,
      economic_transformation_opportunity_costs_of_cfm == "no"            ~ 0L,
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ 0L,
      economic_transformation_opportunity_costs_of_cfm %in% c(
        "Community forest situated very far from community",
        "The Major livelihood for this community is fishing. They don't depend so much on the forest. They can still get materials for construction in the open areas"
      ) ~ 0L,
      TRUE ~ NA_integer_
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M3 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data.opp, family = binomial)

emm.M3      <- emmeans(M3, specs = ~ interview_type, type = "response")
pairwise.M3 <- pairs(emm.M3, adjust = "Tukey")
```

```{r}
#| label: M3_bs_fgd_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Opportunity costs – FGD model results (click to expand)"

summary(M3)
```

```{r}
#| label: pairwise_bs_fgd_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Opportunity costs – FGD pairwise comparisons (click to expand)"

summary(pairwise.M3)
```

Individual interviewees were also asked whether CFM entails opportunity costs. @tbl-bs_individual_opp_costs shows the percentage reporting this, broken down by interviewee role.

```{r}
#| label: tbl-bs_individual_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that CFM entails opportunity costs for the community, by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_opportunity_costs_of_cfm
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_when(
      empl == "yes"           ~ 1,
      empl == "no"            ~ 0,
      str_starts(empl, "No_") ~ 0,
      TRUE                    ~ NA_real_
    )
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n             = n(),
    Opp_costs     = mean(empl, na.rm = TRUE),
    .groups       = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Opp_costs        = "Report opportunity costs"
  ) |>
  fmt_percent(
    columns  = Opp_costs,
    decimals = 0
  ) |>
  tab_header(
    title    = "Opportunity costs of CFM, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses. A positive response indicates that the respondent considers CFM to entail opportunity costs for the community."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Opp_costs)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Opportunity costs – Individual interviewees – Model methods
To examine differences among individual interviewee roles in whether they reported that CFM entails opportunity costs for the community, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CFMG nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_opp_costs
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data.opp.ind <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"           ~ 1,
      economic_transformation_opportunity_costs_of_cfm == "no"            ~ 0,
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ 0,
      TRUE                                                                 ~ NA_real_
    ),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M4 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
               (1 | province) + (1 | province_cfmg),
             data = test.data.opp.ind, family = binomial)

emm.M4      <- emmeans(M4, specs = ~ interviewee_role, type = "response")
pairwise.M4 <- pairs(emm.M4, adjust = "Tukey")
```

```{r}
#| label: M4_bs_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Opportunity costs – Individual model results (click to expand)"

summary(M4)
```

```{r}
#| label: pairwise_bs_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Opportunity costs – Individual pairwise comparisons (click to expand)"

summary(pairwise.M4)
```

The types of opportunity costs experienced are shown in @fig-bs_opp_cost_types.

```{r}
#| label: fig-bs_opp_cost_types
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of opportunity costs experienced under CFM."
#| fig-width: 8
#| fig-height: 5

prefix_oc <- "economic_transformation_main_opportunity_costs_experienced_"

oc_labels <- c(
  "reduced_access_to_forest_resources_e_g_firewood_building_materials"    = "Reduced access",
  "time_spent_on_cfm_activities_instead_of_income_generating_work"        = "Time cost",
  "conflict_with_other_land_resource_uses"                                = "Land use conflict",
  "restrictions_on_land_use_for_farming_or_other_activities"              = "Land restrictions",
  "economic_costs_from_diverting_community_resources_to_cfm"              = "Diverting resources"
)

full_oc_vars <- paste0(prefix_oc, names(oc_labels))

binary_oc <- cfmg |>
  select(all_of(full_oc_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = oc_labels[sub(prefix_oc, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

comment_oc <- cfmg |>
  select(text = economic_transformation_main_opportunity_costs_experienced_comment) |>
  filter(!is.na(text), trimws(text) != "") |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  mutate(word = case_match(
    word,
    c("charcoal", "production", "prohibition", "stopped",
      "restriction", "restrictions") ~ "Charcoal",
    "chitemene"                        ~ "Chitemene",
    "grazing"                          ~ "Grazing",
    "timber"                           ~ "Timber",
    "logging"                          ~ "Logging",
    "fruits"                           ~ "Fruits",
    "medicine"                         ~ "Medicinal plants",
    .default                           = word
  )) |>
  filter(!word %in% c("reduced", "involved", "traditional",
                       "interference", "access", "source",
                       "villages", "moment")) |>
  count(word, name = "freq")

wc_oc <- bind_rows(binary_oc, comment_oc) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(wc_oc, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 18)) +
  scale_colour_gradient(low = "#F4A582", high = "#8B0000") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

