---
title: "CFMG Equitable Benefit Sharing"
author: "Rhett D Harrison"
format:
        html:
                css: custom.css
        pdf: default
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ordinal)
library(tidytext)
library(ggwordcloud)
library(patchwork)

load("cfmg.RData")
```

## Introduction
Here we consider the equitable sharing of benefits arising from community forest management (CFM) among CFMG members.

## Carbon finance

### CFMG involvement in carbon finance

Executive FGDs reported whether their CFMG was involved in carbon finance (@fig-bs_carbon_finance).

```{r}
#| label: fig-bs_carbon_finance
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "CFMG involvement in carbon finance by province, as reported by executive FGDs."
#| fig-width: 8
#| fig-height: 6

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  transmute(
    province,
    response = case_when(
      economic_transformation_cfma_involved_carbon_finance == "yes"                          ~ "Yes",
      str_starts(economic_transformation_cfma_involved_carbon_finance, "Yes")                ~ "Yes",
      economic_transformation_cfma_involved_carbon_finance == "no"                           ~ "No",
      str_starts(economic_transformation_cfma_involved_carbon_finance, "No")                 ~ "No",
      economic_transformation_cfma_involved_carbon_finance %in% c(
        "Not yet", "At the moment_they are not",
        "The cfmg is yet to enter into partnership with the carbon credit organization",
        "The CFMG has not signed form 7 and MOU. They had to cancel the one they had signed earlier due to disagreements in the benefits sharing mechanism for income from carbon credits",
        "Has not signed the agreement with CCS yet due to dispute  over percentages in the benefit sharing mechanism"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(response = factor(response, levels = c("Yes", "No"))) |>
  count(province, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", 0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 1.5),
                     name   = "Number of CFMGs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Community awareness of the carbon credit agreement

FGD groups were also asked whether the community was aware of the carbon credit agreement (@fig-bs_carbon_awareness).

```{r}
#| label: fig-bs_carbon_awareness
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Community awareness of the carbon credit agreement by FGD group."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    group = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    ),
    response = case_when(
      economic_transformation_community_awareness_carbon_credit_agreement == "yes" ~ "Yes",
      economic_transformation_community_awareness_carbon_credit_agreement == "no"  ~ "No",
      economic_transformation_community_awareness_carbon_credit_agreement %in% c(
        "Not aware of the agreement of the Juristiction project we just know what's involved",
        "The agreement is not clear we don't know where they sale the carbon and we don't know how much they get from the money",
        "So far we have just heard the chief signed"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(response = factor(response, levels = c("Yes", "No"))) |>
  count(group, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = group, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", 0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 3),
                     name   = "Number of FGDs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Carbon finance awareness – FGDs – Model methods

To examine differences among FGD types in whether they reported community awareness of the carbon credit agreement, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CFMG nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_bs_fgd_carbon_awareness
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data.ca <- cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    province, cfmg_name,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    ),
    empl = case_when(
      economic_transformation_community_awareness_carbon_credit_agreement == "yes" ~ 1L,
      economic_transformation_community_awareness_carbon_credit_agreement == "no"  ~ 0L,
      economic_transformation_community_awareness_carbon_credit_agreement %in% c(
        "Not aware of the agreement of the Juristiction project we just know what's involved",
        "The agreement is not clear we don't know where they sale the carbon and we don't know how much they get from the money",
        "So far we have just heard the chief signed"
      ) ~ 0L,
      TRUE ~ NA_integer_
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M1 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data.ca, family = binomial)

emm.M1      <- emmeans(M1, specs = ~ interview_type, type = "response")
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: M1_bs_fgd_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Carbon finance awareness – FGD model results (click to expand)"

summary(M1)
```

```{r}
#| label: pairwise_bs_fgd_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Carbon finance awareness – FGD pairwise comparisons (click to expand)"

summary(pairwise.M1)
```

Individual interviewees were also asked whether the community was aware of the carbon credit agreement. @tbl-bs_individual_carbon_awareness shows the percentage reporting awareness, broken down by interviewee role.

```{r}
#| label: tbl-bs_individual_carbon_awareness
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting community awareness of the carbon credit agreement, by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_community_awareness_carbon_credit_agreement
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no"  ~ 0,
                      .default = NA)
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n         = n(),
    Awareness = mean(empl, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Awareness        = "Aware of carbon credit agreement"
  ) |>
  fmt_percent(
    columns  = Awareness,
    decimals = 0
  ) |>
  tab_header(
    title    = "Community awareness of the carbon credit agreement, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Awareness)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Carbon finance awareness – Individual interviewees – Model methods

To examine differences among individual interviewee roles in whether they reported community awareness of the carbon credit agreement, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CFMG nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_bs_individual_carbon_awareness
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data.ca.ind <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl = case_match(
      economic_transformation_community_awareness_carbon_credit_agreement,
      "yes" ~ 1,
      "no"  ~ 0,
      .default = NA
    ),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M2 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data.ca.ind, family = binomial)

emm.M2      <- emmeans(M2, specs = ~ interviewee_role, type = "response")
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")
```

```{r}
#| label: M2_bs_individual_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Carbon finance awareness – Individual model results (click to expand)"

summary(M2)
```

```{r}
#| label: pairwise_bs_individual_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Carbon finance awareness – Individual pairwise comparisons (click to expand)"

summary(pairwise.M2)
```

### Benefit-sharing arrangements

#### Company-community arrangements

Executive FGDs reported which elements were included in their carbon finance benefit-sharing arrangement (@fig-bs_benefit_sharing).

```{r}
#| label: fig-bs_benefit_sharing
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Elements included in the carbon finance benefit-sharing arrangement between company and community, as reported by executive FGDs."
#| fig-width: 9
#| fig-height: 4

prefix_bs <- "economic_transformation_benefit_sharing_arrangement_company_community_"

bs_labels <- c(
  "revenue_sharing_percentage_of_carbon_finance_proceeds_given_to_the_community"              = "Revenue sharing",
  "employment_opportunities_for_community_members"                                             = "Employment",
  "investment_in_community_infrastructure_e_g_schools_clinics_roads"                          = "Infrastructure investment",
  "support_for_sustainable_livelihoods_e_g_alternative_income_generating_activities_training_programs" = "Livelihood support"
)

full_bs_vars <- paste0(prefix_bs, names(bs_labels))

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(full_bs_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n") |>
  mutate(label = factor(bs_labels[sub(prefix_bs, "", variable, fixed = TRUE)],
                        levels = rev(bs_labels))) |>
  ggplot(aes(x = n, y = label)) +
  geom_col(fill = "#2A9D8F", width = 0.55) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, colour = "grey30") +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of CFMGs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

#### Within-community sharing

FGD groups reported how carbon income is shared within the community (@fig-bs_benefit_sharing_community).

```{r}
#| label: fig-bs_benefit_sharing_community
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Benefit-sharing arrangements for carbon income within the community, by FGD group. Note: for 12 executive FGDs, the benefit-sharing arrangement is enshrined in the CFMG constitution; this is not shown in the figure."
#| fig-width: 10
#| fig-height: 5

prefix_bsc <- "economic_transformation_benefit_sharing_arrangement_carbon_income_"

bsc_labels <- c(
  "equally_among_all_members_regardless_of_gender_or_group"                                = "Equal shares",
  "based_on_contributions_or_roles_in_cfmg_activities"                                     = "By contribution",
  "prioritized_for_marginalized_groups_e_g_women_youth_aged_persons_with_disabilities"     = "Marginalised groups",
  "decided_by_the_cfmg_executive"                                                           = "Executive decision",
  "decided_by_the_whole_cfmg"                                                               = "Whole CFMG"
)

full_bsc_vars <- paste0(prefix_bsc, names(bsc_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female", "Executive")
  )) |>
  select(group, all_of(full_bsc_vars)) |>
  pivot_longer(all_of(full_bsc_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    bsc_labels[sub(prefix_bsc, "", variable, fixed = TRUE)],
    levels = bsc_labels
  )) |>
  count(group, mechanism) |>
  ggplot(aes(y = group, x = n, fill = mechanism)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Equal shares"       = "#2A9D8F",
      "By contribution"    = "#57B8B0",
      "Marginalised groups"= "#E9C46A",
      "Executive decision" = "#F4A261",
      "Whole CFMG"         = "#E07B54"
    ),
    guide = guide_legend(reverse = TRUE, nrow = 2, byrow = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(0.85)),
    plot.margin        = margin(10, 15, 10, 10)
  )
```

#### External determination of benefit sharing

FGD groups also reported who determines the benefit-sharing arrangement outside the community (@fig-bs_benefit_sharing_external).

```{r}
#| label: fig-bs_benefit_sharing_external
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Who determines the benefit-sharing arrangement outside the community, by FGD group."
#| fig-width: 10
#| fig-height: 5

prefix_bse <- "economic_transformation_benefit_sharing_mechanism_determined_by_"

bse_labels <- c(
  "the_community"                        = "Community",
  "carbon_trader"                        = "Carbon trader",
  "the_forestry_department_fd"           = "Forestry Dept",
  "an_ngo_or_external_organization"      = "NGO",
  "a_multiple_stakeholder_joint_decision"= "Joint decision",
  "traditional_leaders"                  = "Traditional leaders"
)

full_bse_vars <- paste0(prefix_bse, names(bse_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female", "Executive")
  )) |>
  select(group, all_of(full_bse_vars)) |>
  pivot_longer(all_of(full_bse_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(mechanism = factor(
    bse_labels[sub(prefix_bse, "", variable, fixed = TRUE)],
    levels = bse_labels
  )) |>
  count(group, mechanism) |>
  ggplot(aes(y = group, x = n, fill = mechanism)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Community"          = "#264653",
      "Carbon trader"      = "#2A9D8F",
      "Forestry Dept"      = "#57B8B0",
      "NGO"                = "#E9C46A",
      "Joint decision"     = "#F4A261",
      "Traditional leaders"= "#E07B54"
    ),
    guide = guide_legend(reverse = TRUE, nrow = 2, byrow = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(0.85)),
    plot.margin        = margin(10, 15, 10, 10)
  )
```

### Carbon trader support

The types of CFM activities supported by the carbon trader are shown in @fig-bs_trader_support.

```{r}
#| label: fig-bs_trader_support
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of CFM activities supported by the carbon trader."
#| fig-width: 8
#| fig-height: 5

prefix_ts <- "economic_transformation_cfm_activities_supported_by_carbon_trader_"

ts_labels <- c(
  "patrolling_and_monitoring"                        = "Patrolling",
  "fire_management"                                  = "Fire management",
  "biodiversity_conservation"                        = "Biodiversity",
  "agroforestry"                                     = "Agroforestry",
  "climate_smart_farming"                            = "Climate-smart farming",
  "non_timber_forest_products_ntfps"                 = "NTFPs",
  "eco_tourism"                                      = "Eco-tourism",
  "tree_planting"                                    = "Tree planting",
  "training_and_capacity_building"                   = "Training",
  "carbon_stock_tracking"                            = "Carbon tracking",
  "market"                                           = "Market access",
  "alternative_livelihood_activities"                = "Livelihoods",
  "infractrure"                                      = "Infrastructure",
  "social_gatherings"                                = "Social gatherings",
  "employment"                                       = "Employment",
  "community_sensitization_awareness_engangement"    = "Sensitisation"
)

full_ts_vars <- paste0(prefix_ts, names(ts_labels))

ts_freq <- cfmg |>
  select(all_of(full_ts_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = ts_labels[sub(prefix_ts, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(ts_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 18)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```
