---
title: "CFMG Economic Transformation"
author: "Rhett D Harrison"
format:
        html:
                css: custom.css
        pdf: default
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ordinal)

load("cfmg.RData")
```

## Introduction

Here we consider the effect of CFMG management on indicators of economic transformation.

### Economic benefits

We asked FGD groups whether CFM had delivered each of three types of economic benefit: improved access to natural resources for household use, improved access to natural resources for sale, and employment opportunities (@fig-et_household_resources, @fig-et_resources_for_sale, @fig-et_employment).

```{r}
#| label: fig-et_household_resources
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Responses to whether improved access to natural resources for household use (e.g. firewood, building materials) is an economic benefit of CFM. The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = case_match(
      as.character(economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials),
      "1" ~ "Yes",
      "0" ~ "No"
    ),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

```{r}
#| label: fig-et_resources_for_sale
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Responses to whether improved access to natural resources for sale (e.g. timber, honey, mushrooms) is an economic benefit of CFM. The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = case_match(
      as.character(economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_sale_e_g_timber_honey_mushrooms),
      "1" ~ "Yes",
      "0" ~ "No"
    ),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

```{r}
#| label: fig-et_employment
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Responses to whether employment opportunities (e.g. forest management, eco-tourism, forest patrols) are an economic benefit of CFM. The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = case_match(
      as.character(economic_transformation_economic_benefits_from_cfm_employment_opportunities_e_g_forest_management_eco_tourism_forest_patrols),
      "1" ~ "Yes",
      "0" ~ "No"
    ),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: {.callout-note collapse="true"}
### Economic benefits - FGDs - Model methods

To examine differences among FGD types in whether they reported each economic benefit from CFM, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CFMG nested within province. We used FGD type as the factor. We ran separate models for each of the three economic benefit variables. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_et_fgd_household
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    hh = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(hh) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M1 <- glmer(hh ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M1      <- emmeans(M1, specs = ~ interview_type, type = "response")
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: M1_et_fgd_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Household resources – FGD model results (click to expand)"

summary(M1)
```

```{r}
#| label: pairwise_et_fgd_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Household resources – FGD pairwise comparisons (click to expand)"

summary(pairwise.M1)
```

```{r}
#| label: model_et_fgd_sale
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    sale = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_sale_e_g_timber_honey_mushrooms,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(sale) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M2 <- glmer(sale ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M2      <- emmeans(M2, specs = ~ interview_type, type = "response")
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")
```

```{r}
#| label: M2_et_fgd_sale
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Resources for sale – FGD model results (click to expand)"

summary(M2)
```

```{r}
#| label: pairwise_et_fgd_sale
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Resources for sale – FGD pairwise comparisons (click to expand)"

summary(pairwise.M2)
```

```{r}
#| label: model_et_fgd_employment
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = economic_transformation_economic_benefits_from_cfm_employment_opportunities_e_g_forest_management_eco_tourism_forest_patrols,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M3 <- glmer(empl ~ interview_type + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M3      <- emmeans(M3, specs = ~ interview_type, type = "response")
pairwise.M3 <- pairs(emm.M3, adjust = "Tukey")
```

```{r}
#| label: M3_et_fgd_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Employment – FGD model results (click to expand)"

summary(M3)
```

```{r}
#| label: pairwise_et_fgd_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Employment – FGD pairwise comparisons (click to expand)"

summary(pairwise.M3)
```

FGD groups also reported whether CFM had generated specific income opportunities (@fig-et_income_opportunities).

```{r}
#| label: fig-et_income_opportunities
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Income opportunities reported as economic benefits of CFM, as reported by female and male FGDs. The length of the bars represents the number of FGDs reporting a benefit (Yes) or no benefit (No) for each income source."
#| fig-width: 12
#| fig-height: 6

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male")) |>
  mutate(interview_type = case_match(
    interview_type,
    "cfmg_female" ~ "Female FGD",
    "cfmg_male"   ~ "Male FGD"
  )) |>
  select(
    interview_type,
    "Ecotourism income"          = economic_transformation_economic_benefits_from_cfm_income_from_ecotourism_e_g_guided_tours_camping_fees,
    "Forest product processing"  = economic_transformation_economic_benefits_from_cfm_income_from_forest_product_processing_e_g_honey_bamboo,
    "Conservation revenue"       = economic_transformation_economic_benefits_from_cfm_revenue_from_forest_conservation_programs_e_g_carbon_credits,
    "Alternative livelihoods"    = `economic_transformation_economic_benefits_from_cfm_alternative-livelihoods_local_business_development_e_g_forest_based_crafts_small_scale_industries`,
    "Market linkages"            = economic_transformation_economic_benefits_from_cfm_market_linkages_and_trade_opportunities_for_forest_products,
    "Increased land value"       = economic_transformation_economic_benefits_from_cfm_increased_land_value_due_to_improved_forest_management,
    "Infrastructure development" = economic_transformation_economic_benefits_from_cfm_infrastructure_development_e_g_boreholes_roads_markets_processing_centres
  ) |>
  pivot_longer(-interview_type, names_to = "income_source", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    income_source = factor(income_source, levels = c(
      "Ecotourism income", "Forest product processing", "Conservation revenue",
      "Alternative livelihoods", "Market linkages", "Increased land value",
      "Infrastructure development"
    )),
    interview_type = factor(interview_type, levels = c("Female FGD", "Male FGD"))
  ) |>
  count(interview_type, income_source, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = income_source, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.22),
    name   = "Number of FGDs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  facet_grid(. ~ interview_type) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    strip.text         = element_text(size = rel(1.14), face = "bold"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We also asked individual interviewees whether CFM had delivered each economic benefit. @tbl-et_individual_benefits shows the proportion reporting each benefit, broken down by interviewee role.

```{r}
#| label: tbl-et_individual_benefits
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Proportion of individual interviewees reporting each economic benefit of CFM, by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    hh   = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials,
    sale = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_sale_e_g_timber_honey_mushrooms,
    empl = economic_transformation_economic_benefits_from_cfm_employment_opportunities_e_g_forest_management_eco_tourism_forest_patrols
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "NGO representative"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "NGO representative"
    ))
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n          = n(),
    Household  = mean(hh,   na.rm = TRUE),
    Sale       = mean(sale, na.rm = TRUE),
    Employment = mean(empl, na.rm = TRUE),
    .groups    = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Household        = "Household use",
    Sale             = "For sale",
    Employment       = "Employment"
  ) |>
  fmt_percent(
    columns  = c(Household, Sale, Employment),
    decimals = 0
  ) |>
  tab_header(
    title    = "Economic benefits of CFM reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses. A positive response indicates that CFM has increased access to natural resources for household use or for sale, or has resulted in increased employment."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Household, Sale, Employment)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Economic benefits - Individual interviewees - Model methods

To examine differences among individual interviewee roles in whether they reported each economic benefit from CFM, we employed a generalised linear mixed model (GLMM) with a binomial response and random intercepts for province and CFMG nested within province. We included interviewee role, interviewee gender and interviewee age as factors. We ran separate models for each of the three economic benefit variables. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: model_et_individual_household
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    hh                 = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials,
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(hh) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M4 <- glmer(hh ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M4      <- emmeans(M4, specs = ~ interviewee_role, type = "response")
pairwise.M4 <- pairs(emm.M4, adjust = "Tukey")
```

```{r}
#| label: M4_et_individual_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Household resources – Individual model results (click to expand)"

summary(M4)
```

```{r}
#| label: pairwise_et_individual_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Household resources – Individual pairwise comparisons (click to expand)"

summary(pairwise.M4)
```

```{r}
#| label: model_et_individual_sale
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    sale               = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_sale_e_g_timber_honey_mushrooms,
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(sale) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M5 <- glmer(sale ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M5      <- emmeans(M5, specs = ~ interviewee_role, type = "response")
pairwise.M5 <- pairs(emm.M5, adjust = "Tukey")
```

```{r}
#| label: M5_et_individual_sale
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Resources for sale – Individual model results (click to expand)"

summary(M5)
```

```{r}
#| label: pairwise_et_individual_sale
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Resources for sale – Individual pairwise comparisons (click to expand)"

summary(pairwise.M5)
```

```{r}
#| label: model_et_individual_employment
#| code-fold: true
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = economic_transformation_economic_benefits_from_cfm_employment_opportunities_e_g_forest_management_eco_tourism_forest_patrols,
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M6 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M6      <- emmeans(M6, specs = ~ interviewee_role, type = "response")
pairwise.M6 <- pairs(emm.M6, adjust = "Tukey")
```

```{r}
#| label: M6_et_individual_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Employment – Individual model results (click to expand)"

summary(M6)
```

```{r}
#| label: pairwise_et_individual_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Employment – Individual pairwise comparisons (click to expand)"

summary(pairwise.M6)
```
