---
title: "Economic Transformation"
format:
  html:
    css: custom.css
  pdf:
    include-in-header:
      text: |
        \pagenumbering{arabic}
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ordinal)
library(tidytext)
library(ggwordcloud)
library(ggbreak)
library(patchwork)

load("cfmg.RData")
```

## Summary

Most FGD groups reported that CFM had not improved access to natural resources for either household use or sale, though approximately one third indicated that access had increased. Approximately half of FGD groups agreed that CFM had generated new employment opportunities, with a marked contrast between community youth (who were most likely to agree) and community female FGDs (who were most likely to disagree). Employment was strongly male-biased and the dominant employment category was HFO or Community Scout. Income earning opportunities from CFM remain limited for most communities. Awareness of carbon finance agreements was generally high. However, although around 80% of CFMGs are involved in generating carbon credits, only around one third had received carbon payments. Carbon developer investment was primarily directed at improved forest management and agricultural extension, rather than at developing forest product value chains. However, forest product value chains — particularly honey and mushrooms — dominated both realised and aspirational income sources among CFA communities. Most forest products are sold only at local markets. Partnerships with private enterprises were uncommon and mostly involved honey off-takers. Communities identified timber as the most important unrealised commercial opportunity, although none of the surveyed CFMGs were involved in timber production. Opportunity costs of CFM were recognised by only about one third of FGD groups. Unexpectedly, non-timber forest product (NTFP) harvesting was frequently cited as a precluded activity, suggesting possible misunderstandings around forest use rules in many communities. It is evident from the survey that most CFMG business plans are overly reliant on carbon and potentially valuable NTFP value chains remain under-exploited. Furthermore, the governance index was a significant predictor of economic performance, at least from the perspective of the community FGD groups, which again underscores the importance of ensuring quality governance if the benefits of CFM are to be realised.

::: {.callout-note icon="false"}
## Key conclusions

CFM is generating some economic benefits for communities. However, these are currently very limited and overly focused on carbon, too concentrated in a few low paying roles, such as HFOs, and too unevenly distributed across communities, with women in particular not seeing sufficient economic benefit. To achieve an economic transformation through CFM, requires a diversified, layered forest enterprise model — one that exploits the full range of products, develops value chains and actively connects communities to markets. Such a vision is both achievable and arguably essential for long-term CFM viability, but requires a much greater effort from the FD and CFM partners, such as carbon developers and conseervation NGOs, to identify opportunities and provide capacity development in forest resource management, value addition and business skills. 
:::

## Introduction

In this chapter, we consider the effect of CFM management on indicators of economic transformation.

We hypothesised that CFM would create employment and income-generating opportunities for communities. However, we expected that these would be unevenly distributed — concentrated in carbon-related activities, biased toward men and community leaders, and limited in scope. We also tested the hypothesis from the governance chapter that CFMGs with better governance would achieve better and more equitable economic outcomes. We also invited communities to identify untapped economic opportunities (such as timber and other forest products) that remain largely unexploited due to lack of enterprise development support.

### Access to natural resources

We asked FGD groups whether CFM had improved access to natural resources for household use (@fig-et_household_resources).

```{r}
#| label: fig-et_household_resources
#| echo: false
#| warning: false
#| message: false
#| fig-width: 8
#| fig-height: 5
#| fig-cap: "Has CFM improved access to natural resources for household use? The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

household_col <- "economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials"

# Verify column exists
if (!household_col %in% names(cfmg)) {
  stop("Column not found: ", household_col, "\n",
       "Available columns:\n", 
       paste(grep("household", names(cfmg), value = TRUE, ignore.case = TRUE), collapse = "\n"))
}

# Prepare data with proper validation
plot_data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  select(interview_type, all_of(household_col)) |>
  filter(!is.na(.data[[household_col]])) |>
  mutate(
    response = case_match(
      as.character(.data[[household_col]]),
      "1" ~ "Yes",
      "0" ~ "No",
      .default = NA_character_
    ),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  filter(!is.na(response))

# Check if we have data before plotting
if (nrow(plot_data) == 0) {
  stop("No valid data available for fig-et_household_resources after filtering")
}

# Create plot
plot_data |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Most FGD groups reported no increase in access to natural resources for household use, though approximately one third indicated that access had improved under CFM. Community youth were significantly more likely to report no improvement compared to CFMG executive FGDs, but none of the other pairwise comparisons were significant. Most natural resources are currently accessed for household use and results were similar for those that are collected for sale.

```{r}
#| label: model_et_fgd_household
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    hh = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(hh, gov_index) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M1 <- glmer(hh ~ interview_type + gov_index + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M1      <- emmeans(M1, specs = ~ interview_type)
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: tbl-M1_et_fgd_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on whether communities report improved access to natural resources for household use."

re_M1    <- as.data.frame(VarCorr(M1))
prov_M1  <- round(re_M1$vcov[re_M1$grp == "province"], 3)
cfmg_M1  <- round(re_M1$vcov[re_M1$grp == "province_cfmg"], 3)
r2_M1    <- performance::r2_nakagawa(M1, verbose = FALSE)

tidy(M1, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"          ~ "Intercept (Executive FGD)",
    "interview_typeFemale" ~ "Female FGD",
    "interview_typeMale"   ~ "Male FGD",
    "interview_typeYouth"  ~ "Youth FGD",
    "gov_index"            ~ "Governance index"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for improved access to natural resources for household use — FGD type and governance index",
    subtitle = glue::glue(
      "Estimates are log-odds; reference = Executive FGD. ",
      "Province variance = {prov_M1}; CFA-within-province variance = {cfmg_M1}. ",
      "Marginal R² = {round(r2_M1$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M1$R2_conditional, 3)}. ",
      "n = {nobs(M1)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_et_fgd_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| include: true
#| tbl-cap: "Pairwise comparisons among FGD types for improved access to natural resources for household use."

as.data.frame(pairwise.M1) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast="Contrast", estimate="Estimate", SE="SE", z.ratio="z ratio", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, SE, z.ratio), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Pairwise comparisons — FGD type",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-contrast) |>
  cols_align(align="left", columns=contrast) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=contrast)) |>
  sub_missing()
```

Individual interviewees were also asked whether CFM had improved access to natural resources for household use (@tbl-et_individual_access). Similar to the FGD results, a majority of respondents indicated that they did not think that CFM had improved access to natural resources for household use. However, DFOs were significantly more likely to suggest that CFM had improved access than CFMG executives or traditional leaders. None of the other pairwise comparisons were significant.

```{r}
#| label: tbl-et_individual_access
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting improved access to natural resources as an economic benefit of CFM."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    hh = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "Private enterprise rep"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "Private enterprise rep"
    ))
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n         = n(),
    Household = mean(hh, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Household        = "Household use"
  ) |>
  fmt_percent(
    columns  = Household,
    decimals = 0
  ) |>
  tab_header(
    title    = "Improved access to natural resources for household use as reported by individual interviewees",
    subtitle = "Values represent the % of responses agreeing that CFM has improved access to resources for household use. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Household)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: model_et_individual_household
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    hh                 = economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials,
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(hh) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M4 <- glmer(hh ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M4      <- emmeans(M4, specs = ~ interviewee_role)
pairwise.M4 <- pairs(emm.M4, adjust = "Tukey")
```

```{r}
#| label: tbl-M4_et_individual_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on whether individuals report improved access to natural resources for household use."

re_M4    <- as.data.frame(VarCorr(M4))
prov_M4  <- round(re_M4$vcov[re_M4$grp == "province"], 3)
cfmg_M4  <- round(re_M4$vcov[re_M4$grp == "province_cfmg"], 3)
r2_M4    <- performance::r2_nakagawa(M4, verbose = FALSE)

tidy(M4, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"                        ~ "Intercept (Community member)",
    "interviewee_rolecfmg_exec_member"   ~ "CFMG executive member",
    "interviewee_roletrad_leader"        ~ "Traditional leader",
    "interviewee_roledfo"                ~ "District Forestry Officer",
    "interviewee_roleNGO_rep"            ~ "Private enterprise rep",
    "interviewee_gendermale"             ~ "Male",
    "interviewee_age"                    ~ "Age"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for improved access to natural resources for household use — individual interviewees",
    subtitle = glue::glue(
      "Estimates are log-odds; reference = Community member. ",
      "Province variance = {prov_M4}; CFA-within-province variance = {cfmg_M4}. ",
      "Marginal R² = {round(r2_M4$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M4$R2_conditional, 3)}. ",
      "n = {nobs(M4)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_et_individual_household
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| include: true
#| tbl-cap: "Pairwise comparisons among interviewee roles for improved access to natural resources for household use."

as.data.frame(pairwise.M4) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast="Contrast", estimate="Estimate", SE="SE", z.ratio="z ratio", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, SE, z.ratio), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Pairwise comparisons — interviewee role",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-contrast) |>
  cols_align(align="left", columns=contrast) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=contrast)) |>
  sub_missing()
```

### Employment

We asked FGD groups whether CFM had generated employment opportunities (@fig-et_employment).

```{r}
#| label: fig-et_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Responses to whether CFM has generated employment opportunities. The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    response = economic_transformation_yn_formal_employment_opportunities_created_cfm,
    response = case_match(response,
                          "yes" ~ "Yes",
                          "no" ~ "No",
                          .default = NA),
    response = factor(response, levels = c("Yes", "No")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(response) |>
  count(interview_type, response) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c("Yes" = "#2A9D8F", "No" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#economic-transformation)
:::

As can be seen from the figure, approximately half (but only half) of the FGD groups agreed that CFM had increased employment opportunities. Interestingly, community youth were especially likely to agreed that CFM had created employment opportunities. In contrast, community female FGD groups were much more likely to say that CFM had not increased employment opportunities. All pairwise comparisons were highly significant.

```{r}
#| label: model_et_fgd_employment
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    empl = economic_transformation_yn_formal_employment_opportunities_created_cfm,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(empl, gov_index) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M3 <- glmer(empl ~ interview_type + gov_index + (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M3      <- emmeans(M3, specs = ~ interview_type)
pairwise.M3 <- pairs(emm.M3, adjust = "Tukey")
```

```{r}
#| label: tbl-M3_et_fgd_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on whether communities report that CFM has generated employment opportunities."

re_M3    <- as.data.frame(VarCorr(M3))
prov_M3  <- round(re_M3$vcov[re_M3$grp == "province"], 3)
cfmg_M3  <- round(re_M3$vcov[re_M3$grp == "province_cfmg"], 3)
r2_M3    <- performance::r2_nakagawa(M3, verbose = FALSE)

tidy(M3, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"          ~ "Intercept (Executive FGD)",
    "interview_typeFemale" ~ "Female FGD",
    "interview_typeMale"   ~ "Male FGD",
    "interview_typeYouth"  ~ "Youth FGD",
    "gov_index"            ~ "Governance index"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for employment opportunities generated by CFM — FGD type and governance index",
    subtitle = glue::glue(
      "Estimates are log-odds; reference = Executive FGD. ",
      "Province variance = {prov_M3}; CFA-within-province variance = {cfmg_M3}. ",
      "Marginal R² = {round(r2_M3$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M3$R2_conditional, 3)}. ",
      "n = {nobs(M3)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_et_fgd_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| include: true
#| tbl-cap: "Pairwise comparisons among FGD types for whether CFM has generated employment opportunities."

as.data.frame(pairwise.M3) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast="Contrast", estimate="Estimate", SE="SE", z.ratio="z ratio", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, SE, z.ratio), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Pairwise comparisons — FGD type",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-contrast) |>
  cols_align(align="left", columns=contrast) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=contrast)) |>
  sub_missing()
```

Individual interviewees were also asked whether CFM had generated employment opportunities (@tbl-et_individual_employment). Interestingly, community respondents were much more likely to agree that CFM had created employment opportunities than private enterprise representatives or DFOs. The differences between DFOs and CFMG executives and traditional leaders were significant, but none of the other pairwise comparisons were significant.

```{r}
#| label: tbl-et_individual_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that CFM has generated employment opportunities by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_yn_formal_employment_opportunities_created_cfm
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "Private enterprise rep"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "Private enterprise rep"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA)
    
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n          = n(),
    Employment = mean(empl, na.rm = TRUE),
    .groups    = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Employment       = "Agree CFM has generated employment opportunities"
  ) |>
  fmt_percent(
    columns  = Employment,
    decimals = 0
  ) |>
  tab_header(
    title    = "Employment opportunities reported by individual interviewees",
    subtitle = "Values represent the % of responses agreeing that CFM has generated employment opportunities. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Employment)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: model_et_individual_employment
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl               = economic_transformation_yn_formal_employment_opportunities_created_cfm,
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no" ~ 0,
                      .default = NA),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M6 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data, family = binomial)

emm.M6      <- emmeans(M6, specs = ~ interviewee_role)
pairwise.M6 <- pairs(emm.M6, adjust = "Tukey")
```

```{r}
#| label: tbl-M6_et_individual_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on whether individuals report that CFM has generated employment opportunities."

re_M6    <- as.data.frame(VarCorr(M6))
prov_M6  <- round(re_M6$vcov[re_M6$grp == "province"], 3)
cfmg_M6  <- round(re_M6$vcov[re_M6$grp == "province_cfmg"], 3)
r2_M6    <- performance::r2_nakagawa(M6, verbose = FALSE)

tidy(M6, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"                        ~ "Intercept (Community member)",
    "interviewee_rolecfmg_exec_member"   ~ "CFMG executive member",
    "interviewee_roletrad_leader"        ~ "Traditional leader",
    "interviewee_roledfo"                ~ "District Forestry Officer",
    "interviewee_roleNGO_rep"            ~ "Private enterprise rep",
    "interviewee_gendermale"             ~ "Male",
    "interviewee_age"                    ~ "Age"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for employment opportunities generated by CFM — individual interviewees",
    subtitle = glue::glue(
      "Estimates are log-odds; reference = Community member. ",
      "Province variance = {prov_M6}; CFA-within-province variance = {cfmg_M6}. ",
      "Marginal R² = {round(r2_M6$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M6$R2_conditional, 3)}. ",
      "n = {nobs(M6)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_et_individual_employment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| include: true
#| tbl-cap: "Pairwise comparisons among interviewee roles for whether CFM has generated employment opportunities."

as.data.frame(pairwise.M6) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast="Contrast", estimate="Estimate", SE="SE", z.ratio="z ratio", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, SE, z.ratio), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Pairwise comparisons — interviewee role",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-contrast) |>
  cols_align(align="left", columns=contrast) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=contrast)) |>
  sub_missing()
```


#### Full-time employment by gender, age group and province

CFMG Executive FGDs reported on the number of males, females, and youth in full-time employment in the CFMG (@fig-et_fulltime_gender). Clearly, there are many more men in full-time employment than women. According to the responses across the 75 CFMGs surveyed, 278 men were employed compared to 56 women, and 207 male youths compared to 48 female youths.

```{r}
#| label: fig-et_fulltime_gender
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Number of males, females, male youth, and female youth employed full-time in CFMGs, by province, as reported by executive FGDs."
#| fig-width: 9
#| fig-height: 5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    province,
    "Male"         = economic_transformation_formal_employment_quests_formalemploymentdetails_1__males_employed_cfmg_full_time,
    "Female"       = economic_transformation_formal_employment_quests_formalemploymentdetails_1__females_employed_cfmg_full_time,
    "Male youth"   = economic_transformation_formal_employment_quests_formalemploymentdetails_1__male_youth_employed_cfmg_full_time,
    "Female youth" = economic_transformation_formal_employment_quests_formalemploymentdetails_1__female_youth_employed_cfmg_full_time
  ) |>
  pivot_longer(-province, names_to = "category", values_to = "n_employed") |>
  filter(!is.na(n_employed)) |>
  mutate(category = factor(category, levels = c("Male", "Female", "Male youth", "Female youth"))) |>
  group_by(province, category) |>
  summarise(total = sum(n_employed, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(y = reorder(province, total), x = total, fill = category)) +
  geom_col(
    position = position_dodge(width = 0.85), width = 0.75,
    colour = "white", linewidth = 0.3
  ) +
  geom_text(
    aes(label = total),
    position = position_dodge(width = 0.85),
    hjust = -0.2, size = 3, colour = "grey25", fontface = "bold"
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.18)),
    name   = "Number employed full-time"
  ) +
  coord_cartesian(clip = "off") +
  scale_fill_manual(
    values = c(
      "Male"         = "#2A9D8F",
      "Female"       = "#E07B54",
      "Male youth"   = "#76C8C0",
      "Female youth" = "#F0B49A"
    )
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.4),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

#### Types of formal employment opportunities

Community female, male, and youth FGDs identified the types of formal employment opportunities generated by CFM (@fig-et_employment_types). Clearly, by far the largest type of job created is HFO or Community Scout.

```{r}
#| label: fig-et_employment_types
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of formal employment opportunities reported by female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting that job type."
#| fig-width: 10
#| fig-height: 7

prefix <- "economic_transformation_formal_employment_opportunities_from_cfm_"

job_labels <- c(
  "forest_managers"                                          = "Forest managers",
  "forest_rangers"                                           = "Forest rangers",
  "community_scouts"                                         = "Community scouts",
  "hfos"                                                     = "HFOs",
  "monitoring_and_evaluation_officers"                       = "M&E officers",
  "administrative_staff"                                     = "Admin staff",
  "program_coordinators"                                     = "Program coordinators",
  "training_coordinators_facilitators"                       = "Training coordinators",
  "agroforestry_extension_officers"                          = "Agroforestry extension",
  "nursery_managers"                                         = "Nursery managers",
  "nursery_workers"                                          = "Nursery workers",
  "technical_support_roles_e_g_soil_scientists_agronomists"  = "Technical support",
  "eco_tourism_operators_e_g_tour_guides_lodge_staff"        = "Eco-tourism operators",
  "cultural_heritage_and_craft_trainers"                     = "Cultural heritage & craft",
  "craftsmen_and_woodcarvers"                                = "Craftsmen & woodcarvers",
  "processing_plant_staff_e_g_honey_timber_charcoal"         = "Processing natural resources",
  "marketing_and_sales_agents"                               = "Marketing & sales",
  "researchers"                                              = "Researchers",
  "data_analysts"                                            = "Data analysts",
  "surveyors"                                                = "Surveyors",
  "field_data_collectors"                                    = "Field data collectors",
  "biodiversity_monitors"                                    = "Biodiversity monitors",
  "forest_extension_workers"                                 = "Forest extension workers",
  "forest_protection_and_fire_mnagement"                     = "Forest protection & fire",
  "enterprise_and_processing_roles"                          = "Enterprise",
  "training_and_facilitation"                                = "Training & facilitation",
  "security_services"                                        = "Security services",
  "livestock_and_resource_management"                        = "Livestock & resource mgmt",
  "interns_trainees"                                         = "Interns & trainees",
  "technical_extension_roles"                                = "Technical extension"
)

full_var_names <- paste0(prefix, names(job_labels))

binary_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_var_names)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = job_labels[sub(prefix, "", variable)]) |>
  filter(freq > 0) |>
  select(word, freq)

comment_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(text = economic_transformation_formal_employment_opportunities_from_cfm_comment) |>
  filter(!is.na(text), trimws(text) != "") |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  mutate(word = case_match(word,
    "hammer" ~ "Hammer mill",
    "mill"   ~ "Hammer mill",
    "cows"   ~ "Cattle",
    .default = word
  )) |>
  count(word, name = "freq")

wc_data <- bind_rows(binary_freq, comment_freq) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(wc_data, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 30)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

### Income earning opportunities

#### Carbon finance

We asked the CFMG executive FGDs whether their CFMG had received income from carbon finance (@fig-et_carbon_finance). According to the responses, around one third of CFAs have received carbon payments. This is an under-representation for the number of CFMGs involved in carbon finance, which is around 80% of CFMGs, and reflects only those that have actually started to receive payments.

```{r}
#| label: fig-et_carbon_finance
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "CFMG involvement in carbon finance by province, as reported by CFMG executives."
#| fig-width: 8
#| fig-height: 6

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  transmute(
    province,
    response = case_when(
      economic_transformation_cfma_involved_carbon_finance == "yes"                          ~ "Yes",
      str_starts(economic_transformation_cfma_involved_carbon_finance, "Yes")                ~ "Yes",
      economic_transformation_cfma_involved_carbon_finance == "no"                           ~ "No",
      str_starts(economic_transformation_cfma_involved_carbon_finance, "No")                 ~ "No",
      economic_transformation_cfma_involved_carbon_finance %in% c(
        "Not yet", "At the moment_they are not",
        "The cfmg is yet to enter into partnership with the carbon credit organization",
        "The CFMG has not signed form 7 and MOU. They had to cancel the one they had signed earlier due to disagreements in the benefits sharing mechanism for income from carbon credits",
        "Has not signed the agreement with CCS yet due to dispute  over percentages in the benefit sharing mechanism"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(response = factor(response, levels = c("Yes", "No"))) |>
  count(province, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 1.5),
                     name   = "Number of CFMGs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#economic-transformation)
:::

FGD groups were also asked whether the community was aware of the carbon credit agreement (@fig-et_carbon_awareness). While most CFMG executive groups stated that the community was aware of the carbon finance agreement, interestingly two stated that they were not. Most community female and male groups indicated they were aware of the carbon finance agreement, but more than half of the community youth groups stated they were not aware of the carbon finance agreement. Community youth were significantly less likely to know about the carbon finance agreement than CFMG Executives, but other pairwise comparisons were not significant.

```{r}
#| label: fig-et_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Community awareness of the carbon credit agreement by FGD group."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    group = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    ),
    response = case_when(
      economic_transformation_community_awareness_carbon_credit_agreement == "yes" ~ "Yes",
      economic_transformation_community_awareness_carbon_credit_agreement == "no"  ~ "No",
      economic_transformation_community_awareness_carbon_credit_agreement %in% c(
        "Not aware of the agreement of the Juristiction project we just know what's involved",
        "The agreement is not clear we don't know where they sale the carbon and we don't know how much they get from the money",
        "So far we have just heard the chief signed"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(response = factor(response, levels = c("Yes", "No"))) |>
  count(group, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = group, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 3),
                     name   = "Number of FGDs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#economic-transformation)
:::

```{r}
#| label: model_et_fgd_carbon_awareness
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data.ca <- cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    province, cfmg_name, gov_index,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    ),
    empl = case_when(
      economic_transformation_community_awareness_carbon_credit_agreement == "yes" ~ 1L,
      economic_transformation_community_awareness_carbon_credit_agreement == "no"  ~ 0L,
      economic_transformation_community_awareness_carbon_credit_agreement %in% c(
        "Not aware of the agreement of the Juristiction project we just know what's involved",
        "The agreement is not clear we don't know where they sale the carbon and we don't know how much they get from the money",
        "So far we have just heard the chief signed"
      ) ~ 0L,
      TRUE ~ NA_integer_
    )
  ) |>
  drop_na(empl, gov_index) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M11 <- glmer(empl ~ interview_type + gov_index + (1 | province) + (1 | province_cfmg),
            data = test.data.ca, family = binomial)

emm.M11      <- emmeans(M11, specs = ~ interview_type)
pairwise.M11 <- pairs(emm.M11, adjust = "Tukey")
```

```{r}
#| label: tbl-M11_et_fgd_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on community awareness of the carbon credit agreement."

re_M11    <- as.data.frame(VarCorr(M11))
prov_M11  <- round(re_M11$vcov[re_M11$grp == "province"], 3)
cfmg_M11  <- round(re_M11$vcov[re_M11$grp == "province_cfmg"], 3)
r2_M11    <- performance::r2_nakagawa(M11, verbose = FALSE)

tidy(M11, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"          ~ "Intercept (Executive FGD)",
    "interview_typeFemale" ~ "Female FGD",
    "interview_typeMale"   ~ "Male FGD",
    "interview_typeYouth"  ~ "Youth FGD",
    "gov_index"            ~ "Governance index"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for community awareness of carbon credit agreement — FGD type and governance index",
    subtitle = glue::glue(
      "Estimates are log-odds; reference = Executive FGD. ",
      "Province variance = {prov_M11}; CFA-within-province variance = {cfmg_M11}. ",
      "Marginal R² = {round(r2_M11$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M11$R2_conditional, 3)}. ",
      "n = {nobs(M11)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_et_fgd_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| include: true
#| tbl-cap: "Pairwise comparisons among FGD types for community awareness of the carbon credit agreement."

as.data.frame(pairwise.M11) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast="Contrast", estimate="Estimate", SE="SE", z.ratio="z ratio", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, SE, z.ratio), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Pairwise comparisons — FGD type",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-contrast) |>
  cols_align(align="left", columns=contrast) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=contrast)) |>
  sub_missing()
```

Individual interviewees were also asked whether the community was aware of the carbon credit agreement (@tbl-et_individual_carbon_awareness). Most respondents indicated that the community was aware, but the percentage of positive responses varied across interviewee roles. Interestingly, private enterprise representatives overwhelmingly (\~95%) suggested the community was informed, but the only 70% of DFOs thought the communities were informed, although the difference was not significant.

```{r}
#| label: tbl-et_individual_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting community awareness of the carbon credit agreement, by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_community_awareness_carbon_credit_agreement
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "Private enterprise rep"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "Private enterprise rep"
    )),
    empl = case_match(empl,
                      "yes" ~ 1,
                      "no"  ~ 0,
                      .default = NA)
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n         = n(),
    Awareness = mean(empl, na.rm = TRUE),
    .groups   = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Awareness        = "Aware of carbon credit agreement"
  ) |>
  fmt_percent(
    columns  = Awareness,
    decimals = 0
  ) |>
  tab_header(
    title    = "Community awareness of the carbon credit agreement, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Awareness)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: model_et_individual_carbon_awareness
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data.ca.ind <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl = case_match(
      economic_transformation_community_awareness_carbon_credit_agreement,
      "yes" ~ 1,
      "no"  ~ 0,
      .default = NA
    ),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M12 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
              (1 | province) + (1 | province_cfmg),
            data = test.data.ca.ind, family = binomial)

emm.M12      <- emmeans(M12, specs = ~ interviewee_role)
pairwise.M12 <- pairs(emm.M12, adjust = "Tukey")
```

```{r}
#| label: tbl-M12_et_individual_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on community awareness of the carbon credit agreement."

re_M12    <- as.data.frame(VarCorr(M12))
prov_M12  <- round(re_M12$vcov[re_M12$grp == "province"], 3)
cfmg_M12  <- round(re_M12$vcov[re_M12$grp == "province_cfmg"], 3)
r2_M12    <- performance::r2_nakagawa(M12, verbose = FALSE)

tidy(M12, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"                        ~ "Intercept (Community member)",
    "interviewee_rolecfmg_exec_member"   ~ "CFMG executive member",
    "interviewee_roletrad_leader"        ~ "Traditional leader",
    "interviewee_roledfo"                ~ "District Forestry Officer",
    "interviewee_roleNGO_rep"            ~ "Private enterprise rep",
    "interviewee_gendermale"             ~ "Male",
    "interviewee_age"                    ~ "Age"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for community awareness of carbon credit agreement — individual interviewees",
    subtitle = glue::glue(
      "Estimates are log-odds; reference = Community member. ",
      "Province variance = {prov_M12}; CFA-within-province variance = {cfmg_M12}. ",
      "Marginal R² = {round(r2_M12$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M12$R2_conditional, 3)}. ",
      "n = {nobs(M12)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_et_individual_carbon_awareness
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| include: false
#| tbl-cap: "Pairwise comparisons among interviewee roles for community awareness of the carbon credit agreement."

as.data.frame(pairwise.M12) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast="Contrast", estimate="Estimate", SE="SE", z.ratio="z ratio", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, SE, z.ratio), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Pairwise comparisons — interviewee role",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-contrast) |>
  cols_align(align="left", columns=contrast) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=contrast)) |>
  sub_missing()
```

#### Carbon developer support

The types of CFM activities supported by carbon developers are shown in @fig-et_trader_support. As can be seen, support focused on aspects of forest management or agricultural extension. Other potential elements of support, such as NTFPs value chains, infrastructure development or market access, featured substantially less.

```{r}
#| label: fig-et_trader_support
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of CFM activities supported by the carbon trader."
#| fig-width: 8
#| fig-height: 5

prefix_ts <- "economic_transformation_cfm_activities_supported_by_carbon_trader_"

ts_labels <- c(
  "patrolling_and_monitoring"                        = "Patrolling",
  "fire_management"                                  = "Fire management",
  "biodiversity_conservation"                        = "Biodiversity",
  "agroforestry"                                     = "Agroforestry",
  "climate_smart_farming"                            = "Climate-smart farming",
  "non_timber_forest_products_ntfps"                 = "NTFPs",
  "eco_tourism"                                      = "Eco-tourism",
  "tree_planting"                                    = "Tree planting",
  "training_and_capacity_building"                   = "Training",
  "carbon_stock_tracking"                            = "Carbon tracking",
  "market"                                           = "Market access",
  "alternative_livelihood_activities"                = "Livelihoods",
  "infractrure"                                      = "Infrastructure",
  "social_gatherings"                                = "Social gatherings",
  "employment"                                       = "Employment",
  "community_sensitization_awareness_engangement"    = "Sensitisation"
)

full_ts_vars <- paste0(prefix_ts, names(ts_labels))

ts_freq <- cfmg |>
  select(all_of(full_ts_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = ts_labels[sub(prefix_ts, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(ts_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 18)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

#### Other income earning opportunities

FGD groups also reported whether CFM had generated income earning opportunities through specific interventions, including market linkages, alternative livelihood interventions, conservation revenue, forest product processing and ecotourism (@fig-et_income_opportunities). Although some community FGDs indicated improvement in income earning opportunities under some categories, especially conservation revenue and forest product processing, most indicated that CFM had not generated income earning opportunities in these areas.

```{r}
#| label: fig-et_income_opportunities
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Income earning opportunities generated by CFM, as reported by female, male, and youth FGDs. The length of the bars represents the number of FGDs reporting a benefit (Yes) or no benefit (No) for each income source."
#| fig-width: 7
#| fig-height: 5

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  mutate(interview_type = case_match(
    interview_type,
    "cfmg_female" ~ "Female",
    "cfmg_male"   ~ "Male",
    "cfmg_youth"  ~ "Youth"
  )) |>
  select(
    interview_type,
    "Ecotourism income"         = economic_transformation_economic_benefits_from_cfm_income_from_ecotourism_e_g_guided_tours_camping_fees,
    "Forest product processing" = economic_transformation_economic_benefits_from_cfm_income_from_forest_product_processing_e_g_honey_bamboo,
    "Conservation revenue"      = economic_transformation_economic_benefits_from_cfm_revenue_from_forest_conservation_programs_e_g_carbon_credits,
    "Alternative livelihoods"   = `economic_transformation_economic_benefits_from_cfm_alternative-livelihoods_local_business_development_e_g_forest_based_crafts_small_scale_industries`,
    "Market linkages"           = economic_transformation_economic_benefits_from_cfm_market_linkages_and_trade_opportunities_for_forest_products
  ) |>
  pivot_longer(-interview_type, names_to = "income_source", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    income_source = factor(income_source, levels = c(
      "Ecotourism income", "Forest product processing", "Conservation revenue",
      "Alternative livelihoods", "Market linkages"
    )),
    interview_type = factor(interview_type, levels = c("Female", "Male", "Youth"))
  ) |>
  count(interview_type, income_source, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  (\(d) ggplot(d, aes(y = income_source, x = n_signed, fill = response)) +
    geom_blank(data = tibble(
      interview_type = factor(
        c("Female", "Female", "Male", "Male", "Youth", "Youth"),
        levels = c("Female", "Male", "Youth")
      ),
      income_source = factor(
        rep("Ecotourism income", 6),
        levels = c("Ecotourism income", "Forest product processing",
                   "Conservation revenue", "Alternative livelihoods", "Market linkages")
      ),
      n_signed = c(-60, 0, -60, 10, -60, 10),
      response = factor(c("No", "Yes", "No", "Yes", "No", "Yes"), levels = c("No", "Yes"))
    )) +
    geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
    geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
    geom_text(
      aes(label  = n,
          hjust  = if_else(response == "No", -0.4, -0.4),
          colour = response),
      size = 3, fontface = "bold"
    ) +
    scale_colour_manual(
      values = c("No" = "white", "Yes" = "grey25"),
      guide  = "none"
    ) +
    scale_x_continuous(
      labels = ~abs(.x),
      breaks = seq(-60, 10, by = 20),
      expand = expansion(add = 0),
      name   = "Number of FGDs"
    ) +
    coord_cartesian(clip = "off") +
    scale_y_discrete(expand = expansion(add = 0.5)) +
    scale_fill_manual(
      values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
      guide  = guide_legend(reverse = FALSE)
    ) +
    facet_wrap(~ interview_type, scales = "free_x") +
    labs(y = NULL, fill = NULL) +
    theme_minimal(base_size = 18) +
    theme(
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
      axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
      axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
      axis.title.x       = element_text(size = rel(1), colour = "grey40"),
      legend.position    = "bottom",
      legend.text        = element_text(size = rel(1)),
      strip.text         = element_text(size = rel(1.14), face = "bold"),
      plot.margin        = margin(10, 25, 10, 10)
    )
  )()
```

However, the community groups did identified other additional sources of income generated by CFM (@fig-et_income_sources). Natural resource based value chains featured prominently among these, but interestingly community business was also mentioned.

```{r}
#| label: fig-et_income_sources
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Additional sources of income from CFM reported by female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each income source."
#| fig-width: 10
#| fig-height: 7

prefix_income <- "economic_transformatioNAdditional_sources_of_income_additional_income_sources_from_cfm_"

income_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(starts_with(prefix_income)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(
    word = variable |>
      sub(prefix_income, "", x = _, fixed = TRUE) |>
      gsub("_", " ", x = _) |>
      str_to_title(),
    word = case_match(word,
      "Community Businesses And Enterprises" ~ "Community business",
      "Wildlife Based And Safari Income"  ~ "Wildlife based",
      .default = word
    )
  ) |>
  filter(
    freq > 0,
    !str_starts(word, "Rank"),
    !word %in% c("None", "Dont Know")
  ) |>
  select(word, freq)

set.seed(42)
ggplot(income_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

To better understand the importance of different income earning opportunities we asked the community female, male, and youth FGD groups to rank their top three additional sources of income from CFM (@fig-et_income_ranks). Interestingly, among the 14 value chains identified, 12 were mentioned as 1st rank at least once. This immediately points to the diversity of forest products being harvested from CFAs across the country, but also to the fact that for most CFAs only one product features strongly as an addition source of income. Among these, mushrooms and honey are clearly the most frequently mentioned.

```{r}
#| label: fig-et_income_ranks
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Ranking of additional income sources from CFM by female, male, and youth FGDs. Bars show the number of FGDs assigning each rank. Only income sources receiving at least one top-three ranking are shown."
#| fig-width: 9
#| fig-height: 7

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(matches("Additional_sources_of_income.*rank_")) |>
  pivot_longer(everything(), names_to = "variable", values_to = "rank_val") |>
  mutate(
    rank_val = as.numeric(rank_val),
    source   = sub(".*rank_", "", variable) |>
      gsub("_", " ", x = _) |>
      str_to_title(),
    source = case_match(source,
      "Wildlife Based And Safari Income"  ~ "Wildlife Based",
      "Community Businesses And Enterprises" ~ "Community Business",
      .default = source
    )
  ) |>
  filter(rank_val %in% c(1, 2, 3), source != "Carbon Credits") |>
  mutate(rank = factor(paste0("Rank ", rank_val), levels = c("Rank 3", "Rank 2", "Rank 1"))) |>
  count(source, rank) |>
  group_by(source) |>
  mutate(total = sum(n)) |>
  ungroup() |>
  ggplot(aes(y = reorder(source, total), x = n, fill = rank)) +
  geom_col(width = 0.7, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    size = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of FGDs"
  ) +
  scale_fill_manual(
    values = c(
      "Rank 1" = "#264653",
      "Rank 2" = "#E9C46A",
      "Rank 3" = "#E07B54"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.4),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Female, male, and youth FGDs also reported which forest products were harvested for household use (@fig-et_household_products). Mushrooms again featured prominently, but other natural resources are also clearly important, such as wild fruits, medicinal plants, thatching grass and caterpillars.

```{r}
#| label: fig-et_household_products
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Forest products harvested for household use reported by female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each product."
#| fig-width: 10
#| fig-height: 7

prefix_hh <- "economic_transformation_products_harvested_household_use_"

hh_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(starts_with(prefix_hh)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(
    word = variable |>
      sub(prefix_hh, "", x = _, fixed = TRUE) |>
      gsub("_", " ", x = _) |>
      str_to_title()
  ) |>
  filter(
    freq > 0,
    !word %in% c("Carbon Credits", "Dont Know", "None")
  ) |>
  select(word, freq)

set.seed(42)
ggplot(hh_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

CFMG executive FGDs were asked whether the CFMG was collaborating with any private enterprises (@fig-et_collaboration). The answers indicate that only a small number of CFMGs were collaborating with a private enterprise.

```{r}
#| label: fig-et_collaboration
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "CFMG collaboration with private enterprise as reported by executive FGDs by province."
#| fig-width: 8
#| fig-height: 6

collab_theme <- list(
  theme_minimal(base_size = 18),
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.title         = element_text(size = rel(1.14), face = "bold"),
    plot.margin        = margin(10, 25, 10, 10)
  )
)

# Plot 1: private enterprise collaboration (Yes / No)
p1 <- cfmg |>
  filter(interview_type == "cfmg_executive") |>
  transmute(
    province,
    response = case_match(
      economic_transformation_cfmg_private_enterprise_collaboration,
      "yes" ~ "Yes", "no" ~ "No"
    )
  ) |>
  drop_na(response) |>
  mutate(response = factor(response, levels = c("Yes", "No"))) |>
  count(province, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 3.5, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"), guide = "none") +
  scale_x_continuous(labels = ~abs(.x), expand = expansion(add = 1),
                     name = "Number of CFMGs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F")) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL,
       title = "Collaboration with private enterprise") +
  collab_theme

p1
```

::: return-to-summary
[← Return to Summary](Summary.html#economic-transformation)
:::

The value chains that CFMGs reported working with through private enterprise partnerships are shown in @fig-et_value_chains. Sixteen out of 17 involved partnerships with honey off-takers and just one involved mushrooms.

```{r}
#| label: fig-et_value_chains
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Value chains that CFMGs collaborate with through private enterprise partnerships, as reported by executive FGDs."
#| fig-width: 8
#| fig-height: 5

prefix_vc <- "economic_transformation_private_enterprises_cfmg_works_with_"

vc_labels <- c(
  "timber_companies"               = "Timber",
  "sawmills"                       = "Sawmills",
  "furniture_manufacturers"        = "Furniture",
  "wood_processing_enterprises"    = "Wood processing",
  "eco_tourism_operators"          = "Eco-tourism",
  "mushroom_processors_traders"    = "Mushrooms",
  "wild_fruit_processors_traders"  = "Wild fruit",
  "insect_processors_traders"      = "Insects",
  "chikanda_processors_traders"    = "Chikanda",
 # "carbon_credit_traders"          = "Carbon credits",
  "beekeeping_and_honey_producers" = "Honey",
  "charcoal"                       = "Charcoal",
  "botanicals_harvesting"          = "Botanicals"
)

full_vc_vars <- paste0(prefix_vc, names(vc_labels))

vc_freq <- cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(full_vc_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(label = vc_labels[sub(prefix_vc, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(label, freq)

ggplot(vc_freq, aes(x = freq, y = reorder(label, freq))) +
  geom_col(fill = "#2A9D8F", width = 0.6) +
  geom_text(aes(label = freq), hjust = -0.3, size = 4.5, colour = "grey30") +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of CFMGs"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We asked the community female, male, and youth FGD groups to identify economic opportunities that remained unrealised within their CFAs (@fig-et_unrealised). Interestingly, although a wide range of different natural resource related value chains were mentioned, timber was mention the most often. This is noteworthy as none of the CFMGs we surveyed are involved in timber production.

```{r}
#| label: fig-et_unrealised
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Unrealised economic opportunities identified by FGD groups."
#| fig-width: 8
#| fig-height: 5

prefix_ur <- "economic_transformation_economic_oppportunities_unrealized_economic_opportunities_"

ur_labels <- c(
  "timber"                    = "Timber",
  "firewood"                  = "Firewood",
  "charcoal"                  = "Charcoal",
  "honey"                     = "Honey",
  "mushrooms"                 = "Mushrooms",
  "wild_fruits"               = "Wild fruits",
  "caterpillars"              = "Caterpillars",
  "chikanda"                  = "Chikanda",
  "munkoyo"                   = "Munkoyo",
  "lusala_busala"             = "Lusala/Busala",
  "mungongo"                  = "Mungongo",
  "water"                     = "Water",
  "devils_claw"               = "Devil's claw",
  "fish"                      = "Fish",
  "wildlife"                  = "Wildlife",
  "minerals_mining"           = "Minerals",
  "clay"                      = "Clay",
  "sand"                      = "Sand",
  "grass"                     = "Grass",
  "livestock"                 = "Livestock",
  "crops"                     = "Crops",
  "tourism_education_tours"   = "Tourism",
  "permitting"                = "Permitting",
  "infrastruture_utilities"   = "Infrastructure",
  "bamboo"                    = "Bamboo",
  "medicinal_plants"          = "Medicinal plants",
  "bushmeat"                  = "Bushmeat",
  "resins_gums"               = "Resins",
  "thatching_grass"           = "Thatching grass",
  "fibers"                    = "Fibres",
  "eco_tourism"               = "Eco-tourism"
)

full_ur_vars <- paste0(prefix_ur, names(ur_labels))

ur_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_ur_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = ur_labels[sub(prefix_ur, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(ur_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

::: return-to-summary
[← Return to Summary](Summary.html#economic-transformation)
:::

The ranking of these unrealised opportunities is shown in @fig-et_unrealised_ranks. Again a wide range of resources are mentioned, but the top three are timber, honey and mushrooms. Again, it is interesting that community groups consider their timber resources to have high unrealised potential for commercial development.

```{r}
#| label: fig-et_unrealised_ranks
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Ranking of unrealised economic opportunities by FGD groups. Bars show the frequency of rank 1 (highest priority), rank 2, and rank 3 responses."
#| fig-width: 9
#| fig-height: 7

prefix_ur_r <- "economic_transformation_economic_oppportunities_rank_"

ur_rank_labels <- c(
  "timber"                    = "Timber",
  "firewood"                  = "Firewood",
  "charcoal"                  = "Charcoal",
  "honey"                     = "Honey",
  "mushrooms"                 = "Mushrooms",
  "wild_fruits"               = "Wild fruits",
  "caterpillars"              = "Caterpillars",
  "chikanda"                  = "Chikanda",
  "munkoyo"                   = "Munkoyo",
  "lusala_busala"             = "Lusala/Busala",
  "mungongo"                  = "Mungongo",
  "water"                     = "Water",
  "devils_claw"               = "Devil's claw",
  "fish"                      = "Fish",
  "wildlife"                  = "Wildlife",
  "minerals_mining"           = "Minerals",
  "clay"                      = "Clay",
  "sand"                      = "Sand",
  "grass"                     = "Grass",
  "livestock"                 = "Livestock",
  "crops"                     = "Crops",
  "tourism_education_tours"   = "Tourism",
  "permitting"                = "Permitting",
  "bamboo"                    = "Bamboo",
  "medicinal_plants"          = "Medicinal plants",
  "bushmeat"                  = "Bushmeat",
  "resins_gums"               = "Resins",
  "thatching_grass"           = "Thatching grass",
  "fibers"                    = "Fibres",
  "eco_tourism"               = "Eco-tourism"
)

full_ur_rank_vars <- paste0(prefix_ur_r, names(ur_rank_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_ur_rank_vars)) |>
  pivot_longer(everything(), names_to = "variable", values_to = "rank_val") |>
  mutate(
    rank_val = as.numeric(rank_val),
    source   = ur_rank_labels[sub(prefix_ur_r, "", variable, fixed = TRUE)]
  ) |>
  filter(rank_val %in% c(1, 2, 3)) |>
  mutate(rank = factor(paste0("Rank ", rank_val), levels = c("Rank 3", "Rank 2", "Rank 1"))) |>
  count(source, rank) |>
  group_by(source) |>
  mutate(total = sum(n)) |>
  ungroup() |>
  ggplot(aes(y = reorder(source, total), x = n, fill = rank)) +
  geom_col(width = 0.7, colour = "white", linewidth = 0.3) +
  geom_text(
    aes(label = n),
    position  = position_stack(vjust = 0.5),
    size      = 3.5, colour = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c("Rank 1" = "#264653", "Rank 2" = "#E9C46A", "Rank 3" = "#E07B54"),
    guide  = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.05)),
    name   = "Number of responses"
  ) +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 15, 10, 10)
  )
```

FGD groups also identified potential value-added opportunities from CFM resources (@fig-et_value_added). The overlap with the previous two questions is substantial, but it is interesting that the communities specifically identify honey processing and drying mushrooms, as well as timber processing. This suggests that the communities are aware that there are value addition opportunities that they are currently not realising.

```{r}
#| label: fig-et_value_added
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Potential value-added opportunities identified by FGD groups."
#| fig-width: 8
#| fig-height: 5

prefix_va <- "economic_transformation_potential_value_added_opportunities_"

va_labels <- c(
  "timber_processing_e_g_furniture_carpentry_woodcrafts"                                                    = "Timber processing",
  "charcoal_briquetting_eco_friendly_alternatives_to_traditional_charcoal"                                  = "Charcoal briquettes",
  "honey_processing_and_packaging_e_g_branded_honey_beeswax_products"                                      = "Honey processing",
  "mushroom_drying_and_packaging_for_extended_shelf_life_and_market_access"                                 = "Dried mushrooms",
  "wild_fruit_processing_e_g_juices_jams_dried_fruits"                                                     = "Wild fruits",
  "caterpillar_preservation_and_packaging_e_g_dried_or_canned_for_sale"                                    = "Caterpillars",
  "chikanda"                                                                                                 = "Chikanda",
  "munkoyo"                                                                                                  = "Munkoyo",
  "lusala_busala"                                                                                            = "Lusala/Busala",
  "mungongo"                                                                                                 = "Mungongo",
  "devils_claw"                                                                                              = "Devil's claw",
  "palm_leaves"                                                                                              = "Palm leaves",
  "grass"                                                                                                    = "Grass",
  "fish"                                                                                                     = "Fish",
  "minerals"                                                                                                 = "Minerals",
  "bamboo_products_e_g_furniture_handicrafts_construction_materials"                                        = "Bamboo products",
  "medicinal_plant_extraction_e_g_herbal_teas_essential_oils_medicinal_creams"                              = "Medicinal plants",
  "handicrafts_and_woven_products_e_g_baskets_mats_textiles"                                                = "Handicrafts",
  "resin_and_gum_processing_e_g_adhesives_fragrances"                                                       = "Resins",
  "ecotourism_development_e_g_eco_lodges_cultural_experiences_tour_guiding"                                 = "Eco-tourism",
  "agroforestry_based_businesses_e_g_tree_nurseries_sustainable_agriculture"                                = "Agroforestry"
)

full_va_vars <- paste0(prefix_va, names(va_labels))

va_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_va_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = va_labels[sub(prefix_va, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(va_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

We asked the CFMG Executive FGDs to report which primary markets CFM products are sold into (@fig-et_markets). Evidently, most products are sold only at the local market level, with some also being transported to national markets.

```{r}
#| label: fig-et_markets
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Primary markets for CFM products, as reported by executive FGDs."
#| fig-width: 7
#| fig-height: 4

market_labels <- c(
  "economic_transformation_cfm_products_primary_market_local_markets"    = "Local",
  "economic_transformation_cfm_products_primary_market_national_markets" = "National",
  "economic_transformation_cfm_products_primary_market_regional_markets" = "Regional",
  "economic_transformation_cfm_products_primary_market_eu_markets"       = "EU"
)

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(all_of(names(market_labels))) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n") |>
  mutate(market = factor(market_labels[variable],
                         levels = c("Local", "National", "Regional", "EU"))) |>
  ggplot(aes(x = n, y = market)) +
  geom_col(fill = "#2A9D8F", width = 0.55) +
  geom_text(aes(label = n), hjust = -0.3, size = 5, colour = "grey30") +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of CFMGs"
  ) +
  scale_y_discrete(limits = rev) +
  labs(y = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Opportunity costs

#### Precluded activities

The community female, male, and youth FGD groups identified economic activities that are precluded under CFM regulations (@fig-et_precluded). While many of these are self evident (e.g. charcoal burning), the fact that NTFP harvesting is listed prominently is both surprising and concerning. Whether this reflects a misunderstanding on the part of the communities or indeed that they are being told they cannot collect NTFPs is not clear.

```{r}
#| label: fig-et_precluded
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Economic activities precluded by CFM, as reported by FGD groups."
#| fig-width: 8
#| fig-height: 5

prefix_p <- "economic_transformation_economic_activities_precluded_cfm_"

precluded_labels <- c(
  "logging"                              = "Logging",
  "timber_harvesting"                    = "Timber harvesting",
  "hunting"                              = "Hunting",
  "charcoal_production"                  = "Charcoal",
  "commercial_mining"                    = "Mining",
  "farming"                              = "Farming",
  "livestock_grazing"                    = "Livestock grazing",
  "land_conversion"                      = "Land conversion",
  "introduction_of_invasive_species"     = "Invasive species",
  "ntfp_harvesting_of_forest_products"   = "NTFP harvesting",
  "fishing"                              = "Fishing",
  "beekeping_tradtional"                 = "Beekeeping",
  "fire_based_livelihood_activities"     = "Fire livelihoods"
)

full_p_vars <- paste0(prefix_p, names(precluded_labels))

p_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_p_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = precluded_labels[sub(prefix_p, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(p_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#F4A582", high = "#8B0000") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

#### Other opportunity costs

FGD groups were asked whether CFM entails opportunity costs for the community (@fig-et_opp_costs). Interestingly, many of the FGDs did not recognise any opportunity costs. However, about a third of the FGDs did. The proportion of FGDs recognising opportunity costs was highest among community male FGDs and lowest among community female FGDs. However, none of the pairwise comparisons were significant.

```{r}
#| label: fig-et_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFM entails opportunity costs for the community, by FGD group."
#| fig-width: 8
#| fig-height: 5

cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    group = case_match(
      interview_type,
      "cfmg_executive" ~ "Executive",
      "cfmg_female"    ~ "Female",
      "cfmg_male"      ~ "Male",
      "cfmg_youth"     ~ "Youth"
    ),
    response = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"          ~ "Yes",
      economic_transformation_opportunity_costs_of_cfm == "no"           ~ "No",
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ "No",
      economic_transformation_opportunity_costs_of_cfm %in% c(
        "Community forest situated very far from community",
        "The Major livelihood for this community is fishing. They don't depend so much on the forest. They can still get materials for construction in the open areas"
      ) ~ "No",
      TRUE ~ NA_character_
    )
  ) |>
  drop_na(response) |>
  mutate(
    group    = factor(group, levels = c("Youth", "Male", "Female", "Executive")),
    response = factor(response, levels = c("Yes", "No"))
  ) |>
  count(group, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = group, x = n_signed, fill = response)) +
  geom_col(width = 0.6, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label  = n,
        hjust  = if_else(response == "No", -0.4, -0.4),
        colour = response),
    size = 4, fontface = "bold"
  ) +
  scale_colour_manual(values = c("No" = "white", "Yes" = "grey25"),
                      guide = "none") +
  scale_fill_manual(values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
                    guide  = guide_legend(reverse = TRUE)) +
  scale_x_continuous(labels = ~abs(.x),
                     expand = expansion(add = 3),
                     name   = "Number of FGDs") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  coord_cartesian(clip = "off") +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 18) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text          = element_text(colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#economic-transformation)
:::

```{r}
#| label: model_et_fgd_opp_costs
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data.opp <- cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_female",
                                "cfmg_male", "cfmg_youth")) |>
  transmute(
    province, cfmg_name, gov_index,
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    ),
    empl = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"           ~ 1L,
      economic_transformation_opportunity_costs_of_cfm == "no"            ~ 0L,
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ 0L,
      economic_transformation_opportunity_costs_of_cfm %in% c(
        "Community forest situated very far from community",
        "The Major livelihood for this community is fishing. They don't depend so much on the forest. They can still get materials for construction in the open areas"
      ) ~ 0L,
      TRUE ~ NA_integer_
    )
  ) |>
  drop_na(empl, gov_index) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M9 <- glmer(empl ~ interview_type + gov_index + (1 | province) + (1 | province_cfmg),
            data = test.data.opp, family = binomial)

emm.M9      <- emmeans(M9, specs = ~ interview_type)
pairwise.M9 <- pairs(emm.M9, adjust = "Tukey")
```

```{r}
#| label: tbl-M9_et_fgd_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on whether communities report opportunity costs of CFM."

re_M9    <- as.data.frame(VarCorr(M9))
prov_M9  <- round(re_M9$vcov[re_M9$grp == "province"], 3)
cfmg_M9  <- round(re_M9$vcov[re_M9$grp == "province_cfmg"], 3)
r2_M9    <- performance::r2_nakagawa(M9, verbose = FALSE)
r2_M9_marg <- if (!is.null(r2_M9$R2_marginal)   && !is.na(r2_M9$R2_marginal))   round(r2_M9$R2_marginal,   3) else "NA"
r2_M9_cond <- if (!is.null(r2_M9$R2_conditional) && !is.na(r2_M9$R2_conditional)) round(r2_M9$R2_conditional, 3) else "not available (singular model)"

tidy(M9, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"          ~ "Intercept (Executive FGD)",
    "interview_typeFemale" ~ "Female FGD",
    "interview_typeMale"   ~ "Male FGD",
    "interview_typeYouth"  ~ "Youth FGD",
    "gov_index"            ~ "Governance index"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for opportunity costs of CFM — FGD type and governance index",
    subtitle = glue::glue(
      "Estimates are log-odds; reference = Executive FGD. ",
      "Province variance = {prov_M9}; CFA-within-province variance = {cfmg_M9}. ",
      "Marginal R² = {r2_M9_marg}; ",
      "conditional R² = {r2_M9_cond}. ",
      "n = {nobs(M9)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_et_fgd_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| include: false
#| tbl-cap: "Pairwise comparisons among FGD types for whether communities report opportunity costs of CFM."

as.data.frame(pairwise.M9) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast="Contrast", estimate="Estimate", SE="SE", z.ratio="z ratio", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, SE, z.ratio), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Pairwise comparisons — FGD type",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-contrast) |>
  cols_align(align="left", columns=contrast) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=contrast)) |>
  sub_missing()
```

Individual interviewees were also asked whether CFM entails opportunity costs. @tbl-et_individual_opp_costs shows the percentage reporting this broken down by interviewee role.

```{r}
#| label: tbl-et_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Percentage of individual interviewees reporting that CFM entails opportunity costs for the community, by interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  select(
    interviewee_role,
    empl = economic_transformation_opportunity_costs_of_cfm
  ) |>
  mutate(
    interviewee_role = case_match(
      as.character(interviewee_role),
      "cfmg_member"      ~ "Community member",
      "cfmg_exec_member" ~ "CFMG executive member",
      "trad_leader"      ~ "Traditional leader",
      "dfo"              ~ "District Forestry Officer",
      "NGO_rep"          ~ "Private enterprise rep"
    ),
    interviewee_role = factor(interviewee_role, levels = c(
      "Community member", "CFMG executive member", "Traditional leader",
      "District Forestry Officer", "Private enterprise rep"
    )),
    empl = case_when(
      empl == "yes"           ~ 1,
      empl == "no"            ~ 0,
      str_starts(empl, "No_") ~ 0,
      TRUE                    ~ NA_real_
    )
  ) |>
  group_by(interviewee_role) |>
  summarise(
    n             = n(),
    Opp_costs     = mean(empl, na.rm = TRUE),
    .groups       = "drop"
  ) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(
    interviewee_role = "Role",
    n                = "n",
    Opp_costs        = "Report opportunity costs"
  ) |>
  fmt_percent(
    columns  = Opp_costs,
    decimals = 0
  ) |>
  tab_header(
    title    = "Opportunity costs of CFM, as reported by individual interviewees",
    subtitle = "Values represent the % of positive responses. n = the total number of interview responses. A positive response indicates that the respondent considers CFM to entail opportunity costs for the community."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "left",   columns = interviewee_role) |>
  cols_align(align = "center", columns = c(n, Opp_costs)) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: model_et_individual_opp_costs
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data.opp.ind <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    empl = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"           ~ 1,
      economic_transformation_opportunity_costs_of_cfm == "no"            ~ 0,
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ 0,
      TRUE                                                                 ~ NA_real_
    ),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(empl) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M10 <- glmer(empl ~ interviewee_role + interviewee_gender + interviewee_age +
               (1 | province) + (1 | province_cfmg),
             data = test.data.opp.ind, family = binomial)

emm.M10      <- emmeans(M10, specs = ~ interviewee_role)
pairwise.M10 <- pairs(emm.M10, adjust = "Tukey")
```

```{r}
#| label: tbl-M10_et_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on whether individuals report opportunity costs of CFM."

re_M10    <- as.data.frame(VarCorr(M10))
prov_M10  <- round(re_M10$vcov[re_M10$grp == "province"], 3)
cfmg_M10  <- round(re_M10$vcov[re_M10$grp == "province_cfmg"], 3)
r2_M10    <- performance::r2_nakagawa(M10, verbose = FALSE)

tidy(M10, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"                        ~ "Intercept (Community member)",
    "interviewee_rolecfmg_exec_member"   ~ "CFMG executive member",
    "interviewee_roletrad_leader"        ~ "Traditional leader",
    "interviewee_roledfo"                ~ "District Forestry Officer",
    "interviewee_roleNGO_rep"            ~ "Private enterprise rep",
    "interviewee_gendermale"             ~ "Male",
    "interviewee_age"                    ~ "Age"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for opportunity costs of CFM — individual interviewees",
    subtitle = glue::glue(
      "Estimates are log-odds; reference = Community member. ",
      "Province variance = {prov_M10}; CFA-within-province variance = {cfmg_M10}. ",
      "Marginal R² = {round(r2_M10$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_M10$R2_conditional, 3)}. ",
      "n = {nobs(M10)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_et_individual_opp_costs
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| include: false
#| tbl-cap: "Pairwise comparisons among interviewee roles for whether individuals report opportunity costs of CFM."

as.data.frame(pairwise.M10) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast="Contrast", estimate="Estimate", SE="SE", z.ratio="z ratio", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, SE, z.ratio), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Pairwise comparisons — interviewee role",
    subtitle = "Estimates are log-odds differences on the emmeans scale. Tukey-adjusted p-values."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-contrast) |>
  cols_align(align="left", columns=contrast) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=contrast)) |>
  sub_missing()
```

The types of opportunity costs identified are shown in @fig-et_opp_cost_types. Mentions of reduced access, land restrictions and land use conflict are by some margin the most frequently identified opportunity costs. However, it is also interesting that time costs and diverting community resources are also quite prevalent. CFM requires that community members meet which entails a time cost and community resources are diverted to managing the forest that might otherwise be spent on other activities, such as agriculture.

```{r}
#| label: fig-et_opp_cost_types
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of opportunity costs experienced under CFM."
#| fig-width: 8
#| fig-height: 5

prefix_oc <- "economic_transformation_main_opportunity_costs_experienced_"

oc_labels <- c(
  "reduced_access_to_forest_resources_e_g_firewood_building_materials"    = "Reduced access",
  "time_spent_on_cfm_activities_instead_of_income_generating_work"        = "Time cost",
  "conflict_with_other_land_resource_uses"                                = "Land use conflict",
  "restrictions_on_land_use_for_farming_or_other_activities"              = "Land restrictions",
  "economic_costs_from_diverting_community_resources_to_cfm"              = "Diverting resources"
)

full_oc_vars <- paste0(prefix_oc, names(oc_labels))

binary_oc <- cfmg |>
  select(all_of(full_oc_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = oc_labels[sub(prefix_oc, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

comment_oc <- cfmg |>
  select(text = economic_transformation_main_opportunity_costs_experienced_comment) |>
  filter(!is.na(text), trimws(text) != "") |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  mutate(word = case_match(
    word,
    c("charcoal", "production", "prohibition", "stopped",
      "restriction", "restrictions") ~ "Charcoal",
    "chitemene"                        ~ "Chitemene",
    "grazing"                          ~ "Grazing",
    "timber"                           ~ "Timber",
    "logging"                          ~ "Logging",
    "fruits"                           ~ "Fruits",
    "medicine"                         ~ "Medicinal plants",
    .default                           = word
  )) |>
  filter(!word %in% c("reduced", "involved", "traditional",
                       "interference", "access", "source",
                       "villages", "moment")) |>
  count(word, name = "freq")

wc_oc <- bind_rows(binary_oc, comment_oc) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(wc_oc, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 18)) +
  scale_colour_gradient(low = "#F4A582", high = "#8B0000") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

## Economic transformation index

To summarise the overall economic transformation performance of each CFMG, we constructed an economic transformation score (0–5) by summing five binary indicators reported by FGD groups — improved access to natural resources for household use, new employment opportunities created, carbon payments received (executive FGDs only), collaboration with a private enterprise (executive FGDs only), and absence of opportunity costs — scored as 1 (positive outcome reported) or 0 (no positive outcome or worsening). The distribution of economic transformation scores across CFMGs is shown in @fig-et_index_dist (Appendix III). We then modelled this score using a linear mixed model with the governance index as the predictor, and random intercepts for province and CFA within province, to test whether governance was a significant predictor of overall economic transformation performance (@tbl-et_index_model, @fig-et_index).

```{r}
#| label: data_et_index
#| include: false
#| warning: false
#| message: false
#| error:   false

et_data <- cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_male", "cfmg_female", "cfmg_youth")) |>
  mutate(
    si_hh = case_when(
      as.character(economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials) == "1" ~ 1,
      as.character(economic_transformation_economic_benefits_from_cfm_improved_access_to_natural_resources_for_household_use_e_g_firewood_building_materials) == "0" ~ 0,
      TRUE ~ NA_real_
    ),
    si_carbon = case_when(
      economic_transformation_cfma_involved_carbon_finance == "yes"                          ~ 1,
      str_starts(economic_transformation_cfma_involved_carbon_finance, "Yes")                ~ 1,
      economic_transformation_cfma_involved_carbon_finance == "no"                           ~ 0,
      str_starts(economic_transformation_cfma_involved_carbon_finance, "No")                 ~ 0,
      economic_transformation_cfma_involved_carbon_finance %in% c(
        "Not yet", "At the moment_they are not",
        "The cfmg is yet to enter into partnership with the carbon credit organization",
        "The CFMG has not signed form 7 and MOU. They had to cancel the one they had signed earlier due to disagreements in the benefits sharing mechanism for income from carbon credits",
        "Has not signed the agreement with CCS yet due to dispute  over percentages in the benefit sharing mechanism"
      ) ~ 0,
      TRUE ~ NA_real_
    ),
    si_empl = case_when(
      economic_transformation_yn_formal_employment_opportunities_created_cfm == "yes" ~ 1,
      economic_transformation_yn_formal_employment_opportunities_created_cfm == "no"  ~ 0,
      TRUE ~ NA_real_
    ),
    si_enterprise = case_match(
      economic_transformation_cfmg_private_enterprise_collaboration,
      "yes" ~ 1,
      "no"  ~ 0,
      .default = NA_real_
    ),
    si_opp_costs = case_when(
      economic_transformation_opportunity_costs_of_cfm == "yes"           ~ 0,
      economic_transformation_opportunity_costs_of_cfm == "no"            ~ 1,
      str_starts(economic_transformation_opportunity_costs_of_cfm, "No_") ~ 1,
      economic_transformation_opportunity_costs_of_cfm %in% c(
        "Community forest situated very far from community",
        "The Major livelihood for this community is fishing. They don't depend so much on the forest. They can still get materials for construction in the open areas"
      ) ~ 1,
      TRUE ~ NA_real_
    ),
    et_score = ifelse(
      rowSums(!is.na(cbind(si_hh, si_carbon, si_empl, si_enterprise, si_opp_costs))) == 0,
      NA_real_,
      rowSums(cbind(si_hh, si_carbon, si_empl, si_enterprise, si_opp_costs), na.rm = TRUE)
    ),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    ),
    province_cfmg = interaction(province, cfmg_name)
  ) |>
  drop_na(et_score, gov_index)

M_et <- lmer(
  et_score ~ gov_index + (1 | province) + (1 | province_cfmg),
  data = et_data,
  REML = TRUE
)
```

```{r}
#| label: fig-et_index_dist
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Distribution of mean economic transformation scores across CFMGs (0–5). Each bar represents the number of CFMGs with that mean score, averaged across all FGD types. The score sums five binary indicators: improved household resource access, new employment opportunities, carbon payments received, collaboration with a private enterprise, and absence of opportunity costs."
#| fig-height: 3

et_data |>
  group_by(cfmg_name) |>
  summarise(et_score = mean(et_score, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = et_score)) +
  geom_histogram(binwidth = 0.5, fill = "#2A9D8F", colour = "white", linewidth = 0.4) +
  scale_x_continuous(breaks = 0:5, name = "Economic transformation score (out of 5)") +
  scale_y_continuous(name = "Number of CFMGs") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5)
  )
```

```{r}
#| label: tbl-et_index_model
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of the governance index on economic transformation performance."

re_et    <- as.data.frame(VarCorr(M_et))
prov_et  <- round(re_et$vcov[re_et$grp == "province"], 3)
cfmg_et  <- round(re_et$vcov[re_et$grp == "province_cfmg"], 3)
resid_et <- round(re_et$vcov[re_et$grp == "Residual"], 3)
r2_et    <- performance::r2_nakagawa(M_et, verbose = FALSE)

tidy(M_et, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)" ~ "Intercept",
    "gov_index"   ~ "Governance index"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term="Factor", estimate="Estimate", std.error="SE", p.value=md("*P* value")) |>
  fmt_number(columns=c(estimate, std.error), decimals=3) |>
  fmt_number(columns=p.value, decimals=4) |>
  tab_header(
    title    = "Model results for economic transformation performance — governance index",
    subtitle = glue::glue(
      "Linear mixed model with random intercepts for province and CFA within province. ",
      "Province variance = {prov_et}; CFA-within-province variance = {cfmg_et}; ",
      "residual variance = {resid_et}. ",
      "Marginal R² = {round(r2_et$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_et$R2_conditional, 3)}. ",
      "n = {nobs(M_et)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align="center", columns=-term) |>
  cols_align(align="left", columns=term) |>
  tab_style(style=cell_text(align="center"), locations=cells_column_labels(columns=term)) |>
  sub_missing()
```

```{r}
#| label: fig-et_index
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Relationship between the CFMG governance index and economic transformation score. Each point represents one FGD session. The line shows the fitted relationship from the linear mixed model, with the shaded band representing ± 1 SE."
#| fig-height: 4.5
#| fig-width: 6

et_data |>
  ggplot(aes(x = gov_index, y = et_score)) +
  geom_jitter(width = 0.15, height = 0.1, colour = "#2A9D8F", alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", colour = "#1A6B63", fill = "#76B7A0", alpha = 0.25,
              linewidth = 0.8) +
  scale_x_continuous(name = "Governance index (out of 10)", breaks = 0:10) +
  scale_y_continuous(name = "Economic transformation score (out of 5)", breaks = 0:5,
                     limits = c(-0.2, 5.2)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5)
  )
```

As can be seen, the governance index was a highly significant predictor of economic transformaion score, indicating that in better governance CFMGs the community also perceived better economic benefit from CFM.

## Discussion

Through the FGDs and the individual interviews, we investigated a diverse set of indicators of economic transformation. We covered access to natural resources for household use and sale, employment, income earning opportunities, including carbon finance and other sources of income, and opportunity costs.

Most FGD groups and individual interviewees indicated that CFM had not increased access to natural resources for either household use or sale. However, given that communities  have customary rights of access and were involved in harvesting forest products before the instigation of CFM, it is not surprising that most respondents do not recognise any change. Nevertheless, around one third of respondent groups and individuals felt that access had improved under CFM, presumably reflecting formalisation of community rights to access forest resources.

A larger proportion of FGD groups considered that CFM had had a positive effect on full-time employment. However, the differences among FGD types were highly significant. More community female groups suggested there was no change in employment and more community youth groups felt full-time formal employment opportunities had increased. We also showed that the numbers employed included many more men than women, which would explain these differences in perception. Moreover, while CFM does appear to have increased formal employment, the quality of jobs was poor — most being HFOs or community scouts — and the numbers are paltry. In part, this may reflect the fact that employment in forest product harvesting and processing is rarely formal or full-time but, as we also found, there has been very limited development of forest resource based enterprises. Thus, while CMF does appear to have resulted in a small number of new employment opportunities, these have been largely limited to low paying jobs and have not benefited many women.

A large proportion of CFAs in Zambia were formed on the promise of carbon dollars. However, only approximately one third of the CFAs in our sample appeared to have benefited from carbon payments at the time of the survey. This number reflects the actual disbursement of carbon derived income to the community, as there clearly has been investment in communities by carbon developers. The fact that a number of the FGD groups, especially youth groups, were unaware of the carbon agreement is concerning. It suggests that the carbon developers may have failed to conduct comprehensive FPIC processes involving the entire community and that the CFMG executives may not be communicating properly with their communities. Nevertheless, many carbon developers are clearly supporting communities. Most activities concern forest protection and agricultural extension for sustainable intensification, such as climate smart agriculture. Both of which may be considered as essential to a carbon business plan. However, NTFPs and forest resource based enterprise development were far less often mentioned. We suggest this is an oversight on the part of carbon developers. Enhancing forest product value chains would increase the value of forest conservation for local communities. Furthermore, carbon finance alone is unlikely to fund development in forest margin communities sufficiently. Hence, there is a significant risk that communities will become disillusioned with carbon finance, especially among sections of the community that feel there is an inequitable sharing of benefits. Investing in forest based enterprises should be considered essential to a carbon business plan.

The above results are backed up by the response to questions concerning other income earning opportunities. It seems CFM has not led to increased income from market linkages, alternative livelihoods, conservation revenue, forest product processing or ecotourism except for in a very few cases. Nevertheless, a wide variety of forest products clearly contribute to household use and some additional income. Although honey and mushrooms were most often identified as the most important source of additional income, another ten NTFPs were mention as the most important at least once. However, these additional sources of income were identified in less than half of the CFAs and only 13 CFAs identified a second ranked forest product. Thus, forest products are only rarely being exploited as a source of additional income and if they are it is usually just one. Only a small number of CFMGs (23%) are partnering with an enterprise, such as an off-taker, and among those that do it almost exclusively concerns honey. The communities identified a wide diversity of forest products that could potentially be harvested and processed to provide an additional source of income. Prominent among these were timber and wild fruit, as well as mushrooms and honey. In some cases timber harvesting may conflict with carbon business plans, but not if plantations were established on degraded land. The potential of CFM to support forest product enterprises is clearly massively under realised.

To get a more accurate picture of forest product values a more in depth PEN type survey, involving surveys of household forest product use repeated multiple times over a year, is recommended. The intensity of the survey would necessitate sub-sampling communities from our survey. However, the results of our survey could be used to stratify sampling, so that results could be extrapolated to the national level.

The local communities identified a number of opportunity costs to CFM. Most of these pertained to activities that are precluded, such as those involving forest clearance. However, many groups also identified other opportunity costs, such as time costs and diverted resources. Recognition of opportunity costs is important to understand the cost-benefit of CFM for communities, especially as these costs are likely to be unevenly distributed among community members. It was concerning that many groups mentioned restricted access to NTFPs. Whether this results from a misunderstanding of the regulations surrounding CFM or is because community members are actually being barred from harvesting NTFPs needs to be resolved. CFM should facilitate access to resources and improve their management, not prevent access.

It is noteworthy that the governance index was a highly significant predictor of our economic transformation index. Thus, as with our results from the chapter on SFM, performance is predicated on good governance. It is imperative that the FD and other CFMG partners step up their oversight of CFMG governance to ensure long term positive outcomes from CFM.

Overall, a somewhat positive picture of economic transformation emerges from the responses of the FGDs and individual interviews. However, the responses also indicate that there are huge missed opportunities to diversity income sources and develop forest based enterprises. While changes in indicators, such as employment or additional income from forest products, are not substantial, they are in a positive direction. In the future, most communities will also receive carbon income, even if this has only benefited around one third of CFAs so far. However, CFMG business plans are clearly over-reliant on carbon or sometimes carbon and honey. Substantial opportunities for developing much more diversified business plans exist. These could include a range of different NTFPs and potentially also timber or sustainable charcoal. In addition to simply increasing income earning opportunities, diversified business plans would also provide a buffer against fluctuating commodity prices and because different products are harvested at different times of the year, they can help stabilise income throughout the year.
