---
title: "Sustainable Forest Management"
author: "Rhett D Harrison"
format:
  html:
    css: custom.css
  pdf:
    include-in-header:
      text: |
        \pagenumbering{arabic}
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ordinal)
library(ggwordcloud)

load("cfmg.RData")
```

## Summary

Most FGD respondents indicated that deforestation, charcoal production and fire frequency were decreasing and wildlife populations were increasing with their CFs, and natural resource management was improving within their CFs. In addition, most FGD groups indicated that both deforestation and charcoal production were decreasing outside their CF, suggesting there is no leakage or potentially a positive leakage effect. Community female FGDs tended to have a more positive view of the impact of CFM on SFM compared to the other groups, which is noteworthy given it is primarily the women who harvest forest products, but who often do not have voice in traditional natural resource management. However, it is also important to note that a minority of FGD groups and individual interviewees suggested the situation was worsening in their CFs. Furthermore, the governance index was a significant predictor of SFM performance, at least according to the perceptions expressed by the FGDs. Not many CFMGs were practising restoration or establishing plantations. Existing CFs should be expanded and new ones designed so that they incorporate degraded areas and contribute to national restoration and plantation targets.

::: {.callout-note icon="false"}
## Key conclusions

According to the perceptions expressed by the FGDs, CFM in Zambia is improving SFM across a substantial majority of surveyed communities, with positive spillover effects in the surrounding landscape and encouraging indications of women's engagement in forest governance. However, the survey identified a minority of CFs in which the situation was worsening and the governance index was a significant predictor of SFM performance. The identified link between governance quality and SFM performance is critical and underscores the need for improving governance across the CFM portfolio to ensure Zambia can solidify existing gains and extend them. Zambia could also benefit from a systematic approach to restoration and plantation through CFM.
:::

## Introduction

In this chapter, we consider the effect of community forestry on indicators of SFM. The information analysed is based entirely only on the responses of interviewees to our survey tool. In the next chapter, we will consider the effects of CF on SMF as measured through remote sensing derived metrics.

### Deforestation

We asked interviewees whether CFM has reduced deforestation inside their CF and whether or not deforestation has increased outside their CF.

#### Deforestation inside the community forest

```{r}
#| label: fig-reduced_forest_clearing
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: " 'Has CFM reduced deforestation inside the community forest?' The length of each bar segment represents the proportion of FGD groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    deforestation = case_match(
            impact_on_sfm_cfm_reduced_forest_clearing_cfma,
            "no_there_has_never_been_clearing_for_agriculture" ~ "No change",
            "Agricultural practices ceased way before CFM" ~ "No change",
            "One field for cassava" ~ "No change",
            "Yes there no more fields_  it's all forests" ~ "Significant",
            "Yes_ significant reduction. There was chitemene system before CFM" ~ "Significant",
            "no_increase" ~ "No change",
            "yes_moderate" ~ "Moderate",
            "yes_significant" ~ "Significant",
            .default = NA
    ),
    deforestation = factor(deforestation, levels = c("Significant","Moderate","No change")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(deforestation) |>
  count(interview_type, deforestation) |>
  group_by(interview_type) |>
  mutate(
    prop = n / sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = deforestation)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "No change"         = "#C0392B",
      "Moderate" = "#F0C040",
      "Significant"        = "#2A9D8F"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#sustainable-forest-management)
:::

A majority of groups reported either a moderate or significant reduction in deforestation inside the CF (@fig-reduced_forest_clearing). Moreover, these answers were consistent across the FGD groups. There were no significant differences in response among FGD types (@tbl-M1_deforestation_fgd). The quality of governance, as measured by the governance index we generated in the previous chapter, was a highly significant predictor of the perceived decrease in deforestation within the CF (@tbl-M1_deforestation_fgd).

```{r}
#| label: model_deforestation_fgd
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_cfm_reduced_forest_clearing_cfma,
      "no_there_has_never_been_clearing_for_agriculture" ~ "No change",
      "Agricultural practices ceased way before CFM"     ~ "No change",
      "One field for cassava"                            ~ "No change",
      "Yes there no more fields_  it's all forests"      ~ "Significant",
      "Yes_ significant reduction. There was chitemene system before CFM" ~ "Significant",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("No change", "Moderate", "Significant"),
                           ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(deforestation) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M1 <- clmm(deforestation ~
              interview_type +
              gov_index +
              (1 | province) +
              (1 | province_cfmg),
            data = test.data)

emm.M1     <- emmeans(M1, specs = ~ interview_type) #FGD type effect was not significant so remove pairwise comparison
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: tbl-M1_deforestation_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on perceived reduction in deforestation inside the CF."

tidy(M1) |>
  filter(coef.type == "location") |>
  mutate(term = case_match(term,
    "interview_typeFemale" ~ "Female",
    "interview_typeMale"   ~ "Male",
    "interview_typeYouth"  ~ "Youth",
    "gov_index"            ~ "Governance index",
    .default = term
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
             p.value = md("*P* value")) |>
  fmt_number(columns = c(estimate, std.error), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Model results for perceived reduction in deforestation inside the CF — FGD type and governance index",
    subtitle = glue::glue("We used a CLMM model with random intercepts for province and CF within province. Reference = Executive. n = {nobs(M1)}.")
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -term) |>
  cols_align(align = "left",   columns =  term) |>
  tab_style(style = cell_text(align = "center"),
            locations = cells_column_labels(columns = term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_deforestation_fgd
#| include: false
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived reduction in deforestation inside the CF — FGD type."

as.data.frame(summary(pairwise.M1)) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
             z.ratio = "z ratio", p.value = md("*P* value")) |>
  fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Pairwise comparisons for perceived reduction in deforestation inside the CF — FGD type",
    subtitle = "Tukey HSD adjustment."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -contrast) |>
  cols_align(align = "left",   columns =  contrast) |>
  tab_style(style = cell_text(align = "center"),
            locations = cells_column_labels(columns = contrast)) |>
  sub_missing()
```

Among the individual interviewees (@tbl-deforestation_individual), a majority of interviewees also indicated that CFM had reduced deforestation inside the CF. The differences among the interviewee roles were not significant (@tbl-M2_deforestation_individual). However, there was a significant effect of gender. Male interviewees were significantly more likely to state that CFM had reduced deforestation (@tbl-M2_deforestation_individual).

```{r}
#| label: tbl-deforestation_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived reduction in deforestation inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_cfm_reduced_forest_clearing_cfma,
      "no_there_has_never_been_clearing_for_agriculture" ~ "No change",
      "Agricultural practices ceased way before CFM"     ~ "No change",
      "One field for cassava"                            ~ "No change",
      "Yes there no more fields_  it's all forests"      ~ "Significant",
      "Yes_ significant reduction. There was chitemene system before CFM" ~ "Significant",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "Private enterprise rep"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "Private enterprise rep")
    )
  ) |>
  filter(!is.na(deforestation)) |>
  count(interviewee_role, deforestation, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, deforestation, prop) |>
  pivot_wider(names_from = deforestation, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(`No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived reduction in deforestation inside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: test_data_deforestation_individual
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_cfm_reduced_forest_clearing_cfma,
      "no_there_has_never_been_clearing_for_agriculture" ~ "No change",
      "Agricultural practices ceased way before CFM"     ~ "No change",
      "One field for cassava"                            ~ "No change",
      "Yes there no more fields_  it's all forests"      ~ "Significant",
      "Yes_ significant reduction. There was chitemene system before CFM" ~ "Significant",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("No change", "Moderate", "Significant"),
                           ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(deforestation) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M2 <- clmm(deforestation ~
              interviewee_role +
              interviewee_gender +
              interviewee_age +
              (1 | province) +
              (1 | province_cfmg),
            data = test.data)

emm.M2     <- emmeans(M2, specs = ~ interviewee_role)
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")
```

```{r}
#| label: tbl-M2_deforestation_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on perceived reduction in deforestation inside the CF."

tidy(M2) |>
  filter(coef.type == "location") |>
  mutate(term = case_match(term,
    "interviewee_rolecfmg_exec_member" ~ "CFMG executive member",
    "interviewee_roletrad_leader"      ~ "Traditional leader",
    "interviewee_roledfo"              ~ "DFO",
    "interviewee_roleNGO_rep"          ~ "Private enterprise rep",
    "interviewee_gendermale"           ~ "Gender: male",
    "interviewee_age"                  ~ "Age",
    .default = term
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
             p.value = md("*P* value")) |>
  fmt_number(columns = c(estimate, std.error), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Model results for perceived reduction in deforestation inside the CF — interviewee role, gender and age",
    subtitle = glue::glue("We used a CLMM model with random intercepts for province and CF within province. Reference role = Community member; reference gender = female. n = {nobs(M2)}.")
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -term) |>
  cols_align(align = "left",   columns =  term) |>
  tab_style(style = cell_text(align = "center"),
            locations = cells_column_labels(columns = term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_deforestation_individual
#| include: false
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived reduction in deforestation inside the CF — interviewee role."

as.data.frame(summary(pairwise.M2)) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  mutate(contrast = gsub("cfmg_exec_member", "CFMG exec member", contrast),
         contrast = gsub("cfmg_member",      "Community member",  contrast),
         contrast = gsub("trad_leader",      "Trad leader",       contrast),
         contrast = gsub("NGO_rep",          "NGO rep",           contrast)) |>
  gt() |>
  cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
             z.ratio = "z ratio", p.value = md("*P* value")) |>
  fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Pairwise comparisons for perceived reduction in deforestation inside the CF — interviewee role",
    subtitle = "Tukey HSD adjustment."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -contrast) |>
  cols_align(align = "left",   columns =  contrast) |>
  tab_style(style = cell_text(align = "center"),
            locations = cells_column_labels(columns = contrast)) |>
  sub_missing()
```

#### Deforestation outside the community forest

Sometimes when activities are proscribed in one area they may be displaced other areas; an effect known as leakage. To investigate possible leakage effects, we also asked interviewees whether CFM had led to an increase in deforestation outside their community forest (@fig-increased_deforestation_outside).

```{r}
#| label: fig-increased_deforestation_outside
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: " 'Has CFM increased deforestation outside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_forest_clearing_increased_outside_cfma,
      "Clearing only done for field opening" ~ "No change",
      "In Mboshya there are no agricultural activities" ~ "No change",
      "It increases according to seasons_ especially during the tobacco harvesting" ~ "Moderate",
      "No change_ community members' concentration is on farming. They have appreciation for gover…" ~ "No change",
      "There has never been  cutting of trees out side the cfma" ~ "No change",
      "There has never been any forest clearing outside the cfma." ~ "No change",
      "There has never been cutting of trees" ~ "No change",
      "no_decrease" ~ "Decrease",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("Significant", "Moderate", "No change", "Decrease")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(deforestation) |>
  count(interview_type, deforestation) |>
  group_by(interview_type) |>
  mutate(
    prop = n / sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = deforestation)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Decrease" = "#2A9D8F",
      "No change"   = "#76B7A0",
      "Moderate"    = "#F0C040",
      "Significant" = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1),    colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#sustainable-forest-management)
:::

Contrary to the expectations of leakage, a majority of the FGD groups responded that CFM had actually reduced deforestation outside their CF. This result was consistent across FGD groups and the effect of governance index was not significant (@tbl-M3_deforestation_outside_fgd; (@tbl-pairwise_deforestation_outside_fgd).

```{r}
#| label: model_deforestation_outside_fgd
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_forest_clearing_increased_outside_cfma,
      "Clearing only done for field opening" ~ "No change",
      "In Mboshya there are no agricultural activities" ~ "No change",
      "It increases according to seasons_ especially during the tobacco harvesting" ~ "Moderate",
      "No change_ community members' concentration is on farming. They have appreciation for gover…" ~ "No change",
      "There has never been  cutting of trees out side the cfma" ~ "No change",
      "There has never been any forest clearing outside the cfma." ~ "No change",
      "There has never been cutting of trees" ~ "No change",
      "no_decrease" ~ "Decrease",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("Significant", "Moderate", "No change", "Decrease")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(deforestation) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M3 <- clmm(deforestation ~
              interview_type +
              gov_index +
              (1 | province) +
              (1 | province_cfmg),
            data = test.data)

emm.M3     <- emmeans(M3, specs = ~ interview_type)
pairwise.M3 <- pairs(emm.M3, adjust = "Tukey")
```

```{r}
#| label: tbl-M3_deforestation_outside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on perceived increase in deforestation outside the CF."

tidy(M3) |>
  filter(coef.type == "location") |>
  mutate(term = case_match(term,
    "interview_typeFemale" ~ "Female",
    "interview_typeMale"   ~ "Male",
    "interview_typeYouth"  ~ "Youth",
    "gov_index"            ~ "Governance index",
    .default = term
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
             p.value = md("*P* value")) |>
  fmt_number(columns = c(estimate, std.error), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Model results for perceived increase in deforestation outside the CF — FGD type and governance index",
    subtitle = glue::glue("We used a CLMM model with random intercepts for province and CF within province. Reference = Executive. n = {nobs(M3)}.")
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -term) |>
  cols_align(align = "left",   columns =  term) |>
  tab_style(style = cell_text(align = "center"),
            locations = cells_column_labels(columns = term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_deforestation_outside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived increase in deforestation outside the CF — FGD type."

as.data.frame(summary(pairwise.M3)) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  gt() |>
  cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
             z.ratio = "z ratio", p.value = md("*P* value")) |>
  fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Pairwise comparisons for perceived increase in deforestation outside the CF — FGD type",
    subtitle = "Tukey HSD adjustment."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -contrast) |>
  cols_align(align = "left",   columns =  contrast) |>
  tab_style(style = cell_text(align = "center"),
            locations = cells_column_labels(columns = contrast)) |>
  sub_missing()
```

Among the individual interviewees (@tbl-deforestation_outside_individual), the individuals from the communities agreed that deforestation outside their CF had decreased. However, DFOs were less likely to suggest deforestation had decreased outside the CF.

```{r}
#| label: tbl-deforestation_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived increase in deforestation outside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_forest_clearing_increased_outside_cfma,
      "Clearing only done for field opening" ~ "No change",
      "In Mboshya there are no agricultural activities" ~ "No change",
      "It increases according to seasons_ especially during the tobacco harvesting" ~ "Moderate",
      "No change_ community members' concentration is on farming. They have appreciation for gover…" ~ "No change",
      "There has never been  cutting of trees out side the cfma" ~ "No change",
      "There has never been any forest clearing outside the cfma." ~ "No change",
      "There has never been cutting of trees" ~ "No change",
      "no_decrease" ~ "Decrease",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("Decrease", "No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "Private enterprise rep"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "Private enterprise rep")
    )
  ) |>
  filter(!is.na(deforestation)) |>
  count(interviewee_role, deforestation, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, deforestation, prop) |>
  pivot_wider(names_from = deforestation, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Decrease, `No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived increase in deforestation outside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: test_data_deforestation_outside_individual
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_forest_clearing_increased_outside_cfma,
      "Clearing only done for field opening" ~ "No change",
      "In Mboshya there are no agricultural activities" ~ "No change",
      "It increases according to seasons_ especially during the tobacco harvesting" ~ "Moderate",
      "No change_ community members' concentration is on farming. They have appreciation for gover…" ~ "No change",
      "There has never been  cutting of trees out side the cfma" ~ "No change",
      "There has never been any forest clearing outside the cfma." ~ "No change",
      "There has never been cutting of trees" ~ "No change",
      "no_decrease" ~ "Decrease",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("Significant", "Moderate", "No change", "Decrease")),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(deforestation) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M4 <- clmm(deforestation ~
              interviewee_role +
              interviewee_gender +
              interviewee_age +
              (1 | province) +
              (1 | province_cfmg),
            data = test.data)

emm.M4     <- emmeans(M4, specs = ~ interviewee_role)
pairwise.M4 <- pairs(emm.M4, adjust = "Tukey")
```

The effects of the gender and age of the interviewee were not a significant determinant of their response. However, the differences in opinion between the CFMG executive members and DFOs, and between traditional leaders and DFOs, were significant (@tbl-M4_deforestation_outside_individual, @tbl-pairwise_deforestation_outside_individual).

```{r}
#| label: tbl-M4_deforestation_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on perceived increase in deforestation outside the CF."

tidy(M4) |>
  filter(coef.type == "location") |>
  mutate(term = case_match(term,
    "interviewee_rolecfmg_exec_member" ~ "CFMG executive member",
    "interviewee_roletrad_leader"      ~ "Traditional leader",
    "interviewee_roledfo"              ~ "DFO",
    "interviewee_roleNGO_rep"          ~ "Private enterprise rep",
    "interviewee_gendermale"           ~ "Gender: male",
    "interviewee_age"                  ~ "Age",
    .default = term
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
             p.value = md("*P* value")) |>
  fmt_number(columns = c(estimate, std.error), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Model results for perceived increase in deforestation outside the CF — interviewee role, gender and age",
    subtitle = glue::glue("We used a CLMM model with random intercepts for province and CF within province. Reference role = Community member; reference gender = female. n = {nobs(M4)}.")
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -term) |>
  cols_align(align = "left",   columns =  term) |>
  tab_style(style = cell_text(align = "center"),
            locations = cells_column_labels(columns = term)) |>
  sub_missing()
```

```{r}
#| label: tbl-pairwise_deforestation_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived increase in deforestation outside the CF — interviewee role."

as.data.frame(summary(pairwise.M4)) |>
  select(contrast, estimate, SE, z.ratio, p.value) |>
  mutate(contrast = gsub("cfmg_exec_member", "CFMG exec member", contrast),
         contrast = gsub("cfmg_member",      "Community member",  contrast),
         contrast = gsub("trad_leader",      "Trad leader",       contrast),
         contrast = gsub("NGO_rep",          "NGO rep",           contrast)) |>
  gt() |>
  cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
             z.ratio = "z ratio", p.value = md("*P* value")) |>
  fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Pairwise comparisons for perceived increase in deforestation outside the CF — interviewee role",
    subtitle = "Tukey HSD adjustment."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -contrast) |>
  cols_align(align = "left",   columns =  contrast) |>
  tab_style(style = cell_text(align = "center"),
            locations = cells_column_labels(columns = contrast)) |>
  sub_missing()
```

### Charcoal production

We asked interviewees whether CFM had reduced charcoal production inside their CF and whether it had increased immediately outside their CF.

#### Charcoal production inside the community forest

```{r}
#| label: fig-reduced_charcoal_inside
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: " 'Has CFM reduced charcoal production inside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "No. There have never been  charcoal  production  in this forest"    ~ "No change",
      "No_ charcoal production is not a common practice for community members"    ~ "No change",
      "Some areas are encroached by charcoal production"    ~ "No change",
      "There has never been  charcoal  production"    ~ "No change",
      "There has never been  charcoal  production in this forest"    ~ "No change",
      "yes_reduced"   ~ "Moderate",
      "yes_high" ~ "Significant",
      "no" ~ "Significant",
      "no_stopped" ~ "Significant",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Significant", "Moderate", "No change")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(charcoal) |>
  count(interview_type, charcoal) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = charcoal)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "No change"   = "#C0392B",
      "Moderate"    = "#F0C040",
      "Significant" = "#2A9D8F"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1),    colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Almost all the FGD groups agreed that charcoal making inside their CF had decreased, with most indicating that charcoal making had decreased significantly (@fig-reduced_charcoal_inside). The small differences in responses among FGD groups were not significant (@tbl-M5_charcoal_inside_fgd). The effect of governance index was also not significant, likely reflecting the lack of variance in responses.

```{r}
#| label: model_charcoal_inside_fgd
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "No. There have never been  charcoal  production  in this forest"    ~ "No change",
      "No_ charcoal production is not a common practice for community members"    ~ "No change",
      "Some areas are encroached by charcoal production"    ~ "No change",
      "There has never been  charcoal  production"    ~ "No change",
      "There has never been  charcoal  production in this forest"    ~ "No change",
      "yes_reduced"   ~ "Moderate",
      "yes_high" ~ "Significant",
      "no" ~ "Significant",
      "no_stopped" ~ "Significant",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(charcoal) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$charcoal) >= 2) {
  M5 <- clmm(charcoal ~
                interview_type +
                gov_index +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M5     <- emmeans(M5, specs = ~ interview_type)
  pairwise.M5 <- pairs(emm.M5, adjust = "Tukey")
} else {
  M5 <- NULL
  cat("Model not fitted: only", nlevels(test.data$charcoal),
      "level(s) present in charcoal after filtering.\n")
}
```

```{r}
#| label: tbl-M5_charcoal_inside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on perceived reduction in charcoal production inside the CF."

if (!is.null(M5)) {
  tidy(M5) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interview_typeFemale" ~ "Female",
      "interview_typeMale"   ~ "Male",
      "interview_typeYouth"  ~ "Youth",
      "gov_index"            ~ "Governance index",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived reduction in charcoal production inside the CF — FGD type and governance index",
      subtitle = glue::glue("We used a CLMM model with random intercepts for province and CF within province. Reference = Executive. n = {nobs(M5)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_charcoal_inside_fgd
#| include: false
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived reduction in charcoal production inside the CF — FGD type."

if (!is.null(pairwise.M5)) {
  as.data.frame(summary(pairwise.M5)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived reduction in charcoal production inside the CF — FGD type",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

Likewise, among the individual interviewees (@tbl-charcoal_inside_individual), almost all interviewees agreed that there had been a substantial decrease in charcoal making inside their CF. Again, this result was consistent across interviewee roles (there were no significant differences among roles) (@tbl-M6_charcoal_inside_individual).

```{r}
#| label: tbl-charcoal_inside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived reduction in charcoal production inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "No. There have never been  charcoal  production  in this forest"    ~ "No change",
      "No_ charcoal production is not a common practice for community members"    ~ "No change",
      "Some areas are encroached by charcoal production"    ~ "No change",
      "There has never been  charcoal  production"    ~ "No change",
      "There has never been  charcoal  production in this forest"    ~ "No change",
      "yes_reduced"   ~ "Moderate",
      "yes_high" ~ "Significant",
      "no" ~ "Significant",
      "no_stopped" ~ "Significant",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "Private enterprise rep"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "Private enterprise rep")
    )
  ) |>
  filter(!is.na(charcoal)) |>
  count(interviewee_role, charcoal, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, charcoal, prop) |>
  pivot_wider(names_from = charcoal, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(`No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived reduction in charcoal production inside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: test_data_charcoal_inside_individual
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "No. There have never been  charcoal  production  in this forest"    ~ "No change",
      "No_ charcoal production is not a common practice for community members"    ~ "No change",
      "Some areas are encroached by charcoal production"    ~ "No change",
      "There has never been  charcoal  production"    ~ "No change",
      "There has never been  charcoal  production in this forest"    ~ "No change",
      "yes_reduced"   ~ "Moderate",
      "yes_high" ~ "Significant",
      "no" ~ "Significant",
      "no_stopped" ~ "Significant",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(charcoal) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$charcoal) >= 2) {
  M6 <- clmm(charcoal ~
                interviewee_role +
                interviewee_gender +
                interviewee_age +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M6     <- emmeans(M6, specs = ~ interviewee_role)
  pairwise.M6 <- pairs(emm.M6, adjust = "Tukey")
} else {
  M6 <- NULL
  cat("Model not fitted: only", nlevels(test.data$charcoal),
      "level(s) present in charcoal after filtering.\n")
}
```

```{r}
#| label: tbl-M6_charcoal_inside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on perceived reduction in charcoal production inside the CF."

if (!is.null(M6)) {
  tidy(M6) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interviewee_rolecfmg_exec_member" ~ "CFMG executive member",
      "interviewee_roletrad_leader"      ~ "Traditional leader",
      "interviewee_roledfo"              ~ "DFO",
      "interviewee_roleNGO_rep"          ~ "Private enterprise rep",
      "interviewee_gendermale"           ~ "Gender: male",
      "interviewee_age"                  ~ "Age",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived reduction in charcoal production inside the CF — interviewee role, gender and age",
      subtitle = glue::glue("We used a CLMM model random intercepts for province and CF within province. Reference role = Community member; reference gender = female. n = {nobs(M6)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_charcoal_inside_individual
#| include: false
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived reduction in charcoal production inside the CF — interviewee role."

if (!is.null(pairwise.M6)) {
  as.data.frame(summary(pairwise.M6)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    mutate(contrast = gsub("cfmg_exec_member", "CFMG exec member", contrast),
           contrast = gsub("cfmg_member",      "Community member",  contrast),
           contrast = gsub("trad_leader",      "Trad leader",       contrast),
           contrast = gsub("NGO_rep",          "NGO rep",           contrast)) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived reduction in charcoal production inside the CF — interviewee role",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

#### Charcoal production outside the community forest

To examine for the possibility of leakage, we also asked interviewees whether CFM has led to an increase in charcoal production outside their CF (@fig-increased_charcoal_outside).

```{r}
#| label: fig-increased_charcoal_outside
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: " 'Has CFM increased charcoal production outside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_increased_outside_cfma,
      "no_reduced"    ~ "Decrease",
      "no_change"    ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      "Yes_ moderate increase. Only for household use"   ~ "Moderate",
      "Moderate decrease"    ~ "Decrease",
      "Due to increase in population_ some charcoal production now being observed in some open are…"   ~ "Moderate",
      "CFMG members not involved in charcoal production"    ~ "No change",
      "Charcoal  production  is not allowed in the chiefdom"    ~ "No change",
      "Charcoal  production is not allowed in the chiefdom"    ~ "No change",
      "Charcoal production  is not allowed  in the chiefdom"    ~ "No change",
      "Charcoal production has been banned from the chiefdom"    ~ "No change",
      "Charcoal production is not common in this part of the district"    ~ "No change",
      "Charcoal production is not commonly practiced. There is no market for charcoal production"    ~ "No change",
      "Charcoal production is not practiced in the community"    ~ "No change",
      "Charcoal production not a common practice in the community"    ~ "No change",
      "Charcoal production only done when clearing an area for new agricultural fields"    ~ "No change",
      "No charcoal  production"    ~ "No change",
      "No charcoal  production In this area"    ~ "No change",
      "No charcoal  production in this community"    ~ "No change",
      "No charcoal production"    ~ "No change",
      "No charcoal production near CFMA"    ~ "No change",
      "No. There has never been charcoal  production"    ~ "No change",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Significant", "Moderate", "No change", "Decrease")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(charcoal) |>
  count(interview_type, charcoal) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = charcoal)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Decrease"    = "#2A9D8F",
      "No change"   = "#76B7A0",
      "Moderate"    = "#F0C040",
      "Significant" = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1),    colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen, the vast majority of groups indicated that charcoal making had either decreased outside their CF or there was no change. However, there were a small number that indicated charcoal making had increased, with some indicating it had increased significantly. Female FGD groups were significantly more likely to indicate that charcoal making had decreased outside the CF than other groups (@tbl-M7_charcoal_outside_fgd; @tbl-pairwise_charcoal_outside_fgd). There were no significant differences among these other groups and the effect of the governance index was also not significant.

```{r}
#| label: model_charcoal_outside_fgd
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_increased_outside_cfma,
      "no_reduced"    ~ "Decrease",
      "no_change"    ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      "Yes_ moderate increase. Only for household use"   ~ "Moderate",
      "Moderate decrease"    ~ "Decrease",
      "Due to increase in population_ some charcoal production now being observed in some open are…"   ~ "Moderate",
      "CFMG members not involved in charcoal production"    ~ "No change",
      "Charcoal  production  is not allowed in the chiefdom"    ~ "No change",
      "Charcoal  production is not allowed in the chiefdom"    ~ "No change",
      "Charcoal production  is not allowed  in the chiefdom"    ~ "No change",
      "Charcoal production has been banned from the chiefdom"    ~ "No change",
      "Charcoal production is not common in this part of the district"    ~ "No change",
      "Charcoal production is not commonly practiced. There is no market for charcoal production"    ~ "No change",
      "Charcoal production is not practiced in the community"    ~ "No change",
      "Charcoal production not a common practice in the community"    ~ "No change",
      "Charcoal production only done when clearing an area for new agricultural fields"    ~ "No change",
      "No charcoal  production"    ~ "No change",
      "No charcoal  production In this area"    ~ "No change",
      "No charcoal  production in this community"    ~ "No change",
      "No charcoal production"    ~ "No change",
      "No charcoal production near CFMA"    ~ "No change",
      "No. There has never been charcoal  production"    ~ "No change",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Decrease", "No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(charcoal) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$charcoal) >= 2) {
  M7 <- clmm(charcoal ~
                interview_type +
                gov_index +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M7     <- emmeans(M7, specs = ~ interview_type)
  pairwise.M7 <- pairs(emm.M7, adjust = "Tukey")
} else {
  M7 <- NULL
  cat("Model not fitted: only", nlevels(test.data$charcoal),
      "level(s) present in charcoal after filtering.\n")
}
```

```{r}
#| label: tbl-M7_charcoal_outside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on perceived increase in charcoal production outside the CF."

if (!is.null(M7)) {
  tidy(M7) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interview_typeFemale" ~ "Female",
      "interview_typeMale"   ~ "Male",
      "interview_typeYouth"  ~ "Youth",
      "gov_index"            ~ "Governance index",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived increase in charcoal production outside the CF — FGD type and governance index",
      subtitle = glue::glue("We used a CLMM with random intercepts for province and CF within province. Reference = Executive. n = {nobs(M7)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_charcoal_outside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived increase in charcoal production outside the CF — FGD type."

if (!is.null(pairwise.M7)) {
  as.data.frame(summary(pairwise.M7)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived increase in charcoal production outside the CF — FGD type",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

Among the individual interviewees (@tbl-charcoal_outside_individual), similar to the results on deforestation, most community interviewees suggested there had been a decrease or no change in charcoal production outside the CF. However, private enterprise representatives and DFOs were significantly more likely to indicate that there had been an increase in charcoal production outside the CF (@tbl-M8_charcoal_outside_individual, @tbl-pairwise_charcoal_outside_individual). Interviewee age was also significant — older respondents were more likely to state that there had been an increase in charcoal production outside the CF.

```{r}
#| label: tbl-charcoal_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived increase in charcoal production outside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_increased_outside_cfma,
      "no_reduced"    ~ "Decrease",
      "no_change"    ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      "Yes_ moderate increase. Only for household use"   ~ "Moderate",
      "Moderate decrease"    ~ "Decrease",
      "Due to increase in population_ some charcoal production now being observed in some open are…"   ~ "Moderate",
      "CFMG members not involved in charcoal production"    ~ "No change",
      "Charcoal  production  is not allowed in the chiefdom"    ~ "No change",
      "Charcoal  production is not allowed in the chiefdom"    ~ "No change",
      "Charcoal production  is not allowed  in the chiefdom"    ~ "No change",
      "Charcoal production has been banned from the chiefdom"    ~ "No change",
      "Charcoal production is not common in this part of the district"    ~ "No change",
      "Charcoal production is not commonly practiced. There is no market for charcoal production"    ~ "No change",
      "Charcoal production is not practiced in the community"    ~ "No change",
      "Charcoal production not a common practice in the community"    ~ "No change",
      "Charcoal production only done when clearing an area for new agricultural fields"    ~ "No change",
      "No charcoal  production"    ~ "No change",
      "No charcoal  production In this area"    ~ "No change",
      "No charcoal  production in this community"    ~ "No change",
      "No charcoal production"    ~ "No change",
      "No charcoal production near CFMA"    ~ "No change",
      "No. There has never been charcoal  production"    ~ "No change",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Decrease", "No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "Private enterprise rep"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "Private enterprise rep")
    )
  ) |>
  filter(!is.na(charcoal)) |>
  count(interviewee_role, charcoal, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, charcoal, prop) |>
  pivot_wider(names_from = charcoal, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Decrease, `No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived increase in charcoal production outside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: test_data_charcoal_outside_individual
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_increased_outside_cfma,
      "no_reduced"    ~ "Decrease",
      "no_change"    ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      "Yes_ moderate increase. Only for household use"   ~ "Moderate",
      "Moderate decrease"    ~ "Decrease",
      "Due to increase in population_ some charcoal production now being observed in some open are…"   ~ "Moderate",
      "CFMG members not involved in charcoal production"    ~ "No change",
      "Charcoal  production  is not allowed in the chiefdom"    ~ "No change",
      "Charcoal  production is not allowed in the chiefdom"    ~ "No change",
      "Charcoal production  is not allowed  in the chiefdom"    ~ "No change",
      "Charcoal production has been banned from the chiefdom"    ~ "No change",
      "Charcoal production is not common in this part of the district"    ~ "No change",
      "Charcoal production is not commonly practiced. There is no market for charcoal production"    ~ "No change",
      "Charcoal production is not practiced in the community"    ~ "No change",
      "Charcoal production not a common practice in the community"    ~ "No change",
      "Charcoal production only done when clearing an area for new agricultural fields"    ~ "No change",
      "No charcoal  production"    ~ "No change",
      "No charcoal  production In this area"    ~ "No change",
      "No charcoal  production in this community"    ~ "No change",
      "No charcoal production"    ~ "No change",
      "No charcoal production near CFMA"    ~ "No change",
      "No. There has never been charcoal  production"    ~ "No change",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Decrease", "No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(charcoal) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$charcoal) >= 2) {
  M8 <- clmm(charcoal ~
                interviewee_role +
                interviewee_gender +
                interviewee_age +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M8     <- emmeans(M8, specs = ~ interviewee_role)
  pairwise.M8 <- pairs(emm.M8, adjust = "Tukey")
} else {
  M8 <- NULL
  cat("Model not fitted: only", nlevels(test.data$charcoal),
      "level(s) present in charcoal after filtering.\n")
}
```

```{r}
#| label: tbl-M8_charcoal_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on perceived increase in charcoal production outside the CF."

if (!is.null(M8)) {
  tidy(M8) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interviewee_rolecfmg_exec_member" ~ "CFMG executive member",
      "interviewee_roletrad_leader"      ~ "Traditional leader",
      "interviewee_roledfo"              ~ "DFO",
      "interviewee_roleNGO_rep"          ~ "Private enterprise rep",
      "interviewee_gendermale"           ~ "Gender: male",
      "interviewee_age"                  ~ "Age",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived increase in charcoal production outside the CF — interviewee role, gender and age",
      subtitle = glue::glue("We used a CLMM with random intercepts for province and CF within province. Reference role = Community member; reference gender = female. n = {nobs(M8)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_charcoal_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived increase in charcoal production outside the CF — interviewee role."

if (!is.null(pairwise.M8)) {
  as.data.frame(summary(pairwise.M8)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    mutate(contrast = gsub("cfmg_exec_member", "CFMG exec member", contrast),
           contrast = gsub("cfmg_member",      "Community member",  contrast),
           contrast = gsub("trad_leader",      "Trad leader",       contrast),
           contrast = gsub("NGO_rep",          "NGO rep",           contrast)) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived increase in charcoal production outside the CF — interviewee role",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

### Fire

We asked interviewees about fire monitoring practices and whether CFM had reduced fire incidence inside their CF. We did not ask about changes in fire frequency outside the CF, because clearing fields for agriculture or improving grazing often involves burning.

#### Fire monitoring

```{r}
#| label: fig-fire_monitoring
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFMGs have established fire monitoring in the community forest by province. The length of the bars represents the number of CFMGs either monitoring fires (Yes) or not (No). The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(impact_on_sfm_fire_monitoring_in_cfma, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen in the figure (@fig-fire_monitoring), most CFMGs have instigated a system of fire monitoring. However, ten CFMGs across four provinces from our sample of 75 did not have any system of fire monitoring.

#### Fire incidence inside the community forest

We asked interviewees whether CFM has reduced fire incidence inside their CF (@fig-reduced_fire_inside).

```{r}
#| label: fig-reduced_fire_inside
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "'Has CFM reduced fire incidence inside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "no_incidents_have_increased" ~ "Increase",
      "no_change"     ~ "No change",
      "yes_moderate_reduction"    ~ "Moderate",
      "yes_significant_reduction" ~ "Significant",
      "No change observed as at now" ~ "No change",
      .default = NA
    ),
    fire = factor(fire, levels = c("Significant", "Moderate", "No change","Increase")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(fire) |>
  count(interview_type, fire) |>
  group_by(interview_type) |>
  mutate(
    prop = n / sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = fire)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Increase" = "#C0392B",
      "No change"   = "#F0C040",
      "Moderate"    = "#2A9D8F",
      "Significant" = "#76B7A0"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#sustainable-forest-management)
:::

As can be seen, a majority of groups reported a moderate or significant reduction in fire incidence in the CF. Female FGDs reported more significant declines in fire frequency compared to the other groups (@tbl-M9_fire_fgd). These differences between female FGD groups and the other FGD groups were significant (@tbl-pairwise_fire_fgd). The effect of the governance index was also significant, indicating that CFs with higher governance scores had greater perceived declines in fire incidence.

```{r}
#| label: model_fire_fgd
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "no_incidents_have_increased" ~ "Increase",
      "no_change"     ~ "No change",
      "yes_moderate_reduction"    ~ "Moderate",
      "yes_significant_reduction" ~ "Significant",
      "No change observed as at now" ~ "No change",
      .default = NA
    ),
    fire = factor(fire,
                  levels = c("Increase","No change", "Moderate", "Significant"),
                  ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(fire) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$fire) >= 2) {
  M9 <- clmm(fire ~
                interview_type +
                gov_index +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M9      <- emmeans(M9, specs = ~ interview_type)
  pairwise.M9 <- pairs(emm.M9, adjust = "Tukey")
} else {
  M9 <- NULL
  cat("Model not fitted: only", nlevels(test.data$fire),
      "level(s) present in fire after filtering.\n")
}
```

```{r}
#| label: tbl-M9_fire_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on perceived reduction in fire incidence inside the CF."

if (!is.null(M9)) {
  tidy(M9) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interview_typeFemale" ~ "Female",
      "interview_typeMale"   ~ "Male",
      "interview_typeYouth"  ~ "Youth",
      "gov_index"            ~ "Governance index",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived reduction in fire incidence inside the CF — FGD type and governance index",
      subtitle = glue::glue("We use a CLMM with random intercepts for province and CF within province. Reference = Executive. n = {nobs(M9)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_fire_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived reduction in fire incidence inside the CF — FGD type."

if (!is.null(pairwise.M9)) {
  as.data.frame(summary(pairwise.M9)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived reduction in fire incidence inside the CF — FGD type",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

Among the individual interviewees, there were also some differences in opinion (@tbl-fire_individual). However, these differences were not significant (@tbl-M10_fire_individual). The effects of interviewee gender and age were also not significant.

```{r}
#| label: tbl-fire_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived reduction in fire incidence inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "no_incidents_have_increased" ~ "Increase",
      "no_change"     ~ "No change",
      "yes_moderate_reduction"    ~ "Moderate",
      "yes_significant_reduction" ~ "Significant",
      "No change observed as at now" ~ "No change",
      .default = NA
    ),
    fire = factor(fire,
                  levels = c("Significant", "Moderate", "No change", "Increase"),
                  ordered = TRUE),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "Private enterprise rep"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "Private enterprise rep")
    )
  ) |>
  filter(!is.na(fire)) |>
  count(interviewee_role, fire, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, fire, prop) |>
  pivot_wider(names_from = fire, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Significant, Moderate, `No change`, Increase), decimals = 0) |>
  tab_header(
    title    = "Perceived reduction in fire incidence inside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: test_data_fire_individual
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "no_incidents_have_increased" ~ "Increase",
      "no_change"     ~ "No change",
      "yes_moderate_reduction"    ~ "Moderate",
      "yes_significant_reduction" ~ "Significant",
      "No change observed as at now" ~ "No change",
      .default = NA
    ),
    fire = factor(fire,
                  levels = c("Increase","No change", "Moderate", "Significant"),
                  ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(fire) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$fire) >= 2) {
  M10 <- clmm(fire ~
                 interviewee_role +
                 interviewee_gender +
                 interviewee_age +
                 (1 | province) +
                 (1 | province_cfmg),
               data = test.data)
  emm.M10      <- emmeans(M10, specs = ~ interviewee_role)
  pairwise.M10 <- pairs(emm.M10, adjust = "Tukey")
} else {
  M10 <- NULL
  cat("Model not fitted: only", nlevels(test.data$fire),
      "level(s) present in fire after filtering.\n")
}
```

```{r}
#| label: tbl-M10_fire_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on perceived reduction in fire incidence inside the CF."

if (!is.null(M10)) {
  tidy(M10) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interviewee_rolecfmg_exec_member" ~ "CFMG executive member",
      "interviewee_roletrad_leader"      ~ "Traditional leader",
      "interviewee_roledfo"              ~ "DFO",
      "interviewee_roleNGO_rep"          ~ "Private enterprise rep",
      "interviewee_gendermale"           ~ "Gender: male",
      "interviewee_age"                  ~ "Age",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived reduction in fire incidence inside the CF — interviewee role, gender and age",
      subtitle = glue::glue("CLMM; random intercepts for province and CF within province. Reference role = Community member; reference gender = female. n = {nobs(M10)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_fire_individual
#| include: false
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived reduction in fire incidence inside the CF — interviewee role."

if (!is.null(pairwise.M10)) {
  as.data.frame(summary(pairwise.M10)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    mutate(contrast = gsub("cfmg_exec_member", "CFMG exec member", contrast),
           contrast = gsub("cfmg_member",      "Community member",  contrast),
           contrast = gsub("trad_leader",      "Trad leader",       contrast),
           contrast = gsub("NGO_rep",          "NGO rep",           contrast)) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived reduction in fire incidence inside the CF — interviewee role",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

### Wildlife

We asked interviewees whether CFM had led to an increase in wildlife inside their CF (@fig-wildlife_increase).

```{r}
#| label: fig-wildlife_increase
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "'Has CFM led to an increase in wildlife in the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "no_reduced"              ~ "Decrease",
      "no_change"               ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      .default = NA
    ),
    wildlife = factor(wildlife,
                      levels = c("Significant", "Moderate", "No change", "Decrease")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(wildlife) |>
  count(interview_type, wildlife) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = wildlife)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Significant" = "#2A9D8F",
      "Moderate"    = "#F0C040",
      "No change"   = "#E07B54",
      "Decrease"    = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#sustainable-forest-management)
:::

As can be seen, a majority of groups indicated a significant or moderate increase in wildlife in their CF. However, there were some that indicated there had been a decrease. The responses among FGD types were not significantly different (@tbl-M11_wildlife_fgd). The effect of governance index was also not significant.

```{r}
#| label: model_wildlife_fgd
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "no_reduced"              ~ "Decrease",
      "no_change"               ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      .default = NA
    ),
    wildlife = factor(wildlife,
                      levels = c("Decrease", "No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(wildlife) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$wildlife) >= 2) {
  M11 <- clmm(wildlife ~
                interview_type +
                gov_index +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M11      <- emmeans(M11, specs = ~ interview_type)
  pairwise.M11 <- pairs(emm.M11, adjust = "Tukey")
} else {
  M11 <- NULL
  cat("Model not fitted: only", nlevels(test.data$wildlife),
      "level(s) present in wildlife after filtering.\n")
}
```

```{r}
#| label: tbl-M11_wildlife_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on perceived increase in wildlife inside the CF."

if (!is.null(M11)) {
  tidy(M11) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interview_typeFemale" ~ "Female",
      "interview_typeMale"   ~ "Male",
      "interview_typeYouth"  ~ "Youth",
      "gov_index"            ~ "Governance index",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived increase in wildlife inside the CF — FGD type and governance index",
      subtitle = glue::glue("We used a CLMM with random intercepts for province and CF within province. Reference = Executive. n = {nobs(M11)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_wildlife_fgd
#| include: false
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived increase in wildlife inside the CF — FGD type."

if (!is.null(pairwise.M11)) {
  as.data.frame(summary(pairwise.M11)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived increase in wildlife inside the CF — FGD type",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

Among the individual interviewees (@tbl-wildlife_individual), there were some differences in opinion, but again these differences were not significant (@tbl-M12_wildlife_individual). The effects of interviewee gender and age were also not significant.

```{r}
#| label: tbl-wildlife_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived increase in wildlife inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "no_reduced"              ~ "Decrease",
      "no_change"               ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      .default = NA
    ),
    wildlife = factor(wildlife,
                      levels = c("Decrease", "No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "Private enterprise rep"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "Private enterprise rep")
    )
  ) |>
  filter(!is.na(wildlife)) |>
  count(interviewee_role, wildlife, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, wildlife, prop) |>
  pivot_wider(names_from = wildlife, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Decrease, `No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived increase in wildlife by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: test_data_wildlife_individual
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "no_reduced"              ~ "Decrease",
      "no_change"               ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      .default = NA
    ),
    wildlife = factor(wildlife,
                      levels = c("Decrease", "No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(wildlife) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$wildlife) >= 2) {
  M12 <- clmm(wildlife ~
                 interviewee_role +
                 interviewee_gender +
                 interviewee_age +
                 (1 | province) +
                 (1 | province_cfmg),
               data = test.data)
  emm.M12      <- emmeans(M12, specs = ~ interviewee_role)
  pairwise.M12 <- pairs(emm.M12, adjust = "Tukey")
} else {
  M12 <- NULL
  cat("Model not fitted: only", nlevels(test.data$wildlife),
      "level(s) present in wildlife after filtering.\n")
}
```

```{r}
#| label: tbl-M12_wildlife_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on perceived increase in wildlife inside the CF."

if (!is.null(M12)) {
  tidy(M12) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interviewee_rolecfmg_exec_member" ~ "CFMG executive member",
      "interviewee_roletrad_leader"      ~ "Traditional leader",
      "interviewee_roledfo"              ~ "DFO",
      "interviewee_roleNGO_rep"          ~ "Private enterprise rep",
      "interviewee_gendermale"           ~ "Gender: male",
      "interviewee_age"                  ~ "Age",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived increase in wildlife inside the CF — interviewee role, gender and age",
      subtitle = glue::glue("We used a CLMM with random intercepts for province and CF within province. Reference role = Community member; reference gender = female. n = {nobs(M12)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_wildlife_individual
#| include: false
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived increase in wildlife inside the CF — interviewee role."

if (!is.null(pairwise.M12)) {
  as.data.frame(summary(pairwise.M12)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    mutate(contrast = gsub("cfmg_exec_member", "CFMG exec member", contrast),
           contrast = gsub("cfmg_member",      "Community member",  contrast),
           contrast = gsub("trad_leader",      "Trad leader",       contrast),
           contrast = gsub("NGO_rep",          "NGO rep",           contrast)) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived increase in wildlife inside the CF — interviewee role",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

### Forest resources management

We asked interviewees whether CFM had improved the management of forest resources inside their CF (@fig-forest_resources_management).

```{r}
#| label: fig-forest_resources_management
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "'Has CFM improved forest resources management inside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    frm = case_match(
      impact_on_sfm_forest_resources_management,
      "no_management_has_worsened"    ~ "Management worse",
      "no_change"                     ~ "No change",
      "yes_moderate_improvement"      ~ "Moderate improvement",
      "yes_significant_improvement"   ~ "Significant improvement",
      .default = NA
    ),
    frm = factor(frm, levels = c("Significant improvement", "Moderate improvement",
                                 "No change", "Management worse")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(frm) |>
  count(interview_type, frm) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = frm)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Significant improvement" = "#2A9D8F",
      "Moderate improvement"    = "#F0C040",
      "No change"               = "#E07B54",
      "Management worse"        = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE, nrow = 2)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen, while responses vary among FGD groups, the vast majority of groups agreed that the management of forest resources had improved moderately or significantly under CFM, and only a small number of primarily youth groups considered that it had gotten worse. A very high proportion of community female FGD groups considered that the management of forest resources was significantly better under CFM, a difference that was statistically significant (@tbl-M13_frm_fgd, @tbl-pairwise_frm_fgd). The effect of governance index was also significant.

```{r}
#| label: model_frm_fgd
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    frm = case_match(
      impact_on_sfm_forest_resources_management,
      "no_management_has_worsened"    ~ "Management worse",
      "no_change"                     ~ "No change",
      "yes_moderate_improvement"      ~ "Moderate improvement",
      "yes_significant_improvement"   ~ "Significant improvement",
      .default = NA
    ),
    frm = factor(frm,
                 levels = c("Management worse", "No change",
                            "Moderate improvement", "Significant improvement"),
                 ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(frm) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$frm) >= 2) {
  M13 <- clmm(frm ~
                interview_type +
                gov_index +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M13      <- emmeans(M13, specs = ~ interview_type)
  pairwise.M13 <- pairs(emm.M13, adjust = "Tukey")
} else {
  M13 <- NULL
  cat("Model not fitted: only", nlevels(test.data$frm),
      "level(s) present in frm after filtering.\n")
}
```

```{r}
#| label: tbl-M13_frm_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on perceived improvement in forest resources management."

if (!is.null(M13)) {
  tidy(M13) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interview_typeFemale" ~ "Female",
      "interview_typeMale"   ~ "Male",
      "interview_typeYouth"  ~ "Youth",
      "gov_index"            ~ "Governance index",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived improvement in forest resources management — FGD type and governance index",
      subtitle = glue::glue("We used a CLMM with random intercepts for province and CF within province. Reference = Executive. n = {nobs(M13)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_frm_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived improvement in forest resources management — FGD type."

if (!is.null(pairwise.M13)) {
  as.data.frame(summary(pairwise.M13)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived improvement in forest resources management — FGD type",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

We also asked which specific forest resources had improved under CFM (@fig-harvested_resources).

```{r}
#| label: fig-harvested_resources
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Forest resources that have improved under CFM, as reported by female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting an improvement for each resource."
#| fig-width: 10
#| fig-height: 7

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(
    "Pupwe"            = impact_on_sfm_harvested_forest_resources_pupwe,
    "Devil's claw"     = impact_on_sfm_harvested_forest_resources_devils_claw,
    "Fish"             = impact_on_sfm_harvested_forest_resources_fish,
    "Lusala"           = impact_on_sfm_harvested_forest_resources_lusala,
    "Mungongo"         = impact_on_sfm_harvested_forest_resourcesmungongo,
    "Bamboo"           = impact_on_sfm_harvested_forest_resources_bamboo,
    "Munkoyo"          = impact_on_sfm_harvested_forest_resources_munkoyo,
    "Poles"            = impact_on_sfm_harvested_forest_resources_poles,
    "Bushmeat"         = impact_on_sfm_harvested_forest_resources_bushmeat,
    "Timber"           = impact_on_sfm_harvested_forest_resources_timber,
    "Fibres"           = impact_on_sfm_harvested_forest_resources_fibers,
    "Chikanda"         = impact_on_sfm_harvested_forest_resources_chikanda,
    "Firewood"         = interview_grp_impact_on_sfm_harvested_forest_resources_firewood,
    "Thatching grass"  = impact_on_sfm_harvested_forest_resources_thatching_grass,
    "Medicinal plants" = impact_on_sfm_harvested_forest_resources_medicinal_plants,
    "Caterpillars"     = impact_on_sfm_harvested_forest_resources_caterpillars,
    "Honey"            = impact_on_sfm_harvested_forest_resources_honey,
    "Wild fruits"      = impact_on_sfm_harvested_forest_resources_wild_fruits,
    "Mushrooms"        = impact_on_sfm_harvested_forest_resources_mushrooms
  ) |>
  pivot_longer(everything(), names_to = "word", values_to = "value") |>
  filter(!is.na(value)) |>
  group_by(word) |>
  summarise(freq = sum(as.integer(value) == 1, na.rm = TRUE), .groups = "drop") |>
  filter(freq > 0) |>
  ggplot(aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.8, seed = 9) +
  scale_size(range = c(3, 20)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

Among the individual interviewees (@tbl-frm_individual), the vast majority of interviewees considered that the management of forest resources had improved under CFM. There was a significant difference in responses between community members and DFOs. The DFOs were more likely to suggest there had been no change or only a moderate improvement in forest resource management under CFM, compared to the much higher proportion of community members, CFMG executive members or traditional leaders indicating a substantial improvement in forest resources management (@tbl-M14_frm_individual, @tbl-pairwise_frm_individual). Interestingly, male interviewees were significantly less likely to perceive improvements in forest resources management (@tbl-M14_frm_individual).

```{r}
#| label: tbl-frm_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived change in forest resources management inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    frm = case_match(
      impact_on_sfm_forest_resources_management,
      "no_management_has_worsened"    ~ "Management worse",
      "no_change"                     ~ "No change",
      "yes_moderate_improvement"      ~ "Moderate improvement",
      "yes_significant_improvement"   ~ "Significant improvement",
      .default = NA
    ),
    frm = factor(frm,
                 levels = c("Management worse", "No change",
                            "Moderate improvement", "Significant improvement")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "Private enterprise rep"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "Private enterprise rep")
    )
  ) |>
  filter(!is.na(frm)) |>
  count(interviewee_role, frm, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, frm, prop) |>
  pivot_wider(names_from = frm, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  select(interviewee_role, n,
         `Management worse`, `No change`, `Moderate improvement`, `Significant improvement`) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(`Management worse`, `No change`, `Moderate improvement`, `Significant improvement`), decimals = 0) |>
  tab_header(
    title    = "Perceived improvement in forest resources management by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

```{r}
#| label: test_data_frm_individual
#| include: false
#| warning: false
#| message: false
#| error:   false

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    frm = case_match(
      impact_on_sfm_forest_resources_management,
      "no_management_has_worsened"    ~ "Management worse",
      "no_change"                     ~ "No change",
      "yes_moderate_improvement"      ~ "Moderate improvement",
      "yes_significant_improvement"   ~ "Significant improvement",
      .default = NA
    ),
    frm = factor(frm,
                 levels = c("Management worse", "No change",
                            "Moderate improvement", "Significant improvement"),
                 ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(frm) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$frm) >= 2) {
  M14 <- clmm(frm ~
                 interviewee_role +
                 interviewee_gender +
                 interviewee_age +
                 (1 | province) +
                 (1 | province_cfmg),
               data = test.data)
  emm.M14      <- emmeans(M14, specs = ~ interviewee_role)
  pairwise.M14 <- pairs(emm.M14, adjust = "Tukey")
} else {
  M14 <- NULL
  cat("Model not fitted: only", nlevels(test.data$frm),
      "level(s) present in frm after filtering.\n")
}
```

```{r}
#| label: tbl-M14_frm_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of interviewee role, gender and age on perceived improvement in forest resources management."

if (!is.null(M14)) {
  tidy(M14) |>
    filter(coef.type == "location") |>
    mutate(term = case_match(term,
      "interviewee_rolecfmg_exec_member" ~ "CFMG executive member",
      "interviewee_roletrad_leader"      ~ "Traditional leader",
      "interviewee_roledfo"              ~ "DFO",
      "interviewee_roleNGO_rep"          ~ "Private enterprise rep",
      "interviewee_gendermale"           ~ "Gender: male",
      "interviewee_age"                  ~ "Age",
      .default = term
    )) |>
    select(term, estimate, std.error, p.value) |>
    gt() |>
    cols_label(term = "Factor", estimate = "Estimate", std.error = "SE",
               p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, std.error), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Model results for perceived improvement in forest resources management — interviewee role, gender and age",
      subtitle = glue::glue("We use a CLMM with random intercepts for province and CF within province. Reference role = Community member; reference gender = female. n = {nobs(M14)}.")
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -term) |>
    cols_align(align = "left",   columns =  term) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = term)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

```{r}
#| label: tbl-pairwise_frm_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Pairwise comparisons for perceived improvement in forest resources management — interviewee role."

if (!is.null(pairwise.M14)) {
  as.data.frame(summary(pairwise.M14)) |>
    select(contrast, estimate, SE, z.ratio, p.value) |>
    mutate(contrast = gsub("cfmg_exec_member", "CFMG exec member", contrast),
           contrast = gsub("cfmg_member",      "Community member",  contrast),
           contrast = gsub("trad_leader",      "Trad leader",       contrast),
           contrast = gsub("NGO_rep",          "NGO rep",           contrast)) |>
    gt() |>
    cols_label(contrast = "Contrast", estimate = "Estimate", SE = "SE",
               z.ratio = "z ratio", p.value = md("*P* value")) |>
    fmt_number(columns = c(estimate, SE, z.ratio), decimals = 3) |>
    fmt_number(columns = p.value, decimals = 4) |>
    tab_header(
      title    = "Pairwise comparisons for perceived improvement in forest resources management — interviewee role",
      subtitle = "Tukey HSD adjustment."
    ) |>
    opt_align_table_header(align = "left") |>
    cols_align(align = "center", columns = -contrast) |>
    cols_align(align = "left",   columns =  contrast) |>
    tab_style(style = cell_text(align = "center"),
              locations = cells_column_labels(columns = contrast)) |>
    sub_missing()
} else {
  gt(data.frame(Note = "Model not fitted."))
}
```

### Restoration

We asked CFMG executives whether natural forest restoration is practised inside the community forest (@fig-natural_forest_restoration). We restricted the responses for this question to CFMG executives, because this is the group that should be best informed about the activities of the CFMG (and the differences in responses among groups to this question are not of interest).

```{r}
#| label: fig-natural_forest_restoration
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFMGs undertake natural forest restoration in the community forest by province. The length of the bars represents the number of CFMGs either undertaking restoration (Yes) or not (No). The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(impact_on_sfm_natural_forest_restoration, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

::: return-to-summary
[← Return to Summary](Summary.html#sustainable-forest-management)
:::

Natural forest restoration is being practised across approximately half of the CFs in our sample, mostly in Muchinga, Central and Northern provinces. We also asked CFMG executives about the specific restoration methods applied inside their CF (@fig-restoration_methods). The majority were protecting natural regeneration. Only one CFMG was involved in planting native species and removal of invasives was not practised.

```{r}
#| label: fig-restoration_methods
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Restoration methods applied inside the community forest, as reported by CFMG executives. The length of the bars represents the number of CFMGs either using (Yes) or not using (No) each method. The numbers on the bars represent the actual number of CFMGs in each category."
#| fig-height: 4

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    "Protecting natural regeneration" = impact_on_sfm_restoration_methods_applied_protection,
    "Planting indigenous"             = impact_on_sfm_restoration_methods_applied_planting_with_indigineous_species,
    "Invasives removal"               = impact_on_sfm_restoration_methods_applied_removal
  ) |>
  pivot_longer(everything(), names_to = "method", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    method   = factor(method, levels = c("Invasives removal",
                                         "Planting indigenous",
                                         "Protecting natural regeneration"))
  ) |>
  count(method, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = method, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Plantations

In addition, we asked CFMG executives whether any plantations had been established inside their CF (@fig-plantation_establishment). Only seven out the the 75 CFs were practising any plantation establishment.

```{r}
#| label: fig-plantation_establishment
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFMGs have established a plantation inside the community forest, by province. The length of the bars represents the number of CFMGs either with (Yes) or without (No) a plantation. The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(impact_on_sfm_planted_forest_establishment, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We also asked executives to estimate the number of hectares planted in the plantation (@fig-hectares_planted), but as can bee seen the planted area is very small.

```{r}
#| label: fig-hectares_planted
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Estimated area of plantation established inside the community forest, as reported by CFMG executives (n = 6 CFMGs; total area = 20.5 ha). The horizontal bar is the median and the red dot is the mean. Species planted include *Gliricidia sepium* (n = 3 CFMGs), *Pinus* spp. (n = 2), *Acacia* spp. (n = 1), *Faidherbia albidia* (n = 1), and miombo species (n = 1)."
#| fig-width: 4
#| fig-height: 5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  filter(!is.na(impact_on_sfm_hectares_planted)) |>
  ggplot(aes(x = "", y = impact_on_sfm_hectares_planted)) +
  geom_boxplot(width = 0.4, outlier.shape = NA) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", size = 3, col = "red") +
  theme_bw(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x  = element_text(size = rel(1.14)),
    axis.text.y  = element_text(size = rel(1.14)),
    axis.title.y = element_text(size = rel(1.29))
  ) +
  xlab("") +
  ylab("Hectares planted")
```

### Agroforestry

Last, we asked CFMG executives whether CFM has involved introducing agroforestry practices (@fig-yn_agroforestry). In this context, agroforestry is implemented on agricultural land outside the CF as a means for achieving sustainable intensification and reducing pressure on forests. Around 40% of CFMGs said that agroforestry had been introduced under CFM.

```{r}
#| label: fig-yn_agroforestry
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether agroforestry is employed to enhance SFM, as reported by CFMG executives. The length of the bars represents the number of CFMGs reporting either a positive impact (Yes) or no impact (No). The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(impact_on_sfm_yn_agroforestry, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We asked CFMG executives about the types of agroforestry practices present in the community (@fig-agroforestry_types).

```{r}
#| label: fig-agroforestry_types
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of agroforestry practices reported by CFMG executives (n = 30 CFMGs). The length of the bars represents the number of CFMGs either reporting (Yes) or not reporting (No) each practice."
#| fig-height: 4.5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    "Silvopasture"               = impact_on_sfm_types_agroforestry_practices_Silvopasture,
    "Windbreaks"                 = impact_on_sfm_types_agroforestry_practices_wind_breaks,
    "Home gardens"               = impact_on_sfm_types_agroforestry_practices_home_gardens,
    "Boundary planting/woodlots" = impact_on_sfm_types_agroforestry_practices_boundary_planting_woodlots,
    "Intercropping"              = impact_on_sfm_types_agroforestry_practices_intercropping
  ) |>
  pivot_longer(everything(), names_to = "practice", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    practice = factor(practice, levels = c(
      "Silvopasture", "Windbreaks", "Home gardens",
      "Boundary planting/woodlots", "Intercropping"
    ))
  ) |>
  count(practice, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = practice, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Among the 30 CFMGs practising agroforestry most were employing intercropping with leguminous shrubs for soil health management, but some were planting boundary trees or woodlots and homegardens.

Among the species used (@fig-agroforestry_species), we can see that *Gliricidia* and *Faidherbia*, leguminous shrubs and trees, respectively, that are used for intercropping, were the most popular.

```{r}
#| label: fig-agroforestry_species
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Agroforestry species planted, as reported by CFMG executives (n = 30 CFMGs). The length of the bars represents the number of CFMGs either planting (Yes) or not planting (No) each species."
#| fig-height: 4

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    "Bamboo"             = impact_on_sfm_agroforestry_species_planted_bamboo,
    "Moringa"            = impact_on_sfm_agroforestry_species_planted_moringa,
    "Exotic fruit trees" = impact_on_sfm_agroforestry_species_planted_exotic_fruit_tree,
    "Faidherbia albidia" = impact_on_sfm_agroforestry_species_planted_faidherbia_albidia,
    "Gliricidia sepium"  = impact_on_sfm_agroforestry_species_planted_gliricidia_sepium
  ) |>
  pivot_longer(everything(), names_to = "species", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    species  = factor(species, levels = c(
      "Bamboo", "Moringa", "Exotic fruit trees",
      "Faidherbia albidia", "Gliricidia sepium"
    ))
  ) |>
  count(species, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = species, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20",
                                      face  = "italic"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

## SFM index

To summarise the overall SFM performance of each CFMG, we constructed an SFM score (0–5) by summing five binary indicators of change inside the CF — decrease in deforestation, decrease in charcoal production, decrease in fire incidence, increase in wildlife, and improved natural resources management — scored as 1 (positive change reported) or 0 (no change or worsening). The distribution of SFM scores across CFMGs is shown in @fig-sfm_index_dist (Appendix III). We then modelled this score using a linear mixed model with FGD type and the governance index as predictors, and random intercepts for province and CF within province, to test whether governance and FGD type were significant predictors of overall SFM performance (@tbl-sfm_index_model, @fig-sfm_index).

```{r}
#| label: data_sfm_index
#| include: false
#| warning: false
#| message: false
#| error:   false

sfm_data <- cfmg |>
  filter(interview_type %in% c("cfmg_executive", "cfmg_male", "cfmg_female", "cfmg_youth")) |>
  mutate(
    si_deforest = case_match(
      impact_on_sfm_cfm_reduced_forest_clearing_cfma,
      "yes_significant" ~ 1,
      "yes_moderate"    ~ 1,
      "Yes there no more fields_  it's all forests"                       ~ 1,
      "Yes_ significant reduction. There was chitemene system before CFM" ~ 1,
      "no_increase"                                      ~ 0,
      "no_change"                                        ~ 0,
      "no_there_has_never_been_clearing_for_agriculture" ~ 0,
      "Agricultural practices ceased way before CFM"     ~ 0,
      "One field for cassava"                            ~ 0,
      "dont_know"                                        ~ NA_real_,
      .default = NA_real_
    ),
    si_charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "yes_high"    ~ 1,
      "no"          ~ 1,
      "no_stopped"  ~ 1,
      "yes_reduced" ~ 1,
      "No_ charcoal production is not a common practice for community members" ~ 0,
      "There has never been  charcoal  production"                             ~ 0,
      "dont_know"                                                              ~ NA_real_,
      .default = NA_real_
    ),
    si_fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "yes_significant_reduction" ~ 1,
      "yes_moderate_reduction"    ~ 1,
      "no_change"                                               ~ 0,
      "no_incidents_have_increased"                             ~ 0,
      "No change observed as at now"                            ~ 0,
      "CFMG activities just starting. To be observed this year." ~ 0,
      "Fire management team only recently received training"    ~ 0,
      "Fire monitoring just started this year"                  ~ 0,
      "dont_know"                                               ~ NA_real_,
      .default = NA_real_
    ),
    si_wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "yes_significant_increase" ~ 1,
      "yes_moderate_increase"    ~ 1,
      "Yes_ moderate increase for some species such as bush pig and antelope"                              ~ 1,
      "Yes_ moderate increase. Impala_ waterbuck_ wild pigs_ guinea fowls can now ge seen in the community forest" ~ 1,
      "Yes_ moderate increase. Not one goes to the forest because of distancing thus animals are freely multiplying" ~ 1,
      "Yes_ significant increase. Leading to an increase in human an wildlife conflicts"                   ~ 1,
      "Yes_ significant increase. The CFMA is in the GMA"                                                 ~ 1,
      "Yes_ significant increase. Wild animals now coming closer to households"                            ~ 1,
      "no_change"                                               ~ 0,
      "no_reduced"                                              ~ 0,
      "No wild animals recorded"                                ~ 0,
      "Other wildlife has decreased_ only the hyna population is on the increase" ~ 0,
      "There  are no wildlife animals"                          ~ 0,
      "There has never been  any wildlife recorded"             ~ 0,
      "There has never been wild animals in this community"     ~ 0,
      "There is no wildlife in this community"                  ~ 0,
      "dont_know"                                               ~ NA_real_,
      .default = NA_real_
    ),
    si_nrm = case_match(
      impact_on_sfm_forest_resources_management,
      "yes_significant_improvement" ~ 1,
      "yes_moderate_improvement"    ~ 1,
      "no_change"                   ~ 0,
      "no_management_has_worsened"  ~ 0,
      "dont_know"                   ~ NA_real_,
      .default = NA_real_
    ),
    sfm_score = ifelse(
      rowSums(!is.na(cbind(si_deforest, si_charcoal, si_fire, si_wildlife, si_nrm))) == 0,
      NA_real_,
      rowSums(cbind(si_deforest, si_charcoal, si_fire, si_wildlife, si_nrm), na.rm = TRUE)
    ),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    ),
    province_cfmg = interaction(province, cfmg_name)
  ) |>
  drop_na(sfm_score, gov_index)

M_sfm <- lmer(
  sfm_score ~ interview_type + gov_index + (1 | province) + (1 | province_cfmg),
  data = sfm_data,
  REML = TRUE
)
```

```{r}
#| label: fig-sfm_index_dist
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Distribution of mean SFM scores across CFMGs (0–5). Each bar represents the number of CFMGs with that mean score, averaged across all four FGD types. The score counts the number of five SFM indicators showing improvement: reduced deforestation, reduced charcoal production, reduced fire incidence, increased wildlife, and improved natural resource management."
#| fig-height: 3

sfm_data |>
  group_by(cfmg_name) |>
  summarise(sfm_score = mean(sfm_score, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = sfm_score)) +
  geom_histogram(binwidth = 0.5, fill = "#2A9D8F", colour = "white", linewidth = 0.4) +
  scale_x_continuous(breaks = 0:5, name = "SFM score (out of 5)") +
  scale_y_continuous(name = "Number of CFMGs") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5)
  )
```

```{r}
#| label: tbl-sfm_index_model
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Results of the model examining the effect of FGD type and governance index on SFM performance."

re_sfm    <- as.data.frame(VarCorr(M_sfm))
prov_sfm  <- round(re_sfm$vcov[re_sfm$grp == "province"], 3)
cfmg_sfm  <- round(re_sfm$vcov[re_sfm$grp == "province_cfmg"], 3)
resid_sfm <- round(re_sfm$vcov[re_sfm$grp == "Residual"], 3)
r2_sfm    <- performance::r2_nakagawa(M_sfm)

tidy(M_sfm, effects = "fixed") |>
  mutate(term = case_match(term,
    "(Intercept)"          ~ "Intercept",
    "interview_typeFemale" ~ "Female",
    "interview_typeMale"   ~ "Male",
    "interview_typeYouth"  ~ "Youth",
    "gov_index"            ~ "Governance index"
  )) |>
  select(term, estimate, std.error, p.value) |>
  gt() |>
  cols_label(
    term      = "Factor",
    estimate  = "Estimate",
    std.error = "SE",
    p.value   = md("*P* value")
  ) |>
  fmt_number(columns = c(estimate, std.error), decimals = 3) |>
  fmt_number(columns = p.value, decimals = 4) |>
  tab_header(
    title    = "Model results for SFM performance — FGD type and governance index",
    subtitle = glue::glue(
      "Linear mixed model with random intercepts for province and CF within province. ",
      "Reference = Executive. ",
      "Province variance = {prov_sfm}; CF-within-province variance = {cfmg_sfm}; ",
      "residual variance = {resid_sfm}. ",
      "Marginal R² = {round(r2_sfm$R2_marginal, 3)}; ",
      "conditional R² = {round(r2_sfm$R2_conditional, 3)}. ",
      "n = {nobs(M_sfm)}."
    )
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -term) |>
  cols_align(align = "left", columns = term) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = term)
  ) |>
  sub_missing()
```

As can be seen from the table of model results, FGD type did not have any significant effect on the SFM index. However, the governance index had a highly significant effect. Thus, CFs with higher governance index scores (i.e. those that are better governed) had better outcomes for SFM.

```{r}
#| label: fig-sfm_index
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Relationship between the CFMG governance index and SFM score. Each point represents one FGD session. The line shows the fitted relationship from the linear mixed model, with the shaded band representing ± 1 SE."
#| fig-height: 4.5
#| fig-width: 6

sfm_data |>
  ggplot(aes(x = gov_index, y = sfm_score)) +
  geom_jitter(width = 0.15, height = 0.1, colour = "#2A9D8F", alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", colour = "#1A6B63", fill = "#76B7A0", alpha = 0.25,
              linewidth = 0.8) +
  scale_x_continuous(name = "Governance index (out of 10)", breaks = 0:10) +
  scale_y_continuous(name = "SFM score (out of 5)", breaks = 0:5,
                     limits = c(-0.2, 5.2)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5)
  )
```

## Discussion

All the indicators of SFM point to a situation that is improving across that vast majority of CFs, including deforestation, charcoal production, fire frequency, wildlife abundance and natural resource management. Interestingly, there does not appear to be any indication of leakage for most CFs, as most FGD groups indicated that both deforestation and charcoal production were also decreasing outside their CF. In fact, there is some suggestion of positive leakage, whereby the improved management inside CFs has had a positive effect on their surroundings. There were some differences in responses among the different FGD groups, with community female FGDs tending to have a more positive view of the impact of CFM on SFM compared to the other groups. This is interesting and may indicate that under CFM females have more of a voice in forest and forest resource management. Traditionally, women are highly dependent on forest resources, but usually have little say in forest management decisions. It would be worth following up this study to investigate what is driving this positive take on CFM among community women.

Among the individual interviewees, private enterprise representatives and DFOs tended to have a somewhat more conservative view with respect to positive changes, as compared to individuals from the communities. This may reflect that these professionals have a broader perspective from which to understand changes in forest and forest resource management. However, it could also arise from the fact that they are not as intimately involved in the forest management and hence less aware of the changes. Ultimately, monitoring over a longer period will enable better understanding of how CFM is changing forest management.

Overall, CFM appears to be engendering a change towards improved SFM. However, it is important to note that a small number of FGD groups and individual interviewees suggested the situation was worsening in their CFs. Moreover, the governance index was a highly signficant predictor of the SFM index. For the success of CFM in Zambia, it will be important to identify CFs that are doing less well, and identify and address the governance issues.

Some CFMGs were practising forest restoration, mostly through protecting natural regeneration, and a few were establishing plantations. However, these activities were not widespread among CFMGs. Restoration - especially through protection of natural regeneration - should integral to CFM and could be implement across virtually all CFs in Zambia. As this is an important indicator of SFM and there are government targets to restore 2,000,000 ha by 2030, the FD should integrate restoration targets into forest management plans for CFMGs. A problem is that the boundaries of many existing CFs were drawn around areas containing on relatively good forest (historical deforestation rates are very low in most CFs — see next chapter on remote sensing). The FD and communities should consider expanding CF boundaries to incorporate areas for restoration activities and for new CFs such areas should be incorporated from the outset. Plantations are also an important opportunity for CFM. Plantations can supply local building and fuelwood needs, thereby reducing pressure on natural forests. Furthermore, through sustainable charcoal programmes, they can also supply energy for urban dwellers and again reduce pressure on natural forests. Zambia also has an acute shortage of high quality timber and imports large amounts. A focus on developing plantations through CFM would go some way to meeting these needs, as well as generating a valuable additional source of income for rural communities. As with restoration, the boundaries of CFs should be planned with plantation activities in mind, and these should be focused on heavily degraded area.

In conclusion, our survey data suggest that CFM in Zambia is driving substantial improvements in SFM, reducing deforestation, charcoal production, and fire frequency, increasing wildlife populations, and improving natural resource management. However, it is important to note that these changes were not ubiquitous across all CFs and CFMGs with lower governance index scores were performing significantly more poorly. While some CFMGs are practising restoration and establishing plantations, there is a substantial opportunity to expand the role of CFM in meeting national restoration and plantation targets.
