---
title: "Sustainable Forest Management"
author: "Rhett D Harrison"
format:
        html:
                css: custom.css
        pdf: default
filters:
- collapse-output
editor: visual
---

```{r}
#| label: Load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(grid)
library(gt)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(scales)
library(broom.mixed)
library(modelsummary)
library(ordinal)
library(ggwordcloud)

load("cfmg.RData")
```

## Summary

Most respondents indicated that deforestation, charcoal production, and fire frequency were decreasing, wildlife populations were increasing, and natural resource management was improving within CFs. Most FDG groups also indicated that both deforestation and charcoal production were decreasing outside their CF, suggesting there is no leakage effect. Community female FDGs tended to have a more positive view of the impact of CFM compared to the other groups. However, it is important to note that a small number of FDG groups and individual interviewees suggested the situation was worsening in their CFs. It will be important to identify CFs that are doing less well and address the problems. Few CFMGs were practising natural forest restoration or establishing plantations. Hence, there is a considerable opportunity to expand the contribution of CFM to meeting national restoration and plantation targets.

## Introduction

In this chapter, we consider the effect of community forestry on indicators of SFM. The information is based only on the responses of interviewees to our survey tool. A separate chapter will analyse satellite derived remote sensed data on fire frequency and deforestation.

### Deforestation

We asked interviewees whether CFM has reduced deforestation inside their CF and whether or not deforestation has increased outside their CF. An increase in deforestation outside the community forest might represent leakage, whereby activities involving clearing forest that are proscribed in the CF are displaced to areas around the CF.

#### Deforestation inside the community forest

```{r}
#| label: fig-reduced_forest_clearing
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: " 'Has CFM reduced deforestation inside the community forest?' The length of each bar segment represents the proportion of FGD groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    deforestation = case_match(
            impact_on_sfm_cfm_reduced_forest_clearing_cfma,
            "no_there_has_never_been_clearing_for_agriculture" ~ "No change",
            "Agricultural practices ceased way before CFM" ~ "No change",
            "One field for cassava" ~ "No change",
            "Yes there no more fields_  it's all forests" ~ "Significant",
            "Yes_ significant reduction. There was chitemene system before CFM" ~ "Significant",
            "no_increase" ~ "No change",
            "yes_moderate" ~ "Moderate",
            "yes_significant" ~ "Significant",
            .default = NA
    ),
    deforestation = factor(deforestation, levels = c("Significant","Moderate","No change")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(deforestation) |>
  count(interview_type, deforestation) |>
  group_by(interview_type) |>
  mutate(
    prop = n / sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = deforestation)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "No change"         = "#C0392B",
      "Moderate" = "#F0C040",
      "Significant"        = "#2A9D8F"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

A majority of groups report either a moderate or significant reduction in deforestation inside the CF. Moreover, these answers were consistent across the FGD groups. There were no significant differences in response.

::: {.callout-note collapse="true"}
### Deforestation - FGDs - Model methods

To examine the differences among FGD types in their assessment of whether CFM had reduced deforestation, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_deforestation_fgd
#| include: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_cfm_reduced_forest_clearing_cfma,
      "no_there_has_never_been_clearing_for_agriculture" ~ "No change",
      "Agricultural practices ceased way before CFM"     ~ "No change",
      "One field for cassava"                            ~ "No change",
      "Yes there no more fields_  it's all forests"      ~ "Significant",
      "Yes_ significant reduction. There was chitemene system before CFM" ~ "Significant",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("No change", "Moderate", "Significant"),
                           ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(deforestation) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M1 <- clmm(deforestation ~
              interview_type +
              (1 | province) +
              (1 | province_cfmg),
            data = test.data)

emm.M1     <- emmeans(M1, specs = ~ interview_type)
pairwise.M1 <- pairs(emm.M1, adjust = "Tukey")
```

```{r}
#| label: M1_deforestation_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

summary(M1)
```

```{r}
#| label: pairwise_deforestation_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

summary(pairwise.M1)
```

Among the individual interviewees (@tbl-deforestation_individual), again a majority indicated that CFM had reduced deforestation inside the CF. The differences among the interviewee roles were not significant.

```{r}
#| label: tbl-deforestation_individual
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived reduction in deforestation inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_cfm_reduced_forest_clearing_cfma,
      "no_there_has_never_been_clearing_for_agriculture" ~ "No change",
      "Agricultural practices ceased way before CFM"     ~ "No change",
      "One field for cassava"                            ~ "No change",
      "Yes there no more fields_  it's all forests"      ~ "Significant",
      "Yes_ significant reduction. There was chitemene system before CFM" ~ "Significant",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "NGO representative"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "NGO representative")
    )
  ) |>
  filter(!is.na(deforestation)) |>
  count(interviewee_role, deforestation, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, deforestation, prop) |>
  pivot_wider(names_from = deforestation, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(`No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived reduction in deforestation inside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```


::: {.callout-note collapse="true"}
### Reduced forest clearing - Individual interviewees - Model methods

To examine differences among individual interviewees in their assessment of whether CFM had reduced deforestation, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: test_data_deforestation_individual
#| echo: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_cfm_reduced_forest_clearing_cfma,
      "no_there_has_never_been_clearing_for_agriculture" ~ "No change",
      "Agricultural practices ceased way before CFM"     ~ "No change",
      "One field for cassava"                            ~ "No change",
      "Yes there no more fields_  it's all forests"      ~ "Significant",
      "Yes_ significant reduction. There was chitemene system before CFM" ~ "Significant",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("No change", "Moderate", "Significant"),
                           ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(deforestation) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M2 <- clmm(deforestation ~
              interviewee_role +
              interviewee_gender +
              interviewee_age +
              (1 | province) +
              (1 | province_cfmg),
            data = test.data)

emm.M2     <- emmeans(M2, specs = ~ interviewee_role)
pairwise.M2 <- pairs(emm.M2, adjust = "Tukey")
```

```{r}
#| label: M2_deforestation_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

summary(M2)
```

```{r}
#| label: pairwise_deforestation_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

summary(pairwise.M2)
```

#### Deforestation outside the community forest

To investigate possible leakage effects, we also asked interviewees whether CFM had led to an increase in deforestation outside their community forest.

```{r}
#| label: fig-increased_deforestation_outside
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: " 'Has CFM increased deforestation outside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_forest_clearing_increased_outside_cfma,
      "Clearing only done for field opening" ~ "No change",
      "In Mboshya there are no agricultural activities" ~ "No change",
      "It increases according to seasons_ especially during the tobacco harvesting" ~ "Moderate",
      "No change_ community members' concentration is on farming. They have appreciation for gover…" ~ "No change",
      "There has never been  cutting of trees out side the cfma" ~ "No change",
      "There has never been any forest clearing outside the cfma." ~ "No change",
      "There has never been cutting of trees" ~ "No change",
      "no_decrease" ~ "Decrease",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("Significant", "Moderate", "No change", "Decrease")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(deforestation) |>
  count(interview_type, deforestation) |>
  group_by(interview_type) |>
  mutate(
    prop = n / sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = deforestation)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Decrease" = "#2A9D8F",
      "No change"   = "#76B7A0",
      "Moderate"    = "#F0C040",
      "Significant" = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1),    colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Contrary to the expectations of leakage, a majority of the FGD groups responded that CFM actually reduced deforestation outside their CF. Moreover, the responses were consistent across FGD groups. There were no significant differences in response.

::: {.callout-note collapse="true"}
### Deforestation outside the CF - FGDs - Model methods

To examine the differences among FGD types in their assessment of whether CFM had increased deforestation outside the CF, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_deforestation_outside_fgd
#| include: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_forest_clearing_increased_outside_cfma,
      "Clearing only done for field opening" ~ "No change",
      "In Mboshya there are no agricultural activities" ~ "No change",
      "It increases according to seasons_ especially during the tobacco harvesting" ~ "Moderate",
      "No change_ community members' concentration is on farming. They have appreciation for gover…" ~ "No change",
      "There has never been  cutting of trees out side the cfma" ~ "No change",
      "There has never been any forest clearing outside the cfma." ~ "No change",
      "There has never been cutting of trees" ~ "No change",
      "no_decrease" ~ "Decrease",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("Significant", "Moderate", "No change", "Decrease")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(deforestation) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M3 <- clmm(deforestation ~
              interview_type +
              (1 | province) +
              (1 | province_cfmg),
            data = test.data)

emm.M3     <- emmeans(M3, specs = ~ interview_type)
pairwise.M3 <- pairs(emm.M3, adjust = "Tukey")
```

```{r}
#| label: M3_deforestation_outside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

summary(M3)
```

```{r}
#| label: pairwise_deforestation_outside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

summary(pairwise.M3)
```

Among the individual interviewees (@tbl-deforestation_outside_individual), the individuals from the communities agreed that deforestation outside their CF had decreased. However, NGO representatives and DFOs were less likely to suggest deforestation had decreased outside the CF and more likely to suggest a moderate increase in deforestation.

```{r}
#| label: tbl-deforestation_outside_individual
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived increase in deforestation outside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_forest_clearing_increased_outside_cfma,
      "Clearing only done for field opening" ~ "No change",
      "In Mboshya there are no agricultural activities" ~ "No change",
      "It increases according to seasons_ especially during the tobacco harvesting" ~ "Moderate",
      "No change_ community members' concentration is on farming. They have appreciation for gover…" ~ "No change",
      "There has never been  cutting of trees out side the cfma" ~ "No change",
      "There has never been any forest clearing outside the cfma." ~ "No change",
      "There has never been cutting of trees" ~ "No change",
      "no_decrease" ~ "Decrease",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("Decrease", "No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "NGO representative"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "NGO representative")
    )
  ) |>
  filter(!is.na(deforestation)) |>
  count(interviewee_role, deforestation, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, deforestation, prop) |>
  pivot_wider(names_from = deforestation, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Decrease, `No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived increase in deforestation outside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Deforestation outside the CF - Individual interviewees - Model methods

To examine differences among individual interviewees in their assessment of whether CFM had increased deforestation outside the community forest, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: test_data_deforestation_outside_individual
#| echo: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    deforestation = case_match(
      impact_on_sfm_forest_clearing_increased_outside_cfma,
      "Clearing only done for field opening" ~ "No change",
      "In Mboshya there are no agricultural activities" ~ "No change",
      "It increases according to seasons_ especially during the tobacco harvesting" ~ "Moderate",
      "No change_ community members' concentration is on farming. They have appreciation for gover…" ~ "No change",
      "There has never been  cutting of trees out side the cfma" ~ "No change",
      "There has never been any forest clearing outside the cfma." ~ "No change",
      "There has never been cutting of trees" ~ "No change",
      "no_decrease" ~ "Decrease",
      "no_increase"    ~ "No change",
      "yes_moderate"   ~ "Moderate",
      "yes_significant" ~ "Significant",
      .default = NA
    ),
    deforestation = factor(deforestation,
                           levels = c("Significant", "Moderate", "No change", "Decrease")),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(deforestation) |>
  mutate(province_cfmg = interaction(province, cfmg_name))

M4 <- clmm(deforestation ~
              interviewee_role +
              interviewee_gender +
              interviewee_age +
              (1 | province) +
              (1 | province_cfmg),
            data = test.data)

emm.M4     <- emmeans(M4, specs = ~ interviewee_role)
pairwise.M4 <- pairs(emm.M4, adjust = "Tukey")
```

The differences in opinion between the CFMG executive members and NGO representatives and DFOs, and between traditional leaders and NGO representatives and DFOs, were significant.

```{r}
#| label: M4_deforestation_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

summary(M4)
```

```{r}
#| label: pairwise_deforestation_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

summary(pairwise.M4)
```

### Charcoal production

We asked interviewees whether CFM has reduced charcoal production inside their CF, and whether it has increased immediately outside their CF.

#### Charcoal production inside the community forest

```{r}
#| label: fig-reduced_charcoal_inside
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: " 'Has CFM reduced charcoal production inside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "No. There have never been  charcoal  production  in this forest"    ~ "No change",
      "No_ charcoal production is not a common practice for community members"    ~ "No change",
      "Some areas are encroached by charcoal production"    ~ "No change",
      "There has never been  charcoal  production"    ~ "No change",
      "There has never been  charcoal  production in this forest"    ~ "No change",
      "yes_reduced"   ~ "Moderate",
      "yes_high" ~ "Significant",
      "no" ~ "Significant",
      "no_stopped" ~ "Significant",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Significant", "Moderate", "No change")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(charcoal) |>
  count(interview_type, charcoal) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = charcoal)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "No change"   = "#C0392B",
      "Moderate"    = "#F0C040",
      "Significant" = "#2A9D8F"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1),    colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Almost all the FGD groups agreed that charcoal making inside their CF had decreased, with most indicating that charcoal making had decreased significantly. The small differences in responses among FGD groups were not significant.

::: {.callout-note collapse="true"}
### Charcoal production inside the CF - FGDs - Model methods

To examine the differences among FGD types in their assessment of whether CFM had reduced charcoal production inside the CF, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_charcoal_inside_fgd
#| include: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "No. There have never been  charcoal  production  in this forest"    ~ "No change",
      "No_ charcoal production is not a common practice for community members"    ~ "No change",
      "Some areas are encroached by charcoal production"    ~ "No change",
      "There has never been  charcoal  production"    ~ "No change",
      "There has never been  charcoal  production in this forest"    ~ "No change",
      "yes_reduced"   ~ "Moderate",
      "yes_high" ~ "Significant",
      "no" ~ "Significant",
      "no_stopped" ~ "Significant",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(charcoal) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$charcoal) >= 2) {
  M5 <- clmm(charcoal ~
                interview_type +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M5     <- emmeans(M5, specs = ~ interview_type)
  pairwise.M5 <- pairs(emm.M5, adjust = "Tukey")
} else {
  M5 <- NULL
  cat("Model not fitted: only", nlevels(test.data$charcoal),
      "level(s) present in charcoal after filtering.\n")
}
```

```{r}
#| label: M5_charcoal_inside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M5)) summary(M5) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_charcoal_inside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M5)) summary(pairwise.M5) else cat("Model not fitted.")
```

Likewise, among the individual interviewees (@tbl-charcoal_inside_individual), almost all interviewees agreed that there had been a substantial decrease in charcoal making inside their CF. Again, this result was consistent across interviewee roles (there were no significant differences among roles).

```{r}
#| label: tbl-charcoal_inside_individual
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived reduction in charcoal production inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "No. There have never been  charcoal  production  in this forest"    ~ "No change",
      "No_ charcoal production is not a common practice for community members"    ~ "No change",
      "Some areas are encroached by charcoal production"    ~ "No change",
      "There has never been  charcoal  production"    ~ "No change",
      "There has never been  charcoal  production in this forest"    ~ "No change",
      "yes_reduced"   ~ "Moderate",
      "yes_high" ~ "Significant",
      "no" ~ "Significant",
      "no_stopped" ~ "Significant",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "NGO representative"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "NGO representative")
    )
  ) |>
  filter(!is.na(charcoal)) |>
  count(interviewee_role, charcoal, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, charcoal, prop) |>
  pivot_wider(names_from = charcoal, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(`No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived reduction in charcoal production inside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Charcoal production inside the CF - Individual interviewees - Model methods

To examine differences among individual interviewees in their assessment of whether CFM had reduced charcoal production inside the CF, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: test_data_charcoal_inside_individual
#| echo: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_in_cfma,
      "No. There have never been  charcoal  production  in this forest"    ~ "No change",
      "No_ charcoal production is not a common practice for community members"    ~ "No change",
      "Some areas are encroached by charcoal production"    ~ "No change",
      "There has never been  charcoal  production"    ~ "No change",
      "There has never been  charcoal  production in this forest"    ~ "No change",
      "yes_reduced"   ~ "Moderate",
      "yes_high" ~ "Significant",
      "no" ~ "Significant",
      "no_stopped" ~ "Significant",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(charcoal) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$charcoal) >= 2) {
  M6 <- clmm(charcoal ~
                interviewee_role +
                interviewee_gender +
                interviewee_age +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M6     <- emmeans(M6, specs = ~ interviewee_role)
  pairwise.M6 <- pairs(emm.M6, adjust = "Tukey")
} else {
  M6 <- NULL
  cat("Model not fitted: only", nlevels(test.data$charcoal),
      "level(s) present in charcoal after filtering.\n")
}
```

```{r}
#| label: M6_charcoal_inside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M6)) summary(M6) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_charcoal_inside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M6)) summary(pairwise.M6) else cat("Model not fitted.")
```

#### Charcoal production outside the community forest

We also asked interviewees whether CFM has led to an increase in charcoal production outside their CF.

```{r}
#| label: fig-increased_charcoal_outside
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: " 'Has CFM increased charcoal production outside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_increased_outside_cfma,
      "no_reduced"    ~ "Decrease",
      "no_change"    ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      "Yes_ moderate increase. Only for household use"   ~ "Moderate",
      "Moderate decrease"    ~ "Decrease",
      "Due to increase in population_ some charcoal production now being observed in some open are…"   ~ "Moderate",
      "CFMG members not involved in charcoal production"    ~ "No change",
      "Charcoal  production  is not allowed in the chiefdom"    ~ "No change",
      "Charcoal  production is not allowed in the chiefdom"    ~ "No change",
      "Charcoal production  is not allowed  in the chiefdom"    ~ "No change",
      "Charcoal production has been banned from the chiefdom"    ~ "No change",
      "Charcoal production is not common in this part of the district"    ~ "No change",
      "Charcoal production is not commonly practiced. There is no market for charcoal production"    ~ "No change",
      "Charcoal production is not practiced in the community"    ~ "No change",
      "Charcoal production not a common practice in the community"    ~ "No change",
      "Charcoal production only done when clearing an area for new agricultural fields"    ~ "No change",
      "No charcoal  production"    ~ "No change",
      "No charcoal  production In this area"    ~ "No change",
      "No charcoal  production in this community"    ~ "No change",
      "No charcoal production"    ~ "No change",
      "No charcoal production near CFMA"    ~ "No change",
      "No. There has never been charcoal  production"    ~ "No change",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Significant", "Moderate", "No change", "Decrease")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(charcoal) |>
  count(interview_type, charcoal) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = charcoal)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Decrease"    = "#2A9D8F",
      "No change"   = "#76B7A0",
      "Moderate"    = "#F0C040",
      "Significant" = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1),    colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen, the vast majority of groups indicated that charcoal making had also decreased outside their CF. However, there were a small number that indicated charcoal making had increased, some indicating it had increased significantly. Female FGD groups were significantly more likely to indicate that charcoal making had decreased outside the CF than all other groups (CFMG executive members, male or youth community members). There were no significant differences among these other groups, who most often indicated that there was no change in the amount of charcoal making.  

::: {.callout-note collapse="true"}
### Charcoal production outside the CF - FGDs - Model methods

To examine the differences among FGD types in their assessment of whether CFM had increased charcoal production outside the community forest, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_charcoal_outside_fgd
#| include: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_increased_outside_cfma,
      "no_reduced"    ~ "Decrease",
      "no_change"    ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      "Yes_ moderate increase. Only for household use"   ~ "Moderate",
      "Moderate decrease"    ~ "Decrease",
      "Due to increase in population_ some charcoal production now being observed in some open are…"   ~ "Moderate",
      "CFMG members not involved in charcoal production"    ~ "No change",
      "Charcoal  production  is not allowed in the chiefdom"    ~ "No change",
      "Charcoal  production is not allowed in the chiefdom"    ~ "No change",
      "Charcoal production  is not allowed  in the chiefdom"    ~ "No change",
      "Charcoal production has been banned from the chiefdom"    ~ "No change",
      "Charcoal production is not common in this part of the district"    ~ "No change",
      "Charcoal production is not commonly practiced. There is no market for charcoal production"    ~ "No change",
      "Charcoal production is not practiced in the community"    ~ "No change",
      "Charcoal production not a common practice in the community"    ~ "No change",
      "Charcoal production only done when clearing an area for new agricultural fields"    ~ "No change",
      "No charcoal  production"    ~ "No change",
      "No charcoal  production In this area"    ~ "No change",
      "No charcoal  production in this community"    ~ "No change",
      "No charcoal production"    ~ "No change",
      "No charcoal production near CFMA"    ~ "No change",
      "No. There has never been charcoal  production"    ~ "No change",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Decrease", "No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(charcoal) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$charcoal) >= 2) {
  M7 <- clmm(charcoal ~
                interview_type +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M7     <- emmeans(M7, specs = ~ interview_type)
  pairwise.M7 <- pairs(emm.M7, adjust = "Tukey")
} else {
  M7 <- NULL
  cat("Model not fitted: only", nlevels(test.data$charcoal),
      "level(s) present in charcoal after filtering.\n")
}
```

```{r}
#| label: M7_charcoal_outside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M7)) summary(M7) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_charcoal_outside_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M7)) summary(pairwise.M7) else cat("Model not fitted.")
```

Among the individual interviewees (@tbl-charcoal_outside_individual), similar to the results about deforestation, most community interviewees suggested there had been a decrease or no change in charcoal production outside the CF. However, NGO representatives and DFOs were significantly more likely to indicate that there had been an increase in charcoal production outside the CF. 

```{r}
#| label: tbl-charcoal_outside_individual
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived increase in charcoal production outside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_increased_outside_cfma,
      "no_reduced"    ~ "Decrease",
      "no_change"    ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      "Yes_ moderate increase. Only for household use"   ~ "Moderate",
      "Moderate decrease"    ~ "Decrease",
      "Due to increase in population_ some charcoal production now being observed in some open are…"   ~ "Moderate",
      "CFMG members not involved in charcoal production"    ~ "No change",
      "Charcoal  production  is not allowed in the chiefdom"    ~ "No change",
      "Charcoal  production is not allowed in the chiefdom"    ~ "No change",
      "Charcoal production  is not allowed  in the chiefdom"    ~ "No change",
      "Charcoal production has been banned from the chiefdom"    ~ "No change",
      "Charcoal production is not common in this part of the district"    ~ "No change",
      "Charcoal production is not commonly practiced. There is no market for charcoal production"    ~ "No change",
      "Charcoal production is not practiced in the community"    ~ "No change",
      "Charcoal production not a common practice in the community"    ~ "No change",
      "Charcoal production only done when clearing an area for new agricultural fields"    ~ "No change",
      "No charcoal  production"    ~ "No change",
      "No charcoal  production In this area"    ~ "No change",
      "No charcoal  production in this community"    ~ "No change",
      "No charcoal production"    ~ "No change",
      "No charcoal production near CFMA"    ~ "No change",
      "No. There has never been charcoal  production"    ~ "No change",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Decrease", "No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "NGO representative"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "NGO representative")
    )
  ) |>
  filter(!is.na(charcoal)) |>
  count(interviewee_role, charcoal, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, charcoal, prop) |>
  pivot_wider(names_from = charcoal, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Decrease, `No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived increase in charcoal production outside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Charcoal production outside the CF - Individual interviewees - Model methods

To examine differences among individual interviewees in their assessment of whether CFM had increased charcoal production outside the community forest, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: test_data_charcoal_outside_individual
#| echo: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    charcoal = case_match(
      impact_on_sfm_charcoal_production_increased_outside_cfma,
      "no_reduced"    ~ "Decrease",
      "no_change"    ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      "Yes_ moderate increase. Only for household use"   ~ "Moderate",
      "Moderate decrease"    ~ "Decrease",
      "Due to increase in population_ some charcoal production now being observed in some open are…"   ~ "Moderate",
      "CFMG members not involved in charcoal production"    ~ "No change",
      "Charcoal  production  is not allowed in the chiefdom"    ~ "No change",
      "Charcoal  production is not allowed in the chiefdom"    ~ "No change",
      "Charcoal production  is not allowed  in the chiefdom"    ~ "No change",
      "Charcoal production has been banned from the chiefdom"    ~ "No change",
      "Charcoal production is not common in this part of the district"    ~ "No change",
      "Charcoal production is not commonly practiced. There is no market for charcoal production"    ~ "No change",
      "Charcoal production is not practiced in the community"    ~ "No change",
      "Charcoal production not a common practice in the community"    ~ "No change",
      "Charcoal production only done when clearing an area for new agricultural fields"    ~ "No change",
      "No charcoal  production"    ~ "No change",
      "No charcoal  production In this area"    ~ "No change",
      "No charcoal  production in this community"    ~ "No change",
      "No charcoal production"    ~ "No change",
      "No charcoal production near CFMA"    ~ "No change",
      "No. There has never been charcoal  production"    ~ "No change",
      .default = NA
    ),
    charcoal = factor(charcoal,
                      levels = c("Decrease", "No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(charcoal) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$charcoal) >= 2) {
  M8 <- clmm(charcoal ~
                interviewee_role +
                interviewee_gender +
                interviewee_age +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M8     <- emmeans(M8, specs = ~ interviewee_role)
  pairwise.M8 <- pairs(emm.M8, adjust = "Tukey")
} else {
  M8 <- NULL
  cat("Model not fitted: only", nlevels(test.data$charcoal),
      "level(s) present in charcoal after filtering.\n")
}
```

```{r}
#| label: M8_charcoal_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M8)) summary(M8) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_charcoal_outside_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M8)) summary(pairwise.M8) else cat("Model not fitted.")
```

### Fire

We asked interviewees about fire monitoring practices and whether CFM had reduced fire incidence inside their CF. We did not ask about changes in fire frequency outside the CF, because clearing fields for agriculture or improving grazing often involves burning.

#### Fire monitoring

```{r}
#| label: fig-fire_monitoring
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFMGs have established fire monitoring in the community forest by province. The length of the bars represents the number of CFMGs either monitoring fires (Yes) or not (No). The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(impact_on_sfm_fire_monitoring_in_cfma, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen in the figure, most CFMGs have instigated a system of fire monitoring. However, ten CFMGs across four provinces from our sample of 75 did not have any system of fire monitoring.

#### Fire incidence inside the community forest

We asked interviewees whether CFM has reduced fire incidence inside their CF.

```{r}
#| label: fig-reduced_fire_inside
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "'Has CFM reduced fire incidence inside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "no_incidents_have_increased" ~ "Increase",
      "no_change"     ~ "No change",
      "yes_moderate_reduction"    ~ "Moderate",
      "yes_significant_reduction" ~ "Significant",
      "No change observed as at now" ~ "No change",
      .default = NA
    ),
    fire = factor(fire, levels = c("Significant", "Moderate", "No change","Increase")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(fire) |>
  count(interview_type, fire) |>
  group_by(interview_type) |>
  mutate(
    prop = n / sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = fire)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Increase" = "#C0392B",
      "No change"   = "#F0C040",
      "Moderate"    = "#2A9D8F",
      "Significant" = "#76B7A0"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen, a majority of groups reported a moderate or significant reduction in fire incidence in the CF. Female FDGs reported more significant declines in fire frequency compared to the other groups, who reported more moderate declines. These differences between female FDG groups and the other FDG groups (CFMG Executives, male and youth) were significant.

::: {.callout-note collapse="true"}
### Fire incidence - FGDs - Model methods

To examine the differences among FGD types in their assessment of whether CFM had reduced fire incidence, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_fire_fgd
#| include: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "no_incidents_have_increased" ~ "Increase",
      "no_change"     ~ "No change",
      "yes_moderate_reduction"    ~ "Moderate",
      "yes_significant_reduction" ~ "Significant",
      "No change observed as at now" ~ "No change",
      .default = NA
    ),
    fire = factor(fire,
                  levels = c("Increase","No change", "Moderate", "Significant"),
                  ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(fire) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$fire) >= 2) {
  M9 <- clmm(fire ~
                interview_type +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M9      <- emmeans(M9, specs = ~ interview_type)
  pairwise.M9 <- pairs(emm.M9, adjust = "Tukey")
} else {
  M9 <- NULL
  cat("Model not fitted: only", nlevels(test.data$fire),
      "level(s) present in fire after filtering.\n")
}
```

```{r}
#| label: M9_fire_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M9)) summary(M9) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_fire_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M9)) summary(pairwise.M9) else cat("Model not fitted.")
```

Among the individual interviewees (@tbl-fire_individual), there were also some differences in opinion. Community members tended to say there had been a significant reduction in fire frequency inside their CF, whereas NGO representatives and DFOs more often suggested a moderate reduction or no change. However, these differences were not significant.

```{r}
#| label: tbl-fire_individual
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived reduction in fire incidence inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "no_incidents_have_increased" ~ "Increase",
      "no_change"     ~ "No change",
      "yes_moderate_reduction"    ~ "Moderate",
      "yes_significant_reduction" ~ "Significant",
      "No change observed as at now" ~ "No change",
      .default = NA
    ),
    fire = factor(fire,
                  levels = c("Significant", "Moderate", "No change", "Increase"),
                  ordered = TRUE),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "NGO representative"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "NGO representative")
    )
  ) |>
  filter(!is.na(fire)) |>
  count(interviewee_role, fire, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, fire, prop) |>
  pivot_wider(names_from = fire, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Significant, Moderate, `No change`, Increase), decimals = 0) |>
  tab_header(
    title    = "Perceived reduction in fire incidence inside the community forest by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Fire incidence - Individual interviewees - Model methods

To examine differences among individual interviewees in their assessment of whether CFM had reduced fire incidence, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: test_data_fire_individual
#| echo: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    fire = case_match(
      impact_on_sfm_fire_incidence_reduction_cfma,
      "no_incidents_have_increased" ~ "Increase",
      "no_change"     ~ "No change",
      "yes_moderate_reduction"    ~ "Moderate",
      "yes_significant_reduction" ~ "Significant",
      "No change observed as at now" ~ "No change",
      .default = NA
    ),
    fire = factor(fire,
                  levels = c("Increase","No change", "Moderate", "Significant"),
                  ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(fire) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$fire) >= 2) {
  M10 <- clmm(fire ~
                 interviewee_role +
                 interviewee_gender +
                 interviewee_age +
                 (1 | province) +
                 (1 | province_cfmg),
               data = test.data)
  emm.M10      <- emmeans(M10, specs = ~ interviewee_role)
  pairwise.M10 <- pairs(emm.M10, adjust = "Tukey")
} else {
  M10 <- NULL
  cat("Model not fitted: only", nlevels(test.data$fire),
      "level(s) present in fire after filtering.\n")
}
```

```{r}
#| label: M10_fire_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M10)) summary(M10) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_fire_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M10)) summary(pairwise.M10) else cat("Model not fitted.")
```

### Wildlife

We asked interviewees whether CFM had led to an increase in wildlife inside their CF.

```{r}
#| label: fig-wildlife_increase
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "'Has CFM led to an increase in wildlife in the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "no_reduced"              ~ "Decrease",
      "no_change"               ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      .default = NA
    ),
    wildlife = factor(wildlife,
                      levels = c("Significant", "Moderate", "No change", "Decrease")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(wildlife) |>
  count(interview_type, wildlife) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = wildlife)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Significant" = "#2A9D8F",
      "Moderate"    = "#F0C040",
      "No change"   = "#E07B54",
      "Decrease"    = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen, a magority of groups indicated a significant or moderate increase in wildlife in their CF. However, there were some that indicated there had been a decrease. The responses of the community female FDGs appear somewhat different from the other groups, but these differences were not significant.

::: {.callout-note collapse="true"}
### Wildlife increase - FGDs - Model methods

To examine the differences among FGD types in their assessment of whether CFM had increased wildlife, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_wildlife_fgd
#| include: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "no_reduced"              ~ "Decrease",
      "no_change"               ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      .default = NA
    ),
    wildlife = factor(wildlife,
                      levels = c("Decrease", "No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(wildlife) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$wildlife) >= 2) {
  M11 <- clmm(wildlife ~
                interview_type +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M11      <- emmeans(M11, specs = ~ interview_type)
  pairwise.M11 <- pairs(emm.M11, adjust = "Tukey")
} else {
  M11 <- NULL
  cat("Model not fitted: only", nlevels(test.data$wildlife),
      "level(s) present in wildlife after filtering.\n")
}
```

```{r}
#| label: M11_wildlife_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M11)) summary(M11) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_wildlife_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M11)) summary(pairwise.M11) else cat("Model not fitted.")
```

Among the individual interviewees (@tbl-wildlife_individual), there were also some differences in opinion. For example, only 3% of NGO representatives suggested there had been a decline in wildlife in the CF, compared to between 10% and 22% for the other interviewee roles. However, again these differences were not significant.

```{r}
#| label: tbl-wildlife_individual
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived increase in wildlife inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "no_reduced"              ~ "Decrease",
      "no_change"               ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      .default = NA
    ),
    wildlife = factor(wildlife,
                      levels = c("Decrease", "No change", "Moderate", "Significant")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "NGO representative"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "NGO representative")
    )
  ) |>
  filter(!is.na(wildlife)) |>
  count(interviewee_role, wildlife, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, wildlife, prop) |>
  pivot_wider(names_from = wildlife, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(Decrease, `No change`, Moderate, Significant), decimals = 0) |>
  tab_header(
    title    = "Perceived increase in wildlife by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Wildlife increase - Individual interviewees - Model methods

To examine differences among individual interviewees in their assessment of whether CFM had increased wildlife, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: test_data_wildlife_individual
#| echo: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    wildlife = case_match(
      impact_on_sfm_wildlife_increase,
      "no_reduced"              ~ "Decrease",
      "no_change"               ~ "No change",
      "yes_moderate_increase"   ~ "Moderate",
      "yes_significant_increase" ~ "Significant",
      .default = NA
    ),
    wildlife = factor(wildlife,
                      levels = c("Decrease", "No change", "Moderate", "Significant"),
                      ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(wildlife) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$wildlife) >= 2) {
  M12 <- clmm(wildlife ~
                 interviewee_role +
                 interviewee_gender +
                 interviewee_age +
                 (1 | province) +
                 (1 | province_cfmg),
               data = test.data)
  emm.M12      <- emmeans(M12, specs = ~ interviewee_role)
  pairwise.M12 <- pairs(emm.M12, adjust = "Tukey")
} else {
  M12 <- NULL
  cat("Model not fitted: only", nlevels(test.data$wildlife),
      "level(s) present in wildlife after filtering.\n")
}
```

```{r}
#| label: M12_wildlife_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M12)) summary(M12) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_wildlife_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M12)) summary(pairwise.M12) else cat("Model not fitted.")
```

### Forest resources management

We asked interviewees whether CFM had improved the management of forest resources inside their CF.

```{r}
#| label: fig-forest_resources_management
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "'Has CFM improved forest resources management inside the community forest?' The length of each bar segment represents the proportion of groups in that category. The numbers superimposed on the bars represent the actual number of groups in each category."

cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    frm = case_match(
      impact_on_sfm_forest_resources_management,
      "no_management_has_worsened"    ~ "Management worse",
      "no_change"                     ~ "No change",
      "yes_moderate_improvement"      ~ "Moderate improvement",
      "yes_significant_improvement"   ~ "Significant improvement",
      .default = NA
    ),
    frm = factor(frm, levels = c("Significant improvement", "Moderate improvement",
                                 "No change", "Management worse")),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Youth", "Male", "Female", "Executive")
    )
  ) |>
  drop_na(frm) |>
  count(interview_type, frm) |>
  group_by(interview_type) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  ggplot(aes(y = interview_type, x = prop, fill = frm)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = ifelse(prop >= 0.05, n, "")),
    position = position_stack(vjust = 0.5),
    size = 3.5, fontface = "bold", colour = "white"
  ) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = 0.01),
    name   = "Proportion of groups"
  ) +
  scale_fill_manual(
    values = c(
      "Significant improvement" = "#2A9D8F",
      "Moderate improvement"    = "#F0C040",
      "No change"               = "#E07B54",
      "Management worse"        = "#C0392B"
    ),
    guide = guide_legend(reverse = TRUE, nrow = 2)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

As can be seen, while responses vary among FGD groups, most groups agreed that the management of forest resources had improved moderately or significantly under CFM, and only a small number of primarily youth groups considered that it had gotten worse. A very high proportion of community female FDG groups considered that the management of forest resources was significantly better under CFM, a difference that was statistically significant.

::: {.callout-note collapse="true"}
### Forest resources management - FGDs - Model methods

To examine the differences among FGD types in their assessment of whether CFM had improved forest resources management, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We used FGD type as the factor. We examined the significance of pairwise differences among FGD types using Tukey HSD.
:::

```{r}
#| label: model_frm_fgd
#| include: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type != "individual_interview") |>
  mutate(
    frm = case_match(
      impact_on_sfm_forest_resources_management,
      "no_management_has_worsened"    ~ "Management worse",
      "no_change"                     ~ "No change",
      "yes_moderate_improvement"      ~ "Moderate improvement",
      "yes_significant_improvement"   ~ "Significant improvement",
      .default = NA
    ),
    frm = factor(frm,
                 levels = c("Management worse", "No change",
                            "Moderate improvement", "Significant improvement"),
                 ordered = TRUE),
    interview_type = factor(
      case_match(
        interview_type,
        "cfmg_executive" ~ "Executive",
        "cfmg_female"    ~ "Female",
        "cfmg_male"      ~ "Male",
        "cfmg_youth"     ~ "Youth"
      ),
      levels = c("Executive", "Female", "Male", "Youth")
    )
  ) |>
  drop_na(frm) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$frm) >= 2) {
  M13 <- clmm(frm ~
                interview_type +
                (1 | province) +
                (1 | province_cfmg),
              data = test.data)
  emm.M13      <- emmeans(M13, specs = ~ interview_type)
  pairwise.M13 <- pairs(emm.M13, adjust = "Tukey")
} else {
  M13 <- NULL
  cat("Model not fitted: only", nlevels(test.data$frm),
      "level(s) present in frm after filtering.\n")
}
```

```{r}
#| label: M13_frm_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M13)) summary(M13) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_frm_fgd
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M13)) summary(pairwise.M13) else cat("Model not fitted.")
```

We also asked which specific forest resources had improved under CFM (@fig-harvested_resources).

```{r}
#| label: fig-harvested_resources
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Forest resources that have improved under CFM, as reported by female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting an improvement for each resource."
#| fig-width: 10
#| fig-height: 7

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(
    "Pupwe"            = impact_on_sfm_harvested_forest_resources_pupwe,
    "Devil's claw"     = impact_on_sfm_harvested_forest_resources_devils_claw,
    "Fish"             = impact_on_sfm_harvested_forest_resources_fish,
    "Lusala"           = impact_on_sfm_harvested_forest_resources_lusala,
    "Mungongo"         = impact_on_sfm_harvested_forest_resourcesmungongo,
    "Bamboo"           = impact_on_sfm_harvested_forest_resources_bamboo,
    "Munkoyo"          = impact_on_sfm_harvested_forest_resources_munkoyo,
    "Poles"            = impact_on_sfm_harvested_forest_resources_poles,
    "Bushmeat"         = impact_on_sfm_harvested_forest_resources_bushmeat,
    "Timber"           = impact_on_sfm_harvested_forest_resources_timber,
    "Fibres"           = impact_on_sfm_harvested_forest_resources_fibers,
    "Chikanda"         = impact_on_sfm_harvested_forest_resources_chikanda,
    "Firewood"         = interview_grp_impact_on_sfm_harvested_forest_resources_firewood,
    "Thatching grass"  = impact_on_sfm_harvested_forest_resources_thatching_grass,
    "Medicinal plants" = impact_on_sfm_harvested_forest_resources_medicinal_plants,
    "Caterpillars"     = impact_on_sfm_harvested_forest_resources_caterpillars,
    "Honey"            = impact_on_sfm_harvested_forest_resources_honey,
    "Wild fruits"      = impact_on_sfm_harvested_forest_resources_wild_fruits,
    "Mushrooms"        = impact_on_sfm_harvested_forest_resources_mushrooms
  ) |>
  pivot_longer(everything(), names_to = "word", values_to = "value") |>
  filter(!is.na(value)) |>
  group_by(word) |>
  summarise(freq = sum(as.integer(value) == 1, na.rm = TRUE), .groups = "drop") |>
  filter(freq > 0) |>
  ggplot(aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 30)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

Among the individual interviewees (@tbl-frm_individual) the vast majority of interviewees considered that the management of forest resources had improved under CFM. There was a significant difference in responses between community members and DFOs. The DFOs were more likely to suggest there had been no change or only a moderate improvement in forest resource management under CFM, compared to the much higher proportion of interviewees indicating a significant improvement among community representatives. 

```{r}
#| label: tbl-frm_individual
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| tbl-cap: "Perceived change in forest resources management inside the community forest by individual interviewee role."

cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    frm = case_match(
      impact_on_sfm_forest_resources_management,
      "no_management_has_worsened"    ~ "Management worse",
      "no_change"                     ~ "No change",
      "yes_moderate_improvement"      ~ "Moderate improvement",
      "yes_significant_improvement"   ~ "Significant improvement",
      .default = NA
    ),
    frm = factor(frm,
                 levels = c("Management worse", "No change",
                            "Moderate improvement", "Significant improvement")),
    interviewee_role = factor(
      case_match(
        interviewee_role,
        "cfmg_exec_member" ~ "CFMG executive member",
        "cfmg_member"      ~ "Community member",
        "trad_leader"      ~ "Traditional leader",
        "dfo"              ~ "District Forestry Officer",
        "NGO_rep"          ~ "NGO representative"
      ),
      levels = c("CFMG executive member", "Community member", "Traditional leader",
                 "District Forestry Officer", "NGO representative")
    )
  ) |>
  filter(!is.na(frm)) |>
  count(interviewee_role, frm, name = "obs") |>
  group_by(interviewee_role) |>
  mutate(n = sum(obs), prop = obs / n) |>
  ungroup() |>
  select(interviewee_role, n, frm, prop) |>
  pivot_wider(names_from = frm, values_from = prop, values_fill = 0) |>
  arrange(interviewee_role) |>
  select(interviewee_role, n,
         `Management worse`, `No change`, `Moderate improvement`, `Significant improvement`) |>
  gt() |>
  cols_label(interviewee_role = "Role") |>
  fmt_percent(columns = c(`Management worse`, `No change`, `Moderate improvement`, `Significant improvement`), decimals = 0) |>
  tab_header(
    title    = "Perceived improvement in forest resources management by interviewee role",
    subtitle = "Values represent the % of respondents in each response category. n = the total number of interview responses."
  ) |>
  opt_align_table_header(align = "left") |>
  cols_align(align = "center", columns = -interviewee_role) |>
  cols_align(align = "left", columns = interviewee_role) |>
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = interviewee_role)
  ) |>
  sub_missing()
```

::: {.callout-note collapse="true"}
### Forest resources management - Individual interviewees - Model methods

To examine differences among individual interviewees in their assessment of whether CFM had improved forest resources management, we employed a cumulative link mixed model (CLMM) with random intercepts for province and CF nested within province. We included interviewee role, interviewee gender and interviewee age as factors. To examine pairwise differences among interviewee roles, we employed Tukey HSD.
:::

```{r}
#| label: test_data_frm_individual
#| echo: true
#| warning: false
#| message: false
#| error:   false
#| code-fold: true

test.data <- cfmg |>
  filter(interview_type == "individual_interview") |>
  mutate(
    frm = case_match(
      impact_on_sfm_forest_resources_management,
      "no_management_has_worsened"    ~ "Management worse",
      "no_change"                     ~ "No change",
      "yes_moderate_improvement"      ~ "Moderate improvement",
      "yes_significant_improvement"   ~ "Significant improvement",
      .default = NA
    ),
    frm = factor(frm,
                 levels = c("Management worse", "No change",
                            "Moderate improvement", "Significant improvement"),
                 ordered = TRUE),
    interviewee_gender = demo_information_interviewee_gender,
    interviewee_age    = demo_information_interviewee_age
  ) |>
  drop_na(frm) |>
  mutate(province_cfmg = interaction(province, cfmg_name)) |>
  droplevels()

if (nlevels(test.data$frm) >= 2) {
  M14 <- clmm(frm ~
                 interviewee_role +
                 interviewee_gender +
                 interviewee_age +
                 (1 | province) +
                 (1 | province_cfmg),
               data = test.data)
  emm.M14      <- emmeans(M14, specs = ~ interviewee_role)
  pairwise.M14 <- pairs(emm.M14, adjust = "Tukey")
} else {
  M14 <- NULL
  cat("Model not fitted: only", nlevels(test.data$frm),
      "level(s) present in frm after filtering.\n")
}
```

```{r}
#| label: M14_frm_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Model results (click to expand)"

if (!is.null(M14)) summary(M14) else cat("Model not fitted.")
```

```{r}
#| label: pairwise_frm_individual
#| echo: false
#| warning: false
#| message: false
#| error:   false
#| output-fold: true
#| output-summary: "Pairwise comparisons (click to expand)"

if (!is.null(M14)) summary(pairwise.M14) else cat("Model not fitted.")
```

### Restoration

We asked CFMG executives whether natural forest restoration is practiced inside the community forest. We restrict the responses for this question to CFMG Executives, because this is the groups that should be best informed about the activities of the CFMG (and the differences in responses among groups to this question are no of substantive interest).

```{r}
#| label: fig-natural_forest_restoration
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFMGs undertake natural forest restoration in the community forest by province. The length of the bars represents the number of CFMGs either undertaking restoration (Yes) or not (No). The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(impact_on_sfm_natural_forest_restoration, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(
    y    = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

Natural forest restoration is being practiced across approximately half of the CFs in our sample, mostly in Muchinga, Central and Northern provinces. We also asked CFMG executives about the specific restoration methods applied inside their CF. The majority were conserving natural regeneration. Only one CFMG was involved in planting native species and removal of invasives was not practiced.

```{r}
#| label: fig-restoration_methods
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Restoration methods applied inside the community forest, as reported by CFMG executives. The length of the bars represents the number of CFMGs either using (Yes) or not using (No) each method. The numbers on the bars represent the actual number of CFMGs in each category."
#| fig-height: 4

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    "Protecting natural regeneration" = impact_on_sfm_restoration_methods_applied_protection,
    "Planting indigenous"             = impact_on_sfm_restoration_methods_applied_planting_with_indigineous_species,
    "Invasives removal"               = impact_on_sfm_restoration_methods_applied_removal
  ) |>
  pivot_longer(everything(), names_to = "method", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    method   = factor(method, levels = c("Invasives removal",
                                         "Planting indigenous",
                                         "Protecting natural regeneration"))
  ) |>
  count(method, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = method, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

### Plantations

We also asked CFMG executives whether any plantations has been established inside the community forest.

```{r}
#| label: fig-plantation_establishment
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether CFMGs have established a plantation inside the community forest, by province. The length of the bars represents the number of CFMGs either with (Yes) or without (No) a plantation. The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(impact_on_sfm_planted_forest_establishment, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We also asked executives to estimate the number of hectares planted in the plantation. A small number (~9% of the sample) had established tree plantation inside their CF. However, the hectares planted were small (@fig-hectares_planted).

```{r}
#| label: fig-hectares_planted
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Estimated area of plantation established inside the community forest, as reported by CFMG executives (n = 6 CFMGs; total area = 20.5 ha). The horizontal bar is the median and the red dot is the mean. Species planted include *Gliricidia sepium* (n = 3 CFMGs), *Pinus* spp. (n = 2), *Acacia* spp. (n = 1), *Faidherbia albidia* (n = 1), and miombo species (n = 1)."
#| fig-width: 4
#| fig-height: 5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  filter(!is.na(impact_on_sfm_hectares_planted)) |>
  ggplot(aes(x = "", y = impact_on_sfm_hectares_planted)) +
  geom_boxplot(width = 0.4, outlier.shape = NA) +
  geom_jitter(width = 0.1, alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", size = 3, col = "red") +
  theme_bw(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x  = element_text(size = rel(1.14)),
    axis.text.y  = element_text(size = rel(1.14)),
    axis.title.y = element_text(size = rel(1.29))
  ) +
  xlab("") +
  ylab("Hectares planted")
```

### Agroforestry

Last, we asked CFMG executives whether CFM has involved introducing agroforestry practices. In this context, agroforestry is implemented on agricultural land as a means for achieving sustainable intensification and reducing pressure on forests. Around 40% of CFMGs said that agroforestry had been introduced under CFM.

```{r}
#| label: fig-yn_agroforestry
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Whether agroforestry is employed to enhance SFM, as reported by CFMG executives. The length of the bars represents the number of CFMGs reporting either a positive impact (Yes) or no impact (No). The numbers on the bars represent the actual number of CFMGs in each category."

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  mutate(
    response = substr(impact_on_sfm_yn_agroforestry, 1, 1),
    response = case_match(
      response,
      "y" ~ "Yes",
      "n" ~ "No"
    )
  ) |>
  filter(!is.na(response)) |>
  count(province, response) |>
  mutate(
    n_signed = if_else(response == "No", -n, n),
    response  = factor(response, levels = c("No", "Yes"))
  ) |>
  ggplot(aes(y = reorder(province, n_signed), x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We also asked CFMG executives about the types of agroforestry practices present in the community.

```{r}
#| label: fig-agroforestry_types
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Types of agroforestry practices reported by CFMG executives (n = 30 CFMGs). The length of the bars represents the number of CFMGs either reporting (Yes) or not reporting (No) each practice."
#| fig-height: 4.5

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    "Silvopasture"               = impact_on_sfm_types_agroforestry_practices_Silvopasture,
    "Windbreaks"                 = impact_on_sfm_types_agroforestry_practices_wind_breaks,
    "Home gardens"               = impact_on_sfm_types_agroforestry_practices_home_gardens,
    "Boundary planting/woodlots" = impact_on_sfm_types_agroforestry_practices_boundary_planting_woodlots,
    "Intercropping"              = impact_on_sfm_types_agroforestry_practices_intercropping
  ) |>
  pivot_longer(everything(), names_to = "practice", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    practice = factor(practice, levels = c(
      "Silvopasture", "Windbreaks", "Home gardens",
      "Boundary planting/woodlots", "Intercropping"
    ))
  ) |>
  count(practice, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = practice, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We also asked CFMG executives about the species planted in agroforestry systems. Among the 30 CFMGs practicing agroforestry most were employing intercropping with leguminous shrubs for soil heal management, but planting boundary trees or woodlots and homegardens were also popular interventions. 

Among the species used (@fig-agroforestry_species), we can see that *Gliricidia* and *Faidherbia*, leguminous shrubs and trees, respectively, were the most popular.

```{r}
#| label: fig-agroforestry_species
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Agroforestry species planted, as reported by CFMG executives (n = 30 CFMGs). The length of the bars represents the number of CFMGs either planting (Yes) or not planting (No) each species."
#| fig-height: 4

cfmg |>
  filter(interview_type == "cfmg_executive") |>
  select(
    "Bamboo"             = impact_on_sfm_agroforestry_species_planted_bamboo,
    "Moringa"            = impact_on_sfm_agroforestry_species_planted_moringa,
    "Exotic fruit trees" = impact_on_sfm_agroforestry_species_planted_exotic_fruit_tree,
    "Faidherbia albidia" = impact_on_sfm_agroforestry_species_planted_faidherbia_albidia,
    "Gliricidia sepium"  = impact_on_sfm_agroforestry_species_planted_gliricidia_sepium
  ) |>
  pivot_longer(everything(), names_to = "species", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    response = case_match(as.character(value), "1" ~ "Yes", "0" ~ "No"),
    response = factor(response, levels = c("No", "Yes")),
    species  = factor(species, levels = c(
      "Bamboo", "Moringa", "Exotic fruit trees",
      "Faidherbia albidia", "Gliricidia sepium"
    ))
  ) |>
  count(species, response) |>
  mutate(n_signed = if_else(response == "No", -n, n)) |>
  ggplot(aes(y = species, x = n_signed, fill = response)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.4) +
  geom_vline(xintercept = 0, colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = n,
        hjust = if_else(response == "No", 1.4, -0.4)),
    size = 3.5, fontface = "bold", colour = "grey25"
  ) +
  scale_x_continuous(
    labels = ~abs(.x),
    expand = expansion(mult = 0.18),
    name   = "Number of CFMGs"
  ) +
  scale_fill_manual(
    values = c("No" = "#E07B54", "Yes" = "#2A9D8F"),
    guide  = guide_legend(reverse = FALSE)
  ) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.y        = element_text(size = rel(1.14), colour = "grey20",
                                      face  = "italic"),
    axis.text.x        = element_text(size = rel(1.14), colour = "grey20"),
    axis.title.x       = element_text(size = rel(1), colour = "grey40"),
    legend.position    = "bottom",
    legend.text        = element_text(size = rel(1)),
    plot.margin        = margin(10, 25, 10, 10)
  )
```

## Discussion

All the indicators of SFM point to a situation that is improving under CFM, including deforestation, charcoal production, fire frequency, wildlife abundance and natural resource management. Interestingly, there does not appear to be any indication of leakage for most CFs. Most FDG groups indicated that both deforestation and charcoal production were also decreasing outside their CF. There were some differences in responses among the different FDG groups, with community female FDGs tending to have a more positive view of the impact of CFM compared to the other groups. This is interesting and may indicate that under CFM females have more of a voice in forest and forest resource management. Traditionally, women are highly dependent on forest resources, but usually have little say in forest management decisions. It would be worth following up this study to investigate what is driving this positive take on CFM among community women.

Among the individual interviewees, NGO representatives and DFOs tended to have a somewhat more conservative view with respect to positive changes, as compared to individuals from the communities. This may reflect that these professionals have a broader perspective from which to understand changes in forest and forest resource management. However, it could also arise from the fact that they are not as intimately involved in the forest management and hence less aware of the changes. Ultimately, monitoring over a longer period will enable better understanding of how CFM is changing forest management.

Overall, CFM appears to be engendering a change towards improved SFM. However, it is important to note that a small number of FDG groups and individual interviewees suggested the situation was worsening in their CFs. For the success of CFM in Zambia, it will be important to identify CFs that are doing less well, diagnose the causes and implement changes.

Some CFMGs were practicing forest restoration, mostly through protecting natural regeneration, and a few were establishing plantations. However, these activities were not widespread among CFMGs. Restoration - especially through protection of natural regeneration - should integral to CFM and is probably occurring widely simply through improvements in forest protection. As this is an important indicator of SFM and there are government targets to restore 2,000,000 ha by 2030, the FD should integrate restoration targets into forest management plans for CFMGs and monitor outcomes. Agroforestry was also not widely practiced by CFMGs, but is recognised tool for restoration of degraded agricultural land, and can reduce pressure on forests. Plantations are also an important opportunity for CFM. Plantations can supply local building and fuelwood needs, thereby reducing pressure on natural forests. Through sustainable charcoal programmes, they can also supply energy for urban dwellers and again reduce pressure on natural forests. Last, Zambia has an acute shortage of high quality timber for telegraph poles and construction and imports large amounts of wood. A focus on developing plantations through CFM would go some way to meeting these needs, as well as generating a valuable additional source of income for poor rural communities.

In conclusion, our data suggest that CFM in Zambia is driving substantial improvements in SFM, reducing deforestation, charcoal production, and fire frequency, increasing wildlife populations, and improving natural resource management. However, it is important to note that these changes were not ubiquitous across all CFs. While some CFMGs are practising restoration and establishing plantations, there is a substantial opportunity to expand the role of CFM in meeting national restoration and plantation targets.  

