---
title: "Discussion & Recommentations"
author: "Rhett D Harrison"
format:
        html:
                css: custom.css
        pdf: default
filters:
- collapse-output
editor: visual
---

```{r}
#| label: load_data_and_libraries
#| include: false
#| warning: false
#| message: false
#| error:   false

library(tidyverse)
library(tidytext)
library(ggwordcloud)

load("cfmg.RData")
```

## Discussion

### Feedback from the communities
#### What is working in CFM

We asked the community FGD groups to identify aspects of CFM that they feel are working well in their communities (@fig-conc_working). The most frequently mentioned reflect different aspects of sustainable forest management. Clearly, most communities do feel that CFM has led to improvements in forest management, reducing deforestation and enhancing wildlife conservation. Also mentioned, but less frequently, were indicators of good governance and improved livelihoods. Evidently, CFM is working well for some communities.

```{r}
#| label: fig-conc_working
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Aspects of CFM that are working well, as identified by community female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each aspect."
#| fig-width: 10
#| fig-height: 7

prefix_w <- "opportunitiesb_for_improved_cfm_implementation_cfm_working_aspects_"

working_labels <- c(
  "effective_management_of_forest_resources_e_g_sustainable_harvesting_restoration" = "Forest management",
  "reduction_in_deforestation_and_forest_degradation"                               = "Reduced deforestation",
  "increased_wildlife_conservation_and_biodiversity"                                = "Wildlife conservation",
  "equitable_benefit_sharing_among_community_members"                               = "Equitable benefits",
  "active_participation_of_women_youth_and_marginalized_groups_in_cfm_activities"  = "Inclusive participation",
  "improved_livelihoods_and_income_generating_opportunities_through_cfm_activities" = "Improved livelihoods",
  "successful_partnerships_with_private_enterprises_or_external_institutions"       = "Partnerships",
  "successful_partnerships_with_traditional_leaders"                                = "Traditional leaders",
  "improved_infrastructure"                                                          = "Improved infrastructure",
  "effective_monitoring_and_reporting_of_cfm_outcomes_and_impacts"                 = "Effective monitoring",
  "transparent_and_fair_decision_making_processes_within_the_community"            = "Transparent decision making",
  "strong_community_awareness_and_involvement_in_cfm_activities"                   = "Community awareness"
)

full_w_vars <- paste0(prefix_w, names(working_labels))

# Frequency counts from binary checkbox variables
working_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_w_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = working_labels[sub(prefix_w, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

# Word frequencies from open-ended comments
comment_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(comment = opportunitiesb_for_improved_cfm_implementation_cfm_working_aspects_comment) |>
  drop_na(comment) |>
  unnest_tokens(word, comment) |>
  anti_join(stop_words, by = "word") |>
  filter(nchar(word) > 2, !word %in% c("cfm", "cfmg", "forest", "community", "people")) |>
  count(word, name = "freq") |>
  filter(freq > 1)

# Combine checkbox labels and comment words
working_all <- bind_rows(working_freq, comment_freq) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(working_all, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

#### Areas for improvement

We also asked the community FGD groups to identify the aspects of CFM that they feel need improvement in their communities (@fig-conc_improvement). The diversity of responses was greater and most can be taken as indicators of poor govenance, such as 'community awareness', 'transparent decision-making' and 'inclusive participation'. Economic indictors also featured prominently, such as 'improved livelihoods' and 'partnerships'. Indicators of SFM were also mentioned indicating that the communities in some CFs feel there is scope for improvement there too. Clearly, there is scope for improvement in CFM, especially in governance aspects.

```{r}
#| label: fig-conc_improvement
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Aspects of CFM identified as needing improvement, as identified by community female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each aspect."
#| fig-width: 10
#| fig-height: 7

prefix_i <- "opportunitiesb_for_improved_cfm_implementation_cfm_areas_for_improvement_"

improvement_labels <- c(
  "effective_management_of_forest_resources"                       = "Forest management",
  "reduction_in_deforestation_and_forest_degradation"             = "Reduced deforestation",
  "increased_wildlife_conservation_and_biodiversity"              = "Wildlife conservation",
  "equitable_benefit_sharing_among_community_members"             = "Equitable benefits",
  "active_participation_of_women_youth_and_marginalized_groups"   = "Inclusive participation",
  "improved_livelihoods_and_income_generating_opportunities"      = "Improved livelihoods",
  "partnerships_with_private_enterprises_or_external_institutions" = "Partnerships",
  "effective_monitoring_and_reporting_of_cfm_outcomes"            = "Effective monitoring",
  "transparent_and_fair_decision_making_processes"                = "Transparent decision making",
  "community_awareness_and_involvement_in_cfm_activities"         = "Community awareness",
  "capacity_building"                                             = "Capacity building",
  "infrastructure"                                                = "Infrastructure",
  "alternative_livelihoods"                                       = "Alternative livelihoods",
  "transport"                                                     = "Transport",
  "governanace_leadership_institutional_strengthening"            = "Governance & leadership",
  "external_funding"                                              = "External funding",
  "improved_water_security"                                       = "Water security",
  "human_wildlife_conflict"                                       = "Human-wildlife conflict"
)

full_i_vars <- paste0(prefix_i, names(improvement_labels))

# Frequency counts from binary checkbox variables
improvement_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_i_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = improvement_labels[sub(prefix_i, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

# Word frequencies from open-ended comments
improvement_comment_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(comment = opportunitiesb_for_improved_cfm_implementation_cfm_areas_for_improvement_comment) |>
  drop_na(comment) |>
  unnest_tokens(word, comment) |>
  anti_join(stop_words, by = "word") |>
  mutate(word = case_match(word, "hfos" ~ "HFOs", .default = word)) |>
  filter(nchar(word) > 2,
         !word %in% c("cfm", "cfmg", "forest", "community", "people",
                      "products", "visits", "water", "activities", "provide",
                      "visitation", "production", "increase", "improved",
                      "regular", "change", "continue")) |>
  count(word, name = "freq") |>
  filter(freq > 1)

# Combine checkbox labels and comment words
improvement_all <- bind_rows(improvement_freq, improvement_comment_freq) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(improvement_all, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#F4A582", high = "#8B0000") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

#### External challenges

We asked the community FGD groups to identify the external challenges facing CFM in their communities (@fig-conc_external_challenges). Perhaps unsurprisingly, here the respsonses focus on development issues, such as 'financial constraints', 'poor infrastructure', 'limited market access' and 'insufficient technical expertise'. 

```{r}
#| label: fig-conc_external_challenges
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "External challenges facing CFM, as identified by community female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each challenge."
#| fig-width: 10
#| fig-height: 10

prefix_ec <- "opportunitiesb_for_improved_cfm_implementation_cfm_external_challenges_"

external_challenge_labels <- c(
  "government_policies_and_regulations"                      = "Policy/regulatory barriers",
  "lack_of_external_funding"                                 = "Financial constraints",
  "limited_market_access_for_forest_products"                = "Limited market access",
  "conflicts_with_neighboring_communities"                   = "Conflicts with neighbours",
  "conflicts_within_community"                               = "Internal conflicts",
  "conflicts_with_neighboring_with_partners"                 = "Conflicts with partners",
  "insufficient_technical_expertise"                         = "Insufficient technical expertise",
  "weak_enforcement_of_laws_and_regulations"                 = "Weak law enforcement",
  "political_instability_or_government_changes"              = "Political instability",
  "poor_infrastructure"                                      = "Poor infrastructure",
  "natural_disasters"                                        = "Natural disasters",
  "limited_support_from_govt"                                = "Limited government support",
  "limited_support_from_tradtional_leaders"                  = "Traditional leader challenges",
  "water_social_services_and_basic_needs"                    = "Limited government support",
  "human_wildlife_conflicts"                                 = "Human-wildlife conflict",
  "transport_and_communication"                              = "Transport/communication",
  "chiefdom_succession_disputes"                             = "Chiefdom succession disputes",
  "awareness_information_and_capacity_gaps"                  = "Awareness/capacity gaps",
  "cultural_or_social_resistance_to_cfm_initiatives"        = "Cultural/social resistance",
  "traditional_leadership_and_governance_interference"       = "Traditional leader challenges",
  "social_economic_and_environmnetal_stressors"              = "Socioeconomic stressors",
  "policy_and_regulatory_barriers"                           = "Policy/regulatory barriers",
  "enchroachment_illegal_activities_and_extternal_pressure"  = "Encroachment/illegal activities",
  "partner_related_challenges_and_carbon_market_issues"      = "Partner/carbon market issues",
  "financial_material_and_logistical_constraints"            = "Financial constraints"
)

full_ec_vars <- paste0(prefix_ec, names(external_challenge_labels))

# Frequency counts from binary checkbox variables
external_challenge_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_ec_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = external_challenge_labels[sub(prefix_ec, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

external_challenge_all <- external_challenge_freq |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(external_challenge_all, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#F4A582", high = "#8B0000") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

#### Monitoring methods

We asked the FGD groups to identify the methods they recommend are used monitor CFM progress and outcomes in their communities (@fig-conc_monitoring). While a large number of approaches were mentioned, the community groups overwhelmingly identified community meetings and field visits/patrols as being the preferred approach for monitoring CFM outcomes. However, it is interesting to note that female groups were far more likely to mention community meetings and much less likely to mention field visits/patrols, than the male and youth groups. This likely reflects the gender biases in participation in field visits and patrolling, which from earlier data are almost exclusively a male prerogative.

```{r}
#| label: fig-conc_monitoring
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Methods used to monitor CFM progress and outcomes, as reported by community female, male, and youth FGDs. Bars show the number of FGDs reporting each method."
#| fig-width: 10
#| fig-height: 7

prefix_m <- "opportunitiesb_for_improved_cfm_implementation_cfm_monitoring_methods_"

monitoring_labels <- c(
  "community_meetings_to_review_progress_and_challenges"                                  = "Community meetings",
  "field_visits_or_patrols_to_assess_forest_conditions_and_resource_use"                 = "Field visits/patrols",
  "data_collection_and_surveys_on_forest_health_biodiversity_livelihoods"                = "Data collection/surveys",
  "financial_reporting_and_audits_to_track_resource_allocation_and_benefits"             = "Financial reporting/audits",
  "independent_evaluations_by_external_bodies_ngo_s_government_agencies"                 = "Independent evaluations",
  "participatory_monitoring_involving_community_members_in_data_collection_and_analysis" = "Participatory monitoring",
  "use_of_technology_or_tools_satellite_images_gis_mobile_apps"                          = "Technology/tools",
  "feedback_or_grievance_mechanisms_to_capture_community_concerns"                       = "Feedback/grievance mechanisms"
)

full_m_vars <- paste0(prefix_m, names(monitoring_labels))

cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  mutate(group = factor(
    case_match(
      interview_type,
      "cfmg_female" ~ "Female",
      "cfmg_male"   ~ "Male",
      "cfmg_youth"  ~ "Youth"
    ),
    levels = c("Youth", "Male", "Female")
  )) |>
  select(group, all_of(full_m_vars)) |>
  pivot_longer(all_of(full_m_vars), names_to = "variable", values_to = "val") |>
  filter(val == 1) |>
  mutate(method = factor(
    monitoring_labels[sub(prefix_m, "", variable, fixed = TRUE)],
    levels = rev(monitoring_labels)
  )) |>
  count(group, method) |>
  complete(group, method, fill = list(n = 0)) |>
  ggplot(aes(y = method, x = n, fill = group)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65, colour = "white") +
  geom_text(
    aes(label = ifelse(n > 0, n, "")),
    position = position_dodge(width = 0.75),
    hjust = -0.3, size = 3.5, colour = "grey30"
  ) +
  scale_fill_manual(
    values = c(
      "Female" = "#2A9D8F",
      "Male"   = "#57B8B0",
      "Youth"  = "#E9C46A"
    ),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.15)),
    name   = "Number of FGDs"
  ) +
  scale_y_discrete(labels = \(x) str_wrap(x, width = 25), expand = expansion(add = 0.6)) +
  labs(y = NULL, fill = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.line.x.bottom = element_line(colour = "grey30", linewidth = 0.5),
    axis.text.x        = element_text(colour = "grey20"),
    axis.text.y        = element_text(size = 14, colour = "grey20"),
    legend.position    = "bottom",
    plot.margin        = margin(10, 25, 10, 10)
  )
```

We also asked the community FGD groups how monitoring could be improved (@fig-conc_monitoring_improvement). Unsurpristingly, increased funding and more training featured prominently. However, it was interesting to note that digital tools and community feedback sessions were also mentioned frequently.

```{r}
#| label: fig-conc_monitoring_improvement
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Suggested improvements to CFM monitoring, as identified by community female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each suggestion."
#| fig-width: 10
#| fig-height: 7

prefix_mi <- "opportunitiesb_for_improved_cfm_implementation_cfm_monitoring_improvement_"

monitoring_imp_labels <- c(
  "provide_more_training_for_cfm_group_members_on_monitoring_techniques"                   = "More training",
  "introduce_digital_tools_for_real_time_data_collection_and_reporting"                    = "Digital tools",
  "increase_funding_for_monitoring_activities"                                              = "Increased funding",
  "establish_regular_community_feedback_sessions_to_discuss_monitoring_results"            = "Community feedback sessions",
  "develop_standardized_reporting_templates_for_consistent_data_collection"                = "Standardised reporting",
  "improve_transparency_by_sharing_monitoring_reports_with_the_community"                  = "Improved transparency",
  "increase_participation_of_women_youth_and_marginalized_groups_in_monitoring_activities" = "Inclusive participation",
  "increase_hfos_scouts_guards"                                                             = "More HFOs/scouts/guards",
  "improve_transport_communication"                                                         = "Better transport/communication",
  "ppe"                                                                                     = "PPE",
  "increase_patrols"                                                                        = "More patrols",
  "improve_security_for_hfos"                                                               = "HFO security",
  "collaboration"                                                                           = "Collaboration"
)

full_mi_vars <- paste0(prefix_mi, names(monitoring_imp_labels))

# Frequency counts from binary checkbox variables
monitoring_imp_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_mi_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = monitoring_imp_labels[sub(prefix_mi, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

# Word frequencies from open-ended comments
monitoring_imp_comment_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(comment = opportunitiesb_for_improved_cfm_implementation_cfm_monitoring_improvement_comment) |>
  drop_na(comment) |>
  unnest_tokens(word, comment) |>
  anti_join(stop_words, by = "word") |>
  mutate(word = case_match(word,
    "hfos"      ~ "More HFOs/scouts/guards",
    "hfo"       ~ "More HFOs/scouts/guards",
    "scouts"    ~ "More HFOs/scouts/guards",
    "guards"    ~ "More HFOs/scouts/guards",
    "employ"    ~ "more employment",
    "employing" ~ "more employment",
    "ppe"       ~ "PPE",
    .default = word
  )) |>
  filter(nchar(word) > 2, !word %in% c("cfm", "cfmg", "forest", "community", "people",
                                        "monitoring", "improve", "improvement")) |>
  count(word, name = "freq") |>
  filter(freq > 1)

# Combine checkbox labels and comment words
monitoring_imp_all <- bind_rows(monitoring_imp_freq, monitoring_imp_comment_freq) |>
  group_by(word) |>
  summarise(freq = sum(freq), .groups = "drop")

set.seed(42)
ggplot(monitoring_imp_all, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#74ADCB", high = "#1B618C") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

#### Training needs

We asked community FGD groups to identify the types of training most needed to improve CFM in their communities (@fig-conc_training). The responses appear to reflect the activities they are most involved with, such as fire management and conservation. However, it was interesting to see others also mentioned frequently, such as forest law and policy, agroforestry, nursery management, climate-smart agriculture and sustainable harvesting. These training needs reflect the interests of the communities and so provide a valuable insight.

```{r}
#| label: fig-conc_training
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Training needs identified by community female, male, and youth FGDs. Word size is proportional to the number of FGDs reporting each training type."
#| fig-width: 10
#| fig-height: 7

prefix_tr <- "opportunitiesb_for_improved_cfm_implementation_cfmg_training_type_"

training_labels <- c(
  "forest_conservation_and_biodiversity"                                    = "Conservation",
  "agroforestry"                                                            = "Agroforestry",
  "tree_planting_and_nursery_management"                                    = "Nursery management",
  "forest_law_and_policy"                                                   = "Forest law & policy",
  "governance_and_leadership"                                               = "Governance & leadership",
  "conflict_resolution_and_community_engagement"                            = "Conflict resolution",
  "climate_smart_agriculture"                                               = "Climate-smart agriculture",
  "livestock_management"                                                    = "Livestock management",
  "aquaculture"                                                             = "Aquaculture",
  "climate_change_adaptation_and_mitigation"                               = "Climate change adaptation",
  "sustainable_forest_product_harvesting_e_g_honey_mushrooms_firewood"     = "Sustainable harvesting",
  "financial_management_and_record_keeping"                                 = "Financial management",
  "business_development_and_entrepreneurship"                               = "Business development",
  "marketing_and_value_addition_of_forest_products"                        = "Marketing & value addition",
  "fire_management_and_prevention"                                          = "Fire management"
)

full_tr_vars <- paste0(prefix_tr, names(training_labels))

training_freq <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(all_of(full_tr_vars)) |>
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "freq") |>
  mutate(word = training_labels[sub(prefix_tr, "", variable, fixed = TRUE)]) |>
  filter(freq > 0) |>
  select(word, freq)

set.seed(42)
ggplot(training_freq, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 22)) +
  scale_colour_gradient(low = "#74ADCB", high = "#1B618C") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

### Comments from the enumerators

Throughout the survey the responses are those of the interviewees. However, after each FGD session, enumerators recorded their general observations in an open-ended comment field. We coded the substantive comments thematically and these are shown in @fig-conc_comments, with size proportional to the number of FGD sessions for which that theme was noted. This is a very informative figure as a general take-home of the issues arising from the implementation of CFM in Zambia. Most prominently featured is 'capacity building', recognising that many communities have limited capacity and are struggling to implement CFM currently. Also, prominent were 're-engagement needed', 'CFMG dormant' and 'partner challenges', reflecting inadequate support from the forest department and other supporting organisations, 'youth marginalisation' potentially reflecting elite capture (which was also mentioned directly) and indicating that benefits may not be being shared equitably, and 'leadership challenges' suggesting the CFMG executive may not be functioning properly.

```{r}
#| label: fig-conc_comments
#| code-fold: true
#| warning: false
#| message: false
#| error:   false
#| fig-cap: "Key themes from enumerator observations recorded after each FGD session. Word size is proportional to the number of FGD sessions for which that theme was noted."
#| fig-width: 10
#| fig-height: 8

cfmg_comments <- cfmg |>
  filter(interview_type %in% c("cfmg_female", "cfmg_male", "cfmg_youth")) |>
  select(comment) |>
  drop_na(comment) |>
  filter(!str_detect(comment, regex(
    "^(none|nil|no comment|no comments|replacement|mixed|nil\\.?)$|^[A-Z][a-z]+(\\s[A-Z]?[a-z]*){0,4}\\s?(VAG|CFMG)$|^(mutimba|from the district)",
    ignore_case = TRUE
  )))

theme_patterns <- c(
  "Youth marginalization"           = "youth.{0,30}sidelin|sidelin.{0,30}youth|youth.{0,20}not.{0,20}includ|youth.{0,20}not.{0,20}involv|youth.{0,20}not.{0,20}aware|youth.{0,10}no voice|youth.{0,15}intimidat|youth.{0,20}left out|youth.{0,20}not engaged|youth.{0,20}no information|youth.{0,30}not fully|youth.{0,20}feel.*sidelin",
  "Information gaps"                = "information gap|not aware of|lack.{0,15}information|information.{0,20}only.{0,15}leader|information.{0,20}only.{0,15}executive|information.{0,20}with.{0,15}leader|information.{0,20}with.{0,15}executive|not previewed|information.{0,15}not avail|not availed.{0,20}information|know what.{0,15}happening|huge information",
  "Capacity building"               = "capacity build",
  "Enterprise development"          = "enterprise develop|develop.{0,15}enterprise|no.{0,10}enterprise.{0,10}running|enterprises.{0,10}running|no forest enterprise|enterprise.{0,10}not.{0,10}running",
  "Re-engagement needed"            = "re.engag|reengag|rengag|reorientation",
  "Benefits not realized"           = "not.{0,10}benefit|yet to.{0,20}benefit|not reaping|not seen.{0,15}benefit|yet to produce.*benefit|not.*started.*benefit|lack.*benefit|no.*benefit",
  "CFMG dormant"                    = "dormant|not been functioning|not been active|not been working|not function since|not.{0,5}started|not fully function|not working.{0,10}since|not working.{0,10}past|not functional since|has not been.{0,10}function|not.{0,10}active since|less active|not.*active.*begin",
  "Women marginalized"              = "women.{0,20}unfairly|women.{0,20}not allowed|women.{0,30}low.{0,15}participat|participat.{0,20}women.{0,20}low|women.{0,15}lay.{0,10}off|women.{0,20}not.{0,20}participat|level.{0,20}involv.{0,20}women.{0,20}low|women.{0,10}sidelin|women.{0,15}not.{0,10}includ|women.{0,15}not.{0,10}aware",
  "Elite capture"                   = "headman.{0,20}benefit|monopol|presiding.{0,10}everything|executive.{0,10}families|one person.{0,10}community|scared.{0,10}executive|intimidat.{0,10}executive",
  "Lack of transparency"            = "not transparent|information.{0,15}only.{0,10}executive|cfmg.{0,10}not transparent|lack.{0,10}transparency|mistrust",
  "Leadership challenges"           = "change.{0,10}leadership|review.{0,10}constitution|constitution.{0,10}review|leadership.{0,10}change|leadership.{0,10}not.{0,10}driv|not.{0,15}engage.{0,15}community.{0,15}update|change.{0,10}executive|leadership.{0,10}affects|change.*in.*leadership",
  "Partner challenges"              = "partner.{0,20}pulled out|partner.{0,20}not fulfil|not fulfilling.{0,10}promise|pulled out|partners.{0,20}not.{0,20}fulfill|partner.{0,20}left|partner.{0,20}promise",
  "User groups non-functional"      = "user group.{0,5}non functional|user group.{0,5}not functional|non functional|user groups.{0,5}non|user groups.{0,5}none",
  "Carbon credit delays"            = "carbon.{0,20}not.{0,15}receiv|not.{0,15}receiv.{0,20}carbon|waiting.{0,10}carbon|carbon.{0,10}credit.{0,5}yet|yet.{0,10}carbon|carbon.{0,10}payment|waiting.{0,10}bcp|waiting.{0,10}kabompo",
  "Traditional leader interference" = "traditional leader.{0,10}interfere|chiefdom.{0,10}succession|traditional.{0,10}leader.{0,10}wrangl|high influence.{0,10}traditional",
  "HFO challenges"                  = "hfo.{0,10}not.{0,10}equip|hfo.{0,10}no.{0,10}protective|hfo.{0,10}stop.{0,5}patrol|hfos.{0,10}protective|hfo.{0,8}not.{0,8}uniform|hfo.{0,5}not.{0,5}train|dissolved.{0,5}hfo",
  "High illiteracy"                 = "illiteracy|illiterate",
  "Market access"                   = "lack.{0,5}market|market.{0,5}challenge|no market|lack of market",
  "CRB/CFMG overlap"               = "crb.{0,5}cfmg|cfmg.{0,5}crb|operating.{0,5}crb"
)

theme_counts <- map_dfr(names(theme_patterns), function(theme) {
  n <- sum(str_detect(cfmg_comments$comment, regex(theme_patterns[[theme]], ignore_case = TRUE)))
  tibble(word = theme, freq = n)
}) |>
  filter(freq > 0) |>
  arrange(desc(freq))

set.seed(42)
ggplot(theme_counts, aes(label = word, size = freq, colour = freq)) +
  geom_text_wordcloud(eccentricity = 0.65, seed = 42) +
  scale_size(range = c(3, 18)) +
  scale_colour_gradient(low = "#6ABFB8", high = "#1A6B63") +
  theme_minimal(base_size = 14) +
  theme(plot.margin = margin(0, 0, 0, 0))
```

### Synopsis of the main findings

*_to_do_*

## Recommendations

Derived from our experiences with this project we make the following recommendations. This is not an exhaustitave list. Moreover, we recognise that many of these this may be in hand to some degree (other to a lesser extent). So this list to be seen as affirming and support these ongoing efforts.

### Governance

- DFOs should visit each CFMG once per quarter and support the CFMGs in developing and implementing their workplans. The should be applying the "Checklists – CFM Step 7: Joint Monitoring, Evaluaon and Lesson Learning". KPIs for DFOs should reflect the delivery of CFM support.

- Budgets need to be mage available for regular DFO visits to CFMGs. These could come from the government carbon funds raised, if a proportion of these are ear-marked to support CFM.

- Increase standardisation of important CFMG processes including; (i) elections of CFMG and executive positions; (ii) tenure of CFMG members and executive position; (iii) term limits; (iv) bank accounts, receipt books, HFO payments and gratuities for CFMG activities; (v) reporting requirements, including presentation of reports to community members;  (vi) grievance and redress mechanisms; (vii) carbon benefit sharing mechanisms; and (viii) other benefit sharing mechanisms. Many of these are already in the CFM agreement and any that are not should probably be added. Guidelines could be added for others, such as different permitting arrangements and benefit sharing arrangements, which always require some local tailoring.

- Require annual implementation plans and budgets and require that the DFOs and CFMG executives sign off on these and submit them to the FD database.

- Require registration and permitting of all forest user-groups, even if these are community groups and there is not intention to charge a permit fee. Formalising the forest use and agreeing on management protocols, such as start dates and off-take levels, is essential for SFM.

- Transition the current CFM database into a digital management dashboard for CFM monitoring that can be integrated with digital data collection tools and file uploads. Ideally, it would have an associated App that would facilitate integrated recording of data on common processes, such as CFMG meetings and elections, permits, workplans, budgets, annual reports, and so on.

### Sustainble forest management

- Define restoration areas within the CF in forest management plans and assess area of (i) natural regeneration, (ii) planting with indigenous species, and (iii) planting with exotic species (agroforestry plantings should not be within the CF). Any areas that are severely degraded that are being protected should be considered potential restoration sites and if only protection is being applied this should be considered natural regeneration. This information can be used to 

- Establish community based monitoring systems for NTFPs. These systems should be developed centrally and standardised, so that data can be entered into a FD managed database but with dashboard for community viewing.

- Consider a small plantation programme for CFs for timber or sustainable charcoal production

### Economic transformation

- Require the production of CF business plans

- Establish independent trade association for forest product producers (note: could be Wildlife Producers Association, if they were willing to branch out from game meat and represent small producers)

- Establish national knowledge hub on forest product management, value addition, and marketing

- Require / encourage carbon traders and NGOs working with CFMGs to engage in a minimum of three forest product value chains in each CF

- Develop programme for community-based eco-tourism through CFs and based on national heritage sites

-- Consider promoting Forest Stewardship Certification (FSC) for CFMGs or groups of CFMGs. FSC certification can improve carbon credit values by up to 100% and open international markets for NTFPs.

### Equity and inclusion

- Require identification of opportunity costs of CF and the community groups most affected

- Encourage development of women's forest product enterprises

- Require consideration of vulnerable groups in benefit sharing arrangements
